### 2022-07-11 manuscript/main.tex
minor updates + backing up manuscript draft

\documentclass[letterpaper,man,natbib]{apa6}  %Leave this stuff alone

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage[export]{adjustbox}

% dashed line (https://tex.stackexchange.com/questions/112343/beautiful-table-samples)
\usepackage{array}
\usepackage{arydshln}
\setlength\dashlinedash{0.2pt}
\setlength\dashlinegap{1.5pt}
\setlength\arrayrulewidth{0.3pt}

\title{Bifactor models of the `Maltreatment and Abuse Chronology of Exposure' (MACE) Scale}
\shorttitle{Bifactor models of MACE}
\author{Samuel Zorowitz$^1$, Lauri Tuominen$^{2}$}
\affiliation{$^1$Princeton Neuroscience Institute, Princeton University, USA\\$^2$The Royal’s Institute of Mental Health Research, University of Ottawa, Canada}

\abstract{Your abstract here.}

\begin{document}
\maketitle

\section{Introduction}

% "Happy families are all alike; every unhappy family is unhappy in its own way." – Leo Tolstoy.\break

Childhood maltreatment is a common experience for children across the world \citep{stoltenborgh2015prevalence}. It leads to an array of poor outcomes later in life, including the onset of mental illness \citep{macmillan2001childhood, green2010childhood, kessler2010childhood}, adverse effects on cognitive functioning \citep{kavanaugh2017neurocognitive, r2018common, su2019does}, and poorer physical health \citep{wegman2009meta, widom2012prospective, goodwin2004association}. Furthermore, childhood maltreatment predicts worse treatment outcomes across a variety of psychiatric disorders \citep{nanni2012childhood, thomas2019childhood, schuckher2019history}. As such, assessment of childhood maltreatment in a reliable and valid manner is important research goal in order to refine our understanding of the link between maltreatment and health outcomes; augment treatment strategies for victims of childhood maltreatment; and ultimately design interventions to prevent maltreatment in the first place.

Many measures of childhood abuse and maltreatment have been developed \citep{saini2019systematic}. One recently introduced measure, the Maltreatment and Abuse Chronology of Exposure (MACE) scale \citep{teicher2015maltreatment}, has a number of notable advantages. The MACE is a 52-item self-report questionnaire of childhood maltreatment that assays 10 unique types of maltreatment. Six of these (i.e. parental physical maltreatment, parental non-verbal emotional abuse, parental verbal abuse, sexual abuse, physical neglect, emotional neglect) share overlap with other common measures of abuse (e.g. CTQ \citep{bernstein1998childhood} ); the remaining four solicit additional information on peer abuse, and witnessing violence at home. A number of previous scales were reviewed to select candidate items for the MACE. Item response theory and Rasch model were then used to for final selection of items for each 10 subscales. The MACE also measures detailed information on timing of exposure, which is critical for investigating how the onset of abuse shapes trajectories of development. Designed under the cumulative risk framework \citep{evans2013cumulative}, the MACE provides an overall severity score and multiplicity score (number of types of maltreatment experienced) with good test-retest reliability, convergent validity, and external validity.

More recently, there have been calls to move away from cumulative risk approach \citep{evans2013cumulative} --- which focuses on the number of adverse experiences, but does not distinguish between distinct types of environmental experience --- towards a dimensional approach, which emphasizes how specific types of maltreatment uniquely shape psycho-biological developmental trajectories \citep{mclaughlin2016beyond, belsky2012beyond}. As such, research groups have begun to use the MACE in other ways. Some groups have studied the unique contributions of all 10 subscales to psychopathology \citep{schalinski2015type, gerke2018childhood, schalinski2019early}. Following the threat-deprivation dimensional model \citep{mclaughlin2014childhood}, others have made composite indices of threat and deprivation using the MACE items measuring abuse and neglect, respectively \citep{schalinski2018defining, schalinski2019environmental, teicher2018differential}.

However, a critical challenge in measuring childhood maltreatment is the high co-occurrence of multiple types of abuse and neglect experiences; that is, children experiencing maltreatment are likely to be exposed to two or more types of maltreatment \citep{green2010childhood, mclaughlin2012childhood, marques2021risk, mersky2017rethinking}. For example, virtually all severely maltreated children report emotional abuse \citep{mcgee1995measurement}. As a consequence, isolating the unique contribution of a specific type of maltreatment (e.g. emotional neglect) to the risk of developing some outcome (e.g. depression) is difficult insofar that any one type of adversity is likely confounded with general maltreatment. Concretely, this means that researchers attempting to study the effects of particular types of maltreatment using subscale scores (e.g. emotional neglect from the MACE scale) should take pains to ensure that those measures largely reflect a construct unique from general maltreatment, as represented by the total score.

Bifactor models, a group of confirmatory factor analyses, are well-suited for addressing this issue \citep{bornovalova2020appropriate}. Bifactor models specify that the covariance among the items can be accounted for by a latent general factor, reflecting common variance among all items, and a set of latent specific factors, reflecting additional common variance for subsets of items \citep{Reise2012-ql}. Crucially bifactor models can be used to calculate model-based reliability statistics, e.g. the omega-hierarchical ($\omega_h$) and omega-subscale ($\omega_s$) indices, which respectively measure the proportion of variance in the total and subscale scores attributable to the general factor and specific factors \citep{reise2013applying, rodriguez2016evaluating}. Applied to MACE scale, these indices can guide researchers as to whether subscale scores for particular types of maltreatment are interpretable as such, or if they instead largely reflect general maltreatment. 
% summary of paper paragraph tk

% The bifactor model hypothesizes a general factor, onto which all items load, and a series of orthogonal (uncorrelated) skill-specific grouping factors
\section{Methods}

\subsection{Participants}

The participants in the current study (N=1633 total) belonged to one of two samples. The first sample of participants (N=1051) are from the original development of the MACE scale \citep{teicher2015maltreatment}. The full recruitment details for this sample have been reported elsewhere \citep{teicher2015maltreatment}. Briefly, these participants were recruited from the greater Boston area between 2010 and 2013. Inclusion criteria included being medically healthy, unmedicated, and between 18–25 years of age.  This sample will hereafter be referred to as the \textit{original} sample.

The second sample of participants (N=687) were recruited from the Prolific Academic platform (\url{https://www.prolific.co}) to participate in an online experiment in August -- October, 2021. Of these, N=582 were retained for analysis (see Exclusion Criteria below). Participants were eligible for participation if they were 18 years or older and currently resided in the United States or Canada. Participants received monetary compensation for their time (rate: \$12 USD/hr). This study was approved by the Institutional Review Board of the University of Ottawa, and all participants provided informed consent. This sample will hereafter be referred to as the \textit{replication} sample.

In both the original and replication samples, the majority of participants identified as women (original: male = 381, female = 670; replication: male = 258, female = 310, nonbinary or other = 9, rather not say = 5). The original sample is comprised of proportionally more woman than the replication sample ($z = 4.142, p < 0.001$). On average, participants in the original sample were younger than participants in the replication sample ($M_1 = 23.1$ yrs, $M_2 = 30.2$ yrs, $t = -19.313, p < 0.001$). Finally, the majority of participants in both sample identified as Caucasian or white (original: 76.7\%; replication: 75.8\%). The proportion of participants identifying as white were not different across samples ($z = 0.417$, $p = 0.677$).

\subsection{Measures}

All participants completed the 52-item version of the MACE scale \citep{teicher2015maltreatment}. The scale is a retrospective measure of the severity and timing of exposure to 10 types of childhood maltreatment. For each item, participants endorse whether they ever experienced a particular event during childhood (e.g. ``Parents or guardians intentionally pushed, grabbed, shoved, slapped, pinched, punched or kicked you'') and, if so, at what ages (up to age 18). Here we only analyze the binary endorsement responses (Yes = 1, No = 0) and not the chronology data. Six of the 52-items --- evenly divided between the emotional and physical neglect subscales --- are positively valenced (e.g. ``One or more individuals in your family helped you feel important or special'') and reverse-scored. For these items we follow the scoring convention set by \cite{teicher2015maltreatment}: a score of 1 is assigned to a response only in the absence of a positive event for all 18 years of childhood. This is in contrast to the remaining 46 items in the scale, where a score of 1 is assigned to a response if maltreatment or abuse occurred in at least one year of childhood.

In both the original and replication samples, participants completed a number of additional self-report symptom measures. We do not include these in our analyses and we will therefore not discuss these measures further in the main text. For completeness, we report these secondary self-report measures in the Supplementary Materials.

\subsection{Exclusion criteria}

To ensure data quality in the replication sample, we relied on attention checks embedded in the self-report measures to identify and remove participants engaging in careless or insufficient effort responding \citep{zorowitz2021inattentive}. The data from N=64 (9.9\%)  participants in the replication sample were excluded for failing one or more of these attention checks, leaving the data from N=582 participants for analysis.

\subsection{Differential item functioning}

Prior to analysis, we investigated whether it was appropriate to combine response data from the original and replication samples. We first compared maltreatment endorsement rates for each item across samples. The Spearman-rank correlation of endorsement rates across samples was large ($\rho$ = 0.951, p < 0.001), indicating excellent correspondence in the relative prevalence of childhood maltreatment across samples. When we compare the distribution of person-level sum scores across samples,  however, we found that the number of maltreatment exposures was greater on average in the replication sample than in the original sample ($p_1$ = 0.185, $p_2$ = 0.219, $t$ = -4.611, p < 0.001).

A mean shift in overall endorsement rates is not necessarily an indication that measurement invariance has been violated. One possibility is simply that participants in the replication sample had simply experienced more childhood maltreatment than participants in the original sample. A second possibility is that the differences in the demographic composition of the two samples (e.g. gender, age) moderate the differences in endorsement rates. To investigate these possibilities, we tested for uniform differential item functioning (DIF) using logistic regression \citep{rogers1993comparison}, in which a participant's endorsement response for an item was predicted by their rest score (i.e. their sum score for all other items), sample membership, gender, and age.

Using the cutoffs recommended by \cite{hidalgo2014binary}, we identified 11 of 52 items on the MACE that exhibited large DIF by study (Figure S1). Interestingly these included all six of the reverse-scored items. We also found an additional 11 items that exhibited large DIF by gender (Figure S2). Five of these items concerned sexual abuse (endorsed by more women than men), and another five concerned peer physical abuse (endorsed by more men than women). We identified no items that exhibited large DIF by age. Given the number of items exhibiting DIF by sample membership, we elected to model each sample separately.

\subsection{Response dependency}

The MACE scale contains 12 sets of two to three items with response dependencies (Table S1). Some item sets share a structural dependency, where a response to one item logically requires or strongly implies a response on a second. For example, a person endorsing item 8 (``Parents or guardians hit you so hard that you received medical attention'') is all but certain to endorse item 7 (``Parents or guardians hit you so hard that it left marks for more than a few minutes''). Other item sets are characterized by redundancy dependency, where the same question is essentially asked twice using slightly different language. One such example is item 36 (``Peers forced you to engage in sexual activity against your will'') and 37 (``Peers forced you to do things sexually that you did not want to do``).

For the current analyses, we were concerned about response dependence for two reasons. First, response dependence can inflate estimates of item discrimination parameters, thereby making items appear more informative than they actually are \citep{marais2008formalizing}. Second, in the context of bifactor models, specific factors may become contaminated by dependent items thereby making them appear more reliable than they actually are \citep{reise2013applying}. Following \cite{marais2008formalizing}, we therefore converted each set of dependent dichotomous items into a single polytomous item, thereby reducing the original 52-item MACE scale to 39 items.

\subsection{Confirmatory item factor models}

In order to evaluate the dimensionality of the MACE scale, we fit a series of four confirmatory item factor models to the endorsement data of each sample separately. The basis of each model was the The basis of each model was the graded response model for ordered polytomous data \citep{samejima1997graded}. The structure of each model is depicted in Figure \ref{fig:models}. The first model is a one-factor (undimensional) model whereby all items load onto a general maltreatment factor (Figure \ref{fig:models}a). The second is a bifactor model with a general maltreatment factor and 10 specific factors corresponding to each subscale of the MACE (Figure \ref{fig:models}b).

The third model --- inspired by the deprivation-threat dimensional model \citep{mclaughlin2014childhood} --- is a bifactor S-1 model with a general maltreatment factor and one specific factor measuring neglect (Figure \ref{fig:models}c). We first attempted to fit this structure using a conventional bifactor model with two specific factors (i.e. threat, deprivation), but we observed vanishing factor loadings on the threat specific factor. Therefore, we instead utilized the bifactor S-1 model structure \citep{eid2017anomalous} where all of the threat (abuse) items comprise the general or reference factor, and the neglect items load onto both the general and specific factors.

The fourth model attempts to provide a more parsimonious explanation of the MACE scale response data. We hypothesized items on the MACE can be summarized by overarching factors: maltreatment by adults in the home and maltreatment by peers. We also propose including a methods factor to account for unexplained variance in endorsements specific to the reverse-scored items. Thus, we tested a bifactor S-1 model with a general maltreatment factor and two specific factors (i.e. peer victimization, reverse-scoring; Figure \ref{fig:models}d). Similar to Model 3, initial fits of this structure using a conventional bifactor model with three specific factors resulted in vanishing factor loadings on the household abuse specific factor.

All the confirmatory item factor models were estimated within a Bayesian framework using Hamiltonian Monte Carlo as implemented in Stan (v2.26; \citealt{carpenter2017stan}). For all models, four separate chains with randomized start values each took 3,000 samples from the posterior. The first 2,000 samples from each chain were discarded, so that 4,000 post-warmup samples from the posterior were retained. The $\hat{R}$ values for all parameters were less than 1.01, indicating acceptable convergence between chains, and there were no divergent transitions in any chain.

\subsection{Goodness of fit \& model comparison}

We relied on several indices to evaluate the fit of the models to the data. First we calculated three traditional goodness-of-fit statistics using the ordinal $M_2$ statistic \citep{cai2013limited, maydeu2013goodness}, which is asymptotically $\chi^2$-distributed with $\kappa − \nu$ degrees of freedom ($df$), where $\kappa$ is the number of reduced first- and second-order marginal residuals (i.e. $\kappa = n(n + 1)/2$ for $n$ items) and $\nu$ denotes the number of free parameters. (To improve the stability of the $M_2$, the polytomous items were binarized, corresponding to no versus any endorsement of maltreatment.)  Based on the $M_2$ statistic, we computed the root mean square error of approximation ($\text{RMSEA}_2$), comparative fit index (CFI), and Tucker-Lewis index (TLI). Following convention \citep{hu1999cutoff}, relatively good fit is indicated by the following indices: RMSEA < 0.06, CFI > 0.95, and TLI > 0.95.

Given the questionable utility of SEM-based fit indices for item response theory applications (e.g. \citealt{reise2014evaluating}), we calculated two additional fit indices. First, we relied on posterior predictive model checking and a $\chi^2_{NC}$ discrepancy measure based on the total score distribution \citep{sinharay2006posterior}. This index compares the observed and expected proportion of participants who endorse maltreatment at each sum-score level. To detect local dependence, we calculated Yen's $Q_3$ statistic for each pair of items \citep{yen1984effects}. Following \cite{christensen2017critical}, we calculated the critical $Q_3$-value (i.e. the value above which local dependence is indicated) for each model and sample. Specifically, for each model and sample, we simulated 1000 locally-independent response datasets using the posterior distribution of item and ability parameters. We then recorded maximum $Q_3$ value per simulated dataset. The critical $Q_3$-value (per sample and dataset) was defined as the 99th percentile of the max $Q_3$ values.

Finally, the goodness-of-fit of the models was compared using Bayesian leave-one out cross-validation (LOO-CV; \citealt{vehtari2017practical}). The LOO-CV measure quantifies the discrepancy between the model and the data while taking into account model complexity. Here LOO-CV values are presented in deviance scale, such that smaller values indicate better fit.

\subsection{Reliability indices}

To evaluate the reliability of the MACE total and subscale scores, we calculated on several model-based reliability indices using the standardized factor loadings from the bifactor models. First, we calculated coefficient omega ($\omega$; \citealt{mcdonald1999test}), which quantifies the proportion of reliable score variance attributable to all factors (i.e. group and specific factors). Next we calculated the coefficient omega hierarchical ($\omega_h$) and coefficient omega subscale ($\omega_s$; \citealt{reise2013applying, rodriguez2016evaluating}). The $\omega_h$ index quantifies the fraction of variance in total scores that can be attributed to the individual differences on the general factor. By contrast, the $\omega_s$ index quantifies the fraction of variance in a subscale score that can be attributed to individual differences on a specific factor after controlling for the general factor. The $\omega_h$/$\omega_s$ values facilitate interpretation of the total and subscale scores. Values of $\omega_h$ approaching 1 indicate the general factor is the most dominant source of reliable variance in a total score. When $\omega_h$ is greater than 0.80, the total score on the scale may be considered essentially unidimensional \citep{rodriguez2016applying}. Similarly, values of $\omega_s$ approaching 1 increasingly indicate that individual differences on a subscale score reflect a specific factor and not the general factor. \cite{canivez2016bifactor} suggests that an acceptable $\omega_s$ value is 0.50 and that >0.70 is desirable. In contrast, $\omega_s$ values substantially smaller than 0.50 prevent meaningful interpretation of a subscale as reflecting a specific factor \citep{gignac2013bifactor}.

In order to assess the unidimensionality of the MACE scale, we also calculated for each bifactor model and sample the explained common variance (ECV; \citealt{sijtsma2009use}) and the percent uncontaminated correlations (PUC; \citealt{reise2013multidimensionality}). The ECV is the ratio of the variance explained by the general factor divided by the total common variance of a scale, and thus quantifies the contribution of the general factor relative to the specific factors. As ECV values approach 1, general factor loadings are increasingly expected to resemble those that would be obtained by a one-dimensional model; ECV values greater than 0.70 are often indicative of essential unidimensionality \citep{rodriguez2016applying}. The PUC --- which quantifies how many inter-item correlations are accounted for only by the general factor --- is an important moderator of the ECV. When the PUC increases, the ECV is less important for evaluating the risk of bias when fitting a unidimensional model to data with a latent bifactor structure; when PUC is greater than 0.80, there is a low risk of bias when a multidimensional scale is treated as unidimensional \citep{reise2013multidimensionality}. As an additional check, we calculated for each bifactor model the relative parameter bias, or the difference between an item's loading in the unidimensional  model and its corresponding general factor loading the bifactor, divided by the general factor loading in the bifactor. Parameter bias less than 10–15\% is acceptable \citep{muthen1987structural}.

Though of secondary interest, we also calculated the H index to assess the construct construct replicability of each factor \citep{hancock2001rethinking}. The H index is an estimate of how well a set of items represents a latent variable. Larger H values (e.g. $> 0.70$) indicate a well-defined latent variable, with factor loadings more likely to be stable across studies; conversely, low H values suggest a poorly defined latent variable, where the pattern of factor loadings are liable to change across studies \citep{hancock2001rethinking}.

\subsection{Exploratory factor analysis}

In addition to confirmatory factor analysis, we also performed exploratory factor analysis (EFA) on the endorsement data from both samples to see what factor structure would emerge with no \emph{a priori} restrictions \citep{schmitt2018selecting}. We elected to extract two- and three-factor solutions based to match the number of factors in the two bifactor S-1 models. That is, we wanted to see if a two-factor solution would resemble threat and deprivation factors; and if a three-factor solution would resemble family maltreatment, peer victimization, and reverse-scoring factors.

EFA was performed using the \textit{lavaan} (v0.6.11; \citealt{lavaan}) package available in R. Factors were extracted using the geomin \citep{yates1987multivariate} and cf-quartimax rotation criteria \citep{crawford1970general}, which were selected in order to extract simple factor structures with few cross-loadings. We used two rotation criteria in order to examine the generalizability of the factor solutions across oblique and orthogonal rotations.

\section{Results}

\subsection{Goodness-of-fit \& model comparison}

The fit indices for each model and sample are summarized in Table \ref{table:diagnostics}. According to the traditional indices, we found all four models provided an acceptable fit to the response data. Across samples, all the models had RMSEA values < 0.06 and the majority exhibited CFI and TLI values > 0.95. The exception were models 1 and 3 in the original sample, which exhibited acceptable fits (CFI and TFI values > 0.90). In general, model fits were better for the replication sample data than for the original sample. Furthermore none of the posterior predictive p-values corresponding to the $\chi^2_{NC}$ discrepancy measure exceeded the critical value, indicating that all models were sufficiently able to replicate the sum score distribution for each sample.

Next we inspected the $Q_3$ indices for evidence of local dependence. For all models and samples, the maximum observed $Q_3$ index was greater than the 99\% percentile critical value indicating the presence of local dependence. To understand the extent of the local dependence, we visualized the $Q_3$ values for item pairs exceeding the critical value for each model and sample (Figure S3). For both samples the unidimensional model exhibited many item pairs with local dependence, predominantly among the peer victimization and reverse-scored items, suggesting possible underfactorization. By comparison, the remaining models exhibited far fewer locally dependent item pairs, the majority of which were easy to understand. For example, the large $Q_3$ value for item 17 (``Parents made inappropriate sexual comments or suggestions to your sibling or stepsibling'') and item 18 (``Parents touched or fondled your sibling or stepsibling in a sexual way'') is easily understood as surface local dependence. Regardless, the estimated item discrimination parameters for all models and samples seldom exceeded the usual range (i.e. $\alpha \leq 4$) suggesting that whatever local dependence was present is exerting minimal bias on parameter estimates \citep{edwards2018diagnostic}. Thus, we conclude that all of the models provide at least acceptable fits to the response data from both samples.

The results of the model comparison are also summarized in Table \ref{table:diagnostics}. In both the original and replication datasets, models 2 and 4 exhibited better fit to the data than models 1 and 3. Between models 2 and 4, the former provided a numerically better fit in the original sample ($\Delta$ LOO = 7.9, se = 56.3) but a worse fit in the replication sample ($\Delta$ LOO = -24.0, se = 52.0). The differences in LOO-CV values are both within four standard errors of the mean, however, indicating only weak predictive improvements \citep{vehtari2022cv}. Thus, the model comparison suggests that models 2 and 4 yield approximately equal fits to the data, which is noteworthy insofar that model 4 assumes a much simpler and parsimonious factor structure of the MACE scale.

\subsection{Confirmatory item factor models}

The standardized factor loadings from the unidimensional model fit to the original and replication datasets are are summarized in the first column of Figure \ref{fig:loadings_original} and Figure \ref{fig:loadings_online} respectively.

\subsection{Model 1: Unidimensional model}

To begin we fit a unidimensional model to the MACE response data, in which all items load onto a single general maltreatment factor (Figure 1a). In the original dataset, the item discrimination parameters were between 0.506 and and 2.944, with an average value of 1.357 (95\% HDI = 1.295 -- 1.414). For the replication dataset, item discrimination ranged between 0.657 and 2.407 with an average value of  1.349 (95\% HDI = 1.272 -- 1.422). Across datasets, all but one item exhibited medium ($\alpha \in$ 0.65 -- 1.34) to high ($\alpha \in$ 1.35 -- 1.69) levels of discrimination \citep{baker2017basics}. Differences in item discrimination estimates between datasets were marginal (RMSE = 0.038, 95\% HDI = 0.014 -- 0.068), indicating consistency of discrimination across samples.

In the original dataset, factor loadings ranged between 0.279 and 0.864, with an average value of 0.597 (95\% HDI = 0.579 -- 0.614). Similarly, in the replication dataset, factor loadings ranged between 0.359 and 0.814 with an average value of 0.600 (95\% HDI = 0.578 -- 0.621). With the exception of item 42, all items across datasets exhibited consistent and moderate-to-large loadings on the unitary factor. In other words, all but one item appear to measure, at least in part, general maltreatment.

\subsection{Model 2: Bifactor model with 10 specific factors}

The standardized factor loadings from the fit of the bifactor model to the original and replication datasets are are summarized in the second column of Figure \ref{fig:loadings_original} and Figure \ref{fig:loadings_online} respectively. For both datasets, the factor loadings on the general factor largely resembled those from the unidimensional models. The average magnitude of the factor loadings was 0.575 (95\% HDI = 0.556 -- 0.593) for the original dataset and 0.580 (95\% HDI = 0.560 -- 0.601) for the replication dataset. The relative parameter bias\footnote{The relative parameter bias is defined as the difference between an item’s loading in the unidimensional solution and its general factor  loading in the bifactor, divided by the general factor loading in the bifactor model.} was 2.1\% for the original dataset and 2.7\% in the replication dataset, indicating little bias in the factor loadings in a unidimensional solution due to multidimensionality.

In contrast loadings on the specific factors were predominantly small-to-moderate in magnitude, and seldom exceeded loadings on the general factor. In the original dataset, specific factor loadings ranged 0.014 to 0.797 with an average value of 0.313 (95\% HDI = 0.285 -- 0.341). For the replication dataset, the range was 0.043 to 0.698 with an average value of 0.343 (95\% HDI = 0.315 -- 0.369). Only items from the peer victimization subscales (i.e. PeerVA, PeerPA) had specific factor loadings of roughly equal or greater magnitude than those of general factor.

Next we inspected the reliability of the MACE total score. The $\omega$ values indicated the total score had excellent reliability in both the original ($\omega$ = 0.963) and replication ($\omega$ = 0.965) datasets. Across datasets, the omega hierarchical values was 0.924 indicating that 96.0\% and 95.8\% of the reliable variance in total scores are attributable to general maltreatment. Moreover, the explained common variance values were 0.716 for the original dataset and 0.697 for the replication dataset. This indicates the presence of a strong general maltreatment factor in the dataset, and suggests MACE response data is largely undimensional.

Next we inspected the reliability of the MACE subscale scores. The model-based reliability indices were variable in both the original (mean $\omega$ = 0.750, range = 0.590 -- 0.909) and replication (mean $\omega$ = 0.756, range = 0.595 -- 0.893) datasets, though most subscale scores exhibited adequate reliability. Crucially, the omega subscale were typically small in both the original (mean $\omega_s$ = 0.181, range = 0.027 -- 0.621) and (mean $\omega_s$ = 0.205, range = 0.037 -- 0.565) replication datasets. This indicates that most subscale scores contain little reliable variance apart from the general factor; that is, the majority of MACE subscales scores cannot be meaningfully interpreted as such. The only exception was for the peer verbal abuse subscale, where more than half the reliable variance in subscale scores were independent of general maltreatment (original = 70.7\%; replication = 66.2\%). These findings were echoed in the H index values, which were unacceptably small for all subscales across both datasets (with an exception again for the peer verbal abuse subscale).

\subsection{Model 3: Bifactor S-1 model with one specific factor (neglect)}

As for the full bifactor model, the factor loadings on the threat-deprivation bifactor S-1 model resembled those from the unidimensional models. The average magnitude of the factor loadings was 0.575 (95\% HDI = 0.556 -- 0.593) for the original dataset and 0.580 (95\% HDI = 0.560 -- 0.601) for the replication dataset. The standardized factor loadings for the specific factor (i.e. deprivation) were on average [stats here]. The loadings, however, exhibited a striking pattern where only the reverse-scored items exhibited moderate loadings; the the remaining four items exhibited negligible loadings. That is, the "deprivation" factor appears to reflect a methods factor instead.

Though the subscale reliability was adequate (original: $\omega$ = 0.877; replication: $\omega$ = 0.921), the omega subscale values indicated that neglect scores contained less than half reliable variance unique from the general factor (original: $\omega_h$ = 0.384; replication: $\omega_h$ = 0.375). Not only is the validity of the deprivation subscale questionable, but so too is its reliability.

\subsection{Model 4: Bifactor S-1 with two specific factors (peer victimization, reverse-scoring)}

Again the standardized factor loadings on general factor of the parsimonious bifactor S-1 model resembled those from the unidimensional models. The average magnitude of the factor loadings was 0.575 (95\% HDI = 0.556 -- 0.593) for the original dataset and 0.580 (95\% HDI = 0.560 -- 0.601) for the replication dataset. Notably, the factor loadings of the peer victimization and methods factor were also moderate in magnitude and on average larger than those observed on the general factor, lending credence for their independence.

In contrast, the peer victimization factor of the second bifactor S-1 model (Model 4) exhibited good subscore reliability in both datasets original: $\omega$ = 0.889; replication: $\omega$ = 0.868) and adequate omega subscale reliability (original: $\omega_h$ = 0.569; replication: $\omega_h$ = 0.493). In other words, more half of its reliable variance was unique from the general factor (original = 0.640; replication = 0.568).

\subsection{Exploratory factor analysis}

The standardized factor loadings for the two- and three-factor solutions generated by the cf-quartimax rotation are summarized in Figure \ref{fig:efa}. (The corresponding factor loadings yielded by the geomin rotation were largely similar and are presented Figure S1.) In the two-factor solutions, two unique factor structures emerged by dataset. In the original dataset, there was a dominant household maltreatment factor and a secondary peer victimization factor. In contrast, the replication dataset exhibited a primary factor comprised of household and peer maltreatment, and a secondary reverse-scoring methods factor.

In the three-factor solutions, the recovered factor structures were largely similar across datasets. There was a household maltreatment factor, a peer victimization factor, and a reverse-scoring methods factor. Of note, the sexual abuse items seemed to load more strongly on the peer victimization factor in the replication dataset than in the original dataset, warranting some caution in interpreting the stability of the factor structures. In sum, the structure of both the two- and three-factor EFA solutions resembled the bifactor S-1 model with specific peer victimization and reverse-scoring methods factors (Model 4), providing additional support for this model of covariation in childhood maltreatment.

\section{Discussion}

%------------------------------------%
% Discussion: summary
%------------------------------------%

% essentially unidimensional, sans a smaller social factor

%------------------------------------%
% Discussion: previous factor models of maltreatment scales
%------------------------------------%

Our results are consistent with a number of other studies that have found a dominant general factor in measures of childhood maltreatment. Applying a bifactor model to the childhood trauma questionnaire, \cite{spinhoven2014childhood} identified a robust general maltreatment factor that explained away the majority of reliable variance in subscale scores (with the exception of sexual abuse). \cite{meinck2021factor} found similar results when investigating the bifactor structure of the International Child Abuse Screening Tool (ICAST). In a recent investigation of the adverse childhood experience (ACE) questionnaire, \cite{dobson2021latent} found that on average items loaded more strongly onto the general factor than onto their respective specific factors. Collectively these results echo the co-occurrence of types of childhood maltreatment across measurement tools.

In turn, our results are worth interpreting in light of recent studies using MACE subscale scores to investigate the relationship between particular types of childhood maltreatment and various psychopathological or neurobiological outcomes. For example, recent studies have used the 10 MACE subscales to look at the relationship of particular types of childhood maltreatment to lifetime risk of depression \citep{gerke2018childhood}, dissociative symptoms in schizophrenia \citep{schalinski2015type}, and cortisol concentration \citep{schalinski2019early}. Others have used the MACE to form indices of threat and deprivation to study... Our results indicate that, reflecting the cofrequency of maltreatment, most of the reliable variance in these subscale scores are attributable to general maltreatment and not any of the specific types of maltreatment intended to be assessed. Many of these studies attempt to bypass these issues by use of more sophisticated analytic techniques, such as random forest regression. One issue is that these methods are not necessarily protective against collinearity \citep{gregorutti2017correlation}. But this point belies the deeper problem that it is difficult to imagine how a researcher would make substantive sense of a study reporting differential correlates for total and subscale scores, given the redundancy, as well as the low reliability of the part of the subscale not redundant with the total score. As such, results from these earlier studies should be interpreted with caution. As discussed above, general factor may account for most of the variance independent of the measurement tool. Thus, these issues may be prevalent throughout the childhood maltreatment research literature that has explored contributions of a single maltreatment type (e.g. emotional neglect) to an outcome.

 In sum, bifactor models fit to MACE response data revealed the presence of a strong general maltreatment factor that explained the majority of variance in the response data. Moreover, an interrogation of subscale scores revealed these to be largely unreliable measures of specific types of maltreatment. In other words, MACE subscale scores predominantly reflect overall maltreatment and cannot be interpreted as unique measures of specific maltreatment. The sum total of these results further suggest the MACE is essentially unidimensional, reflecting general severity of maltreatment.

%------------------------------------%
% Discussion: formative vs. reflective indicators
%------------------------------------%

% not clear to me this topic is relevant anymore

% These issues reflect a more fundamental question: whether to consider childhood maltreatment a formative or reflective variable \citep{bollen2011three}. On the one hand, maltreatment and adversity events define trauma exposure and can thus be understood as a formative variable; indeed others have argued that trauma events should be understood as formative \citep{netland2001assessment, layne2010unpacking}. On the other hand, certain types of maltreatment likely reflect a common generative process (e.g. parental style correlated neglect and physical abuse) \citep{hall2010measuring}. Childhood maltreatment may also be better understood as reflecting a general process in the context of retrospective recall, where psychological processes may shape the recall and reporting of childhood events \citep{dovran2013psychometric}. This issue is not merely semantic; it has important consequences for how we understand, measure, and interpret maltreatment data.

%------------------------------------%
% Discussion: differential item functioning (gender, reverse-scoring, sample?)
%------------------------------------%

In both data sets, women endorsed more sexual abuse, and men reported more physical abuse by peers. These results are in line with previous studies showing that girls are more often victims of sexual abuse \citep{stoltenborgh2015prevalence}, whereas boys are more often victims of bullying, especially physical bullying \citep{scheithauer2006physical}. We also found evidence that a fifth of items in the MACE exhibited some differential item functioning across the two samples. That is, the response rate for XX, YY was higher in ... . These differences may be attributable to different populations (Boston vs. across Canada \& USA), recruitment (local advertisement vs. prolific.co), exclusion criteria (required to be healthy and unmedicated vs. none), generational effect (data collected in 2010-2013 vs. 2021), technical differences in the online assessment platform ("HIPAA-compliant online enrollment system" vs. website built specifically to collect MACE data with jspsych \citep{de2015jspsych}), or differences in excluding inattentive or careless responders (none vs. infrequency items \citep{zorowitz2021inattentive}).

%------------------------------------%
% Discussion: peer victimization, bullying
%------------------------------------%

The best fitting model suggested that the MACE contains an independent peer victimization (bullying) factor comprised of items in the peer verbal abuse and peer physical abuse subscales. Correlation between the bullying factor and the general factor was moderate. In accordance, a previous report on two prospective cohorts estimated that around 40\% of children who were maltreated were also bullied \citep{lereya2015adult}. The authors of that study speculated that there may be common risk factors for childhood maltreatment and bullying such as family instability and poor social skills. On the other hand, the authors also hypothesized that maltreatment may make children more susceptible to bullying by, for instance, disrupting the child's ability to regulate emotions. Similar to childhood maltreatment by household members, bullying by peers is a common childhood experience \citep{world2020spotlight} and has detrimental effects on mental health independent of other types of childhood maltreatment \citep{singham2017concurrent, lereya2015adult}. Separating peer bullying as its own factor in the MACE may therefore offer a more granular view of the harmful childhood experiences of the responders. In the future, the peer bullying factor should be validated against other restrospective bullying measures such as the California Bullying Victimization Scale \citep{green2018initial}. Finally, the MACE does not assess cyberbullying which is an increasingly common form of bullying.

%------------------------------------%
% limitations
%------------------------------------%

%------------------------------------%
% future directions: age data
%------------------------------------%

%------------------------------------%
% conclusions / recommendations
%------------------------------------%

\bibliography{main}

\begin{table}
\small
\centering
\begin{adjustbox}{center}
\begin{tabular}{ccccccccccr}
\toprule
Sample & Model & $\chi^2 (df)$ &  RMSEA & CFI & TLI & $\chi^2_{NC}$ (PPP) & $Q3_{\max}$ &  $Q3_{\text{crit}}$ & LOO-CV & \multicolumn{1}{c}{$\Delta$LOO (se)} \\
\midrule
Original & 1 &  1858 (702) &  0.040 &  0.927 &  0.923 &  0.051 (0.230) &  0.589 &  0.239 &  30967.2 &  1748.0 (76.7) \\
& 2 &  1373 (663) &  0.032 &  0.955 &  0.950 &  0.046 (0.356) &  0.573 &  0.300 &  29211.3 & \multicolumn{1}{c}{-} \\
& 3 &  1571 (692) &  0.035 &  0.944 &  0.940 &  0.051 (0.214) &  0.591 &  0.255 &  30531.1 &  1312.0 (68.0) \\
& 4 &  1253 (687) &  0.028 &  0.964 &  0.961 &  0.046 (0.346) &  0.591 &  0.279 &  29219.2 & 7.9 (56.3) \\
\midrule
Replication & 1 &  1086 (702) &  0.031 &  0.954 &  0.952 &  0.065 (0.629) &  0.513 &  0.250 &  19771.6 &  1089.3 (58.4) \\
& 2 &   956 (663) &  0.028 &  0.965 &  0.961 &  0.058 (0.766) &  0.456 &  0.310 &  18706.2 &    24.0 (52.0) \\
& 3 &   881 (692) &  0.022 &  0.977 &  0.976 &  0.062 (0.684) &  0.468 &  0.316 &  19177.7 &   495.5 (45.0) \\
& 4 &   842 (687) &  0.020 &  0.982 &  0.980 &  0.058 (0.756) &  0.437 &  0.318 &  18682.2 & \multicolumn{1}{c}{-} \\
\bottomrule
\end{tabular}
\end{adjustbox}
\captionsetup{width=1.1\textwidth}
\caption{\normalfont Fit statistics for the confirmatory factor models by sample. LOO-CV values are presented in deviance scale (i.e. smaller values indicate better fit). Abbreviations: RMSEA = root mean square error of approximation; CFI = comparative fit index; TLI = Tucker-Lewis index; PPP = posterior predictive p-value; $Q3_{\max}$ = maximum $Q_3$ value in the sample for the model; $Q3_{\text{crit}}$ = critical $Q_3$ value for the model; LOO-CV = leave-one-out cross-validation.}
\label{table:diagnostics}
\end{table}

\begin{table}
\centering
\begin{tabular*}{\textwidth}{crccccccccc}
\toprule
& & & \multicolumn{4}{c}{Original} & \multicolumn{4}{c}{Replication} \\
\cmidrule(lr){4-7}\cmidrule(lr){8-11}
Model & Subscale & PUC & ECV & $\omega$ & $\omega_{h/s}$ & H & ECV & $\omega$ & $\omega_{h/s}$ & H \\
\midrule
2 & General  &  0.912&0.716 &  0.963 &   0.924 &  0.964&0.697 &  0.965 &   0.924 &  0.961 \\
& VA       &       &      &  0.909 &   0.069 &  0.273&      &  0.893 &   0.162 &  0.478 \\
& PA       &       &      &  0.590 &   0.037 &  0.052&      &  0.595 &   0.148 &  0.194 \\
& NVEA     &       &      &  0.848 &   0.136 &  0.525&      &  0.827 &   0.117 &  0.424 \\
& SA       &       &      &  0.710 &   0.059 &  0.134&     &  0.749 &   0.290 &  0.452 \\
& EN       &       &      &  0.715 &   0.162 &  0.304&     &  0.874 &   0.203 &  0.542 \\
& PN       &       &      &  0.800 &   0.217 &  0.479&     &  0.812 &   0.114 &  0.284 \\
& WSV      &       &      &  0.695 &   0.027 &  0.053&     &  0.651 &   0.037 &  0.067 \\
& WIPV     &       &      &  0.655 &   0.146 &  0.197&      &  0.612 &   0.077 &  0.107 \\
& PeerVA   &       &      &  0.878 &   0.621 &  0.818&     &  0.853 &   0.565 &  0.766 \\
& PeerPA   &       &      &  0.699 &   0.335 &  0.443&     &  0.694 &   0.337 &  0.446 \\
\midrule
3 & General &  0.939&0.864 &  0.958 &   0.928 &  0.964&0.841 &  0.959 &   0.922 &  0.959 \\
& neglect &       &      &  0.877 &   0.384 &  0.766&      &  0.921 &   0.375 &  0.814 \\
\midrule
4 & General &  0.931&0.738 &  0.961 &   0.896 &  0.965&0.751 &  0.961 &   0.904 &  0.960 \\
& Peer    &       &      &  0.889 &   0.569 &  0.842&      &  0.868 &   0.493 &  0.781 \\
& Reverse &       &      &  0.839 &   0.584 &  0.739&     &  0.915 &   0.498 &  0.773 \\
\bottomrule
\end{tabular*}
\captionsetup{width=1.\textwidth}
\caption{\normalfont Model-based reliability indices for the three bifactor models. Abbreviations: VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; EN = emotional neglect; PN = physical neglect; WSV = witnessing sibling violence; WIPV = witnessing inter-parental violence; PeerVA = peer verbal abuse; PeerPA = peer physical abuse; ECV = explained common variance.}
\label{table:reliability}
\end{table}

\begin{figure}
    \centering
    \includegraphics[width=1.1\textwidth,center]{figures/fig00.png}
    \caption{Caption}
    \label{fig:models}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=1.1\textwidth,center]{figures/fig01.png}
    \caption{Standardized factor loadings from the four models estimated from the original dataset. VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; WSV = witnessing sibling vioence; WIPV = witnessing inter-parental violence; EN = emotional neglect; PN = physical neglect; PeerVA = peer verbal abuse; PeerPA = peer physical abuse. ** Reverse-scored items.}
    \label{fig:loadings_original}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=1.1\textwidth,center]{figures/fig02.png}
    \caption{Standardized factor loadings from the four models estimated from the online dataset. VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; WSV = witnessing sibling vioence; WIPV = witnessing inter-parental violence; EN = emotional neglect; PN = physical neglect; PeerVA = peer verbal abuse; PeerPA = peer physical abuse. ** Reverse-scored items.}
    \label{fig:loadings_online}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=1\textwidth,center]{figures/fig03.png}
    \caption{Caption}
    \label{fig:efa}
\end{figure}

\end{document}

### 2022-07-11 manuscript/supplement.tex
minor updates + backing up manuscript draft

\documentclass[letterpaper,man,natbib]{apa6}  %Leave this stuff alone

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage[export]{adjustbox}

% dashed line (https://tex.stackexchange.com/questions/112343/beautiful-table-samples)
\usepackage{array}
\usepackage{arydshln}
\setlength\dashlinedash{0.2pt}
\setlength\dashlinegap{1.5pt}
\setlength\arrayrulewidth{0.3pt}

\title{Supplemental materials for "Bifactor models of the ``Maltreatment and Abuse Chronology of Exposure' (MACE) Scale''}
\shorttitle{Bifactor models of MACE (Suppelment)}
\author{Samuel Zorowitz$^1$, Lauri Tuominen$^{2}$}
\affiliation{$^1$Princeton Neuroscience Institute, Princeton University, USA\\$^2$The Royal’s Institute of Mental Health Research, University of Ottawa, Canada}

\begin{document}
\maketitle

\begin{table}
\begin{tabular*}{\textwidth}{clccc}
\centering
& & \multicolumn{3}{c}{Tetrachoric Corr.} \\
\cmidrule(lr){3-5}
\# & Item & Pair & 1 & 2 \\
\toprule
6 & {\small Intentionally pushed, grabbed, shoved, slapped, pinched, punched} & \\
  & {\small or kicked you.} & \\
7 & {\small Hit you so hard that it left marks for more than a few minutes.} & (7,8) & 0.774 & 0.763 \\
8 & {\small Hit you so hard, or intentionally harmed you in some way, that} & \\
& {\small you received or should have received medical attention.} & \\
\midrule
9	& {\small Spanked you on your buttocks, arms or legs.} & (9,10) & 0.977 & 0.789 \\
10	& {\small Spanked you on your bare (unclothed) buttocks.} & (9,11) & 0.980 & 0.757 \\
11 & {\small Spanked you with an object such as a strap, belt, brush, paddle,} & (10,11) & 0.474 & 0.615 \\
 & {\small rod, etc.} & \\
\midrule
13 & {\small Touched or fondled your body in a sexual way.} & (13,14) & 0.949 & 0.961 \\
14 & {\small Had you touch their body in a sexual way.} & \\
\midrule
15 & {\small Hit your sibling (stepsibling) so hard that it left marks for more} & (15,16) & 0.961 & 0.976 \\
& {\small than a few minutes.} & \\
16 & {\small Hit your sibling (stepsibling) so hard, or intentionally harmed him} & \\
& {\small or her in some way, that he/she received or should have received} & \\
& {\small medical attention.} & \\
\midrule
19 & {\small Had you touch their body in a sexual way.} & (19,20) & 0.920 & 0.916 \\
20 & {\small Actually had sexual intercourse (oral, anal or vaginal) with you.} & \\
\midrule
21 & {\small Saw adults living in the household push, grab, slap or throw} & (21,22) & 0.920 & 0.917 \\
& {\small something at your mother (stepmother, grandmother).} & \\
22 & {\small Saw adults living in the household hit your mother, stepmother, or} & (21,23) & 0.831 & 0.775 \\
& {\small grandmother so hard that it left marks for more than a few minutes.} & \\
23 & {\small Saw adults living in the household hit your mother, stepmother, or} & (22,23) & 0.948 & 0.916 \\
& {\small grandmother so hard, or intentionally harm her in some way, that} & \\
& {\small she received or should have received medical attention.} & \\
\midrule
\end{tabular*}
\end{table}

\begin{table}[ht!]
\begin{tabular}{clccc}
\centering
24 & {\small Saw adults living in the household push, grab, slap or throw} & (24,25) & 0.991 & 0.824 \\
& {\small something at your father (stepfather, grandfather).} & \\
25 & {\small Saw adults living in the household hit your father, stepfather, or} & \\
& {\small grandfather so hard that it left marks for more than a few minutes.} & \\
\midrule
33 & {\small Intentionally pushed, grabbed, shoved, slapped, pinched, punched,} & (33, 34) & 0.833 & 0.833 \\
& {\small or kicked you.} & \\
34 & {\small Hit you so hard that it left marks for more than a few minutes.} & (33, 35) & 0.677 & 0.795 \\
35 & {\small Hit you so hard, or intentionally harmed you in some way, that} & (34, 35) & 0.902 & 0.915 \\
& {\small you received or should have received medical attention.} & \\
\midrule
36 & {\small Forced you to engage in sexual activity against your will.} & (36,37) & 0.980 & 0.967 \\
37 & {\small Forced you to do things sexually that you did not want to do.} & \\
\bottomrule
\end{tabular}
\end{table}

\end{document}

### 2022-07-26 manuscript/main.tex
backing up before refactorization
manuscript updates

@@
 
+% override apa6 section headers
+% https://tex.stackexchange.com/questions/125537/how-to-modify-subsubsection-header-apa6-cls
+\makeatletter
+\renewcommand{\subsubsection}{\@startsection{subsubsection}{3}
+  {\z@}%
+  {\b@level@two@skip}{\e@level@two@skip}%
+  {\normalfont\normalsize\bfseries}}
+\makeatother
+
 % dashed line (https://tex.stackexchange.com/questions/112343/beautiful-table-samples)
@@
 
-\title{Bifactor models of the `Maltreatment and Abuse Chronology of Exposure' (MACE) Scale}
+\title{The `Maltreatment and Abuse Chronology of Exposure' (MACE) scale is essentially unidimensional}
 \shorttitle{Bifactor models of MACE}
@@
 \begin{document}
+
 \maketitle
@@
 
-% "Happy families are all alike; every unhappy family is unhappy in its own way." – Leo Tolstoy.\break
+Childhood maltreatment is an unfortunately common experience for children across the globe \citep{stoltenborgh2015prevalence}.
+It is a risk factor for multiple poor outcomes later in life, including the onset of mental illness \citep{macmillan2001childhood, green2010childhood, kessler2010childhood}, challenges in cognitive functioning \citep{kavanaugh2017neurocognitive, r2018common, su2019does}, and poorer physical health \citep{wegman2009meta, widom2012prospective, goodwin2004association}.
+Childhood maltreatment also predicts worse treatment outcomes for many psychiatric disorders \citep{nanni2012childhood, thomas2019childhood, schuckher2019history}.
+Thus, the reliable and valid assessment of childhood maltreatment is an important research goal in order to refine our understanding of the link between maltreatment and health outcomes; augment treatment strategies for victims of childhood maltreatment; and ultimately design better interventions to prevent maltreatment in the first place.
 
-Childhood maltreatment is a common experience for children across the world \citep{stoltenborgh2015prevalence}.
-It leads to an array of poor outcomes later in life, including the onset of mental illness \citep{macmillan2001childhood, green2010childhood, kessler2010childhood}, adverse effects on cognitive functioning \citep{kavanaugh2017neurocognitive, r2018common, su2019does}, and poorer physical health \citep{wegman2009meta, widom2012prospective, goodwin2004association}.
-Furthermore, childhood maltreatment predicts worse treatment outcomes across a variety of psychiatric disorders \citep{nanni2012childhood, thomas2019childhood, schuckher2019history}.
-As such, assessment of childhood maltreatment in a reliable and valid manner is important research goal in order to refine our understanding of the link between maltreatment and health outcomes; augment treatment strategies for victims of childhood maltreatment; and ultimately design interventions to prevent maltreatment in the first place.
+There already exist many measures of childhood abuse and maltreatment for researchers to choose from \citep{saini2019systematic}.
+One recently developed measure, the Maltreatment and Abuse Chronology of Exposure (MACE) scale \citep{teicher2015maltreatment}, has a number of notable advantages.
+The MACE is a 52-item retrospective self-report questionnaire comprised of 10 short subscales that each assess a unique type of childhood maltreatment.
+Six of these (e.g.
+parental verbal abuse, physical abuse, nonverbal emotional abuse, sexual abuse, emotional neglect, physical neglect) overlap with other common measures of abuse; the remaining four assess forms of maltreatment that are less frequently represented in other measures (e.g.
+peer victimization, witnessing domestic violence).
+The items in the MACE were carefully selected to measure adverse childhood experiences of increasing severity in order to more finely differentiate people of varying exposure levels.
+The MACE also measures the timing of maltreatment experiences, which is critical for investigating how the onset and duration of abuse shapes trajectories of development.
+The MACE provides two measures, an overall severity score (number of maltreatment experiences) and a multiplicity score (number of types of maltreatment experienced), with excellent temporal stability, good convergent validity with other standard measures, and improved predictive validity in predicting psychiatric symptom ratings \citep{teicher2015maltreatment}.
+For these reasons, the MACE has been listed as among the best measures of childhood maltreatment currently available \citep{saini2019systematic, georgieva2022systematic}.
 
-Many measures of childhood abuse and maltreatment have been developed \citep{saini2019systematic}.
-One recently introduced measure, the Maltreatment and Abuse Chronology of Exposure (MACE) scale \citep{teicher2015maltreatment}, has a number of notable advantages.
-The MACE is a 52-item self-report questionnaire of childhood maltreatment that assays 10 unique types of maltreatment.
-Six of these (i.e.
-parental physical maltreatment, parental non-verbal emotional abuse, parental verbal abuse, sexual abuse, physical neglect, emotional neglect) share overlap with other common measures of abuse (e.g.
-CTQ \citep{bernstein1998childhood} ); the remaining four solicit additional information on peer abuse, and witnessing violence at home.
-A number of previous scales were reviewed to select candidate items for the MACE.
-Item response theory and Rasch model were then used to for final selection of items for each 10 subscales.
-The MACE also measures detailed information on timing of exposure, which is critical for investigating how the onset of abuse shapes trajectories of development.
-Designed under the cumulative risk framework \citep{evans2013cumulative}, the MACE provides an overall severity score and multiplicity score (number of types of maltreatment experienced) with good test-retest reliability, convergent validity, and external validity.
+In recent years, there have been calls for childhood maltreatment research to move away from the cumulative risk framework \citep{evans2013cumulative} --- which focuses only on the number of adverse experiences --- towards a dimensional approach, which emphasizes how distinct types of maltreatment uniquely shape developmental trajectories \citep{mclaughlin2016beyond, belsky2012beyond}.
+In response, some researchers have begun to use the MACE to form new indices of maltreatment in lieu of or in addition to the severity and multiplicity scores.
+For example, some have investigated the unique contributions of individual differences on all 10 MACE subscale scores to the development of psychopathology \citep{schalinski2015type, gerke2018childhood, schalinski2019early}.
+Following a prominent dimensional model of maltreatment \citep{mclaughlin2014childhood}, other groups have made composite indices of threat and deprivation experiences using items in the MACE that respectively assess abuse and neglect experiences \citep{schalinski2018defining, schalinski2019environmental, teicher2018differential}.
+Though these alternatives maltreatment scores may yield new insight, their psychometric properties have not been extensively investigated.
 
-More recently, there have been calls to move away from cumulative risk approach \citep{evans2013cumulative} --- which focuses on the number of adverse experiences, but does not distinguish between distinct types of environmental experience --- towards a dimensional approach, which emphasizes how specific types of maltreatment uniquely shape psycho-biological developmental trajectories \citep{mclaughlin2016beyond, belsky2012beyond}.
-As such, research groups have begun to use the MACE in other ways.
-Some groups have studied the unique contributions of all 10 subscales to psychopathology \citep{schalinski2015type, gerke2018childhood, schalinski2019early}.
-Following the threat-deprivation dimensional model \citep{mclaughlin2014childhood}, others have made composite indices of threat and deprivation using the MACE items measuring abuse and neglect, respectively \citep{schalinski2018defining, schalinski2019environmental, teicher2018differential}.
+One critical challenge in measuring childhood maltreatment is the high co-occurrence adverse experiences.
+That is, children experiencing one form of maltreatment (e.g.
+physical abuse) are likely to experience one or more other types (e.g.
+emotional abuse; \citealt{green2010childhood, mclaughlin2012childhood, marques2021risk, mersky2017rethinking}).
+As a consequence, isolating the unique contribution of one type of maltreatment to the development of some negative outcome is difficult insofar as any one type of adversity is likely confounded with general maltreatment.
+Concretely, this means that researchers who want to study the effects of particular types of maltreatment using subscale scores or composite indices need to verify those measures reflect a construct mostly independent of overall maltreatment, lest they draw spurious conclusions.
 
-However, a critical challenge in measuring childhood maltreatment is the high co-occurrence of multiple types of abuse and neglect experiences; that is, children experiencing maltreatment are likely to be exposed to two or more types of maltreatment \citep{green2010childhood, mclaughlin2012childhood, marques2021risk, mersky2017rethinking}.
-For example, virtually all severely maltreated children report emotional abuse \citep{mcgee1995measurement}.
-As a consequence, isolating the unique contribution of a specific type of maltreatment (e.g.
-emotional neglect) to the risk of developing some outcome (e.g.
-depression) is difficult insofar that any one type of adversity is likely confounded with general maltreatment.
-Concretely, this means that researchers attempting to study the effects of particular types of maltreatment using subscale scores (e.g.
-emotional neglect from the MACE scale) should take pains to ensure that those measures largely reflect a construct unique from general maltreatment, as represented by the total score.
+Bifactor models, one type of hierarchical factor models, are well-suited for addressing this issue \citep{bornovalova2020appropriate}.
+They assume that the covariance structure of a set of items can be accounted for by a general factor, reflecting shared variance across all items, and one or more specific factors, reflecting additional shared variance among subsets of items \citep{Reise2012-ql}.
+Crucially, bifactor models can be used to calculate model-based reliability statistics, including the omega hierarchical ($\omega_h$) and omega subscale ($\omega_s$) indices, which respectively measure the proportion of variance in total and subscale scores attributable to the general factor and specific factors \citep{reise2013applying, rodriguez2016evaluating}.
+Applied to the MACE scale, these indices can inform researchers as to whether subscale scores or composite indices for particular types of maltreatment are interpretable as such, or if they instead predominantly reflect overall maltreatment.
 
-Bifactor models, a group of confirmatory factor analyses, are well-suited for addressing this issue \citep{bornovalova2020appropriate}.
-Bifactor models specify that the covariance among the items can be accounted for by a latent general factor, reflecting common variance among all items, and a set of latent specific factors, reflecting additional common variance for subsets of items \citep{Reise2012-ql}.
-Crucially bifactor models can be used to calculate model-based reliability statistics, e.g.
-the omega-hierarchical ($\omega_h$) and omega-subscale ($\omega_s$) indices, which respectively measure the proportion of variance in the total and subscale scores attributable to the general factor and specific factors \citep{reise2013applying, rodriguez2016evaluating}.
-Applied to MACE scale, these indices can guide researchers as to whether subscale scores for particular types of maltreatment are interpretable as such, or if they instead largely reflect general maltreatment.
 % summary of paper paragraph tk
 
-% The bifactor model hypothesizes a general factor, onto which all items load, and a series of orthogonal (uncorrelated) skill-specific grouping factors
 \section{Methods}
@@
 Participants received monetary compensation for their time (rate: \$12 USD/hr).
-This study was approved by the Institutional Review Board of the University of Ottawa, and all participants provided informed consent.
+This study was approved by the Institutional Review Board of the University of Ottawa (\#2021003), and all participants provided informed consent.
 This sample will hereafter be referred to as the \textit{replication} sample.
@@
 
-The fit indices for each model and sample are summarized in Table \ref{table:diagnostics}.
-According to the traditional indices, we found all four models provided an acceptable fit to the response data.
-Across samples, all the models had RMSEA values < 0.06 and the majority exhibited CFI and TLI values > 0.95.
-The exception were models 1 and 3 in the original sample, which exhibited acceptable fits (CFI and TFI values > 0.90).
+The goodness-of-fit indices for each model and sample are summarized in Table \ref{table:diagnostics}.
+According to the traditional indices, each model provided at least an acceptable fit to the response data.
+All RMSEA value were <0.06 and the majority of CFI and TLI values were >0.95.
+The only exception were for the fits of models 1 and 3 to the original sample data, which exhibited only acceptable fits (i.e.
+CFI/TLI > 0.90).
 In general, model fits were better for the replication sample data than for the original sample.
-Furthermore none of the posterior predictive p-values corresponding to the $\chi^2_{NC}$ discrepancy measure exceeded the critical value, indicating that all models were sufficiently able to replicate the sum score distribution for each sample.
+Furthermore, all of the posterior predictive p-values corresponding to the $\chi^2_{NC}$ discrepancy measure were within the acceptable range.
+Thus all models were sufficiently able to reproduce the distribution of overall severity scores for each sample.
 
 Next we inspected the $Q_3$ indices for evidence of local dependence.
-For all models and samples, the maximum observed $Q_3$ index was greater than the 99\% percentile critical value indicating the presence of local dependence.
-To understand the extent of the local dependence, we visualized the $Q_3$ values for item pairs exceeding the critical value for each model and sample (Figure S3).
-For both samples the unidimensional model exhibited many item pairs with local dependence, predominantly among the peer victimization and reverse-scored items, suggesting possible underfactorization.
-By comparison, the remaining models exhibited far fewer locally dependent item pairs, the majority of which were easy to understand.
-For example, the large $Q_3$ value for item 17 (``Parents made inappropriate sexual comments or suggestions to your sibling or stepsibling'') and item 18 (``Parents touched or fondled your sibling or stepsibling in a sexual way'') is easily understood as surface local dependence.
-Regardless, the estimated item discrimination parameters for all models and samples seldom exceeded the usual range (i.e.
-$\alpha \leq 4$) suggesting that whatever local dependence was present is exerting minimal bias on parameter estimates \citep{edwards2018diagnostic}.
-Thus, we conclude that all of the models provide at least acceptable fits to the response data from both samples.
+For all models and samples, the maximum observed $Q_3$ index was greater than the 99\% percentile critical value.
+To understand the extent of local dependence for each model and sample, we visualized the $Q_3$ values for item pairs exceeding the critical value (Figure S3).
+The unidimensional model fits exhibited many locally dependent item pairs, especially among the peer victimization and reverse-scored items, suggesting possible underfactorization.
+By comparison, models 2 -- 4 exhibited many fewer locally dependent item pairs.
+Of these, the reasons for dependence were clear.
+For example, the dependence between items 17 (``Parents made inappropriate sexual comments or suggestions to your sibling or stepsibling'') and 18 (``Parents touched or fondled your sibling or stepsibling in a sexual way'') present for all models is easily understood as surface local dependence.
+Regardless, the estimated item discrimination parameters for each model fit seldom exceeded the normal range ($\alpha \leq 4$).
+That is, the degree of local dependence present in the models is not yielding substantial bias in parameter estimation \citep{edwards2018diagnostic}.
+Thus, we conclude that all model fits exhibit at least an acceptable fit to the response data.
 
 The results of the model comparison are also summarized in Table \ref{table:diagnostics}.
-In both the original and replication datasets, models 2 and 4 exhibited better fit to the data than models 1 and 3.
-Between models 2 and 4, the former provided a numerically better fit in the original sample ($\Delta$ LOO = 7.9, se = 56.3) but a worse fit in the replication sample ($\Delta$ LOO = -24.0, se = 52.0).
-The differences in LOO-CV values are both within four standard errors of the mean, however, indicating only weak predictive improvements \citep{vehtari2022cv}.
-Thus, the model comparison suggests that models 2 and 4 yield approximately equal fits to the data, which is noteworthy insofar that model 4 assumes a much simpler and parsimonious factor structure of the MACE scale.
+Across samples, models 2 and 4 provided better fits to the data than models 1 and 3.
+Model 2 provided a numerically better fit to the data than model 4 in the original sample ($\Delta \text{LOO}$ = 7.9, se = 56.3), but the opposite was observed in the replication sample ($\Delta \text{LOO}$ = -24.0, se = 52.0).
+These differences in LOO-CV values, however, are both within four standard errors of the mean indicating only weak predictive improvements \citep{vehtari2022cv}.
+Thus the results suggest that models 2 and 4 provide approximately equal fits to the data, which is notable given that model 4 assumes a more parsimonious factor structure of the MACE.
 
-\subsection{Confirmatory item factor models}
+\subsection{Confirmatory item response models}
 
-The standardized factor loadings from the unidimensional model fit to the original and replication datasets are are summarized in the first column of Figure \ref{fig:loadings_original} and Figure \ref{fig:loadings_online} respectively.
+In the following sections, we interpret the fit of each confirmatory item response model.
+To ease interpretation, we have reproduced the standardized factor loadings of each model for the original sample in Figure \ref{fig:loadings_original} and for the replication sample in Figure \ref{fig:loadings_online}.
+The model-based reliability indices for the bifactor models are summarized in Table \ref{table:reliability}.
 
-\subsection{Model 1: Unidimensional model}
+\subsubsection{Model 1: Unidimensional model of the MACE}
 
-To begin we fit a unidimensional model to the MACE response data, in which all items load onto a single general maltreatment factor (Figure 1a).
-In the original dataset, the item discrimination parameters were between 0.506 and and 2.944, with an average value of 1.357 (95\% HDI = 1.295 -- 1.414).
-For the replication dataset, item discrimination ranged between 0.657 and 2.407 with an average value of  1.349 (95\% HDI = 1.272 -- 1.422).
-Across datasets, all but one item exhibited medium ($\alpha \in$ 0.65 -- 1.34) to high ($\alpha \in$ 1.35 -- 1.69) levels of discrimination \citep{baker2017basics}.
-Differences in item discrimination estimates between datasets were marginal (RMSE = 0.038, 95\% HDI = 0.014 -- 0.068), indicating consistency of discrimination across samples.
+The item discrimination parameters of the unidimensional model were, on average, large for both the original sample (mean $\alpha$ = 1.357, 95\% HDI = 1.295 -- 1.414) and replication sample (mean $\alpha$ = 1.349, 95\% HDI = 1.272 -- 1.422).
+The standardized factor loadings were correspondingly large in magnitude.
+In the original sample, factor loadings ranged between 0.279 and 0.864, with an average value of 0.597 (95\% HDI = 0.579 -- 0.614); in the replication sample, factor loadings ranged between 0.359 and 0.814, with an average value of 0.600 (95\% HDI = 0.578 -- 0.621).
+Most of the items had loadings greater than 0.50 (original: 80.2\%, replication: 80.0\%), and thereby appear to be good indicators of general maltreatment.
+The rank correlation of factor loadings between samples was $\rho$ = 0.565 (95\% HDI = 0.414 -- 0.711), indicating only moderate agreement on the relative loading of items onto the unitary factor across samples.
 
-In the original dataset, factor loadings ranged between 0.279 and 0.864, with an average value of 0.597 (95\% HDI = 0.579 -- 0.614).
-Similarly, in the replication dataset, factor loadings ranged between 0.359 and 0.814 with an average value of 0.600 (95\% HDI = 0.578 -- 0.621).
-With the exception of item 42, all items across datasets exhibited consistent and moderate-to-large loadings on the unitary factor.
-In other words, all but one item appear to measure, at least in part, general maltreatment.
+To provide one perspective on the dimensionality of the MACE, we next inspected the first six eigenvalues of the polychoric correlation matrix for each sample.
+In the original sample, these values were 14.62, 3.77, 2.64, 2.11, 1.79, and 1.35; in the replication sample, they are 13.27, 4.43, 2.48, 2.04, 1.81, and 1.56.
+The first-to-second eigenvalue ratios are 3.84 and 3.02, respectively, just above the 3-to-1 criterion often cited in evaluating whether item response data are essentially unidimensional \citep{embretson2013item}.
 
-\subsection{Model 2: Bifactor model with 10 specific factors}
+In sum, the results provide evidence on average in favor a unidimensional model of the MACE.
+The factor loadings suggest the presence of a single factor running through the items, as do the eigenvalues of the polychoric correlation matrix.
+Conversely, the degree of local dependence present among in the unidimensional model fits --- especially among the peer victimization items --- suggest there may be secondary dimensions to the MACE.
+We turn next to the multidimensional confirmatory models to explore these possibilities.
 
-The standardized factor loadings from the fit of the bifactor model to the original and replication datasets are are summarized in the second column of Figure \ref{fig:loadings_original} and Figure \ref{fig:loadings_online} respectively.
-For both datasets, the factor loadings on the general factor largely resembled those from the unidimensional models.
-The average magnitude of the factor loadings was 0.575 (95\% HDI = 0.556 -- 0.593) for the original dataset and 0.580 (95\% HDI = 0.560 -- 0.601) for the replication dataset.
-The relative parameter bias\footnote{The relative parameter bias is defined as the difference between an item’s loading in the unidimensional solution and its general factor  loading in the bifactor, divided by the general factor loading in the bifactor model.} was 2.1\% for the original dataset and 2.7\% in the replication dataset, indicating little bias in the factor loadings in a unidimensional solution due to multidimensionality.
+\subsubsection{Model 2: Bifactor model with 10 specific factors}
 
-In contrast loadings on the specific factors were predominantly small-to-moderate in magnitude, and seldom exceeded loadings on the general factor.
-In the original dataset, specific factor loadings ranged 0.014 to 0.797 with an average value of 0.313 (95\% HDI = 0.285 -- 0.341).
-For the replication dataset, the range was 0.043 to 0.698 with an average value of 0.343 (95\% HDI = 0.315 -- 0.369).
+Next we inspect the bifactor model with 10 specific factors, one per MACE subscale (model 2).
+In the original sample, the loadings on the general maltreatment factor ranged between 0.235 and 0.864, with an average value of 0.575 (95\% HDI = 0.556 -- 0.593).
+In the replication sample, the general factor loadings ranged between 0.299 and 0.827, with an average value of 0.580 (95\% HDI = 0.560 -- 0.601).
+The relative parameter bias was marginal for both samples (original = 2.1\%; replication = 2.7\%), suggesting little bias in the factor loadings of the unidimensional models due to multidimensionality.
+
+By contrast the standardized loadings on the specific factors were much smaller in magnitude, seldom exceeding loadings on the general factor.
+In the original dataset, the specific factor loadings ranged 0.014 to 0.797 with an average value of 0.313 (95\% HDI = 0.285 -- 0.341).
+In the replication dataset, the range was between 0.043 and 0.698, with an average value of 0.343 (95\% HDI = 0.315 -- 0.369).
 Only items from the peer victimization subscales (i.e.
-PeerVA, PeerPA) had specific factor loadings of roughly equal or greater magnitude than those of general factor.
+peer verbal \& physical abuse) had specific factor loadings of roughly equal or greater magnitude to those on the general factor.
 
-Next we inspected the reliability of the MACE total score.
-The $\omega$ values indicated the total score had excellent reliability in both the original ($\omega$ = 0.963) and replication ($\omega$ = 0.965) datasets.
-Across datasets, the omega hierarchical values was 0.924 indicating that 96.0\% and 95.8\% of the reliable variance in total scores are attributable to general maltreatment.
-Moreover, the explained common variance values were 0.716 for the original dataset and 0.697 for the replication dataset.
-This indicates the presence of a strong general maltreatment factor in the dataset, and suggests MACE response data is largely undimensional.
+Given these, we found that the MACE overall severity score had excellent reliability (original: $\omega$ = 0.963; replication: $\omega$ = 0.965).
+Further, the $\omega_h$ value for both samples was 0.924 indicating that 96.0\% and 95.8\% of the reliable variance in the overall severity score is attributable to the general maltreatment factor.
+(This is not in itself so surprising given the large number of subscales and relative paucity of items per subscale.) Notably, the ECV values were 0.716 and 0.697 for the original and replication samples, respectively.
+In conjunction with a large PUC value of 0.912, these results provide clear evidence for a dominant general maltreatment factor in the MACE.
 
-Next we inspected the reliability of the MACE subscale scores.
-The model-based reliability indices were variable in both the original (mean $\omega$ = 0.750, range = 0.590 -- 0.909) and replication (mean $\omega$ = 0.756, range = 0.595 -- 0.893) datasets, though most subscale scores exhibited adequate reliability.
-Crucially, the omega subscale were typically small in both the original (mean $\omega_s$ = 0.181, range = 0.027 -- 0.621) and (mean $\omega_s$ = 0.205, range = 0.037 -- 0.565) replication datasets.
-This indicates that most subscale scores contain little reliable variance apart from the general factor; that is, the majority of MACE subscales scores cannot be meaningfully interpreted as such.
-The only exception was for the peer verbal abuse subscale, where more than half the reliable variance in subscale scores were independent of general maltreatment (original = 70.7\%; replication = 66.2\%).
-These findings were echoed in the H index values, which were unacceptably small for all subscales across both datasets (with an exception again for the peer verbal abuse subscale).
+The reliability of the MACE subscale scores were more variable in comparison.
+The $\omega$ values ranged between 0.590 and 0.909 in the original sample, and between 0.595 and 0.893 in the replication sample, with smaller values for subscales with fewer items.
+Crucially, the $\omega_s$ values were mostly small for both the original sample (mean $\omega_s$ = 0.181, range = 0.027 -- 0.621) and replication sample (mean $\omega_s$ = 0.205, range = 0.037 -- 0.565).
+That is, the MACE subscale scores contain almost no reliable variance independent of the general maltreatment factor.
+Therefore they should not be interpreted as reflecting specific forms of maltreatment.
+The only exception was for the peer verbal abuse subscale, where more than half the reliable variance in subscale scores (original = 70.7\%; replication = 66.2\%) was independent of general maltreatment.
 
-\subsection{Model 3: Bifactor S-1 model with one specific factor (neglect)}
+The results of the bifactor model were unequivocal.
+Responses on the MACE are characterized by a dominant general maltreatment factor with strong loadings from the majority of items.
+In stark contrast, the specific factors corresponding to each MACE subscale are weak and characterized mostly by small factor loadings.
+Whereas the MACE overall severity score is reliable and a valid measure of general maltreatment, the MACE subscale scores are less reliable and cannot be meaningfully interpreted as uniquely reflecting exposure to specific types of abuse.
+The question remains, however, if there are other valid composite scores to be derived from the MACE in addition to the overall severity score.
 
-As for the full bifactor model, the factor loadings on the threat-deprivation bifactor S-1 model resembled those from the unidimensional models.
-The average magnitude of the factor loadings was 0.575 (95\% HDI = 0.556 -- 0.593) for the original dataset and 0.580 (95\% HDI = 0.560 -- 0.601) for the replication dataset.
-The standardized factor loadings for the specific factor (i.e.
-deprivation) were on average [stats here].
-The loadings, however, exhibited a striking pattern where only the reverse-scored items exhibited moderate loadings; the the remaining four items exhibited negligible loadings.
-That is, the "deprivation" factor appears to reflect a methods factor instead.
+\subsubsection{Model 3: Bifactor S-1 model with one specific factor (deprivation)}
 
-Though the subscale reliability was adequate (original: $\omega$ = 0.877; replication: $\omega$ = 0.921), the omega subscale values indicated that neglect scores contained less than half reliable variance unique from the general factor (original: $\omega_h$ = 0.384; replication: $\omega_h$ = 0.375).
-Not only is the validity of the deprivation subscale questionable, but so too is its reliability.
+We turn our attention next to the threat-deprivation bifactor S-1 model (model 3).
+The loadings on the general factor largely resembled those of the bifactor model.
+In the original sample, the loadings on the general maltreatment factor ranged between 0.150 and 0.864, with an average value of 0.577 (95\% HDI = 0.558 -- 0.594).
+In the replication sample, the general factor loadings ranged between 0.395 and 0.831, with an average value of 0.580 (95\% HDI = 0.558 -- 0.601).
 
-\subsection{Model 4: Bifactor S-1 with two specific factors (peer victimization, reverse-scoring)}
+The loadings on the specific (neglect) factor were smaller on average.
+In the original sample, the average specific loading was 0.410 (95\% HDI = 0.373 -- 0.454); in the replication sample, it was 0.454 (95\% HDI = 0.409 -- 0.499).
+The specific loadings exhibited a notable pattern where only the six reverse-scored items exhibited moderate loadings; the remaining four item exhibited negligible factor loadings.
+In other words, the ``neglect'' factor appears instead to reflect a reverse-scoring methods factor.
 
-Again the standardized factor loadings on general factor of the parsimonious bifactor S-1 model resembled those from the unidimensional models.
-The average magnitude of the factor loadings was 0.575 (95\% HDI = 0.556 -- 0.593) for the original dataset and 0.580 (95\% HDI = 0.560 -- 0.601) for the replication dataset.
-Notably, the factor loadings of the peer victimization and methods factor were also moderate in magnitude and on average larger than those observed on the general factor, lending credence for their independence.
+As observed with the bifactor model, the overall severity score under the threat- deprivation bifactor S-1 model exhibited excellent reliability (original: $\omega$ = 0.958; replication: $\omega$ = 0.959).
+Unsurprisingly, the $\omega_h$ values indicated the majority of reliable variance in severity score is attributable to the general factor (original: $\omega_h$ = 0.928; replication: $\omega_h$ = 0.922).
+The reliability of the ``neglect'' composite score was similarly good (original: $\omega$ = 0.877; replication: $\omega$ = 0.921).
+The $\omega_s$ values, however, indicated the composite score contained less than half reliable variance independent of the general factor (original: $\omega_h$ = 0.384; replication: $\omega_h$ = 0.375).
 
-In contrast, the peer victimization factor of the second bifactor S-1 model (Model 4) exhibited good subscore reliability in both datasets original: $\omega$ = 0.889; replication: $\omega$ = 0.868) and adequate omega subscale reliability (original: $\omega_h$ = 0.569; replication: $\omega_h$ = 0.493).
-In other words, more half of its reliable variance was unique from the general factor (original = 0.640; replication = 0.568).
+In summary then, the results do not support the use of composite threat and deprivation scores as derived from the abuse and neglect items on the MACE.
+The minority of reliable variance unique from general maltreatment prevents the deprivation score from being interpreted as such.
+Furthermore, the items comprising deprivation composite index seem to be confounded by a reverse-scoring methods factor.
+
+\subsubsection{Model 4: Bifactor S-1 with two specific factors (peer victimization, reverse-scoring)}
+
+Finally, we inspect the bifactor S-1 model with two specific factors measuring peer victimization and reverse-scoring (model 4).
+In the original sample, the loadings on the general maltreatment factor ranged between 0.158 and 0.864, with an average value of 0.563 (95\% HDI = 0.545 -- 0.580).
+In the replication sample, the general factor loadings ranged between 0.298 and 0.852, with an average value of 0.572 (95\% HDI = 0.550 -- 0.593).
+Interestingly, the loadings on the specific factors were of approximately equal magnitude.
+In the original dataset, the specific factor loadings ranged 0.327 to 0.789 with an average value of 0.550 (95\% HDI = 0.522 -- 0.576).
+In the replication dataset, the range was between 0.183 and 0.684, with an average value of 0.526 (95\% HDI = 0.492 -- 0.556).
+
+As with the other bifactor models, the overall severity score under this bifactor S-1 model exhibited excellent reliability (original: $\omega$ = 0.961; replication: $\omega$ = 0.961).
+Again the majority of reliable variance in severity score was attributable to the general factor.
+The peer victimization composite score exhibited good reliability across samples (original: $\omega$ = 0.889; replication: $\omega$ = 0.868).
+Notably, the corresponding $\omega_s$ values were 0.569 for the original sample and 0.493 for the replication sample.
+That is, slightly more than half of the variance in peer victimization scores (original: 64.0\%; replication: 56.8\%) is unique from the general maltreatment factor.
+These results suggest there exists a secondary factor in the MACE, peer victimization, independent of general maltreatment (albeit marginally so).
+
+Briefly, we note that the composite scores formed by the reverse-scored items was also reliable (original: $\omega$ = 0.839; replication: $\omega$ = 0.915).
+Moreover, the corresponding $omega_s$ values were similar in magnitude to those observed for the peer victimization scores (original: $\omega_s$ = 0.584; replication: $\omega_s$ = 0.498).
+As such, these scores are also have majority reliable variance independent of general maltreatment.
+It is difficult to interpret these scores, however, as it is unclear if they primarily reflect neglect experiences, a wording effect, or a scoring-induced methods artifact.
 
@@
 The standardized factor loadings for the two- and three-factor solutions generated by the cf-quartimax rotation are summarized in Figure \ref{fig:efa}.
-(The corresponding factor loadings yielded by the geomin rotation were largely similar and are presented Figure S1.) In the two-factor solutions, two unique factor structures emerged by dataset.
-In the original dataset, there was a dominant household maltreatment factor and a secondary peer victimization factor.
-In contrast, the replication dataset exhibited a primary factor comprised of household and peer maltreatment, and a secondary reverse-scoring methods factor.
+(The corresponding factor loadings produced by the geomin rotation were similar and are presented in Figure S4.) In the two-factor solution for the original sample, there was a primary adult maltreatment factor and a secondary peer victimization factor.
+In contrast, the replication sample produced a primary adult/peer maltreatment factor with a secondary reverse-scoring factor.
 
-In the three-factor solutions, the recovered factor structures were largely similar across datasets.
-There was a household maltreatment factor, a peer victimization factor, and a reverse-scoring methods factor.
-Of note, the sexual abuse items seemed to load more strongly on the peer victimization factor in the replication dataset than in the original dataset, warranting some caution in interpreting the stability of the factor structures.
-In sum, the structure of both the two- and three-factor EFA solutions resembled the bifactor S-1 model with specific peer victimization and reverse-scoring methods factors (Model 4), providing additional support for this model of covariation in childhood maltreatment.
+The three-factor solution for each sample were similar.
+The analysis produced a primary adult maltreatment factor with secondary peer victimization and reverse-scoring factors.
+(To note items concerning sexual abuse loaded more strongly on the peer victimization scale in the replication sample, warranting some caution in interpreting the replicability of factor structures.) With the oblique rotation, we observed only modest inter-factor correlations (original: mean $r$ = 0.371; replication: mean $r$ = 0.269).
+To conclude, the structure of both the two- and three-factor exploratory factor solutions resembled the structure of the bifactor S-1 model with two specific factors (model 4), providing further support for this model of responses on the MACE.
 
@@
 
-%------------------------------------%
-% Discussion: summary
-%------------------------------------%
+The MACE is an increasingly popular retrospective self-report measure of childhood maltreatment.
+Despite this, there has been little investigation into its factor structure \citep{saini2019systematic}, raising questions about the validity and interpretability of MACE subscale scores and composite indices.
+Here we employed a series of confirmatory and exploratory item factor models to investigate the dimensionality of the MACE.
+Simply put, we found repeated evidence across two independent samples that the MACE is essentially unidimensional.
+First, one-factor (unidimensional) models of the MACE provided adequate (albeit worse) fits to both datasets.
+Second, the omega hierarchical ($\omega_h$) for the full bifactor model was 0.924, indicating that the overwhelming majority of variance  in MACE severity scores is attributable to general maltreatment and only trivially affected by multidimensionality of the specific factors.
+Third, the explained common variance was, on average, approximately 0.70, suggesting that unidimensional models of the MACE are not expected to suffer from seriously biased parameter estimates.
 
-% essentially unidimensional, sans a smaller social factor
+As a corollary, we found little evidence for the reliability of MACE subscale scores or composite indices.
+Indeed, the omega subscale ($\omega_s$) values indicated that there was little reliable variance in almost all of the MACE subscale scores after accounting for the general factor.
+That is, MACE subscale scores cannot be interpreted as such but should instead be understood as predominantly reflecting overall maltreatment.
+The same can be said of the deprivation composite index, which had little reliable variance unique from the general factor and is seemingly contaminated by a reverse-scoring methods factor.
+Notably, we did find weak evidence for a secondary peer victimization factor in the MACE, whose composite scores possess a small majority of unique variance independent of general maltreatment.
+The existence of this peer victimization factor was supposed by exploratory factor modeling.
+Thus, it may be possible to derive two measures from the MACE, maltreatment by parents or other adults and maltreatment by peers, but further research is required to support this conclusion.
 
-%------------------------------------%
-% Discussion: previous factor models of maltreatment scales
-%------------------------------------%
+Our results are consistent with a number of studies investigating the factor structure of other childhood maltreatment measures.
+Bifactor models of the Childhood Trauma Questionnaire have similarly found a dominant general maltreatment factor with smaller specific factors \citep{spinhoven2014childhood, hollerbach2018main, stagaki2022mediating}.
+Similar results were obtained in recent applications of a bifactor model to the International Child Abuse Screening Tool \citep{meinck2021factor} and the Adverse Childhood Experience questionnaire \citep{dobson2021latent}.
+Together, these and our current results further highlight the high co-occurrence of childhood maltreatment experinces and the challenge that poses for investigating how particular forms of abuse shape development.
 
-Our results are consistent with a number of other studies that have found a dominant general factor in measures of childhood maltreatment.
-Applying a bifactor model to the childhood trauma questionnaire, \cite{spinhoven2014childhood} identified a robust general maltreatment factor that explained away the majority of reliable variance in subscale scores (with the exception of sexual abuse).
-\cite{meinck2021factor} found similar results when investigating the bifactor structure of the International Child Abuse Screening Tool (ICAST).
-In a recent investigation of the adverse childhood experience (ACE) questionnaire, \cite{dobson2021latent} found that on average items loaded more strongly onto the general factor than onto their respective specific factors.
-Collectively these results echo the co-occurrence of types of childhood maltreatment across measurement tools.
+Our results are also worth considering alongside other recent studies using the MACE to investigate the link between particular types of childhood maltreatment and various psycho-biological outcomes.
+For example, some groups have looked at the association between MACE subscale scores and lifetime risk of depression \citep{gerke2018childhood}, dissociative symptoms \citep{schalinski2015type}, and cortisol concentration \citep{schalinski2019early}.
+Others have used the MACE to form composite indices of threat and deprivation experiences to study their respective links to cognitive functioning \citep{schalinski2018defining} and hippocampal volume \citep{teicher2018differential}.
+Our results are clear that the majority of reliable variance in these subscale and composite scores are attributable to overall maltreatment, thereby precluding meaningful interpretation of these measures of specific types of maltreatment as such.
+This in turn makes it difficult to make substantive sense of these studies highlighting the unique contribution of a subscale over the total severity score, given their redundancy, as well as the low reliability of the part of the subscale not redundant with the total score.
 
-In turn, our results are worth interpreting in light of recent studies using MACE subscale scores to investigate the relationship between particular types of childhood maltreatment and various psychopathological or neurobiological outcomes.
-For example, recent studies have used the 10 MACE subscales to look at the relationship of particular types of childhood maltreatment to lifetime risk of depression \citep{gerke2018childhood}, dissociative symptoms in schizophrenia \citep{schalinski2015type}, and cortisol concentration \citep{schalinski2019early}.
-Others have used the MACE to form indices of threat and deprivation to study...
-Our results indicate that, reflecting the cofrequency of maltreatment, most of the reliable variance in these subscale scores are attributable to general maltreatment and not any of the specific types of maltreatment intended to be assessed.
-Many of these studies attempt to bypass these issues by use of more sophisticated analytic techniques, such as random forest regression.
-One issue is that these methods are not necessarily protective against collinearity \citep{gregorutti2017correlation}.
-But this point belies the deeper problem that it is difficult to imagine how a researcher would make substantive sense of a study reporting differential correlates for total and subscale scores, given the redundancy, as well as the low reliability of the part of the subscale not redundant with the total score.
-As such, results from these earlier studies should be interpreted with caution.
-As discussed above, general factor may account for most of the variance independent of the measurement tool.
-Thus, these issues may be prevalent throughout the childhood maltreatment research literature that has explored contributions of a single maltreatment type (e.g.
-emotional neglect) to an outcome.
-
- In sum, bifactor models fit to MACE response data revealed the presence of a strong general maltreatment factor that explained the majority of variance in the response data.
-Moreover, an interrogation of subscale scores revealed these to be largely unreliable measures of specific types of maltreatment.
-In other words, MACE subscale scores predominantly reflect overall maltreatment and cannot be interpreted as unique measures of specific maltreatment.
-The sum total of these results further suggest the MACE is essentially unidimensional, reflecting general severity of maltreatment.
-
-%------------------------------------%
-% Discussion: formative vs. reflective indicators
-%------------------------------------%
-
-% not clear to me this topic is relevant anymore
-
-% These issues reflect a more fundamental question: whether to consider childhood maltreatment a formative or reflective variable \citep{bollen2011three}.
-On the one hand, maltreatment and adversity events define trauma exposure and can thus be understood as a formative variable; indeed others have argued that trauma events should be understood as formative \citep{netland2001assessment, layne2010unpacking}.
-On the other hand, certain types of maltreatment likely reflect a common generative process (e.g.
-parental style correlated neglect and physical abuse) \citep{hall2010measuring}.
-Childhood maltreatment may also be better understood as reflecting a general process in the context of retrospective recall, where psychological processes may shape the recall and reporting of childhood events \citep{dovran2013psychometric}.
-This issue is not merely semantic; it has important consequences for how we understand, measure, and interpret maltreatment data.
-
-%------------------------------------%
-% Discussion: differential item functioning (gender, reverse-scoring, sample?)
-%------------------------------------%
-
-In both data sets, women endorsed more sexual abuse, and men reported more physical abuse by peers.
-These results are in line with previous studies showing that girls are more often victims of sexual abuse \citep{stoltenborgh2015prevalence}, whereas boys are more often victims of bullying, especially physical bullying \citep{scheithauer2006physical}.
-We also found evidence that a fifth of items in the MACE exhibited some differential item functioning across the two samples.
-That is, the response rate for XX, YY was higher in ...
-.
-These differences may be attributable to different populations (Boston vs.
-across Canada \& USA), recruitment (local advertisement vs.
-prolific.co), exclusion criteria (required to be healthy and unmedicated vs.
-none), generational effect (data collected in 2010-2013 vs.
-2021), technical differences in the online assessment platform ("HIPAA-compliant online enrollment system" vs.
-website built specifically to collect MACE data with jspsych \citep{de2015jspsych}), or differences in excluding inattentive or careless responders (none vs.
-infrequency items \citep{zorowitz2021inattentive}).
-
-%------------------------------------%
-% Discussion: peer victimization, bullying
-%------------------------------------%
-
-The best fitting model suggested that the MACE contains an independent peer victimization (bullying) factor comprised of items in the peer verbal abuse and peer physical abuse subscales.
+Worth further consideration is the peer victimization factor.
 Correlation between the bullying factor and the general factor was moderate.
@@
 
-%------------------------------------%
-% limitations
-%------------------------------------%
+As part of current analyses, we also discovered other features of the MACE.
+We identified sources of differential item functioning between men and women.
+Across datasets, women endorsed more sexual abuse, and men reported more physical abuse by peers.
+These results are in line with previous studies showing that girls are more often victims of sexual abuse \citep{stoltenborgh2015prevalence}, whereas boys are more often victims of bullying, especially physical bullying \citep{scheithauer2006physical}.
+We also identified a number of other items that exhibited some differential item functioning across samples.
+These differences may be attributable to different populations (Boston vs.
+across Canada \& USA), recruitment (local advertisement vs.
+prolific.co), exclusion criteria (required to be healthy and unmedicated vs.
+none), generational effect (data collected in 2010-2013 vs.
+2021), technical differences in the online assessment platform ("HIPAA-compliant online enrollment system" vs.
+website built specifically to collect MACE data with jspsych \citep{de2015jspsych}), or differences in excluding inattentive or careless responders (none vs.
+infrequency items \citep{zorowitz2021inattentive}).
+The causes and consequences of measurement invariance in the MACE deserves further study.
 
-%------------------------------------%
-% future directions: age data
-%------------------------------------%
+We also identified others methods features of the MACE.
+For one thing, we observed a consistent reverse-scoring methods factor in the MACE.
+One possibility is that this is driven by response styles, as has been seen in other questionnaires.
+A second possibility is that this is induced by the scoring procedure.
+Future studies should investigate alternative scoring or remove them altogether.
+A second thing, the MACE has a large number of items with response dependence.
+That is, the MACE has far fewer than 52 independent items.
+It is possible then that the MACE could be abbreviated, though this may lose sensitivity by dropping items with high/low difficulty.
+Again, future work.
 
-%------------------------------------%
-% conclusions / recommendations
-%------------------------------------%
+The present study is, of course, not without its own limitations.
+One obvious limitation is that we only study the binary endorsement data, but do not investigate the chronology data also sampled as part of the MACE.
+Though the endorsement data is predominantly unidimensional, this may belie other important patterns present only in the chronology data.
+In actuality, there are two dimensions governing responses in the MACE: susceptibility to maltreatment, or how many types or forms of maltreatment a person experiences, and severity, or how much or how long a person experienced maltreatment.
+Such a process may be well modeled by a multidimensional hurdle response model \citep{magnus2021symptom}, modeling a person's probability of endorsing a maltreatment event and, if endorsed, the frequency of the maltreatment.
+This would allow study of the independence of these two dimensions, but could also provide insight into the structural validity of the chronology data.
+Fodder for future work.
+
+A second limitation is the sample.
+Though our sample studied responses in the MACE in a more diverse sample with respect to gender and age, our sample is still lacking.
+Our sample is not racially or ethnically diverse.
+To the extent to which other studies have identified measurement invariance in more diverse samples (https://doi.org/10.1016/j.jpsychores.2007.04.010, https://doi.org/10.1016/j.chiabu.2019.104030), it is not clear to what extent the factor structure explored here holds.
+More research is necessary to ensure that the MACE is valid cross-culturally.
+
+To conclude, we found that the MACE is essentially unidimensional.
+Maltreatment endorsements are, for the most part, governed by a unitary maltreatment susceptibility factor.
+The MACE total score is valid and reliable.
+The MACE subscale scores, however, are not and should not be treated as interpretable measures of maltreatment subtypes.
+Researchers should be careful not to use subscale scores before they are validated, lest they draw spurious inferences.
 
@@
 3 & General &  0.939&0.864 &  0.958 &   0.928 &  0.964&0.841 &  0.959 &   0.922 &  0.959 \\
-& neglect &       &      &  0.877 &   0.384 &  0.766&      &  0.921 &   0.375 &  0.814 \\
+& Neglect &       &      &  0.877 &   0.384 &  0.766&      &  0.921 &   0.375 &  0.814 \\
 \midrule
@@
     \includegraphics[width=1.1\textwidth,center]{figures/fig01.png}
-    \caption{Standardized factor loadings from the four models estimated from the original dataset.
+    \caption{Standardized factor loadings from the four models estimated from the original sample.
 VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; WSV = witnessing sibling vioence; WIPV = witnessing inter-parental violence; EN = emotional neglect; PN = physical neglect; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.
@@
     \includegraphics[width=1.1\textwidth,center]{figures/fig02.png}
-    \caption{Standardized factor loadings from the four models estimated from the online dataset.
+    \caption{Standardized factor loadings from the four models estimated from the online sample.
 VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; WSV = witnessing sibling vioence; WIPV = witnessing inter-parental violence; EN = emotional neglect; PN = physical neglect; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.

### 2022-07-28 manuscript/main.tex
updates to manuscript

@@
 
-The MACE scale contains 12 sets of two to three items with response dependencies (Table S1).
-Some item sets share a structural dependency, where a response to one item logically requires or strongly implies a response on a second.
+The MACE contains a total of 12 item pairs and triplets characterized by response dependence (Table S1).
+The items in some of these sets are structurally dependent, where a response to one item requires, or strongly implies, a certain response on the second.
 For example, a person endorsing item 8 (``Parents or guardians hit you so hard that you received medical attention'') is all but certain to endorse item 7 (``Parents or guardians hit you so hard that it left marks for more than a few minutes'').
-Other item sets are characterized by redundancy dependency, where the same question is essentially asked twice using slightly different language.
-One such example is item 36 (``Peers forced you to engage in sexual activity against your will'') and 37 (``Peers forced you to do things sexually that you did not want to do``).
+Other items are dependent by virtue of redundancy.
+One such example is item 36 (``Peers forced you to engage in sexual activity against your will'') and 37 (``Peers forced you to do things sexually that you did not want to do``); here the same question is asked twice using slightly different language.
 
-For the current analyses, we were concerned about response dependence for two reasons.
-First, response dependence can inflate estimates of item discrimination parameters, thereby making items appear more informative than they actually are \citep{marais2008formalizing}.
-Second, in the context of bifactor models, specific factors may become contaminated by dependent items thereby making them appear more reliable than they actually are \citep{reise2013applying}.
-Following \cite{marais2008formalizing}, we therefore converted each set of dependent dichotomous items into a single polytomous item, thereby reducing the original 52-item MACE scale to 39 items.
+For this study, we were concerned about response dependence for two reasons.
+First, it can inflate the estimates of item discrimination parameters, making items appear more informative than they actually are \citep{marais2008formalizing}.
+Second, response dependence can inflate the estimates of item loadings on specific factors in bifactor models, giving false support for a larger number of factors than necessary \citep{reise2013applying}.
+Following \cite{marais2008formalizing}, we therefore converted each set of dependent items into a single polytomous item.
+Thus, the MACE was reduced from 52 to 39 items in total.
 
@@
 
-In order to evaluate the dimensionality of the MACE scale, we fit a series of four confirmatory item factor models to the endorsement data of each sample separately.
-The basis of each model was the The basis of each model was the graded response model for ordered polytomous data \citep{samejima1997graded}.
-The structure of each model is depicted in Figure \ref{fig:models}.
-The first model is a one-factor (undimensional) model whereby all items load onto a general maltreatment factor (Figure \ref{fig:models}a).
-The second is a bifactor model with a general maltreatment factor and 10 specific factors corresponding to each subscale of the MACE (Figure \ref{fig:models}b).
+To begin our investigation of the MACE factor structure, we fit a series of confirmatory item factor models to the response data from each sample separately.
+The basis of each model is the graded response model for ordered polytomous data \citep{samejima1997graded}.
+The factor structure for each of the four models is depicted in Figure \ref{fig:models}.
+The first model we fit is a one-factor (undimensional) model whereby all items load onto a general maltreatment factor (Figure \ref{fig:models}a).
+The second is a bifactor model, in which all items load onto a general maltreatment factor and also one of 10 specific factors corresponding to their subscale membership (Figure \ref{fig:models}b).
 
-The third model --- inspired by the deprivation-threat dimensional model \citep{mclaughlin2014childhood} --- is a bifactor S-1 model with a general maltreatment factor and one specific factor measuring neglect (Figure \ref{fig:models}c).
-We first attempted to fit this structure using a conventional bifactor model with two specific factors (i.e.
-threat, deprivation), but we observed vanishing factor loadings on the threat specific factor.
-Therefore, we instead utilized the bifactor S-1 model structure \citep{eid2017anomalous} where all of the threat (abuse) items comprise the general or reference factor, and the neglect items load onto both the general and specific factors.
+The third model we fit investigated a conceptual model of maltreatment that distinguishes between threat and deprivation as distinct forms of childhood adversity \citep{mclaughlin2014childhood}.
+Initially we fit a bifactor model in which all items loaded onto a general maltreatment factor one either the threat or deprivation specific factors.
+However, we observed vanishing item loadings on the threat specific factor.
+As such, we instead fit a bifactor S-1 model \citep{eid2017anomalous} where all items load onto a general maltreatment (or reference) factor, and only the items measuring neglect experiences load onto their own specific factor (Figure \ref{fig:models}c).
 
-The fourth model attempts to provide a more parsimonious explanation of the MACE scale response data.
-We hypothesized items on the MACE can be summarized by overarching factors: maltreatment by adults in the home and maltreatment by peers.
-We also propose including a methods factor to account for unexplained variance in endorsements specific to the reverse-scored items.
-Thus, we tested a bifactor S-1 model with a general maltreatment factor and two specific factors (i.e.
-peer victimization, reverse-scoring; Figure \ref{fig:models}d).
-Similar to Model 3, initial fits of this structure using a conventional bifactor model with three specific factors resulted in vanishing factor loadings on the household abuse specific factor.
+The fourth and final model we fit was designed provide a more parsimonious factor structure for the MACE.
+Items in the MACE can be neatly divided according to the source of adversity: maltreatment by adults at home (e.g.
+parents) or by peers.
+We hypothesized then that there would be two principal factors underlying responses on the MACE, maltreatment by adults and peer victimization.
+Based on preliminary model fits, we also suspected the presence of a secondary methods factor affecting the 6 reverse-scored items.
+Initially we fit a bifactor model in which all items loaded onto a general maltreatment factor and also one of three specific factors: maltreatment by adults, peer victimization, or reverse-scoring.
+However, we also observed vanishing loadings for the maltreatment by adults specific factor.
+Thus, we instead fit a bifactor S-1 model where all items load onto a general maltreatment factor; in addition, the peer victimization and reverse-scoring items load onto their own respective specific factors (Figure \ref{fig:models}d).
 
-All the confirmatory item factor models were estimated within a Bayesian framework using Hamiltonian Monte Carlo as implemented in Stan (v2.26; \citealt{carpenter2017stan}).
-For all models, four separate chains with randomized start values each took 3,000 samples from the posterior.
+All confirmatory item factor models were estimated within a Bayesian framework using Hamiltonian Monte Carlo as implemented in Stan (v2.26; \citealt{carpenter2017stan}).
+For each model, four separate chains with randomized start values each took 3,000 samples from the posterior.
 The first 2,000 samples from each chain were discarded, so that 4,000 post-warmup samples from the posterior were retained.
@@
 
-We relied on several indices to evaluate the fit of the models to the data.
-First we calculated three traditional goodness-of-fit statistics using the ordinal $M_2$ statistic \citep{cai2013limited, maydeu2013goodness}, which is asymptotically $\chi^2$-distributed with $\kappa − \nu$ degrees of freedom ($df$), where $\kappa$ is the number of reduced first- and second-order marginal residuals (i.e.
+We relied on multiple indices to evaluate the fit of the models to the data.
+First we calculated three traditional goodness-of-fit statistics using the ordinal $M_2$ statistic \citep{cai2013limited, maydeu2013goodness}.
+$M_2$ is asymptotically $\chi^2$-distributed with $\kappa − \nu$ degrees of freedom ($df$), where $\kappa$ is the reduced number of first- and second-order marginal residuals (i.e.
 $\kappa = n(n + 1)/2$ for $n$ items) and $\nu$ denotes the number of free parameters.
-(To improve the stability of the $M_2$, the polytomous items were binarized, corresponding to no versus any endorsement of maltreatment.)  Based on the $M_2$ statistic, we computed the root mean square error of approximation ($\text{RMSEA}_2$), comparative fit index (CFI), and Tucker-Lewis index (TLI).
-Following convention \citep{hu1999cutoff}, relatively good fit is indicated by the following indices: RMSEA < 0.06, CFI > 0.95, and TLI > 0.95.
+To improve the stability of the $M_2$ statistic, the responses to polytomous items were binarized (i.e.
+none vs.
+some maltreatment).
+Using the $M_2$ statistic, we calculated the root mean square error of approximation (RMSEA), comparative fit index (CFI), and Tucker-Lewis index (TLI) for each model fit.
+Following convention \citep{hu1999cutoff}, the benchmarks for adequate model fit are RMSEA < 0.08, CFI > 0.90, and TLI > 0.90; in turn, the benchmarks for good model fit are RMSEA < 0.05, CFI > 0.95, and TLI > 0.95.
 
-Given the questionable utility of SEM-based fit indices for item response theory applications (e.g.
+Given the questionable utility of SEM-based fit indices for item response models (e.g.
 \citealt{reise2014evaluating}), we calculated two additional fit indices.
-First, we relied on posterior predictive model checking and a $\chi^2_{NC}$ discrepancy measure based on the total score distribution \citep{sinharay2006posterior}.
-This index compares the observed and expected proportion of participants who endorse maltreatment at each sum-score level.
-To detect local dependence, we calculated Yen's $Q_3$ statistic for each pair of items \citep{yen1984effects}.
-Following \cite{christensen2017critical}, we calculated the critical $Q_3$-value (i.e.
-the value above which local dependence is indicated) for each model and sample.
-Specifically, for each model and sample, we simulated 1000 locally-independent response datasets using the posterior distribution of item and ability parameters.
-We then recorded maximum $Q_3$ value per simulated dataset.
-The critical $Q_3$-value (per sample and dataset) was defined as the 99th percentile of the max $Q_3$ values.
+First, we employed posterior predictive model checking and the $\chi^2_{NC}$ discrepancy measure based on the overall severity score distribution \citep{sinharay2006posterior}.
+This measure compares the observed and model-predicted proportion of participants at each overall severity score level.
+Second, we calculated Yen's $Q_3$ statistic for each pair of items to detect local dependence \citep{yen1984effects}.
+For each model and sample, we also calculated the critical $Q_3$-value above which local dependence is indicated  \citep{christensen2017critical}.
+Briefly, for each model and sample, we simulated 1000 locally independent datasets using the posterior distribution of the model parameters.
+We then recorded the maximum $Q_3$ value per simulation.
+The critical $Q_3$-value was defined as the 99th percentile of these max $Q_3$ values.
 
@@
 
-\subsection{Reliability indices}
+\subsection{Model-based reliability indices}
 

### 2022-07-28 manuscript/main.tex
manuscript updates (pt 2, today)

@@
 
-\title{The `Maltreatment and Abuse Chronology of Exposure' (MACE) scale is essentially unidimensional}
+\title{The MACE is essentially unidimensional: Bifactor models of the `Maltreatment and Abuse Chronology of Exposure' scale}
 \shorttitle{Bifactor models of MACE}
@@
 
-To evaluate the reliability of the MACE total and subscale scores, we calculated on several model-based reliability indices using the standardized factor loadings from the bifactor models.
+To evaluate the reliability of the MACE total and subscale/composite scores, we computed several model-based reliability indices using the factor loading estimates from the bifactor models.
 First, we calculated coefficient omega ($\omega$; \citealt{mcdonald1999test}), which quantifies the proportion of reliable score variance attributable to all factors (i.e.
-group and specific factors).
-Next we calculated the coefficient omega hierarchical ($\omega_h$) and coefficient omega subscale ($\omega_s$; \citealt{reise2013applying, rodriguez2016evaluating}).
-The $\omega_h$ index quantifies the fraction of variance in total scores that can be attributed to the individual differences on the general factor.
-By contrast, the $\omega_s$ index quantifies the fraction of variance in a subscale score that can be attributed to individual differences on a specific factor after controlling for the general factor.
-The $\omega_h$/$\omega_s$ values facilitate interpretation of the total and subscale scores.
-Values of $\omega_h$ approaching 1 indicate the general factor is the most dominant source of reliable variance in a total score.
-When $\omega_h$ is greater than 0.80, the total score on the scale may be considered essentially unidimensional \citep{rodriguez2016applying}.
-Similarly, values of $\omega_s$ approaching 1 increasingly indicate that individual differences on a subscale score reflect a specific factor and not the general factor.
+both group and specific factors).
+Next we calculated coefficient omega hierarchical ($\omega_h$) and coefficient omega subscale ($\omega_s$;  \citealt{reise2013applying, rodriguez2016evaluating}).
+Whereas $\omega_h$ measures the fraction of total score variance attributable to the general factor, $\omega_s$ measures the fraction of subscale score variance attributable to a specific factor after controlling for the general factor.
+The values of $\omega_h$/$\omega_s$ facilitate interpretation of total and subscale scores.
+Values of $\omega_h$ approaching 1 indicate the general factor is the primary source of reliable variance in a total score.
+When $\omega_h$ > 0.80, the total score may be considered an essentially unidimensional reflection of the general factor (even if the data are multidimensional;  \citealt{rodriguez2016applying}).
+So too values of $\omega_s$ approaching 1 indicate a specific factor, and not the general factor, is the primary source of reliable variance in a subscale score.
 \cite{canivez2016bifactor} suggests that an acceptable $\omega_s$ value is 0.50 and that >0.70 is desirable.
-In contrast, $\omega_s$ values substantially smaller than 0.50 prevent meaningful interpretation of a subscale as reflecting a specific factor \citep{gignac2013bifactor}.
+In contrast, values of $\omega_s$ smaller than 0.50 preclude meaningful interpretation of subscale scores as reflecting a specific factor \citep{gignac2013bifactor}.
 
-In order to assess the unidimensionality of the MACE scale, we also calculated for each bifactor model and sample the explained common variance (ECV; \citealt{sijtsma2009use}) and the percent uncontaminated correlations (PUC; \citealt{reise2013multidimensionality}).
-The ECV is the ratio of the variance explained by the general factor divided by the total common variance of a scale, and thus quantifies the contribution of the general factor relative to the specific factors.
-As ECV values approach 1, general factor loadings are increasingly expected to resemble those that would be obtained by a one-dimensional model; ECV values greater than 0.70 are often indicative of essential unidimensionality \citep{rodriguez2016applying}.
-The PUC --- which quantifies how many inter-item correlations are accounted for only by the general factor --- is an important moderator of the ECV.
-When the PUC increases, the ECV is less important for evaluating the risk of bias when fitting a unidimensional model to data with a latent bifactor structure; when PUC is greater than 0.80, there is a low risk of bias when a multidimensional scale is treated as unidimensional \citep{reise2013multidimensionality}.
-As an additional check, we calculated for each bifactor model the relative parameter bias, or the difference between an item's loading in the unidimensional  model and its corresponding general factor loading the bifactor, divided by the general factor loading in the bifactor.
+To assess the unidimensionality of the MACE scale, we also calculated the explained common variance (ECV; \citealt{sijtsma2009use}) and the proportion of uncontaminated correlations (PUC; \citealt{reise2013multidimensionality}).
+The ECV is the ratio of the variance explained by the general factor divided by the total common variance of a scale; it therefore measures the contribution of the general factor relative to the specific factors.
+As ECV values approach 1, the general factor loadings of a bifactor model are expected to resemble those that would be obtained by a one-dimensional model.
+ECV values greater than 0.70 often indicate a scale is essential unidimensional \citep{rodriguez2016applying}.
+In turn, the PUC quantifies how many inter-item correlations are accounted for only by the general factor and is an important moderator of the ECV.
+As the PUC increases, ECV is less important for evaluating the risk of bias when fitting a unidimensional model to data with a latent bifactor structure.
+When PUC > 0.80, there is a low risk of bias when a multidimensional scale is treated as unidimensional \citep{reise2013multidimensionality}.
+As an additional check, we calculated the relative parameter bias (RPB), which is the difference between an item's loading in a unidimensional model and its corresponding general factor loading the bifactor, divided by the general factor loading in the bifactor.
 Parameter bias less than 10–15\% is acceptable \citep{muthen1987structural}.
 
-Though of secondary interest, we also calculated the H index to assess the construct construct replicability of each factor \citep{hancock2001rethinking}.
+Finally, we calculated the H index to assess the construct replicability of each factor \citep{hancock2001rethinking}.
 The H index is an estimate of how well a set of items represents a latent variable.
-Larger H values (e.g.
-$> 0.70$) indicate a well-defined latent variable, with factor loadings more likely to be stable across studies; conversely, low H values suggest a poorly defined latent variable, where the pattern of factor loadings are liable to change across studies \citep{hancock2001rethinking}.
+H values closer to 1 indicate a well-defined latent variable with factor loadings more likely to be stable across studies.
+In contrast, small H values suggest a poorly defined latent variable where factor loadings are liable to change across studies.
+Because we are primarily focused on measuring the reliability of MACE total and subscale/composite scores, we report the H values below but do not discuss them further.
 
@@
 
-In addition to confirmatory factor analysis, we also performed exploratory factor analysis (EFA) on the endorsement data from both samples to see what factor structure would emerge with no \emph{a priori} restrictions \citep{schmitt2018selecting}.
-We elected to extract two- and three-factor solutions based to match the number of factors in the two bifactor S-1 models.
-That is, we wanted to see if a two-factor solution would resemble threat and deprivation factors; and if a three-factor solution would resemble family maltreatment, peer victimization, and reverse-scoring factors.
+In addition to confirmatory factor analysis, we also we also performed exploratory factor analysis (EFA) on the response data from each sample separately.
+The goal of the EFA was to observe what factor structure would emerge with no \emph{a priori} restrictions \citep{schmitt2018selecting}.
+We decided to extract two- and three- factor solutions based on the total number of factors in the bifactor S-1 models.
+We were particularly interested in seeing if a two-factor solution would produce the factors from model 3 (i.e.
+threat and deprivation); and if a three-factor solution would produce the three factors from model 4 (i.e.
+maltreatment by adults, peer victimization, reverse-scoring).
 
@@
 Factors were extracted using the geomin \citep{yates1987multivariate} and cf-quartimax rotation criteria \citep{crawford1970general}, which were selected in order to extract simple factor structures with few cross-loadings.
-We used two rotation criteria in order to examine the generalizability of the factor solutions across oblique and orthogonal rotations.
+We used two rotation criteria in order to examine the generalizability of the factor solutions across multiple rotations.
 
@@
 According to the traditional indices, each model provided at least an acceptable fit to the response data.
-All RMSEA value were <0.06 and the majority of CFI and TLI values were >0.95.
+All RMSEA value were <0.05 and the majority of CFI and TLI values were >0.95.
 The only exception were for the fits of models 1 and 3 to the original sample data, which exhibited only acceptable fits (i.e.

### 2022-07-30 manuscript/main.tex
manuscript updates (2022-07-30)

@@
 As ECV values approach 1, the general factor loadings of a bifactor model are expected to resemble those that would be obtained by a one-dimensional model.
-ECV values greater than 0.70 often indicate a scale is essential unidimensional \citep{rodriguez2016applying}.
+ECV values greater than 0.70 often indicate a scale is essentially unidimensional \citep{rodriguez2016applying}.
 In turn, the PUC quantifies how many inter-item correlations are accounted for only by the general factor and is an important moderator of the ECV.
@@
 
-The MACE is an increasingly popular retrospective self-report measure of childhood maltreatment.
-Despite this, there has been little investigation into its factor structure \citep{saini2019systematic}, raising questions about the validity and interpretability of MACE subscale scores and composite indices.
-Here we employed a series of confirmatory and exploratory item factor models to investigate the dimensionality of the MACE.
-Simply put, we found repeated evidence across two independent samples that the MACE is essentially unidimensional.
-First, one-factor (unidimensional) models of the MACE provided adequate (albeit worse) fits to both datasets.
-Second, the omega hierarchical ($\omega_h$) for the full bifactor model was 0.924, indicating that the overwhelming majority of variance  in MACE severity scores is attributable to general maltreatment and only trivially affected by multidimensionality of the specific factors.
-Third, the explained common variance was, on average, approximately 0.70, suggesting that unidimensional models of the MACE are not expected to suffer from seriously biased parameter estimates.
+The MACE is a promising alternative measure of childhood maltreatment \citep{georgieva2022systematic}.
+However, there has been little study of its structural validity \citep{saini2019systematic}, leaving unanswered questions about the scale's underlying factor structure and the validity of its total/subscale scores.
+Here we investigated the factor structure of the MACE using both confirmatory and exploratory item response modeling.
+The fits of a bifactor model (model 2) to the response data from both samples revealed a dominant general factor with uniformly strong loadings from the majority of items, and substantially weaker loadings on the specific subscale factors.
+The general maltreatment factor explained approximately 70\% of the reliable variance in responses across samples.
+When comparing the loadings on the general factor to those from a unidimensonal model, the relative parameter bias was marginal.
+Together, these results support the conclusion that the MACE is essentially unidimensional.
 
-As a corollary, we found little evidence for the reliability of MACE subscale scores or composite indices.
-Indeed, the omega subscale ($\omega_s$) values indicated that there was little reliable variance in almost all of the MACE subscale scores after accounting for the general factor.
-That is, MACE subscale scores cannot be interpreted as such but should instead be understood as predominantly reflecting overall maltreatment.
-The same can be said of the deprivation composite index, which had little reliable variance unique from the general factor and is seemingly contaminated by a reverse-scoring methods factor.
-Notably, we did find weak evidence for a secondary peer victimization factor in the MACE, whose composite scores possess a small majority of unique variance independent of general maltreatment.
-The existence of this peer victimization factor was supposed by exploratory factor modeling.
-Thus, it may be possible to derive two measures from the MACE, maltreatment by parents or other adults and maltreatment by peers, but further research is required to support this conclusion.
+We also calculated model-based reliability indices of the MACE total and subscale/composite scores using the bifactor model fits.
+We found that the MACE total score has excellent reliability and is a valid measure of general maltreatment.
+In contrast, we found that the MACE subscale scores were not valid measures of specific forms of maltreatment; in all but one example, the majority of reliable variance in subscale scores was attributable to the general factor, thereby precluding meaningful interpretation of the MACE subscale scores as such.
+The same was true of the composite deprivation scores, derived from the neglect items on the MACE, which also possessed little reliable variance independent of the general factor and was contaminated by a reverse-scoring methods factor.
+In sum, whereas the MACE total score is reliable and valid, composite scores derived from a subset of items are seldom interpretable as such and instead reflect general maltreatment.
 
-Our results are consistent with a number of studies investigating the factor structure of other childhood maltreatment measures.
-Bifactor models of the Childhood Trauma Questionnaire have similarly found a dominant general maltreatment factor with smaller specific factors \citep{spinhoven2014childhood, hollerbach2018main, stagaki2022mediating}.
-Similar results were obtained in recent applications of a bifactor model to the International Child Abuse Screening Tool \citep{meinck2021factor} and the Adverse Childhood Experience questionnaire \citep{dobson2021latent}.
-Together, these and our current results further highlight the high co-occurrence of childhood maltreatment experinces and the challenge that poses for investigating how particular forms of abuse shape development.
+Of the four confirmatory factor models we tested, the best-fitting and most parsimonious was a bifactor S-1 model with dominant general maltreatment factor and two smaller specific factors reflecting peer victimization and reverse-scoring (model 4).
+This model of responses on the MACE was further supported by the exploratory factor analyses, where the 3-factor solution found a similar partitioning of the items.
+Notably, across samples, composite scores derived from the peer victimization items were both reliable and interpretable insofar that the majority of their variance were attributable to the peer victimization factor even after accounting for the general maltreatment factor.
+It therefore may be valid to derive two score measures from the MACE, maltreatment by adults and by peers, but further research is needed to properly validate these measures.
 
-Our results are also worth considering alongside other recent studies using the MACE to investigate the link between particular types of childhood maltreatment and various psycho-biological outcomes.
-For example, some groups have looked at the association between MACE subscale scores and lifetime risk of depression \citep{gerke2018childhood}, dissociative symptoms \citep{schalinski2015type}, and cortisol concentration \citep{schalinski2019early}.
-Others have used the MACE to form composite indices of threat and deprivation experiences to study their respective links to cognitive functioning \citep{schalinski2018defining} and hippocampal volume \citep{teicher2018differential}.
-Our results are clear that the majority of reliable variance in these subscale and composite scores are attributable to overall maltreatment, thereby precluding meaningful interpretation of these measures of specific types of maltreatment as such.
-This in turn makes it difficult to make substantive sense of these studies highlighting the unique contribution of a subscale over the total severity score, given their redundancy, as well as the low reliability of the part of the subscale not redundant with the total score.
+Our results are consistent with a number of previous studies investigating the factor structure of other childhood maltreatment measures.
+For example, bifactor models of the Childhood Trauma Questionnaire have also revealed a dominant general maltreatment factor and weaker specific factors \citep{spinhoven2014childhood, hollerbach2018main, stagaki2022mediating}.
+Similar results were found in recent applications of a bifactor model to the International Child Abuse Screening Tool \citep{meinck2021factor} and the Adverse Childhood Experience questionnaire \citep{dobson2021latent}.
+Together, these and our current results underscore the challenge of developing versatile measures of childhood maltreatment.
+When the experience of distinct forms of abuse are correlated, it is difficult to measure any particular type of abuse independent of overall maltreatment.
 
-Worth further consideration is the peer victimization factor.
+Our results are also worth considering alongside previous studies that used subscale and/or composite scores from the MACE to investigate the link between specific forms of childhood maltreatment and assorted psycho-biological outcomes.
+For example, researchers have looked at the relationship between MACE subscale scores and risk of depression \citep{gerke2018childhood}, dissociative symptoms, \citep{schalinski2015type}, and cortisol concentration \citep{schalinski2019early}.
+Still others have derived composite threat and deprivation scores from the MACE to study their associations with cognitive functioning \citep{schalinski2018defining} and hippocampal volume \citep{teicher2018differential}.
+Our results are unequivocal that these measures principally reflect general maltreatment and not particular forms of adversity.
+This presents a challenge for interpreting the results of those studies.
+It is worth remarking that several of these studies attempted to overcome these issues by simultaneously included all of the MACE subscale/composite scores in their models using more sophisticated multivariate analyses (e.g.
+random forest regression).
+This approach is unsatisfactory for two reasons.
+First, random forest models can return spurious results when handling collinear predictors \citep{gregorutti2017correlation}.
+Second, even if one were to properly control for general maltreatment, our results indicate that the remaining reliable variance in subscale scores is so low as to render any inference of questionable utility.
+
+%Worth further consideration is the peer victimization factor.
 Correlation between the bullying factor and the general factor was moderate.
@@
 
-As part of current analyses, we also discovered other features of the MACE.
-We identified sources of differential item functioning between men and women.
-Across datasets, women endorsed more sexual abuse, and men reported more physical abuse by peers.
-These results are in line with previous studies showing that girls are more often victims of sexual abuse \citep{stoltenborgh2015prevalence}, whereas boys are more often victims of bullying, especially physical bullying \citep{scheithauer2006physical}.
-We also identified a number of other items that exhibited some differential item functioning across samples.
-These differences may be attributable to different populations (Boston vs.
-across Canada \& USA), recruitment (local advertisement vs.
-prolific.co), exclusion criteria (required to be healthy and unmedicated vs.
-none), generational effect (data collected in 2010-2013 vs.
-2021), technical differences in the online assessment platform ("HIPAA-compliant online enrollment system" vs.
-website built specifically to collect MACE data with jspsych \citep{de2015jspsych}), or differences in excluding inattentive or careless responders (none vs.
-infrequency items \citep{zorowitz2021inattentive}).
-The causes and consequences of measurement invariance in the MACE deserves further study.
+The current study revealed other features of the MACE worth remarking on.
+For example, we identified several sources of DIF on the MACE.
+Across samples we found that women were more likely to endorse having been sexually abused, whereas men were more likely to report being physically abused by peers.
+These findings are in line with previous studies that found girls are more frequently victims of sexual abuse \citep{stoltenborgh2015prevalence} and boys are more frequently victims of bullying (especially physical bullying; \citealt{scheithauer2006physical}).
+We also identified a number of items that exhibited DIF across samples.
+The cause of this is unclear, and may reflect differences in sample population (greater Boston area vs.
+US \& Canada), study location (in clinic vs.
+online), recruitment years (early 2010s vs.
+early 2020s), inclusion criteria (healthy \& unmedicated vs.
+none), and/or exclusion criteria (none vs.
+attention checks).
+The causes and impact of violations of measurement invariance on the MACE deserve further study.
 
-We also identified others methods features of the MACE.
-For one thing, we observed a consistent reverse-scoring methods factor in the MACE.
-One possibility is that this is driven by response styles, as has been seen in other questionnaires.
-A second possibility is that this is induced by the scoring procedure.
-Future studies should investigate alternative scoring or remove them altogether.
-A second thing, the MACE has a large number of items with response dependence.
-That is, the MACE has far fewer than 52 independent items.
-It is possible then that the MACE could be abbreviated, though this may lose sensitivity by dropping items with high/low difficulty.
-Again, future work.
+We also identified a reverse-scoring methods factor in the MACE.
+In our view, the most plausible explanation is that, under the \cite{teicher2015maltreatment} scoring procedure, affirmative responses on the regular and reverse-scored items on the MACE represent very different responses.
+For a regular item on the MACE, an endorsement indicates that a person experienced an adverse event during one year of childhood at least.
+In contrast, an endorsement on a reverse-scored item indicates that a person experienced experienced the absence of a positive event for all 18 years of childhood.
+Alternative causes of the reverse-scoring factor, such as participant confusion, acquiescence, or carelessness \citep{weijters2013reversed}, are also possible.
+Additional research is needed to see if alternative scoring procedures eliminates this factor.
 
-The present study is, of course, not without its own limitations.
-One obvious limitation is that we only study the binary endorsement data, but do not investigate the chronology data also sampled as part of the MACE.
-Though the endorsement data is predominantly unidimensional, this may belie other important patterns present only in the chronology data.
-In actuality, there are two dimensions governing responses in the MACE: susceptibility to maltreatment, or how many types or forms of maltreatment a person experiences, and severity, or how much or how long a person experienced maltreatment.
-Such a process may be well modeled by a multidimensional hurdle response model \citep{magnus2021symptom}, modeling a person's probability of endorsing a maltreatment event and, if endorsed, the frequency of the maltreatment.
-This would allow study of the independence of these two dimensions, but could also provide insight into the structural validity of the chronology data.
-Fodder for future work.
+The current study is, of course, not without limitations.
+The most apparent limitation is that here we only studied the dichotomous response data but not the chronology data also solicited as part of the MACE.
+Though we found that the factor structure of the dichotomous responses are essentially unidimensional, this may belie more complex patterns present in the maltreatment onset and duration data.
+Indeed, modeling the timeseries of adverse events in childhood --- perhaps using dynamic factor analysis \citep{zhang2007bayesian} --- might be more sensitive to more nuanced patterns of maltreatment co-occurrence that are lost when collapsing the data into a measure of any vs.
+no maltreatment.
+Another possibility is to jointly model the MACE endorsement and chronology data using a multidimensional hurdle model \citep{magnus2021symptom}, which measures both a person's susceptibility to maltreatment (i.e.
+number of unique types of maltreatment experienced) and severity of maltreatment (i.e.
+for how long maltreatment was experienced).
+More research is needed to develop valid and reliable MACE scores that integrate both the endorsement and chronology data.
 
-A second limitation is the sample.
-Though our sample studied responses in the MACE in a more diverse sample with respect to gender and age, our sample is still lacking.
-Our sample is not racially or ethnically diverse.
-To the extent to which other studies have identified measurement invariance in more diverse samples (https://doi.org/10.1016/j.jpsychores.2007.04.010, https://doi.org/10.1016/j.chiabu.2019.104030), it is not clear to what extent the factor structure explored here holds.
-More research is necessary to ensure that the MACE is valid cross-culturally.
+A second limitation concerns our sample.
+Though the inclusion of the replication sample increased the overall diversity of our participants with respect to gender and age, other aspects of diversity are notably lacking.
+For example, our combined sample was not racially or ethnically diverse.
+This is important as previous studies have identified DIF for childhood trauma questionnaires by race \citep{thombs2007evaluation, rodriguez2019identification}, which raises the possibility that similar issues could be found for the MACE with more diverse samples.
+Moreover, as noted above, we are not currently able to explain why DIF was identified between the original and replication samples.
+Further research is needed to identify and unpack violations of measurement invariance in the MACE in order to effectively assess childhood maltreatment in many different populations.
 
-To conclude, we found that the MACE is essentially unidimensional.
-Maltreatment endorsements are, for the most part, governed by a unitary maltreatment susceptibility factor.
-The MACE total score is valid and reliable.
-The MACE subscale scores, however, are not and should not be treated as interpretable measures of maltreatment subtypes.
-Researchers should be careful not to use subscale scores before they are validated, lest they draw spurious inferences.
+To conclude, we found that the factor structure of the MACE is essentially unidimensional.
+Despite the presence of smaller secondary factors, bifactor modeling revealed that the large majority of reliable variance in the dichotomous MACE responses are attributable to a general maltreatment factor.
+Thus, we found that the MACE total score is a valid and reliable of overall childhood maltreatment.
+In contrast, we found that MACE subscale and composite scores are not valid measures of particular types of childhood maltreatment, but instead primarily reflect general maltreatment.
+Moving forward, childhood abuse and neglect researchers utilizing the MACE (or otherwise) should take pains to ensure that any subscale or composite scores they employ are valid, lest they draw spurious inferences.
 

### 2022-07-30 manuscript/main.tex
manuscript updates (supplement folded into main doc; table s1 formatted; tables/figs placed inline) + reverted to stan tetrachoric

@@
-\documentclass[letterpaper,man,natbib]{apa6}  %Leave this stuff alone
+\documentclass[letterpaper,man,natbib,floatsintext,longtable]{apa6}
 
@@
 
-% dashed line (https://tex.stackexchange.com/questions/112343/beautiful-table-samples)
+% center contents of longtable
 \usepackage{array}
-\usepackage{arydshln}
-\setlength\dashlinedash{0.2pt}
-\setlength\dashlinegap{1.5pt}
-\setlength\arrayrulewidth{0.3pt}
+\newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}}
 
@@
 
+Data availability statement to conclude
+
 \section{Methods}
@@
 
-In both the original and replication samples, the majority of participants identified as women (original: male = 381, female = 670; replication: male = 258, female = 310, nonbinary or other = 9, rather not say = 5).
-The original sample is comprised of proportionally more woman than the replication sample ($z = 4.142, p < 0.001$).
-On average, participants in the original sample were younger than participants in the replication sample ($M_1 = 23.1$ yrs, $M_2 = 30.2$ yrs, $t = -19.313, p < 0.001$).
-Finally, the majority of participants in both sample identified as Caucasian or white (original: 76.7\%; replication: 75.8\%).
-The proportion of participants identifying as white were not different across samples ($z = 0.417$, $p = 0.677$).
+\begin{table}[t!]
+    \centering
+    \begin{tabular*}{\textwidth}{lccc}
+    \toprule
+    Variable & Original (N=1051) & Replication (N=582) & p-value \\
+    \midrule
+    \textbf{Gender}, N (\%)                & & & p < 0.001 \\
+    \hspace{1em} Women                     & 670 (63.7\%) & 310 (53.3\%) & \\
+    \hspace{1em} Men                       & 381 (36.3\%) & 258 (44.3\%) & \\
+    \hspace{1em} Trans or NB               &   0  (0.0\%) &   9  (1.5\%) & \\
+    \hspace{1em} Rather not say            &   0  (0.0\%) &   5  (0.9\%) & \\
+    \midrule
+    \textbf{Age}, yrs                      & & & p < 0.001 \\
+    \hspace{1em} Mean (range)              & 23.1 (18 -- 27) & 30.2 (18 -- 80) & \\
+    \midrule
+    \textbf{Ethnicity}, N (\%)             & & & p = 0.214 \\
+    \hspace{1em} Not Hispanic or Latino    & 963 (91.6\%) & 518 (89.0\%) & \\
+    \hspace{1em} Hispanic or Latino        &  84  (8.0\%) &  57  (9.8\%) & \\
+    \hspace{1em} Rather not say            &   4  (0.4\%) &   7  (1.2\%) & \\
+    \midrule
+    \textbf{Race}, N (\%)                  & & & p = 0.677 \\
+    \hspace{1em} White                     & 806 (76.7\%) & 441 (75.8\%) & \\ 
+    \hspace{1em} Asian                     & 103  (9.8\%) &  65 (11.2\%) & \\
+    \hspace{1em} Black or African American &  81  (7.7\%) &  31  (5.3\%) & \\
+    \hspace{1em} Other                     &  61  (5.8\%) &  28  (4.8\%) & \\
+    \hspace{1em} Rather not say            &   0  (0.0\%) &  17  (2.9\%) & \\
+    \bottomrule
+    \end{tabular*}
+    \caption{\normalfont Demographics of the two samples included in this study.
+The original sample was recruited and first reported by \cite{teicher2015maltreatment}.
+The replication sample was recruited by the current authors for the purposes of this study.}
+    \label{tab:demographics}
+\end{table}
+
+The demographics of the two samples are summarized in Table \ref{tab:demographics}.
+In both samples, the majority of participants identified as women; however, the original sample was comprised of proportionally more woman than the replication sample ($z$ = 4.142, $p$ < 0.001).
+Moreover, participants in the original sample were younger on average than participants in the replication sample ($t$ = -19.313, $p$ < 0.001).
+Finally, the majority of participants in both samples identified as White, and not as Hispanic or Latino.
+The proportion of participants identifying as White was not statistically different across samples ($z = 0.417$, $p = 0.677$); neither was the proportion of participants identifying as Hispanic or Latino ($z$ = -1.241, $p$ = 0.214).
 
@@
 
-\subsection{Response dependency}
+\subsection{Response dependence}
 
-The MACE contains a total of 12 item pairs and triplets characterized by response dependence (Table S1).
+The MACE contains a total of 12 item pairs and triplets characterized by response dependence (Table \ref{tab:dependence}).
 The items in some of these sets are structurally dependent, where a response to one item requires, or strongly implies, a certain response on the second.
@@
 As such, we instead fit a bifactor S-1 model \citep{eid2017anomalous} where all items load onto a general maltreatment (or reference) factor, and only the items measuring neglect experiences load onto their own specific factor (Figure \ref{fig:models}c).
+
+\begin{figure}[t!]
+    \centering
+    \includegraphics[width=1.1\textwidth,center]{figures/fig01.png}
+    \caption{Caption}
+    \label{fig:models}
+\end{figure}
 
@@
 
+\begin{table}[t!]
+\small
+\centering
+\begin{adjustbox}{center}
+\begin{tabular}{ccccccccccr}
+\toprule
+Sample & Model & $\chi^2 (df)$ &  RMSEA & CFI & TLI & $\chi^2_{NC}$ (PPP) & $Q3_{\max}$ &  $Q3_{\text{crit}}$ & LOO-CV & \multicolumn{1}{c}{$\Delta$LOO (se)} \\
+\midrule
+Original & 1 &  1858 (702) &  0.040 &  0.927 &  0.923 &  0.051 (0.230) &  0.589 &  0.239 &  30967.2 &  1748.0 (76.7) \\
+& 2 &  1373 (663) &  0.032 &  0.955 &  0.950 &  0.046 (0.356) &  0.573 &  0.300 &  29211.3 & \multicolumn{1}{c}{-} \\
+& 3 &  1571 (692) &  0.035 &  0.944 &  0.940 &  0.051 (0.214) &  0.591 &  0.255 &  30531.1 &  1312.0 (68.0) \\
+& 4 &  1253 (687) &  0.028 &  0.964 &  0.961 &  0.046 (0.346) &  0.591 &  0.279 &  29219.2 & 7.9 (56.3) \\
+\midrule
+Replication & 1 &  1086 (702) &  0.031 &  0.954 &  0.952 &  0.065 (0.629) &  0.513 &  0.250 &  19771.6 &  1089.3 (58.4) \\
+& 2 &   956 (663) &  0.028 &  0.965 &  0.961 &  0.058 (0.766) &  0.456 &  0.310 &  18706.2 &    24.0 (52.0) \\
+& 3 &   881 (692) &  0.022 &  0.977 &  0.976 &  0.062 (0.684) &  0.468 &  0.316 &  19177.7 &   495.5 (45.0) \\
+& 4 &   842 (687) &  0.020 &  0.982 &  0.980 &  0.058 (0.756) &  0.437 &  0.318 &  18682.2 & \multicolumn{1}{c}{-} \\
+\bottomrule
+\end{tabular}
+\end{adjustbox}
+\captionsetup{width=1.1\textwidth}
+\caption{\normalfont Fit statistics for the confirmatory factor models by sample.
+LOO-CV values are presented in deviance scale (i.e.
+smaller values indicate better fit).
+Abbreviations: RMSEA = root mean square error of approximation; CFI = comparative fit index; TLI = Tucker-Lewis index; PPP = posterior predictive p-value; $Q3_{\max}$ = maximum $Q_3$ value in the sample for the model; $Q3_{\text{crit}}$ = critical $Q_3$ value for the model; LOO-CV = leave-one-out cross-validation.}
+\label{table:diagnostics}
+\end{table}
+
 Next we inspected the $Q_3$ indices for evidence of local dependence.
@@
 The model-based reliability indices for the bifactor models are summarized in Table \ref{table:reliability}.
+
+\begin{figure}[t!]
+    \centering
+    \includegraphics[width=1.1\textwidth,center]{figures/fig02.png}
+    \caption{Standardized factor loadings from the four models estimated from the original sample.
+VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; WSV = witnessing sibling vioence; WIPV = witnessing inter-parental violence; EN = emotional neglect; PN = physical neglect; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.
+** Reverse-scored items.}
+    \label{fig:loadings_original}
+\end{figure}
 
@@
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1.1\textwidth,center]{figures/fig03.png}
+    \caption{Standardized factor loadings from the four models estimated from the online sample.
+VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; WSV = witnessing sibling vioence; WIPV = witnessing inter-parental violence; EN = emotional neglect; PN = physical neglect; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.
+** Reverse-scored items.}
+    \label{fig:loadings_online}
+\end{figure}
+
 By contrast the standardized loadings on the specific factors were much smaller in magnitude, seldom exceeding loadings on the general factor.
@@
 peer verbal \& physical abuse) had specific factor loadings of roughly equal or greater magnitude to those on the general factor.
+
+\begin{table}[h!]
+\centering
+\begin{tabular*}{\textwidth}{crccccccccc}
+\toprule
+& & & \multicolumn{4}{c}{Original} & \multicolumn{4}{c}{Replication} \\
+\cmidrule(lr){4-7}\cmidrule(lr){8-11}
+Model & Subscale & PUC & ECV & $\omega$ & $\omega_{h/s}$ & H & ECV & $\omega$ & $\omega_{h/s}$ & H \\
+\midrule
+2 & General &  0.912 & 0.716 &  0.963 &   0.924 &  0.964 & 0.697 & 0.965 & 0.924 & 0.961 \\
+& VA        &        &       &  0.909 &   0.069 &  0.273 &       & 0.893 & 0.162 & 0.478 \\
+& PA        &        &       &  0.590 &   0.037 &  0.052 &       & 0.595 & 0.148 & 0.194 \\
+& NVEA      &        &       &  0.848 &   0.136 &  0.525 &       & 0.827 & 0.117 & 0.424 \\
+& SA        &        &       &  0.710 &   0.059 &  0.134 &       & 0.749 & 0.290 & 0.452 \\
+& EN        &        &       &  0.715 &   0.162 &  0.304 &       & 0.874 & 0.203 & 0.542 \\
+& PN        &        &       &  0.800 &   0.217 &  0.479 &       & 0.812 & 0.114 & 0.284 \\
+& WSV       &        &       &  0.695 &   0.027 &  0.053 &       & 0.651 & 0.037 & 0.067 \\
+& WIPV      &        &       &  0.655 &   0.146 &  0.197 &       & 0.612 & 0.077 & 0.107 \\
+& PeerVA    &        &       &  0.878 &   0.621 &  0.818 &       & 0.853 & 0.565 & 0.766 \\
+& PeerPA    &        &       &  0.699 &   0.335 &  0.443 &       & 0.694 & 0.337 & 0.446 \\
+\midrule
+3 & General &  0.939 & 0.864 &  0.958 &   0.928 &  0.964 & 0.841 & 0.959 & 0.922 & 0.959 \\
+& Neglect   &        &       &  0.877 &   0.384 &  0.766 &       & 0.921 & 0.375 & 0.814 \\
+\midrule
+4 & General &  0.931 & 0.738 &  0.961 &   0.896 &  0.965 & 0.751 & 0.961 & 0.904 & 0.960 \\
+& Peer      &        &       &  0.889 &   0.569 &  0.842 &       & 0.868 & 0.493 & 0.781 \\
+& Reverse   &        &       &  0.839 &   0.584 &  0.739 &       & 0.915 & 0.498 & 0.773 \\
+\bottomrule
+\end{tabular*}
+\captionsetup{width=1.\textwidth}
+\caption{\normalfont Model-based reliability indices for the three bifactor models.
+Abbreviations: VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; EN = emotional neglect; PN = physical neglect; WSV = witnessing sibling violence; WIPV = witnessing inter-parental violence; PeerVA = peer verbal abuse; PeerPA = peer physical abuse; ECV = explained common variance.}
+\label{table:reliability}
+\end{table}
 
@@
 To conclude, the structure of both the two- and three-factor exploratory factor solutions resembled the structure of the bifactor S-1 model with two specific factors (model 4), providing further support for this model of responses on the MACE.
+
+\begin{figure}[t!]
+    \centering
+    \includegraphics[width=1\textwidth,center]{figures/fig04.png}
+    \caption{Caption}
+    \label{fig:efa}
+\end{figure}
 
@@
 Moving forward, childhood abuse and neglect researchers utilizing the MACE (or otherwise) should take pains to ensure that any subscale or composite scores they employ are valid, lest they draw spurious inferences.
-
 \bibliography{main}
 
-\begin{table}
-\small
+\pagebreak
+\section{Supplementary materials}
+
+\setcounter{figure}{0}
+\setcounter{table}{0}
+\renewcommand{\thetable}{S\arabic{table}}
+\renewcommand{\thefigure}{S\arabic{figure}}
+
+\begin{longtable}{P{0.02\linewidth}p{0.7\linewidth}|P{0.05\linewidth}P{0.05\linewidth}P{0.05\linewidth}}
 \centering
-\begin{adjustbox}{center}
-\begin{tabular}{ccccccccccr}
+\# & Item & \multicolumn{3}{c}{Tetrachoric Corr.} \\
 \toprule
-Sample & Model & $\chi^2 (df)$ &  RMSEA & CFI & TLI & $\chi^2_{NC}$ (PPP) & $Q3_{\max}$ &  $Q3_{\text{crit}}$ & LOO-CV & \multicolumn{1}{c}{$\Delta$LOO (se)} \\
+6 & {\small Intentionally pushed, grabbed, shoved, slapped, pinched, punched or kicked you.} & - & & \\
+7 & {\small Hit you so hard that it left marks for more than a few minutes.} & 0.778 & - &  \\
+8 & {\small Hit you so hard, or intentionally harmed you in some way, that you received or should have received medical attention.} & 0.702 & 0.758 & - \\
 \midrule
-Original & 1 &  1858 (702) &  0.040 &  0.927 &  0.923 &  0.051 (0.230) &  0.589 &  0.239 &  30967.2 &  1748.0 (76.7) \\
-& 2 &  1373 (663) &  0.032 &  0.955 &  0.950 &  0.046 (0.356) &  0.573 &  0.300 &  29211.3 & \multicolumn{1}{c}{-} \\
-& 3 &  1571 (692) &  0.035 &  0.944 &  0.940 &  0.051 (0.214) &  0.591 &  0.255 &  30531.1 &  1312.0 (68.0) \\
-& 4 &  1253 (687) &  0.028 &  0.964 &  0.961 &  0.046 (0.346) &  0.591 &  0.279 &  29219.2 & 7.9 (56.3) \\
+9& {\small Spanked you on your buttocks, arms or legs.} & - & \\
+10	& {\small Spanked you on your bare (unclothed) buttocks.} & 0.862 & - & \\
+11 & {\small Spanked you with an object such as a strap, belt, brush, paddle, rod, etc.} & 0.816 & 0.519 & - \\
 \midrule
-Replication & 1 &  1086 (702) &  0.031 &  0.954 &  0.952 &  0.065 (0.629) &  0.513 &  0.250 &  19771.6 &  1089.3 (58.4) \\
-& 2 &   956 (663) &  0.028 &  0.965 &  0.961 &  0.058 (0.766) &  0.456 &  0.310 &  18706.2 &    24.0 (52.0) \\
-& 3 &   881 (692) &  0.022 &  0.977 &  0.976 &  0.062 (0.684) &  0.468 &  0.316 &  19177.7 &   495.5 (45.0) \\
-& 4 &   842 (687) &  0.020 &  0.982 &  0.980 &  0.058 (0.756) &  0.437 &  0.318 &  18682.2 & \multicolumn{1}{c}{-} \\
+13 & {\small Touched or fondled your body in a sexual way.} & - \\
+14 & {\small Had you touch their body in a sexual way.} & 0.954 & - \\
+\midrule
+15 & {\small Hit your sibling (stepsibling) so hard that it left marks for more than a few minutes.} &  \\
+16 & {\small Hit your sibling (stepsibling) so hard, or intentionally harmed him or her in some way, that he/she received or should have received medical attention.} & 0.966 & - \\
+\midrule
+19 & {\small Had you touch their body in a sexual way.} & - \\
+20 & {\small Actually had sexual intercourse (oral, anal or vaginal) with you.} & 0.909 & - \\
+\midrule
+21 & {\small Saw adults living in the household push, grab, slap or throw something at your mother (stepmother, grandmother).} & - \\
+22 & {\small Saw adults living in the household hit your mother, stepmother, or grandmother so hard that it left marks for more than a few minutes.} & 0.920 & -  \\
+23 & {\small Saw adults living in the household hit your mother, stepmother, or grandmother so hard, or intentionally harm her in some way, that she received or should have received medical attention.} & 0.808 & 0.934 & - \\
+\midrule
+24 & {\small Saw adults living in the household push, grab, slap or throw something at your father (stepfather, grandfather).} &  \\
+25 & {\small Saw adults living in the household hit your father, stepfather, or grandfather so hard that it left marks for more than a few minutes.} & 0.900 & -\\
+\midrule
+33 & {\small Peers intentionally pushed, grabbed, shoved, slapped, pinched, punched, or kicked you.} & - \\
+34 & {\small Peers hit you so hard that it left marks for more than a few minutes.} & 0.829 & - \\
+35 & {\small Peers hit you so hard, or intentionally harmed you in some way, that you received or should have received medical attention.} & 0.723 & 0.907 & - \\
+\midrule
+36 & {\small Peers forced you to engage in sexual activity against your will.} &  - & \\
+37 & {\small Peers forced you to do things sexually that you did not want to do.} & 0.975 & - \\
 \bottomrule
-\end{tabular}
-\end{adjustbox}
-\captionsetup{width=1.1\textwidth}
-\caption{\normalfont Fit statistics for the confirmatory factor models by sample.
-LOO-CV values are presented in deviance scale (i.e.
-smaller values indicate better fit).
-Abbreviations: RMSEA = root mean square error of approximation; CFI = comparative fit index; TLI = Tucker-Lewis index; PPP = posterior predictive p-value; $Q3_{\max}$ = maximum $Q_3$ value in the sample for the model; $Q3_{\text{crit}}$ = critical $Q_3$ value for the model; LOO-CV = leave-one-out cross-validation.}
-\label{table:diagnostics}
-\end{table}
-
-\begin{table}
-\centering
-\begin{tabular*}{\textwidth}{crccccccccc}
-\toprule
-& & & \multicolumn{4}{c}{Original} & \multicolumn{4}{c}{Replication} \\
-\cmidrule(lr){4-7}\cmidrule(lr){8-11}
-Model & Subscale & PUC & ECV & $\omega$ & $\omega_{h/s}$ & H & ECV & $\omega$ & $\omega_{h/s}$ & H \\
-\midrule
-2 & General  &  0.912&0.716 &  0.963 &   0.924 &  0.964&0.697 &  0.965 &   0.924 &  0.961 \\
-& VA       &       &      &  0.909 &   0.069 &  0.273&      &  0.893 &   0.162 &  0.478 \\
-& PA       &       &      &  0.590 &   0.037 &  0.052&      &  0.595 &   0.148 &  0.194 \\
-& NVEA     &       &      &  0.848 &   0.136 &  0.525&      &  0.827 &   0.117 &  0.424 \\
-& SA       &       &      &  0.710 &   0.059 &  0.134&     &  0.749 &   0.290 &  0.452 \\
-& EN       &       &      &  0.715 &   0.162 &  0.304&     &  0.874 &   0.203 &  0.542 \\
-& PN       &       &      &  0.800 &   0.217 &  0.479&     &  0.812 &   0.114 &  0.284 \\
-& WSV      &       &      &  0.695 &   0.027 &  0.053&     &  0.651 &   0.037 &  0.067 \\
-& WIPV     &       &      &  0.655 &   0.146 &  0.197&      &  0.612 &   0.077 &  0.107 \\
-& PeerVA   &       &      &  0.878 &   0.621 &  0.818&     &  0.853 &   0.565 &  0.766 \\
-& PeerPA   &       &      &  0.699 &   0.335 &  0.443&     &  0.694 &   0.337 &  0.446 \\
-\midrule
-3 & General &  0.939&0.864 &  0.958 &   0.928 &  0.964&0.841 &  0.959 &   0.922 &  0.959 \\
-& Neglect &       &      &  0.877 &   0.384 &  0.766&      &  0.921 &   0.375 &  0.814 \\
-\midrule
-4 & General &  0.931&0.738 &  0.961 &   0.896 &  0.965&0.751 &  0.961 &   0.904 &  0.960 \\
-& Peer    &       &      &  0.889 &   0.569 &  0.842&      &  0.868 &   0.493 &  0.781 \\
-& Reverse &       &      &  0.839 &   0.584 &  0.739&     &  0.915 &   0.498 &  0.773 \\
-\bottomrule
-\end{tabular*}
-\captionsetup{width=1.\textwidth}
-\caption{\normalfont Model-based reliability indices for the three bifactor models.
-Abbreviations: VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; EN = emotional neglect; PN = physical neglect; WSV = witnessing sibling violence; WIPV = witnessing inter-parental violence; PeerVA = peer verbal abuse; PeerPA = peer physical abuse; ECV = explained common variance.}
-\label{table:reliability}
-\end{table}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=1.1\textwidth,center]{figures/fig00.png}
-    \caption{Caption}
-    \label{fig:models}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=1.1\textwidth,center]{figures/fig01.png}
-    \caption{Standardized factor loadings from the four models estimated from the original sample.
-VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; WSV = witnessing sibling vioence; WIPV = witnessing inter-parental violence; EN = emotional neglect; PN = physical neglect; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.
-** Reverse-scored items.}
-    \label{fig:loadings_original}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=1.1\textwidth,center]{figures/fig02.png}
-    \caption{Standardized factor loadings from the four models estimated from the online sample.
-VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; WSV = witnessing sibling vioence; WIPV = witnessing inter-parental violence; EN = emotional neglect; PN = physical neglect; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.
-** Reverse-scored items.}
-    \label{fig:loadings_online}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=1\textwidth,center]{figures/fig03.png}
-    \caption{Caption}
-    \label{fig:efa}
-\end{figure}
+\caption{\normalfont The 12 item pairs and triplets in the MACE characterized by response dependence.
+The tetrachoric correlation (averaged over samples) for each pair of items in a set is presented in tandem with the item wording.}
+\label{tab:dependence}
+\end{longtable}
 

### 2022-07-30 manuscript/main.tex
updated supp figs; updated naming conventions; more manuscript edits

@@
 \usepackage[export]{adjustbox}
+\usepackage{setspace}
 
@@
 
-Using the cutoffs recommended by \cite{hidalgo2014binary}, we identified 11 of 52 items on the MACE that exhibited large DIF by study (Figure S1).
+Using the cutoffs recommended by \cite{hidalgo2014binary}, we identified 11 of 52 items on the MACE that exhibited large DIF by study (Figure \ref{fig:dif_study}).
 Interestingly these included all six of the reverse-scored items.
-We also found an additional 11 items that exhibited large DIF by gender (Figure S2).
+We also found an additional 11 items that exhibited large DIF by gender (Figure \ref{fig:dif_gender}).
 Five of these items concerned sexual abuse (endorsed by more women than men), and another five concerned peer physical abuse (endorsed by more men than women).
@@
 
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.82\textwidth,center]{figures/figS01.png}
+    \caption{\normalfont \begin{spacing}{1.25} Differential item functioning by sample.
+(A) Proportions of participants endorsing having experienced a maltreatment event by study sample.
+(B) The log odds that the proportion of endorsements are larger in one or the other sample.
+Dotted lines indicate the threshold for large DIF recommended by \cite{hidalgo2014binary}.
+** Reverse-scored items \end{spacing}}
+    \label{fig:dif_study}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.82\textwidth,center]{figures/figS02.png}
+    \caption{\normalfont \begin{spacing}{1.25} Differential item functioning by gender.
+(A) Proportions of participants endorsing having experienced a maltreatment event by study gender.
+(B) The log odds that the proportion of endorsements are larger in one or the other gender.
+Dotted lines indicate the threshold for large DIF recommended by \cite{hidalgo2014binary}.
+** Reverse-scored items \end{spacing}}
+    \label{fig:dif_gender}
+\end{figure}
+
 \end{document}

### 2022-07-30 analysis/scripts/make_fig02.py

@@ -13,7 +13,7 @@ sns.set_style("ticks")
 #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
 
 ## I/O parameters.
-study = 'tuominen2022'
+study = 'teicher2015'
 
 ## Define subscales by model.
 factors = {
@@ -184,4 +184,4 @@ ax.invert_yaxis()
 sns.despine(left=True, right=True, top=True, bottom=True, ax=ax)
     
 ## Save figure.
-plt.savefig(os.path.join(ROOT_DIR, 'figures', 'fig02.png'), dpi=100)
+plt.savefig(os.path.join(ROOT_DIR, '..', 'figures', 'fig02.png'), dpi=100)

### 2022-07-30 analysis/scripts/make_fig03.py

@@ -13,7 +13,7 @@ sns.set_style("ticks")
 #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
 
 ## I/O parameters.
-study = 'teicher2015'
+study = 'tuominen2022'
 
 ## Define subscales by model.
 factors = {
@@ -184,4 +184,4 @@ ax.invert_yaxis()
 sns.despine(left=True, right=True, top=True, bottom=True, ax=ax)
     
 ## Save figure.
-plt.savefig(os.path.join(ROOT_DIR, 'figures', 'fig01.png'), dpi=100)
+plt.savefig(os.path.join(ROOT_DIR, '..', 'figures', 'fig03.png'), dpi=100)

### 2022-07-30 analysis/scripts/make_figS01.py

@@ -1,0 +1,118 @@
+import os
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+from os.path import dirname
+from pandas import read_csv
+import matplotlib.gridspec as gridspec
+sns.set_theme(style='white', context='notebook', font_scale=1.33)
+ROOT_DIR = dirname(dirname(os.path.realpath(__file__)))
+
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+### Define data table.
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+
+## Load data.
+mace = read_csv(os.path.join(ROOT_DIR, 'data', 'mace.csv'))
+covariates = read_csv(os.path.join(ROOT_DIR, 'data', 'covariates.csv'))
+dif = read_csv(os.path.join(ROOT_DIR, 'stan_results', 'logdif.csv'))
+
+## Apply screening.
+covariates = covariates.query('infreq <= 0.5')
+mace = mace[mace.subject.isin(covariates.query('infreq <= 0.5').subject)]
+
+## Restrict to available data.
+table = mace[mace.response.notnull()]
+
+## Construct tables.
+table = table.pivot_table('response',['subscale','item'],'study').reset_index()
+table = table.merge(dif.loc[dif.variable == 'study', ['item','coef']])
+
+## Add fake rows (for padding).
+for i in range(9): table = table.append({'item':-(i+1)},ignore_index=True)
+
+## Sort items.
+order = [1, 2, 3, 4, -1, 6, 7, 8, 9, 10, 11, -2, 5, 40, 41, 48, 49, 50, -3, 12, 13, 14, 19, 20, 36, 37,
+         -4, 38, 39, 42, 43, 52, -5, 44, 45, 46, 47, 51, -6, 15, 16, 17, 18, -7, 21, 22, 23, 24, 25, -8, 
+         26, 27, 28, 29, 30, -9, 31, 32, 33, 34, 35]
+table = table.set_index('item').loc[order].reset_index()
+
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+### Initialize canvas.
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+
+## Initialize canvas.
[…]

### 2022-07-30 analysis/scripts/make_figS02.py

@@ -1,0 +1,122 @@
+import os
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+from os.path import dirname
+from pandas import read_csv
+import matplotlib.gridspec as gridspec
+sns.set_theme(style='white', context='notebook', font_scale=1.33)
+ROOT_DIR = dirname(dirname(os.path.realpath(__file__)))
+
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+### Define data table.
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+
+## Load data.
+mace = read_csv(os.path.join(ROOT_DIR, 'data', 'mace.csv'))
+covariates = read_csv(os.path.join(ROOT_DIR, 'data', 'covariates.csv'))
+dif = read_csv(os.path.join(ROOT_DIR, 'stan_results', 'logdif.csv'))
+
+## Apply screening.
+covariates = covariates.query('infreq <= 0.5')
+mace = mace[mace.subject.isin(covariates.query('infreq <= 0.5').subject)]
+
+## Restrict to available data.
+table = mace[mace.response.notnull()]
+
+## Merge with covariates.
+table = table.merge(covariates[['subject','gender']])
+table = table.query('gender == "Male" or gender == "Female"')
+
+## Construct tables.
+table = table.pivot_table('response',['subscale','item'],'gender').reset_index()
+table = table.merge(dif.loc[dif.variable == 'gender', ['item','coef']])
+
+## Add fake rows (for padding).
+for i in range(9): table = table.append({'item':-(i+1)},ignore_index=True)
+
+## Sort items.
+order = [1, 2, 3, 4, -1, 6, 7, 8, 9, 10, 11, -2, 5, 40, 41, 48, 49, 50, -3, 12, 13, 14, 19, 20, 36, 37,
+         -4, 38, 39, 42, 43, 52, -5, 44, 45, 46, 47, 51, -6, 15, 16, 17, 18, -7, 21, 22, 23, 24, 25, -8, 
+         26, 27, 28, 29, 30, -9, 31, 32, 33, 34, 35]
+table = table.set_index('item').loc[order].reset_index()
+
[…]

### 2022-07-31 manuscript/main.tex
renamed figures; add efa fit indices; clean up efa figures; clean up figure 1; clean up efa section fof manuscript

@@
 \usepackage{graphicx}
-\usepackage{booktabs}
-\usepackage[export]{adjustbox}
-\usepackage{setspace}
+\usepackage{booktabs}             % fancy latex tables
+\usepackage[export]{adjustbox}    % center wide tables on page
+\usepackage{setspace}             % adjust caption linespacing
 
@@
 
-% summary of paper paragraph tk
-
-Data availability statement to conclude
+summary of paper paragraph tk. data availability statement to conclude
 
@@
     \caption{\normalfont Demographics of the two samples included in this study.
-The original sample was recruited and first reported by \cite{teicher2015maltreatment}.
+The original sample was recruited and reported by \cite{teicher2015maltreatment}.
 The replication sample was recruited by the current authors for the purposes of this study.}
@@
     \includegraphics[width=1.1\textwidth,center]{figures/fig01.png}
-    \caption{Caption}
+    \captionsetup{width=1.1\textwidth}
+    \caption{\normalfont The structure of the four competing confirmatory factor models of the MACE.
+(A) One-factor (unidimensional) model.
+(B) Bifactor model with 10 specific factors (one per MACE subscale).
+(C) Bifactor S-1 model with 1 specific factor (neglect).
+(D) Bifactor S-1 model with 2 specific factors (peer victimization, reverse-scoring).
+Notes: G = general maltreatment; VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; WSV = witnessing sibling vioence; WIPV = witnessing inter-parental violence; EN = emotional neglect; PN = physical neglect; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.}
     \label{fig:models}
@@
 We used two rotation criteria in order to examine the generalizability of the factor solutions across multiple rotations.
+We again relied on traditional fit indices (i.e.
+RMSEA, CFI, TLI) to evaluate the goodness-of-fit of the exploratory models to the response data of both samples.
 
@@
 
-The goodness-of-fit indices for each model and sample are summarized in Table \ref{table:diagnostics}.
+The goodness-of-fit indices for each model and sample are summarized in Table \ref{table:cfa_diagnostics}.
 According to the traditional indices, each model provided at least an acceptable fit to the response data.
@@
 smaller values indicate better fit).
-Abbreviations: RMSEA = root mean square error of approximation; CFI = comparative fit index; TLI = Tucker-Lewis index; PPP = posterior predictive p-value; $Q3_{\max}$ = maximum $Q_3$ value in the sample for the model; $Q3_{\text{crit}}$ = critical $Q_3$ value for the model; LOO-CV = leave-one-out cross-validation.}
-\label{table:diagnostics}
+Notes: RMSEA = root mean square error of approximation; CFI = comparative fit index; TLI = Tucker-Lewis index; PPP = posterior predictive p-value; $Q3_{\max}$ = maximum $Q_3$ value in the sample for the model; $Q3_{\text{crit}}$ = critical $Q_3$ value for the model; LOO-CV = leave-one-out cross-validation.}
+\label{table:cfa_diagnostics}
 \end{table}
@@
     \includegraphics[width=1.1\textwidth,center]{figures/fig02.png}
+    \captionsetup{width=1.1\textwidth}
     \caption{Standardized factor loadings from the four models estimated from the original sample.
-VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; WSV = witnessing sibling vioence; WIPV = witnessing inter-parental violence; EN = emotional neglect; PN = physical neglect; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.
+Notes: VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; WSV = witnessing sibling vioence; WIPV = witnessing inter-parental violence; EN = emotional neglect; PN = physical neglect; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.
 ** Reverse-scored items.}
     \label{fig:loadings_original}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1.1\textwidth,center]{figures/fig03.png}
+    \captionsetup{width=1.1\textwidth}
+    \caption{Standardized factor loadings from the four models estimated from the online sample.
+Notes: VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; WSV = witnessing sibling vioence; WIPV = witnessing inter-parental violence; EN = emotional neglect; PN = physical neglect; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.
+** Reverse-scored items.}
+    \label{fig:loadings_online}
 \end{figure}
@@
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=1.1\textwidth,center]{figures/fig03.png}
-    \caption{Standardized factor loadings from the four models estimated from the online sample.
-VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; WSV = witnessing sibling vioence; WIPV = witnessing inter-parental violence; EN = emotional neglect; PN = physical neglect; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.
-** Reverse-scored items.}
-    \label{fig:loadings_online}
-\end{figure}
-
 By contrast the standardized loadings on the specific factors were much smaller in magnitude, seldom exceeding loadings on the general factor.
@@
 \cmidrule(lr){4-7}\cmidrule(lr){8-11}
-Model & Subscale & PUC & ECV & $\omega$ & $\omega_{h/s}$ & H & ECV & $\omega$ & $\omega_{h/s}$ & H \\
+Model & Factor & PUC & ECV & $\omega$ & $\omega_{h/s}$ & H & ECV & $\omega$ & $\omega_{h/s}$ & H \\
 \midrule
@@
 \caption{\normalfont Model-based reliability indices for the three bifactor models.
-Abbreviations: VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; EN = emotional neglect; PN = physical neglect; WSV = witnessing sibling violence; WIPV = witnessing inter-parental violence; PeerVA = peer verbal abuse; PeerPA = peer physical abuse; ECV = explained common variance.}
+Notes: PUC = proportion uncontaminated correlations; ECV = explained common variance; VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; EN = emotional neglect; PN = physical neglect; WSV = witnessing sibling violence; WIPV = witnessing inter-parental violence; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.}
 \label{table:reliability}
@@
 
-The standardized factor loadings for the two- and three-factor solutions generated by the cf-quartimax rotation are summarized in Figure \ref{fig:efa}.
-(The corresponding factor loadings produced by the geomin rotation were similar and are presented in Figure S4.) In the two-factor solution for the original sample, there was a primary adult maltreatment factor and a secondary peer victimization factor.
+\begin{figure}[t!]
+    \centering
+    \includegraphics[width=1\textwidth,center]{figures/fig04.png}
+    \caption{\normalfont Standardized factor loadings for the 2- and 3-factor solutions produced by the cf-quartimax rotation for original sample (top row) and replication sample (bottom row).
+Notes: VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; EN = emotional neglect; PN = physical neglect; WSV = witnessing sibling violence; WIPV = witnessing inter-parental violence; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.
+Only factor loadings $\geq$ 0.30 are shown.
+** Reverse-scored items.}
+    \label{fig:efa_cf}
+\end{figure}
+
+The goodness-of-fit of the exploratory factor models to the response data of both samples are summarized in Table \ref{tab:efa_diagnostics}.
+Briefly, the fit indices indicated that all models provided an adequate fit to the data.
+The standardized factor loadings for the 2- and 3-factor solutions produced by the cf-quartimax rotation are summarized in Figure \ref{fig:efa_cf}.
+(The corresponding factor loadings produced by the geomin rotation were qualitatively similar and are presented in Figure \ref{fig:efa_geomin}.) In the 2-factor solution for the original sample, there was a primary adult maltreatment factor and a secondary peer victimization factor.
 In contrast, the replication sample produced a primary adult/peer maltreatment factor with a secondary reverse-scoring factor.
@@
 To conclude, the structure of both the two- and three-factor exploratory factor solutions resembled the structure of the bifactor S-1 model with two specific factors (model 4), providing further support for this model of responses on the MACE.
-
-\begin{figure}[t!]
-    \centering
-    \includegraphics[width=1\textwidth,center]{figures/fig04.png}
-    \caption{Caption}
-    \label{fig:efa}
-\end{figure}
 
@@
 
+\begin{table}[H]
+    \centering
+    \begin{tabular}{llccccc}
+    \toprule
+    Sample      & Model    & $\chi^2$ ($df$) & RMSEA & CFI & TLI & SRMR \\
+    \midrule
+    Original    & 2-factor & 1876 (664) & 0.049 & 0.962 & 0.957 & 0.122 \\
+                & 3-factor & 1237 (627) & 0.036 & 0.981 & 0.977 & 0.105 \\
+    \midrule
+    Replication & 2-factor & 1737 (664) & 0.052 & 0.961 & 0.957 & 0.108 \\
+                & 3-factor & 1227 (627) & 0.040 & 0.978 & 0.974 & 0.097 \\
+    \bottomrule
+    \end{tabular}
+    \caption{\normalfont Fit statistics for the exploratory factor models by sample.
+Notes: RMSEA = root mean square error of approximation; CFI = comparative fit index; TLI = Tucker-Lewis index; SRMR = standardized root mean square residual.}
+    \label{tab:efa_diagnostics}
+\end{table}
+
 \begin{figure}[H]
@@
 
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=1\textwidth,center]{figures/figS04.png}
+    \caption{\normalfont Standardized factor loadings for the 2- and 3-factor solutions produced by the geomin rotation for original sample (top row) and replication sample (bottom row).
+Notes: VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; EN = emotional neglect; PN = physical neglect; WSV = witnessing sibling violence; WIPV = witnessing inter-parental violence; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.
+Only factor loadings $\geq$ 0.30 are shown.
+** Reverse-scored items.}
+    \label{fig:efa_geomin}
+\end{figure}
+
 \end{document}

### 2022-07-31 analysis/scripts/fit_efa.R

@@ -1,5 +1,6 @@
 library(tidyverse)
 library(lavaan)
+set.seed(47404)
 # https://solomonkurz.netlify.app/post/2021-05-11-yes-you-can-fit-an-exploratory-factor-analysis-with-lavaan/
 # https://vankesteren.github.io/efast/efa_lavaan
 
@@ -14,16 +15,15 @@ data = read.csv('../data/mace.csv')
 studies = list('teicher2015', 'tuominen2022')
 
 ## Define rotations.
-rotations = list('geomin', 'cf-quartimax')
+rotations = list('cf-quartimax', 'geomin')
 
-## Define locally dependent variables.
+## Define locally dependent items
 ld = list(c('x6', 'x7', 'x8'), c('x9', 'x10', 'x11'), c('x13', 'x14'), 
           c('x15', 'x16'), c('x19', 'x20'), c('x21', 'x22', 'x23'), 
           c('x24', 'x25'), c('x33', 'x34', 'x35'), c('x36', 'x37'))
 
 ## Define settings.
-orthogonal = TRUE
-
+orthogonal = FALSE
 #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
 ### Main loop.
 #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
@@ -57,6 +57,14 @@ for (study in studies) {
     efa <- cfa(model = f2, data = df, rotation = rotation, estimator = "WLSMV",
                orthogonal = orthogonal, ordered = TRUE, std.lv = TRUE)
     
+    ## Display model fit.
+    metrics = fitMeasures(efa)
+    print(c(study, 'f2', rotation))
+    print(round(metrics[c(3, 4, 36, 17, 18, 50)], 3))
+    
+    ## Display factor correlations.
+    print(lavInspect(efa, what="cor.lv"))
+    
     ## Extract & save factor loadings.
     loadings = inspect(efa, what="std")$lambda
     write.csv(loadings, paste('..', 'stan_results', study, paste0('efa_f2_', rotation, '.csv'), sep='/'))
@@ -71,8 +79,13 @@ for (study in studies) {
                      x48 + x49 + x50 + x51 + x52'
     
     ## Fit EFA (3-factor).
-    efa <- cfa(model = f3, data = df, rotation = rotation, estimator = "WLSMV",
-               orthogonal = orthogonal, ordered = TRUE, std.lv = TRUE)
[…]

### 2022-07-31 analysis/scripts/make_fig04.py

@@ -1,0 +1,113 @@
+import os
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+import seaborn as sns
+from os.path import dirname
+from pandas import Categorical, DataFrame, concat, read_csv
+from mpl_toolkits.axes_grid1 import make_axes_locatable
+ROOT_DIR = dirname(dirname(os.path.realpath(__file__)))
+sns.set_theme(context='notebook', style='white', font='sans-serif', font_scale=1.33)
+
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+### Define data parameters.
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+
+## Define LD item pairs.
+ld = [[6,7,8], [9,10,11], [13,14], [15,16], [19,20], [21,22,23], [24,25], [33,34,35], [36,37]]
+
+## Define item ordering.
+order = ['VA','PA','NVEA','SA','WSV','WIPV','EN','PN','PeerVA','PeerPA']
+
+## Define loadings threshold.
+thresh = 0.3
+
+## Define rotation.
+rotation = 'cf-quartimax'
+
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+### Load and prepare design matrix.
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+
+## Load design data.
+design = read_csv(os.path.join(ROOT_DIR, 'data', 'design.csv'), index_col=0)
+
+## Collapse across LD items.
+for ix in ld: 
+    design = design.drop(index=ix[1:])
+    design = design.rename(index={ix[0]:'/'.join(['%s' %i for i in ix])})
+
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+### Define plot variables.
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+
+## Define palette.
+cmap = sns.diverging_palette(20, 220, sep=60, as_cmap=True)
+
+## Define labels.
+ylabels = ['Original','Replication']
+
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
[…]

### 2022-07-31 analysis/scripts/make_figS04.py

@@ -1,0 +1,113 @@
+import os
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+import seaborn as sns
+from os.path import dirname
+from pandas import Categorical, DataFrame, concat, read_csv
+from mpl_toolkits.axes_grid1 import make_axes_locatable
+ROOT_DIR = dirname(dirname(os.path.realpath(__file__)))
+sns.set_theme(context='notebook', style='white', font='sans-serif', font_scale=1.33)
+
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+### Define data parameters.
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+
+## Define LD item pairs.
+ld = [[6,7,8], [9,10,11], [13,14], [15,16], [19,20], [21,22,23], [24,25], [33,34,35], [36,37]]
+
+## Define item ordering.
+order = ['VA','PA','NVEA','SA','WSV','WIPV','EN','PN','PeerVA','PeerPA']
+
+## Define loadings threshold.
+thresh = 0.3
+
+## Define rotation.
+rotation = 'geomin'
+
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+### Load and prepare design matrix.
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+
+## Load design data.
+design = read_csv(os.path.join(ROOT_DIR, 'data', 'design.csv'), index_col=0)
+
+## Collapse across LD items.
+for ix in ld: 
+    design = design.drop(index=ix[1:])
+    design = design.rename(index={ix[0]:'/'.join(['%s' %i for i in ix])})
+
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+### Define plot variables.
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+
+## Define palette.
+cmap = sns.diverging_palette(20, 220, sep=60, as_cmap=True)
+
+## Define labels.
+ylabels = ['Original','Replication']
+
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
[…]

### 2022-07-31 manuscript/main.tex
add fig S03; completed draft of supplementary materials; down to editing

@@
 In both the original and replication samples, participants completed a number of additional self-report symptom measures.
-We do not include these in our analyses and we will therefore not discuss these measures further in the main text.
+We do not include these in our analyses and therefore do not discuss these measures further in the main text.
 For completeness, we report these secondary self-report measures in the Supplementary Materials.
@@
 EFA was performed using the \textit{lavaan} (v0.6.11; \citealt{lavaan}) package available in R.
-Factors were extracted using the geomin \citep{yates1987multivariate} and cf-quartimax rotation criteria \citep{crawford1970general}, which were selected in order to extract simple factor structures with few cross-loadings.
+Factors were extracted using the oblique geomin \citep{yates1987multivariate} and cf-quartimax rotation criteria \citep{crawford1970general}, which were selected in order to extract simple factor structures with few cross-loadings.
 We used two rotation criteria in order to examine the generalizability of the factor solutions across multiple rotations.
@@
 For all models and samples, the maximum observed $Q_3$ index was greater than the 99\% percentile critical value.
-To understand the extent of local dependence for each model and sample, we visualized the $Q_3$ values for item pairs exceeding the critical value (Figure S3).
+To understand the extent of local dependence for each model and sample, we visualized the $Q_3$ values for item pairs exceeding the critical value (Figure \ref{fig:local_dependence}).
 The unidimensional model fits exhibited many locally dependent item pairs, especially among the peer victimization and reverse-scored items, suggesting possible underfactorization.
@@
 
-The results of the model comparison are also summarized in Table \ref{table:diagnostics}.
+The results of the model comparison are also summarized in Table \ref{table:cfa_diagnostics}.
 Across samples, models 2 and 4 provided better fits to the data than models 1 and 3.
@@
 
+\subsection*{Additional self-report measures}
+
+In addition to the MACE, participants in both samples completed a number of additional self-report measures.
+Participants in the original sample completed multiple measures concerning psychiatric symptoms; these have already been reported \citep{teicher2015maltreatment}.
+The participants in the replication sample completed three additional self-report measures all assessing negative symptoms and motivation for rewards.
+Specifically, participants completed the Motivation and Pleasure scale \citep{llerena2013motivation}, the Self-evaluation of Negative Symptoms scale \citep{dollfus2016self}, and the revised Behavioral Inhibition / Behavioral Activation scale \citep{pagliaccio2016revising}.
+
+Following best practices for online research \citep{zorowitz2021inattentive}, several attention checks were embedded in the secondary self-report measures filled out by the replication sample.
+Specifically, we used infrequency items which are questions for which all  or virtually all attentive participants should provide the same response.
+Specifically, we used the following questions:
+
+\begin{enumerate}
+    \item I'm able to blink my eyes without difficulty. (all endorse)
+    \item I was motivated to write Mumfred Mumford's biography. (none endorse)
+    \item I worry about the 2001 Cricket World Cup. (none endorse)
+    \item Please select "Yes" and then select age "17". (instructed item)
+\end{enumerate}
+
+\noindent Prior to conducting the study, the infrequency items were piloted on an independent sample of participants to ensure that they elicited one dominant response.
+Participants were excluded from analysis if they responded incorrectly to one or more of these items.
+
+\pagebreak
+\subsection*{Supplementary tables \& figures}
+
 \begin{longtable}{P{0.02\linewidth}p{0.7\linewidth}|P{0.05\linewidth}P{0.05\linewidth}P{0.05\linewidth}}
@@
 Dotted lines indicate the threshold for large DIF recommended by \cite{hidalgo2014binary}.
-** Reverse-scored items \end{spacing}}
+** Reverse-scored items.
+\end{spacing}}
     \label{fig:dif_study}
@@
 Dotted lines indicate the threshold for large DIF recommended by \cite{hidalgo2014binary}.
-** Reverse-scored items \end{spacing}}
+** Reverse-scored items.
+\end{spacing}}
     \label{fig:dif_gender}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=1.05\textwidth,center]{figures/figS03.png}
+    \caption{\normalfont Yen's $Q_3$ values for item pairs exceeding the 99\% percentile critical value for each confirmatory item response model.
+Cells beneath the gray diagonal line correspond to the original sample, whereas cells above the line correspond to the replication sample.
+Notes: VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; EN = emotional neglect; PN = physical neglect; WSV = witnessing sibling violence; WIPV = witnessing inter-parental violence; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.}
+    \label{fig:local_dependence}
 \end{figure}

### 2022-08-01 analysis/scripts/fit_grmq_x2.py
remove extraneous code (almost gave me a heart attack)

@@ -338,11 +338,7 @@ for study in studies:
 
             ## Define item indices.
             ix, = np.where(K == k)
-
-            ## Compute cell proportions (observed).
-            prob_obs = Y[ix].mean()
-            C1_obs.append(prob_obs)
-
+            
             ## Compute cell proportions (predicted).
             prop_hat = Y_hat[::thin,ix].mean(axis=1)
             C1_hat.append(prop_hat)
@@ -359,11 +355,7 @@ for study in studies:
             while ix1.size != ix2.size:
                 if ix1.size > ix2.size: ix1 = ix1[np.in1d(J[ix1], J[ix2])]
                 elif ix2.size > ix1.size: ix2 = ix2[np.in1d(J[ix2], J[ix1])]
-
-            ## Compute cell proportions (observed).
-            prop_obs = binprop(Y[ix1] + 2 * Y[ix2])
-            C2_obs.append(prop_obs[-1])
-
+                
             ## Compute cell proportions (predicted).
             prop_hat = np.apply_along_axis(binprop, 1, Y_hat[::thin,ix1] + 2 * Y_hat[::thin,ix2])
             C2_hat.append(prop_hat[:,-1])

### 2022-08-06 manuscript/main.tex
getting close

@@
 
+% decrease size of references (lol)
+\renewcommand*{\bibfont}{\small}
+
 \title{The MACE is essentially unidimensional: Bifactor models of the `Maltreatment and Abuse Chronology of Exposure' scale}
@@
 
-\abstract{Your abstract here.}
+\abstract{
+
+}
 
@@
 Childhood maltreatment is an unfortunately common experience for children across the globe \citep{stoltenborgh2015prevalence}.
-It is a risk factor for multiple poor outcomes later in life, including the onset of mental illness \citep{macmillan2001childhood, green2010childhood, kessler2010childhood}, challenges in cognitive functioning \citep{kavanaugh2017neurocognitive, r2018common, su2019does}, and poorer physical health \citep{wegman2009meta, widom2012prospective, goodwin2004association}.
-Childhood maltreatment also predicts worse treatment outcomes for many psychiatric disorders \citep{nanni2012childhood, thomas2019childhood, schuckher2019history}.
+It is a risk factor for multiple poor outcomes later in life, including the onset of mental illness \citep{kessler2010childhood}, challenges in cognitive functioning \citep{su2019does}, and poorer physical health \citep{wegman2009meta}.
+Childhood maltreatment also predicts worse treatment outcomes for several psychiatric disorders \citep{nanni2012childhood, thomas2019childhood}.
 Thus, the reliable and valid assessment of childhood maltreatment is an important research goal in order to refine our understanding of the link between maltreatment and health outcomes; augment treatment strategies for victims of childhood maltreatment; and ultimately design better interventions to prevent maltreatment in the first place.
@@
 There already exist many measures of childhood abuse and maltreatment for researchers to choose from \citep{saini2019systematic}.
-One recently developed measure, the Maltreatment and Abuse Chronology of Exposure (MACE) scale \citep{teicher2015maltreatment}, has a number of notable advantages.
-The MACE is a 52-item retrospective self-report questionnaire comprised of 10 short subscales that each assess a unique type of childhood maltreatment.
-Six of these (e.g.
-parental verbal abuse, physical abuse, nonverbal emotional abuse, sexual abuse, emotional neglect, physical neglect) overlap with other common measures of abuse; the remaining four assess forms of maltreatment that are less frequently represented in other measures (e.g.
+One recently developed measure, the Maltreatment and Abuse Chronology of Exposure scale (MACE; \citealt{teicher2015maltreatment}), has a number of notable advantages.
+The MACE is a 52-item retrospective self-report scale composed of 10 subscales, each of which measures a unique type of childhood maltreatment.
+Although six of these subscales overlap with other standard measures of abuse (e.g.
+physical abuse, neglect), the remaining four are less commonly represented in other scales (e.g.
 peer victimization, witnessing domestic violence).
-The items in the MACE were carefully selected to measure adverse childhood experiences of increasing severity in order to more finely differentiate people of varying exposure levels.
-The MACE also measures the timing of maltreatment experiences, which is critical for investigating how the onset and duration of abuse shapes trajectories of development.
-The MACE provides two measures, an overall severity score (number of maltreatment experiences) and a multiplicity score (number of types of maltreatment experienced), with excellent temporal stability, good convergent validity with other standard measures, and improved predictive validity in predicting psychiatric symptom ratings \citep{teicher2015maltreatment}.
-For these reasons, the MACE has been listed as among the best measures of childhood maltreatment currently available \citep{saini2019systematic, georgieva2022systematic}.
+The items in the MACE were carefully selected to measure maltreatment of varying severity in order to finely differentiate people with differing exposures.
+The MACE also asks for the onset and duration of maltreatment experiences, which is necessary for investigating how the timing of abuse shapes development.
+The MACE was originally developed to provide two measures: a total score (number of overall maltreatment experiences) and a multiplicity score (number of types of maltreatment experienced).
+These scores have excellent temporal stability, convergent validity with other childhood maltreatment scales, and good predictive validity for a multitude of psychiatric symptoms \citep{teicher2015maltreatment}.
+For these reasons the MACE has been included as among the best measures of childhood maltreatment currently available \citep{saini2019systematic, georgieva2022systematic}.
 
 In recent years, there have been calls for childhood maltreatment research to move away from the cumulative risk framework \citep{evans2013cumulative} --- which focuses only on the number of adverse experiences --- towards a dimensional approach, which emphasizes how distinct types of maltreatment uniquely shape developmental trajectories \citep{mclaughlin2016beyond, belsky2012beyond}.
-In response, some researchers have begun to use the MACE to form new indices of maltreatment in lieu of or in addition to the severity and multiplicity scores.
-For example, some have investigated the unique contributions of individual differences on all 10 MACE subscale scores to the development of psychopathology \citep{schalinski2015type, gerke2018childhood, schalinski2019early}.
-Following a prominent dimensional model of maltreatment \citep{mclaughlin2014childhood}, other groups have made composite indices of threat and deprivation experiences using items in the MACE that respectively assess abuse and neglect experiences \citep{schalinski2018defining, schalinski2019environmental, teicher2018differential}.
-Though these alternatives maltreatment scores may yield new insight, their psychometric properties have not been extensively investigated.
+In response, some researchers have begun to use the MACE to form new indices of maltreatment in lieu of, or in addition to, the total and multiplicity scores.
+For example, some have used all 10 MACE subscale scores to study the unique contributions of each type of maltreatment to the development of psychopathology \citep{schalinski2015type, gerke2018childhood, schalinski2019early}.
+Inspired by a prominent dimensional model of maltreatment \citep{mclaughlin2014childhood}, others have made new subscale scores indexing threat and deprivation experiences using the items in the MACE that respectively measure abuse and neglect \citep{schalinski2018defining, schalinski2019environmental, teicher2018differential}.
+The psychometric properties of these new subscale scores have not been thoroughly investigated.
 
-One critical challenge in measuring childhood maltreatment is the high co-occurrence adverse experiences.
-That is, children experiencing one form of maltreatment (e.g.
-physical abuse) are likely to experience one or more other types (e.g.
-emotional abuse; \citealt{green2010childhood, mclaughlin2012childhood, marques2021risk, mersky2017rethinking}).
-As a consequence, isolating the unique contribution of one type of maltreatment to the development of some negative outcome is difficult insofar as any one type of adversity is likely confounded with general maltreatment.
-Concretely, this means that researchers who want to study the effects of particular types of maltreatment using subscale scores or composite indices need to verify those measures reflect a construct mostly independent of overall maltreatment, lest they draw spurious conclusions.
+A critical challenge in studying the consequences of childhood maltreatment is the high co-occurrence of adverse experiences.
+That is, children who have experienced any single form of maltreatment are likely to have experienced multiple other types \citep{dong2004interrelatedness, herrenkohl2009assessing, kessler2010childhood}.
+As a consequence, isolating the unique contribution of one type of maltreatment to some outcome is difficult insofar as any one type of adversity is likely confounded with overall maltreatment.
+Concretely, this means that researchers intending to study the effects of particular types of maltreatment using subscale scores need to verify that those measures actually reflect a construct unique from from overall maltreatment, lest they draw spurious conclusions.
 
-Bifactor models, one type of hierarchical factor models, are well-suited for addressing this issue \citep{bornovalova2020appropriate}.
-They assume that the covariance structure of a set of items can be accounted for by a general factor, reflecting shared variance across all items, and one or more specific factors, reflecting additional shared variance among subsets of items \citep{Reise2012-ql}.
-Crucially, bifactor models can be used to calculate model-based reliability statistics, including the omega hierarchical ($\omega_h$) and omega subscale ($\omega_s$) indices, which respectively measure the proportion of variance in total and subscale scores attributable to the general factor and specific factors \citep{reise2013applying, rodriguez2016evaluating}.
-Applied to the MACE scale, these indices can inform researchers as to whether subscale scores or composite indices for particular types of maltreatment are interpretable as such, or if they instead predominantly reflect overall maltreatment.
+Bifactor models, a type of hierarchical factor model, are well-suited for addressing this issue \citep{bornovalova2020appropriate}.
+They assume that the covariation in responses to a set of items can be accounted for by a \emph{general} factor, reflecting shared variance across all items, and one or more \emph{specific} factors, reflecting additional shared variance among subsets of items \citep{Reise2012-ql}.
+Crucially, bifactor models can be used to calculate model-based reliability statistics, including the omega hierarchical ($\omega_h$) and omega subscale ($\omega_s$) indices \citep{reise2013scoring}, which respectively measure the proportion of variance in total and subscale scores attributable to the general factor and specific factors.
+Applied to the MACE, these indices can help determine whether subscale scores measuring particular types of maltreatment are interpretable as such, or if they instead chiefly reflect overall maltreatment.
 
@@
 The participants in the current study (N=1633 total) belonged to one of two samples.
-The first sample of participants (N=1051) are from the original development of the MACE scale \citep{teicher2015maltreatment}.
+The first sample of participants (N=1051) are from the original development of the MACE \citep{teicher2015maltreatment}.
 The full recruitment details for this sample have been reported elsewhere \citep{teicher2015maltreatment}.
@@
 The demographics of the two samples are summarized in Table \ref{tab:demographics}.
-In both samples, the majority of participants identified as women; however, the original sample was comprised of proportionally more woman than the replication sample ($z$ = 4.142, $p$ < 0.001).
-Moreover, participants in the original sample were younger on average than participants in the replication sample ($t$ = -19.313, $p$ < 0.001).
-Finally, the majority of participants in both samples identified as White, and not as Hispanic or Latino.
-The proportion of participants identifying as White was not statistically different across samples ($z = 0.417$, $p = 0.677$); neither was the proportion of participants identifying as Hispanic or Latino ($z$ = -1.241, $p$ = 0.214).
+The majority of participants in both samples identified as women, though the original sample was composed of proportionally more woman than the replication sample ($z$ = 4.142, $p$ < 0.001).
+Participants in the original sample were also younger on average ($t$ = -19.313, $p$ < 0.001).
+The majority of participants in both samples identified as White and not as Hispanic or Latino.
+The proportion of participants identifying as White was not significantly different across samples ($z = 0.417$, $p = 0.677$), nor was the proportion of participants identifying as Hispanic or Latino ($z$ = -1.241, $p$ = 0.214).
 
@@
 
-All participants completed the 52-item version of the MACE scale \citep{teicher2015maltreatment}.
-The scale is a retrospective measure of the severity and timing of exposure to 10 types of childhood maltreatment.
+All participants completed the 52-item version of the MACE \citep{teicher2015maltreatment}.
+The scale is a retrospective self-report measure of the severity and timing of exposure to 10 types of childhood maltreatment.
 For each item, participants endorse whether they ever experienced a particular event during childhood (e.g.
 ``Parents or guardians intentionally pushed, grabbed, shoved, slapped, pinched, punched or kicked you'') and, if so, at what ages (up to age 18).
-Here we only analyze the binary endorsement responses (Yes = 1, No = 0) and not the chronology data.
-Six of the 52-items --- evenly divided between the emotional and physical neglect subscales --- are positively valenced (e.g.
-``One or more individuals in your family helped you feel important or special'') and reverse-scored.
+Here we only analyze the dichotomous endorsement responses (Yes = 1, No = 0) and not the chronology data.
+Six of the 52-items --- evenly divided between the emotional and physical neglect subscales --- are positively valenced and reverse-scored (e.g.
+``One or more individuals in your family helped you feel important or special'').
 For these items we follow the scoring convention set by \cite{teicher2015maltreatment}: a score of 1 is assigned to a response only in the absence of a positive event for all 18 years of childhood.
-This is in contrast to the remaining 46 items in the scale, where a score of 1 is assigned to a response if maltreatment or abuse occurred in at least one year of childhood.
+This is in contrast to the remaining items in the scale, where a score of 1 is assigned to a response if maltreatment or abuse occurred in at least one year of childhood.
+
+% (i.e. verbal abuse, physical abuse, nonverbal emotional abuse, sexual abuse, emotional neglect, and physical neglect by parents or guardians)
 
@@
 We do not include these in our analyses and therefore do not discuss these measures further in the main text.
-For completeness, we report these secondary self-report measures in the Supplementary Materials.
+For completeness, we report these secondary self-report measures in the supplementary materials.
 
@@
 
-To ensure data quality in the replication sample, we relied on attention checks embedded in the self-report measures to identify and remove participants engaging in careless or insufficient effort responding \citep{zorowitz2021inattentive}.
+To ensure data quality in the replication sample, we relied on attention checks embedded in the self-report measures to identify and remove participants engaging in careless or insufficient effort responding (\citealt{zorowitz2021inattentive}; see the supplementary materials for the attention checks).
 The data from N=64 (9.9\%)  participants in the replication sample were excluded for failing one or more of these attention checks, leaving the data from N=582 participants for analysis.
@@
 Prior to analysis, we investigated whether it was appropriate to combine response data from the original and replication samples.
-We first compared maltreatment endorsement rates for each item across samples.
-The Spearman-rank correlation of endorsement rates across samples was large ($\rho$ = 0.951, p < 0.001), indicating excellent correspondence in the relative prevalence of childhood maltreatment across samples.
-When we compare the distribution of person-level sum scores across samples,  however, we found that the number of maltreatment exposures was greater on average in the replication sample than in the original sample ($p_1$ = 0.185, $p_2$ = 0.219, $t$ = -4.611, p < 0.001).
+We first compared maltreatment endorsement rates for each item between samples.
+The Spearman-rank correlation of endorsement rates across samples was large ($\rho$ = 0.951, p < 0.001), indicating excellent correspondence in the relative prevalence of maltreatment types across samples.
+When we compared the distribution of person-level total scores across samples, however, we found that scores were greater on average in the replication sample than in the original sample ($p_1$ = 0.185, $p_2$ = 0.219, $t$ = -4.611, p < 0.001).
 
-A mean shift in overall endorsement rates is not necessarily an indication that measurement invariance has been violated.
-One possibility is simply that participants in the replication sample had simply experienced more childhood maltreatment than participants in the original sample.
-A second possibility is that the differences in the demographic composition of the two samples (e.g.
-gender, age) moderate the differences in endorsement rates.
-To investigate these possibilities, we tested for uniform differential item functioning (DIF) using logistic regression \citep{rogers1993comparison}, in which a participant's endorsement response for an item was predicted by their rest score (i.e.
-their sum score for all other items), sample membership, gender, and age.
+An overall shift in overall endorsement rates by sample is not necessarily evidence for a violation of measurement invariance.
+One possibility is the replication sample had simply experienced more adversity than the original sample.
+A second is that the differences in the demographic composition of the two samples (e.g.
+gender, age) explain the differences in endorsement rates.
+To investigate these possibilities, we tested for uniform differential item functioning (DIF) using logistic regression, in which a participant's endorsement response on an item is regressed against their rest score (i.e.
+their total score for all other items), sample membership, gender, and age.
 
 Using the cutoffs recommended by \cite{hidalgo2014binary}, we identified 11 of 52 items on the MACE that exhibited large DIF by study (Figure \ref{fig:dif_study}).
-Interestingly these included all six of the reverse-scored items.
+Interestingly these included all six reverse-scored items.
 We also found an additional 11 items that exhibited large DIF by gender (Figure \ref{fig:dif_gender}).
-Five of these items concerned sexual abuse (endorsed by more women than men), and another five concerned peer physical abuse (endorsed by more men than women).
+Five of these items measured sexual abuse (endorsed by more women than men), and another five measured peer physical abuse (endorsed by more men than women).
 We identified no items that exhibited large DIF by age.
-Given the number of items exhibiting DIF by sample membership, we elected to model each sample separately.
+Given the number of items exhibiting DIF by sample membership, we decided to analyze each sample separately.
 
@@
 The MACE contains a total of 12 item pairs and triplets characterized by response dependence (Table \ref{tab:dependence}).
-The items in some of these sets are structurally dependent, where a response to one item requires, or strongly implies, a certain response on the second.
+The items in some of these sets are structurally dependent, where a response to one item requires or strongly implies a certain response on the second.
 For example, a person endorsing item 8 (``Parents or guardians hit you so hard that you received medical attention'') is all but certain to endorse item 7 (``Parents or guardians hit you so hard that it left marks for more than a few minutes'').
 Other items are dependent by virtue of redundancy.
-One such example is item 36 (``Peers forced you to engage in sexual activity against your will'') and 37 (``Peers forced you to do things sexually that you did not want to do``); here the same question is asked twice using slightly different language.
+One such example is items 36 (``Peers forced you to engage in sexual activity against your will'') and 37 (``Peers forced you to do things sexually that you did not want to do``); here the same question is asked twice using slightly different language.
 
-For this study, we were concerned about response dependence for two reasons.
+We were concerned about response dependence for two reasons.
 First, it can inflate the estimates of item discrimination parameters, making items appear more informative than they actually are \citep{marais2008formalizing}.
-Second, response dependence can inflate the estimates of item loadings on specific factors in bifactor models, giving false support for a larger number of factors than necessary \citep{reise2013applying}.
-Following \cite{marais2008formalizing}, we therefore converted each set of dependent items into a single polytomous item.
-Thus, the MACE was reduced from 52 to 39 items in total.
+Second, response dependence can inflate the estimates of item loadings on specific factors in bifactor models, giving false support for models with a greater number of factors than needed \citep{reise2013applying}.
+Following \cite{marais2008formalizing}, we therefore converted each set of dependent items into a single polytomous item thereby reducing the MACE from 52 to 39 items in total.
 
@@
 
-To begin our investigation of the MACE factor structure, we fit a series of confirmatory item factor models to the response data from each sample separately.
+To investigate the factor structure of the MACE, we fit a series of confirmatory item factor models separately to the response data from each sample.
 The basis of each model is the graded response model for ordered polytomous data \citep{samejima1997graded}.
 The factor structure for each of the four models is depicted in Figure \ref{fig:models}.
-The first model we fit is a one-factor (undimensional) model whereby all items load onto a general maltreatment factor (Figure \ref{fig:models}a).
-The second is a bifactor model, in which all items load onto a general maltreatment factor and also one of 10 specific factors corresponding to their subscale membership (Figure \ref{fig:models}b).
+The first model is a one-factor (undimensional) model where all items load onto a general maltreatment factor (Figure \ref{fig:models}a).
+The second is a bifactor model where all items load onto a general maltreatment factor and also one of 10 specific factors based on their subscale membership (Figure \ref{fig:models}b).
 
 The third model we fit investigated a conceptual model of maltreatment that distinguishes between threat and deprivation as distinct forms of childhood adversity \citep{mclaughlin2014childhood}.
-Initially we fit a bifactor model in which all items loaded onto a general maltreatment factor one either the threat or deprivation specific factors.
+Initially we fit a bifactor model where all items loaded onto a general maltreatment factor and either the threat or deprivation specific factors.
 However, we observed vanishing item loadings on the threat specific factor.
@@
 
-The fourth and final model we fit was designed provide a more parsimonious factor structure for the MACE.
+The fourth and final model proposes a more parsimonious factor structure for the MACE.
 Items in the MACE can be neatly divided according to the source of adversity: maltreatment by adults at home (e.g.
@@
 
-We relied on multiple indices to evaluate the fit of the models to the data.
-First we calculated three traditional goodness-of-fit statistics using the ordinal $M_2$ statistic \citep{cai2013limited, maydeu2013goodness}.
+We relied on multiple indices to judge the fit of the models to the data.
+First we calculated three traditional goodness-of-fit statistics using the ordinal $M_2$ statistic \citep{cai2013limited}.
 $M_2$ is asymptotically $\chi^2$-distributed with $\kappa − \nu$ degrees of freedom ($df$), where $\kappa$ is the reduced number of first- and second-order marginal residuals (i.e.
@@
 none vs.
-some maltreatment).
+any maltreatment).
 Using the $M_2$ statistic, we calculated the root mean square error of approximation (RMSEA), comparative fit index (CFI), and Tucker-Lewis index (TLI) for each model fit.
@@
 
-Given the questionable utility of SEM-based fit indices for item response models (e.g.
+Given the limitations of SEM-based fit indices for item response models (e.g.
 \citealt{reise2014evaluating}), we calculated two additional fit indices.
-First, we employed posterior predictive model checking and the $\chi^2_{NC}$ discrepancy measure based on the overall severity score distribution \citep{sinharay2006posterior}.
-This measure compares the observed and model-predicted proportion of participants at each overall severity score level.
-Second, we calculated Yen's $Q_3$ statistic for each pair of items to detect local dependence \citep{yen1984effects}.
-For each model and sample, we also calculated the critical $Q_3$-value above which local dependence is indicated  \citep{christensen2017critical}.
+First, we used posterior predictive model checking to calculate a $\chi^2_{NC}$ discrepancy measure based on the total score distribution \citep{sinharay2006posterior}.
+This measure compares the observed and model-predicted proportion of participants at each total score level.
+Second, we calculated Yen's $Q_3$ statistic for each pair of items to detect local dependence \citep{yen1984effects} and also the critical $Q_3$-value above which local dependence is indicated  \citep{christensen2017critical}.
 Briefly, for each model and sample, we simulated 1000 locally independent datasets using the posterior distribution of the model parameters.
@@
 
-To evaluate the reliability of the MACE total and subscale/composite scores, we computed several model-based reliability indices using the factor loading estimates from the bifactor models.
+To evaluate the reliability of the MACE total and subscale scores, we computed several model-based reliability indices using the factor loading estimates from the bifactor models.
 First, we calculated coefficient omega ($\omega$; \citealt{mcdonald1999test}), which quantifies the proportion of reliable score variance attributable to all factors (i.e.
 both group and specific factors).
-Next we calculated coefficient omega hierarchical ($\omega_h$) and coefficient omega subscale ($\omega_s$;  \citealt{reise2013applying, rodriguez2016evaluating}).
+Next we calculated coefficient omega hierarchical ($\omega_h$) and coefficient omega subscale ($\omega_s$;  \citealt{reise2013scoring}).
 Whereas $\omega_h$ measures the fraction of total score variance attributable to the general factor, $\omega_s$ measures the fraction of subscale score variance attributable to a specific factor after controlling for the general factor.
@@
 Values of $\omega_h$ approaching 1 indicate the general factor is the primary source of reliable variance in a total score.
-When $\omega_h$ > 0.80, the total score may be considered an essentially unidimensional reflection of the general factor (even if the data are multidimensional;  \citealt{rodriguez2016applying}).
-So too values of $\omega_s$ approaching 1 indicate a specific factor, and not the general factor, is the primary source of reliable variance in a subscale score.
-\cite{canivez2016bifactor} suggests that an acceptable $\omega_s$ value is 0.50 and that >0.70 is desirable.
-In contrast, values of $\omega_s$ smaller than 0.50 preclude meaningful interpretation of subscale scores as reflecting a specific factor \citep{gignac2013bifactor}.
-
-To assess the unidimensionality of the MACE scale, we also calculated the explained common variance (ECV; \citealt{sijtsma2009use}) and the proportion of uncontaminated correlations (PUC; \citealt{reise2013multidimensionality}).
+When $\omega_h$ > 0.80, the total score may be considered an essentially unidimensional reflection of the general factor even if the data are multidimensional \citep{rodriguez2016applying}.
+So too values of $\omega_s$ approaching 1 indicate the specific factor, and not the general factor, is the primary source of reliable variance in a subscale score.
+\cite{canivez2016bifactor} suggested that an acceptable $\omega_s$ value is 0.50 and that >0.70 is desirable.
+In contrast, values of $\omega_s$ smaller than 0.50 preclude the interpretation of subscale scores as reflecting a specific factor \citep{gignac2013bifactor}.
+To assess the unidimensionality of the MACE, we also calculated the explained common variance (ECV; \citealt{sijtsma2009use}) and the proportion of uncontaminated correlations (PUC; \citealt{reise2013multidimensionality}).
 The ECV is the ratio of the variance explained by the general factor divided by the total common variance of a scale; it therefore measures the contribution of the general factor relative to the specific factors.
@@
 In contrast, small H values suggest a poorly defined latent variable where factor loadings are liable to change across studies.
-Because we are primarily focused on measuring the reliability of MACE total and subscale/composite scores, we report the H values below but do not discuss them further.
+Because we are primarily focused on measuring the reliability of MACE total and subscale scores, we report the H values below but do not discuss them further.
 
@@
 In addition to confirmatory factor analysis, we also we also performed exploratory factor analysis (EFA) on the response data from each sample separately.
-The goal of the EFA was to observe what factor structure would emerge with no \emph{a priori} restrictions \citep{schmitt2018selecting}.
+The goal of the EFA was to observe what factor structure would emerge with no \emph{a priori} restrictions.
 We decided to extract two- and three- factor solutions based on the total number of factors in the bifactor S-1 models.
@@
 The goodness-of-fit indices for each model and sample are summarized in Table \ref{table:cfa_diagnostics}.
-According to the traditional indices, each model provided at least an acceptable fit to the response data.
-All RMSEA value were <0.05 and the majority of CFI and TLI values were >0.95.
-The only exception were for the fits of models 1 and 3 to the original sample data, which exhibited only acceptable fits (i.e.
-CFI/TLI > 0.90).
+According to the traditional indices, each model provided at least an acceptable fit to the data.
+All RMSEA value were <0.05 and most CFI and TLI values were >0.95.
+The only exception were for the fits of models 1 and 3 to the original sample data where CFI/TLI > 0.90.
 In general, model fits were better for the replication sample data than for the original sample.
-Furthermore, all of the posterior predictive p-values corresponding to the $\chi^2_{NC}$ discrepancy measure were within the acceptable range.
-Thus all models were sufficiently able to reproduce the distribution of overall severity scores for each sample.
+Furthermore, none of the posterior predictive p-values corresponding to the $\chi^2_{NC}$ discrepancy measure exceeded the critical values indicating that all models were able to reproduce the observed distribution of total scores for each sample.
 
@@
 Next we inspected the $Q_3$ indices for evidence of local dependence.
-For all models and samples, the maximum observed $Q_3$ index was greater than the 99\% percentile critical value.
-To understand the extent of local dependence for each model and sample, we visualized the $Q_3$ values for item pairs exceeding the critical value (Figure \ref{fig:local_dependence}).
-The unidimensional model fits exhibited many locally dependent item pairs, especially among the peer victimization and reverse-scored items, suggesting possible underfactorization.
-By comparison, models 2 -- 4 exhibited many fewer locally dependent item pairs.
-Of these, the reasons for dependence were clear.
-For example, the dependence between items 17 (``Parents made inappropriate sexual comments or suggestions to your sibling or stepsibling'') and 18 (``Parents touched or fondled your sibling or stepsibling in a sexual way'') present for all models is easily understood as surface local dependence.
+For all models and samples, the maximum observed $Q_3$ index was greater than the critical $Q_3$ value.
+To gauge the extent of local dependence, we visualized the $Q_3$ values for item pairs exceeding the critical value for each model and sample (Figure \ref{fig:local_dependence}).
+The unidimensional model fits exhibited many locally dependent item pairs --- especially among the peer victimization and reverse-scored items --- suggesting possible underfactorization.
+By comparison, the three other models exhibited many fewer locally dependent item pairs.
+Of these pairs, the reasons for dependence were apparent.
+For example, the dependence between items 17 (``Parents made inappropriate sexual comments or suggestions to your sibling'') and 18 (``Parents touched or fondled your sibling in a sexual way'') observed in all models is almost certainly the result of measuring highly similar content.
 Regardless, the estimated item discrimination parameters for each model fit seldom exceeded the normal range ($\alpha \leq 4$).
-That is, the degree of local dependence present in the models is not yielding substantial bias in parameter estimation \citep{edwards2018diagnostic}.
-Thus, we conclude that all model fits exhibit at least an acceptable fit to the response data.
+That is, the level of local dependence present in the models did not cause substantial bias in parameter estimation \citep{edwards2018diagnostic}.
+Thus, we conclude that all model fits provide at least an acceptable fit to the response data.
 
@@
 In the following sections, we interpret the fit of each confirmatory item response model.
-To ease interpretation, we have reproduced the standardized factor loadings of each model for the original sample in Figure \ref{fig:loadings_original} and for the replication sample in Figure \ref{fig:loadings_online}.
+To facilitate interpretation, we have reproduced the standardized factor loadings of each model for the original sample in Figure \ref{fig:loadings_original} and for the replication sample in Figure \ref{fig:loadings_online}.
 The model-based reliability indices for the bifactor models are summarized in Table \ref{table:reliability}.
@@
 
-\subsubsection{Model 1: Unidimensional model of the MACE}
+\subsubsection{Model 1: One-factor (unidimensional) model of the MACE}
 
@@
 The standardized factor loadings were correspondingly large in magnitude.
-In the original sample, factor loadings ranged between 0.279 and 0.864, with an average value of 0.597 (95\% HDI = 0.579 -- 0.614); in the replication sample, factor loadings ranged between 0.359 and 0.814, with an average value of 0.600 (95\% HDI = 0.578 -- 0.621).
-Most of the items had loadings greater than 0.50 (original: 80.2\%, replication: 80.0\%), and thereby appear to be good indicators of general maltreatment.
-The rank correlation of factor loadings between samples was $\rho$ = 0.565 (95\% HDI = 0.414 -- 0.711), indicating only moderate agreement on the relative loading of items onto the unitary factor across samples.
+In the original sample, factor loadings ranged between 0.279 and 0.864 with an average value of 0.597 (95\% HDI = 0.579 -- 0.614); in the replication sample, factor loadings ranged between 0.359 and 0.814 with an average value of 0.600 (95\% HDI = 0.578 -- 0.621).
+Most items had loadings greater than 0.50 (original: 80.2\%, replication: 80.0\%), and thereby appear to be good indicators of general maltreatment.
+The rank correlation of factor loadings between samples was $\rho$ = 0.565 (95\% HDI = 0.414 -- 0.711), indicating only moderate agreement on the relative loading of items onto the common factor across samples.
 
-To provide one perspective on the dimensionality of the MACE, we next inspected the first six eigenvalues of the polychoric correlation matrix for each sample.
+To evaluate the dimensionality of the MACE, we inspected the first six eigenvalues of the polychoric correlation matrix for each sample.
 In the original sample, these values were 14.62, 3.77, 2.64, 2.11, 1.79, and 1.35; in the replication sample, they are 13.27, 4.43, 2.48, 2.04, 1.81, and 1.56.
-The first-to-second eigenvalue ratios are 3.84 and 3.02, respectively, just above the 3-to-1 criterion often cited in evaluating whether item response data are essentially unidimensional \citep{embretson2013item}.
+The first-to-second eigenvalue ratios are 3.84 and 3.02, respectively, just above the 3-to-1 criterion often cited in determining whether item response data are essentially unidimensional \citep{embretson2013item}.
 
-In sum, the results provide evidence on average in favor a unidimensional model of the MACE.
-The factor loadings suggest the presence of a single factor running through the items, as do the eigenvalues of the polychoric correlation matrix.
-Conversely, the degree of local dependence present among in the unidimensional model fits --- especially among the peer victimization items --- suggest there may be secondary dimensions to the MACE.
+In sum, the fits of the one-factor models provide partial evidence in support of a unidimensional account of the MACE.
+The items exhibited uniformly moderate-to-large loadings onto the common factor.
+Moreover, the eigenvalues of the polychoric correlation matrix suggest one principal factor.
+However, the degree of local dependence present among in the unidimensional model fits (especially among the peer victimization items) suggest there may be additional dimensions to the MACE.
 We turn next to the multidimensional confirmatory models to explore these possibilities.
@@
 
-Next we inspect the bifactor model with 10 specific factors, one per MACE subscale (model 2).
+Next we inspect the bifactor model with 10 specific factors, i.e.
+one per MACE subscale (model 2).
 In the original sample, the loadings on the general maltreatment factor ranged between 0.235 and 0.864, with an average value of 0.575 (95\% HDI = 0.556 -- 0.593).
@@
 Whereas the MACE overall severity score is reliable and a valid measure of general maltreatment, the MACE subscale scores are less reliable and cannot be meaningfully interpreted as uniquely reflecting exposure to specific types of abuse.
-The question remains, however, if there are other valid composite scores to be derived from the MACE in addition to the overall severity score.
+The question remains, however, if there are other valid subscale scores to be derived from the MACE in addition to the overall severity score.
 
@@
 Unsurprisingly, the $\omega_h$ values indicated the majority of reliable variance in severity score is attributable to the general factor (original: $\omega_h$ = 0.928; replication: $\omega_h$ = 0.922).
-The reliability of the ``neglect'' composite score was similarly good (original: $\omega$ = 0.877; replication: $\omega$ = 0.921).
-The $\omega_s$ values, however, indicated the composite score contained less than half reliable variance independent of the general factor (original: $\omega_h$ = 0.384; replication: $\omega_h$ = 0.375).
+The reliability of the ``neglect'' subscale score was similarly good (original: $\omega$ = 0.877; replication: $\omega$ = 0.921).
+The $\omega_s$ values, however, indicated the subscale score contained less than half reliable variance independent of the general factor (original: $\omega_h$ = 0.384; replication: $\omega_h$ = 0.375).
 
-In summary then, the results do not support the use of composite threat and deprivation scores as derived from the abuse and neglect items on the MACE.
+In summary then, the results do not support the use of threat and deprivation subscale scores as derived from the abuse and neglect items on the MACE.
 The minority of reliable variance unique from general maltreatment prevents the deprivation score from being interpreted as such.
-Furthermore, the items comprising deprivation composite index seem to be confounded by a reverse-scoring methods factor.
+Furthermore, the items comprising deprivation subscale seem to be confounded by a reverse-scoring methods factor.
 
@@
 Again the majority of reliable variance in severity score was attributable to the general factor.
-The peer victimization composite score exhibited good reliability across samples (original: $\omega$ = 0.889; replication: $\omega$ = 0.868).
+The peer victimization subscale score exhibited good reliability across samples (original: $\omega$ = 0.889; replication: $\omega$ = 0.868).
 Notably, the corresponding $\omega_s$ values were 0.569 for the original sample and 0.493 for the replication sample.
@@
 
-Briefly, we note that the composite scores formed by the reverse-scored items was also reliable (original: $\omega$ = 0.839; replication: $\omega$ = 0.915).
+Briefly, we note that the subscale scores formed by the reverse-scored items was also reliable (original: $\omega$ = 0.839; replication: $\omega$ = 0.915).
 Moreover, the corresponding $omega_s$ values were similar in magnitude to those observed for the peer victimization scores (original: $\omega_s$ = 0.584; replication: $\omega_s$ = 0.498).
@@
 
-We also calculated model-based reliability indices of the MACE total and subscale/composite scores using the bifactor model fits.
+We also calculated model-based reliability indices of the MACE total and subscale scores using the bifactor model fits.
 We found that the MACE total score has excellent reliability and is a valid measure of general maltreatment.
 In contrast, we found that the MACE subscale scores were not valid measures of specific forms of maltreatment; in all but one example, the majority of reliable variance in subscale scores was attributable to the general factor, thereby precluding meaningful interpretation of the MACE subscale scores as such.
-The same was true of the composite deprivation scores, derived from the neglect items on the MACE, which also possessed little reliable variance independent of the general factor and was contaminated by a reverse-scoring methods factor.
-In sum, whereas the MACE total score is reliable and valid, composite scores derived from a subset of items are seldom interpretable as such and instead reflect general maltreatment.
+The same was true of the deprivation subscale scores, derived from the neglect items on the MACE, which also possessed little reliable variance independent of the general factor and was contaminated by a reverse-scoring methods factor.
+In sum, whereas the MACE total score is reliable and valid, subscale scores derived from a subset of items are seldom interpretable as such and instead reflect general maltreatment.
 
@@
 This model of responses on the MACE was further supported by the exploratory factor analyses, where the 3-factor solution found a similar partitioning of the items.
-Notably, across samples, composite scores derived from the peer victimization items were both reliable and interpretable insofar that the majority of their variance were attributable to the peer victimization factor even after accounting for the general maltreatment factor.
+Notably, across samples, subscale scores derived from the peer victimization items were both reliable and interpretable insofar that the majority of their variance were attributable to the peer victimization factor even after accounting for the general maltreatment factor.
 It therefore may be valid to derive two score measures from the MACE, maltreatment by adults and by peers, but further research is needed to properly validate these measures.
@@
 
-Our results are also worth considering alongside previous studies that used subscale and/or composite scores from the MACE to investigate the link between specific forms of childhood maltreatment and assorted psycho-biological outcomes.
+Our results are also worth considering alongside previous studies that used subscale scores from the MACE to investigate the link between specific forms of childhood maltreatment and assorted psycho-biological outcomes.
 For example, researchers have looked at the relationship between MACE subscale scores and risk of depression \citep{gerke2018childhood}, dissociative symptoms, \citep{schalinski2015type}, and cortisol concentration \citep{schalinski2019early}.
-Still others have derived composite threat and deprivation scores from the MACE to study their associations with cognitive functioning \citep{schalinski2018defining} and hippocampal volume \citep{teicher2018differential}.
+Still others have derived threat and deprivation subscale scores from the MACE to study their associations with cognitive functioning \citep{schalinski2018defining} and hippocampal volume \citep{teicher2018differential}.
 Our results are unequivocal that these measures principally reflect general maltreatment and not particular forms of adversity.
 This presents a challenge for interpreting the results of those studies.
-It is worth remarking that several of these studies attempted to overcome these issues by simultaneously included all of the MACE subscale/composite scores in their models using more sophisticated multivariate analyses (e.g.
+It is worth remarking that several of these studies attempted to overcome these issues by simultaneously included all of the MACE subscale scores in their models using more sophisticated multivariate analyses (e.g.
 random forest regression).
@@
 
-%Worth further consideration is the peer victimization factor.
-Correlation between the bullying factor and the general factor was moderate.
-In accordance, a previous report on two prospective cohorts estimated that around 40\% of children who were maltreated were also bullied \citep{lereya2015adult}.
-The authors of that study speculated that there may be common risk factors for childhood maltreatment and bullying such as family instability and poor social skills.
-On the other hand, the authors also hypothesized that maltreatment may make children more susceptible to bullying by, for instance, disrupting the child's ability to regulate emotions.
-Similar to childhood maltreatment by household members, bullying by peers is a common childhood experience \citep{world2020spotlight} and has detrimental effects on mental health independent of other types of childhood maltreatment \citep{singham2017concurrent, lereya2015adult}.
-Separating peer bullying as its own factor in the MACE may therefore offer a more granular view of the harmful childhood experiences of the responders.
-In the future, the peer bullying factor should be validated against other restrospective bullying measures such as the California Bullying Victimization Scale \citep{green2018initial}.
-Finally, the MACE does not assess cyberbullying which is an increasingly common form of bullying.
+One might question whether it is at all appropriate to predict childhood maltreatment data using common factor models (e.g.
+bifactor models).
+The core assumption of common factor models is that the direction of causality is from the factor to the indicators.
+That is, variation in item responses (e.g.
+low mood, anergia) are caused by or \emph{relfect} variation in the shared factor (e.g.
+depression).
+This is in contrast to composite variable models where the direction of causality is reversed.
+That is, the indicators (e.g.
+income, education) cause or \emph{form} the latent variable (e.g.
+socioeconomic status).
+As a consequence, the models are differentiable in two keys ways \citep{bollen1991conventional, bollen2011three}.
+First, covariation among indicators is expected under a common factor model due to the influence of a shared common factor; conversely, covariation among indicators is not required (though may still occur) under a composite variable model.
+Second, indicators under the common factor model, but not the composite variable model, are expected to share causal antecedents because they reflect a common source.
+Multiple authors have argued that adverse childhood experiences should (often) be treated as formative, not reflective \citep{netland2001assessment, layne2010unpacking}.
+Indeed it is the occurrence of abuse and neglect that defines childhood maltreatment, not the other way around.
+Moreover, theoretical models of the consequences of childhood maltreatment neither require that adverse experiences covary nor assume they share causal antecedents.
+
+For the purposes of explaining patterns of responding on the MACE, we think we are justified in using a common factor model.
+The vast majority of items on the MACE measure maltreatment by parents or other adults in the home.
+Thus, the homogeneity of these indicators makes it plausible that they reflect a common source (e.g.
+maladaptive parenting, household dysfunction), as does their pattern of covariation as evidenced by strong loadings onto a common factor.
+Moreover, there is theoretical support for a model in which a common source like maladaptive parenting, itself caused by a diversity of parent level (e.g.
+emotional lability, impulse control, history of childhood abuse) and child-parent level factors, resulting in a multiple of forms of maltreatment \citep{belsky1993etiology}.
+This is empirically supported insofar that multiple forms of maltreatment share overlapping risk factors \citep{stith2009risk, mulder2018risk, assink2019risk}.
+Furthermore, the MACE is a retrospective self-report measure.
+Prospective and retrospective measures do not have perfect agreement \citep{patten2015retrospective, reuben2016lest}, and thus must reflect other factors related to the participants including individual differences in current state, memory, appraisal, personality, and biases in discloure \citep{susser2012still, reuben2016lest, colman2016consistency}.
+
+This in turn raises another question: whether it is reasonable for indicators of maladaptive parenting and peer victimization on the MACE to share a common factor.
+At first blush, one might reasonably assume that the causes of the former would be independent of the causes of the latter.
+However, there is ample evidence that children who have been maltreated are at greater risk for subsequent peer victimization \citep{bolger2001developmental, benedini2016cycle}.
+The reasons for this association are manifold.
+For example, maltreatment may influence a child's sociocognitive development (e.g.
+attachment style, emotion regulation) in ways that may predispose them to peer victimization \citep{goemans2021child, McCrory2022-cj}.
+Parental neglect, or the absence of adult supervision, is also a risk factor for peer victimization \citep{lereya2013parenting}, potentially by increasing the likelihood of children participating in higher risk social situations where victimization may take place.
+Insofar that maladaptive parenting is an indirect cause of peer victimization (mediated by a number of complex psychosocial pathways), then we believe we are justified specifying indicators of peer victimization to load in part onto the common factor.
+This is also highlights the need for future study of the peer victimzation factor on the MACE, to better understand how it is related to and distinct from childhood maltreatment by parents and what its downstream consequences are.
 
@@
 For example, we identified several sources of DIF on the MACE.
-Across samples we found that women were more likely to endorse having been sexually abused, whereas men were more likely to report being physically abused by peers.
-These findings are in line with previous studies that found girls are more frequently victims of sexual abuse \citep{stoltenborgh2015prevalence} and boys are more frequently victims of bullying (especially physical bullying; \citealt{scheithauer2006physical}).
-We also identified a number of items that exhibited DIF across samples.
+Across samples we found that women experienced more sexual abuse, whereas men experienced more physical abuse by peers.
+These findings are in line with previous research \citep{radford2013prevalence}.
+We also identified a number of items that exhibited DIF by sample.
 The cause of this is unclear, and may reflect differences in sample population (greater Boston area vs.
@@
 
-The current study is, of course, not without limitations.
+The current study is not without limitations.
 The most apparent limitation is that here we only studied the dichotomous response data but not the chronology data also solicited as part of the MACE.
-Though we found that the factor structure of the dichotomous responses are essentially unidimensional, this may belie more complex patterns present in the maltreatment onset and duration data.
+Though we found that the factor structure of the dichotomous responses are essentially unidimensional, this may belie more complex patterns present in the maltreatment timing data.
 Indeed, modeling the timeseries of adverse events in childhood --- perhaps using dynamic factor analysis \citep{zhang2007bayesian} --- might be more sensitive to more nuanced patterns of maltreatment co-occurrence that are lost when collapsing the data into a measure of any vs.
@@
 Thus, we found that the MACE total score is a valid and reliable of overall childhood maltreatment.
-In contrast, we found that MACE subscale and composite scores are not valid measures of particular types of childhood maltreatment, but instead primarily reflect general maltreatment.
-Moving forward, childhood abuse and neglect researchers utilizing the MACE (or otherwise) should take pains to ensure that any subscale or composite scores they employ are valid, lest they draw spurious inferences.
+In contrast, we found that MACE subscale scores are not valid measures of particular types of childhood maltreatment, but instead primarily reflect general maltreatment.
+Moving forward, childhood abuse and neglect researchers utilizing the MACE (or otherwise) should take pains to ensure that any subscale scores they employ are valid, lest they draw spurious inferences.
+
 \bibliography{main}

### 2022-08-07 manuscript/main.tex
manuscipt edits (2022-08-07a)

@@
 All participants completed the 52-item version of the MACE \citep{teicher2015maltreatment}.
-The scale is a retrospective self-report measure of the severity and timing of exposure to 10 types of childhood maltreatment.
+The MACE is a retrospective self-report scale that measures 10 unique types of childhood maltreatment (i.e.
+parental verbal abuse, physical abuse, nonverbal emotional abuse, sexual abuse, emotional neglect, physical neglect, witnessing violence to siblings, witnessing interparental violence, peer verbal abuse, peer physical abuse).
 For each item, participants endorse whether they ever experienced a particular event during childhood (e.g.
@@
 This is in contrast to the remaining items in the scale, where a score of 1 is assigned to a response if maltreatment or abuse occurred in at least one year of childhood.
-
-% (i.e. verbal abuse, physical abuse, nonverbal emotional abuse, sexual abuse, emotional neglect, and physical neglect by parents or guardians)
 
@@
 
-Using the cutoffs recommended by \cite{hidalgo2014binary}, we identified 11 of 52 items on the MACE that exhibited large DIF by study (Figure \ref{fig:dif_study}).
-Interestingly these included all six reverse-scored items.
-We also found an additional 11 items that exhibited large DIF by gender (Figure \ref{fig:dif_gender}).
-Five of these items measured sexual abuse (endorsed by more women than men), and another five measured peer physical abuse (endorsed by more men than women).
+Using the cutoffs recommended by \cite{hidalgo2014binary}, we identified 11 of 52 items on the MACE with large DIF by study (Figure \ref{fig:dif_study}).
+These included all six reverse-scored items.
+We also found an additional 11 items with large DIF by gender (Figure \ref{fig:dif_gender}).
+Five of these items measured sexual abuse (endorsed more frequently by women), and another five measured peer physical abuse (endorsed more frequently by men).
 We identified no items that exhibited large DIF by age.
-Given the number of items exhibiting DIF by sample membership, we decided to analyze each sample separately.
+Given the number of items with DIF by sample membership, we decided to analyze each sample separately.
 
@@
 The second is a bifactor model where all items load onto a general maltreatment factor and also one of 10 specific factors based on their subscale membership (Figure \ref{fig:models}b).
+This model will be used to evaluate if the MACE can produce reliable subscale scores for each of the 10 types of maltreatment assessed by the MACE.
 
-The third model we fit investigated a conceptual model of maltreatment that distinguishes between threat and deprivation as distinct forms of childhood adversity \citep{mclaughlin2014childhood}.
-Initially we fit a bifactor model where all items loaded onto a general maltreatment factor and either the threat or deprivation specific factors.
+The third model was motivated by a theoretical account of childhood maltreatment that distinguishes between threat and deprivation experiences as distinct types of adversity \citep{mclaughlin2014childhood}.
+To explore this model of the MACE, we initially fit a bifactor model where all items loaded onto a general maltreatment factor, and also a threat or neglect specific factor.
 However, we observed vanishing item loadings on the threat specific factor.
-As such, we instead fit a bifactor S-1 model \citep{eid2017anomalous} where all items load onto a general maltreatment (or reference) factor, and only the items measuring neglect experiences load onto their own specific factor (Figure \ref{fig:models}c).
+As such, we instead fit a bifactor S-1 model \citep{eid2017anomalous} where all items load onto a general maltreatment factor, and only items measuring neglect load onto a specific factor (Figure \ref{fig:models}c).
+This model will be used to evaluate if the MACE can produce reliable neglect subscale scores distinguishable from general maltreatment.
 
@@
 
-The fourth and final model proposes a more parsimonious factor structure for the MACE.
-Items in the MACE can be neatly divided according to the source of adversity: maltreatment by adults at home (e.g.
-parents) or by peers.
-We hypothesized then that there would be two principal factors underlying responses on the MACE, maltreatment by adults and peer victimization.
-Based on preliminary model fits, we also suspected the presence of a secondary methods factor affecting the 6 reverse-scored items.
-Initially we fit a bifactor model in which all items loaded onto a general maltreatment factor and also one of three specific factors: maltreatment by adults, peer victimization, or reverse-scoring.
-However, we also observed vanishing loadings for the maltreatment by adults specific factor.
-Thus, we instead fit a bifactor S-1 model where all items load onto a general maltreatment factor; in addition, the peer victimization and reverse-scoring items load onto their own respective specific factors (Figure \ref{fig:models}d).
+The fourth and final model proposes a more parsimonious structure of the MACE based on the content of its items.
+Items on the MACE can be organized according to the source of maltreatment: parents (or other adults in the home) and peers.
+Thus, we hypothesized there would be two factors underlying responses on the MACE: a primary maladaptive parenting factor and a secondary peer victimization factor.
+Based on preliminary model fitting, we also suspected the presence of a tertiary methods factor affecting the six reverse-scored items.
+We initially fit a bifactor model in which all items loaded onto a general maltreatment factor and one of three specific factors (maladaptive parenting, peer victimization, reverse-scoring); however, we observed vanishing loadings on the first specific factor.
+As such, we instead fit a bifactor S-1 model where all items load onto a general maltreatment factor, and the peer victimization \& reverse-scoring items additionally load onto their own specific factors (Figure \ref{fig:models}d).
+This model will be used to evaluate if the MACE can produce peer victimization reliable subscale scores distinguishable from general maltreatment.
 
@@
 This measure compares the observed and model-predicted proportion of participants at each total score level.
-Second, we calculated Yen's $Q_3$ statistic for each pair of items to detect local dependence \citep{yen1984effects} and also the critical $Q_3$-value above which local dependence is indicated  \citep{christensen2017critical}.
-Briefly, for each model and sample, we simulated 1000 locally independent datasets using the posterior distribution of the model parameters.
+Second, to test for local dependence in each pair of items, we calculated Yen's $Q_3$ statistic \citep{yen1984effects} and also the critical $Q_3$-value (i.e.
+the value above which local dependence is indicated; \citealt{christensen2017critical}).
+For each model and sample, we simulated 1000 locally independent datasets using the posterior distribution of the model parameters.
 We then recorded the maximum $Q_3$ value per simulation.
@@
 
-Finally, the goodness-of-fit of the models was compared using Bayesian leave-one out cross-validation (LOO-CV; \citealt{vehtari2017practical}).
-The LOO-CV measure quantifies the discrepancy between the model and the data while taking into account model complexity.
-Here LOO-CV values are presented in deviance scale, such that smaller values indicate better fit.
+The goodness-of-fit of the models was compared using Bayesian leave-one out cross- validation (LOO-CV; \citealt{vehtari2017practical}).
+The LOO-CV measure quantifies the discrepancy between the model and data while taking into account model complexity.
+LOO-CV values are presented in deviance scale, i.e.
+smaller values indicate better fit.
 
@@
 In contrast, values of $\omega_s$ smaller than 0.50 preclude the interpretation of subscale scores as reflecting a specific factor \citep{gignac2013bifactor}.
+
 To assess the unidimensionality of the MACE, we also calculated the explained common variance (ECV; \citealt{sijtsma2009use}) and the proportion of uncontaminated correlations (PUC; \citealt{reise2013multidimensionality}).

### 2022-08-08 manuscript/main.tex
manuscipt edits (2022-08-07b)

@@
 Childhood maltreatment is an unfortunately common experience for children across the globe \citep{stoltenborgh2015prevalence}.
-It is a risk factor for multiple poor outcomes later in life, including the onset of mental illness \citep{kessler2010childhood}, challenges in cognitive functioning \citep{su2019does}, and poorer physical health \citep{wegman2009meta}.
+It is a risk factor for multiple poor outcomes later in life, including the onset of mental illness \citep{kessler2010childhood}, challenges in cognitive functioning \citep{su2019does}, and poor physical health \citep{wegman2009meta}.
 Childhood maltreatment also predicts worse treatment outcomes for several psychiatric disorders \citep{nanni2012childhood, thomas2019childhood}.
-Thus, the reliable and valid assessment of childhood maltreatment is an important research goal in order to refine our understanding of the link between maltreatment and health outcomes; augment treatment strategies for victims of childhood maltreatment; and ultimately design better interventions to prevent maltreatment in the first place.
+Thus, reliable and valid assessment of childhood maltreatment is an important research goal in order to refine our understanding of the link between maltreatment and health outcomes; augment treatment strategies for victims of childhood maltreatment; and ultimately design better interventions to prevent maltreatment in the first place.
 
@@
 An overall shift in overall endorsement rates by sample is not necessarily evidence for a violation of measurement invariance.
-One possibility is the replication sample had simply experienced more adversity than the original sample.
+One possibility is that the replication sample had simply experienced more adversity than the original sample.
 A second is that the differences in the demographic composition of the two samples (e.g.
@@
 
-To evaluate the reliability of the MACE total and subscale scores, we computed several model-based reliability indices using the factor loading estimates from the bifactor models.
+To evaluate the reliability of the MACE total and subscale scores, we calculated several model-based reliability indices using the factor loadings from the bifactor models.
 First, we calculated coefficient omega ($\omega$; \citealt{mcdonald1999test}), which quantifies the proportion of reliable score variance attributable to all factors (i.e.
-both group and specific factors).
+both general and specific factors).
 Next we calculated coefficient omega hierarchical ($\omega_h$) and coefficient omega subscale ($\omega_s$;  \citealt{reise2013scoring}).
@@
 Values of $\omega_h$ approaching 1 indicate the general factor is the primary source of reliable variance in a total score.
-When $\omega_h$ > 0.80, the total score may be considered an essentially unidimensional reflection of the general factor even if the data are multidimensional \citep{rodriguez2016applying}.
-So too values of $\omega_s$ approaching 1 indicate the specific factor, and not the general factor, is the primary source of reliable variance in a subscale score.
+When $\omega_h$ > 0.80, the total score may be considered an essentially unidimensional reflection of the general factor \citep{rodriguez2016applying}.
+Values of $\omega_s$ approaching 1 indicate the specific factor, not the general factor, is the primary source of reliable variance in a subscale score.
 \cite{canivez2016bifactor} suggested that an acceptable $\omega_s$ value is 0.50 and that >0.70 is desirable.
-In contrast, values of $\omega_s$ smaller than 0.50 preclude the interpretation of subscale scores as reflecting a specific factor \citep{gignac2013bifactor}.
+In contrast, values of $\omega_s$ < 0.50 preclude the interpretation of subscale scores as primarily reflecting a specific factor \citep{gignac2013bifactor}.
 
-To assess the unidimensionality of the MACE, we also calculated the explained common variance (ECV; \citealt{sijtsma2009use}) and the proportion of uncontaminated correlations (PUC; \citealt{reise2013multidimensionality}).
-The ECV is the ratio of the variance explained by the general factor divided by the total common variance of a scale; it therefore measures the contribution of the general factor relative to the specific factors.
-As ECV values approach 1, the general factor loadings of a bifactor model are expected to resemble those that would be obtained by a one-dimensional model.
+To assess the unidimensionality of the MACE, we calculated the explained common variance (ECV; \citealt{sijtsma2009use}) and the proportion of uncontaminated correlations (PUC; \citealt{reise2013multidimensionality}).
+The ECV is the ratio of the variance explained by the general factor divided by the total common variance of a scale.
+It thus measures the contribution of the general factor relative to the specific factors.
+As ECV values approach 1, the general factor loadings of a bifactor model are expected to resemble the item loadings that would be obtained by a one-dimensional model.
 ECV values greater than 0.70 often indicate a scale is essentially unidimensional \citep{rodriguez2016applying}.
-In turn, the PUC quantifies how many inter-item correlations are accounted for only by the general factor and is an important moderator of the ECV.
-As the PUC increases, ECV is less important for evaluating the risk of bias when fitting a unidimensional model to data with a latent bifactor structure.
-When PUC > 0.80, there is a low risk of bias when a multidimensional scale is treated as unidimensional \citep{reise2013multidimensionality}.
-As an additional check, we calculated the relative parameter bias (RPB), which is the difference between an item's loading in a unidimensional model and its corresponding general factor loading the bifactor, divided by the general factor loading in the bifactor.
-Parameter bias less than 10–15\% is acceptable \citep{muthen1987structural}.
+In turn, PUC quantifies how many inter-item correlations are accounted for only by the general factor.
+PUC is an important moderator of the ECV; as PUC increases, ECV is less important for evaluating the risk of bias when fitting a unidimensional model to data with a latent bifactor structure.
+When PUC > 0.80, there is a low risk of bias when a multi- dimensional scale is treated as unidimensional \citep{reise2013multidimensionality}.
+We also calculated the relative parameter bias (RPB), which is the difference between an item's loading in a unidimensional model and its corresponding general factor loading in a bifactor model, divided by the general factor loading from the bifactor model.
+RPB less than 10–15\% is acceptable \citep{muthen1987structural}.
 
 Finally, we calculated the H index to assess the construct replicability of each factor \citep{hancock2001rethinking}.
-The H index is an estimate of how well a set of items represents a latent variable.
-H values closer to 1 indicate a well-defined latent variable with factor loadings more likely to be stable across studies.
-In contrast, small H values suggest a poorly defined latent variable where factor loadings are liable to change across studies.
-Because we are primarily focused on measuring the reliability of MACE total and subscale scores, we report the H values below but do not discuss them further.
+The H index is an estimate of how well a set of items represents a latent factor.
+H values closer to 1 indicate a well-defined latent factor with item loadings more likely to be stable across studies.
+In contrast, small H values suggest a poorly defined latent factor where item loadings are liable to change across studies.
+Because we are presently focused on evaluating the reliability of MACE total and subscale scores, we report the H values below but do not interpret them further.
 
@@
 
-In addition to confirmatory factor analysis, we also we also performed exploratory factor analysis (EFA) on the response data from each sample separately.
-The goal of the EFA was to observe what factor structure would emerge with no \emph{a priori} restrictions.
+In addition to confirmatory factor analysis, we also performed exploratory factor analysis (EFA) on the response data from each sample separately.
+The goal of the EFA was to examine what factor structure would emerge with no \emph{a priori} restrictions placed on the data.
 We decided to extract two- and three- factor solutions based on the total number of factors in the bifactor S-1 models.
-We were particularly interested in seeing if a two-factor solution would produce the factors from model 3 (i.e.
-threat and deprivation); and if a three-factor solution would produce the three factors from model 4 (i.e.
-maltreatment by adults, peer victimization, reverse-scoring).
+We were particularly interested to see if a two-factor solution would reproduce the factors from model 3 (threat, deprivation); and if a three- factor solution would reproduce the three factors from model 4 (maladaptive parenting, peer victimization, reverse-scoring).
 
@@
 Factors were extracted using the oblique geomin \citep{yates1987multivariate} and cf-quartimax rotation criteria \citep{crawford1970general}, which were selected in order to extract simple factor structures with few cross-loadings.
-We used two rotation criteria in order to examine the generalizability of the factor solutions across multiple rotations.
+We used two rotation criteria in order to examine the generalizability of the factor solutions across rotations.
 We again relied on traditional fit indices (i.e.
-RMSEA, CFI, TLI) to evaluate the goodness-of-fit of the exploratory models to the response data of both samples.
+RMSEA, CFI, TLI) to evaluate the goodness-of-fit of the exploratory factor models to the response data of both samples.
 
@@
 
+\subsubsection{Model 1: One-factor (unidimensional) model of the MACE}
+
+The item discrimination parameters of the unidimensional model were, on average, large for both the original sample (mean $\alpha$ = 1.357, 95\% HDI = 1.295 -- 1.414) and replication sample (mean $\alpha$ = 1.349, 95\% HDI = 1.272 -- 1.422).
+The standardized factor loadings were correspondingly large in magnitude.
+In the original sample, factor loadings ranged between 0.279 and 0.864 with an average value of 0.597 (95\% HDI = 0.579 -- 0.614); in the replication sample, factor loadings ranged between 0.359 and 0.814 with an average value of 0.600 (95\% HDI = 0.578 -- 0.621).
+Most items had loadings greater than 0.50 (original: 80.0\%, replication: 80.0\%), and thereby appear to be good indicators of general maltreatment.
+The rank correlation of general factor loadings between samples was $\rho$ = 0.565 (95\% HDI = 0.414 -- 0.711), indicating only moderate agreement between samples.
+
+To begin characterizing the dimensionality of the MACE, we inspected the first six eigenvalues of the polychoric item correlation matrix for each sample.
+In the original sample, these values were 14.62, 3.77, 2.64, 2.11, 1.79, and 1.35; in the replication sample, they were 13.27, 4.43, 2.48, 2.04, 1.81, and 1.56.
+The first-to-second eigenvalue ratios of 3.84 and 3.02 are just above the 3-to-1 criterion often cited in determining whether item response data are essentially unidimensional \citep{embretson2013item}.
+
+In sum, the fits of the one-factor models provide partial evidence in support of a unidimensional model of the MACE.
+Virtually all items exhibited moderate-to-large loadings onto the common factor.
+Moreover, the eigenvalues of the polychoric correlation matrix suggested one primary factor.
+However, the degree of local dependence present among in the unidimensional model fits (especially among the peer victimization items) suggest there may be additional dimensions to the MACE.
+We turn next to the multidimensional confirmatory models to explore these possibilities.
+
 \begin{figure}[t!]
@@
 
-\subsubsection{Model 1: One-factor (unidimensional) model of the MACE}
-
-The item discrimination parameters of the unidimensional model were, on average, large for both the original sample (mean $\alpha$ = 1.357, 95\% HDI = 1.295 -- 1.414) and replication sample (mean $\alpha$ = 1.349, 95\% HDI = 1.272 -- 1.422).
-The standardized factor loadings were correspondingly large in magnitude.
-In the original sample, factor loadings ranged between 0.279 and 0.864 with an average value of 0.597 (95\% HDI = 0.579 -- 0.614); in the replication sample, factor loadings ranged between 0.359 and 0.814 with an average value of 0.600 (95\% HDI = 0.578 -- 0.621).
-Most items had loadings greater than 0.50 (original: 80.2\%, replication: 80.0\%), and thereby appear to be good indicators of general maltreatment.
-The rank correlation of factor loadings between samples was $\rho$ = 0.565 (95\% HDI = 0.414 -- 0.711), indicating only moderate agreement on the relative loading of items onto the common factor across samples.
-
-To evaluate the dimensionality of the MACE, we inspected the first six eigenvalues of the polychoric correlation matrix for each sample.
-In the original sample, these values were 14.62, 3.77, 2.64, 2.11, 1.79, and 1.35; in the replication sample, they are 13.27, 4.43, 2.48, 2.04, 1.81, and 1.56.
-The first-to-second eigenvalue ratios are 3.84 and 3.02, respectively, just above the 3-to-1 criterion often cited in determining whether item response data are essentially unidimensional \citep{embretson2013item}.
-
-In sum, the fits of the one-factor models provide partial evidence in support of a unidimensional account of the MACE.
-The items exhibited uniformly moderate-to-large loadings onto the common factor.
-Moreover, the eigenvalues of the polychoric correlation matrix suggest one principal factor.
-However, the degree of local dependence present among in the unidimensional model fits (especially among the peer victimization items) suggest there may be additional dimensions to the MACE.
-We turn next to the multidimensional confirmatory models to explore these possibilities.
-
 \subsubsection{Model 2: Bifactor model with 10 specific factors}
@@
 Next we inspect the bifactor model with 10 specific factors, i.e.
-one per MACE subscale (model 2).
-In the original sample, the loadings on the general maltreatment factor ranged between 0.235 and 0.864, with an average value of 0.575 (95\% HDI = 0.556 -- 0.593).
-In the replication sample, the general factor loadings ranged between 0.299 and 0.827, with an average value of 0.580 (95\% HDI = 0.560 -- 0.601).
-The relative parameter bias was marginal for both samples (original = 2.1\%; replication = 2.7\%), suggesting little bias in the factor loadings of the unidimensional models due to multidimensionality.
+one per subscale on the MACE (model 2).
+In the original sample, the loadings on the general maltreatment factor ranged between 0.235 and 0.864 with an average value of 0.575 (95\% HDI = 0.556 -- 0.593).
+In the replication sample, the general factor loadings ranged between 0.299 and 0.827 with an average value of 0.580 (95\% HDI = 0.560 -- 0.601).
+The relative parameter bias was marginal for both samples (original = 2.1\%; replication = 2.7\%), suggesting there was little bias in the factor loadings of the one-factor models due to multidimensionality.
 
-By contrast the standardized loadings on the specific factors were much smaller in magnitude, seldom exceeding loadings on the general factor.
-In the original dataset, the specific factor loadings ranged 0.014 to 0.797 with an average value of 0.313 (95\% HDI = 0.285 -- 0.341).
+In contrast, the loadings on the specific factors were smaller and seldom exceeded loadings on the general factor.
+In the original dataset, the specific factor loadings ranged between 0.014 to 0.797 with an average value of 0.313 (95\% HDI = 0.285 -- 0.341).
 In the replication dataset, the range was between 0.043 and 0.698, with an average value of 0.343 (95\% HDI = 0.315 -- 0.369).
-Only items from the peer victimization subscales (i.e.
-peer verbal \& physical abuse) had specific factor loadings of roughly equal or greater magnitude to those on the general factor.
+Only items from the peer verbal and physical abuse subscales had specific factor loadings of equal or greater magnitude to those on the general factor.
 
@@
 
-Given these, we found that the MACE overall severity score had excellent reliability (original: $\omega$ = 0.963; replication: $\omega$ = 0.965).
-Further, the $\omega_h$ value for both samples was 0.924 indicating that 96.0\% and 95.8\% of the reliable variance in the overall severity score is attributable to the general maltreatment factor.
-(This is not in itself so surprising given the large number of subscales and relative paucity of items per subscale.) Notably, the ECV values were 0.716 and 0.697 for the original and replication samples, respectively.
-In conjunction with a large PUC value of 0.912, these results provide clear evidence for a dominant general maltreatment factor in the MACE.
+We found that the MACE total score had excellent reliability (original: $\omega$ = 0.963; replication: $\omega$ = 0.965).
+The $\omega_h$ value was 0.924 for both samples indicating that 96.0\% and 95.8\% of the reliable variance in the total score is attributable to general maltreatment.
+(This result is not very surprising given the large number of subscales composed of few items.) Importantly, the ECV values were 0.716 and 0.697 for the original and replication samples, respectively.
+In conjunction with a large PUC value of 0.912, these results provide strong evidence for a dominant general maltreatment factor in the MACE.
 
 The reliability of the MACE subscale scores were more variable in comparison.
-The $\omega$ values ranged between 0.590 and 0.909 in the original sample, and between 0.595 and 0.893 in the replication sample, with smaller values for subscales with fewer items.
-Crucially, the $\omega_s$ values were mostly small for both the original sample (mean $\omega_s$ = 0.181, range = 0.027 -- 0.621) and replication sample (mean $\omega_s$ = 0.205, range = 0.037 -- 0.565).
-That is, the MACE subscale scores contain almost no reliable variance independent of the general maltreatment factor.
-Therefore they should not be interpreted as reflecting specific forms of maltreatment.
-The only exception was for the peer verbal abuse subscale, where more than half the reliable variance in subscale scores (original = 70.7\%; replication = 66.2\%) was independent of general maltreatment.
+The $\omega$ values ranged between 0.590 and 0.909 in the original sample, and between 0.595 and 0.893 in the replication sample, with smaller $\omega$ values for subscales with fewer items.
+Crucially, the $\omega_s$ values were small for both the original sample (mean $\omega_s$ = 0.181, range = 0.027 -- 0.621) and replication sample (mean $\omega_s$ = 0.205, range = 0.037 -- 0.565).
+That is, the MACE subscale scores contain almost no reliable variance after controlling for the general maltreatment factor.
+They should not therefore be interpreted as reflecting specific forms of maltreatment.
+The only exception was for the peer verbal abuse subscale, where more than half the reliable variance in subscale scores was independent of general maltreatment (original = 70.7\%; replication = 66.2\%).
 
-The results of the bifactor model were unequivocal.
-Responses on the MACE are characterized by a dominant general maltreatment factor with strong loadings from the majority of items.
-In stark contrast, the specific factors corresponding to each MACE subscale are weak and characterized mostly by small factor loadings.
-Whereas the MACE overall severity score is reliable and a valid measure of general maltreatment, the MACE subscale scores are less reliable and cannot be meaningfully interpreted as uniquely reflecting exposure to specific types of abuse.
-The question remains, however, if there are other valid subscale scores to be derived from the MACE in addition to the overall severity score.
+The results of the full bifactor model were clear: responses on the MACE are explained by a dominant general maltreatment factor characterized by large loadings from the majority of items.
+In contrast, specific factors corresponding to each MACE subscale are weak and characterized mostly by small factor loadings.
+Whereas the MACE total score is a reliable and valid measure of overall maltreatment, the MACE subscale scores are less reliable and cannot be meaningfully interpreted as uniquely reflecting exposure to particular types of abuse.
+The question remains, however, if there are other subscale scores to be derived from the MACE.
 
-\subsubsection{Model 3: Bifactor S-1 model with one specific factor (deprivation)}
+\subsubsection{Model 3: Bifactor S-1 model with one specific factor (neglect)}
 
-We turn our attention next to the threat-deprivation bifactor S-1 model (model 3).
+We turn our attention next to the bifactor S-1 model with one specific factor (neglect; model 3).
 The loadings on the general factor largely resembled those of the bifactor model.
-In the original sample, the loadings on the general maltreatment factor ranged between 0.150 and 0.864, with an average value of 0.577 (95\% HDI = 0.558 -- 0.594).
-In the replication sample, the general factor loadings ranged between 0.395 and 0.831, with an average value of 0.580 (95\% HDI = 0.558 -- 0.601).
+In the original sample, the loadings on the general maltreatment factor ranged between 0.150 and 0.864 with an average value of 0.577 (95\% HDI = 0.558 -- 0.594).
+In the replication sample, the general factor loadings ranged between 0.395 and 0.831 with an average value of 0.580 (95\% HDI = 0.558 -- 0.601).
 
-The loadings on the specific (neglect) factor were smaller on average.
+The loadings on the neglect specific factor were smaller on average.
 In the original sample, the average specific loading was 0.410 (95\% HDI = 0.373 -- 0.454); in the replication sample, it was 0.454 (95\% HDI = 0.409 -- 0.499).
-The specific loadings exhibited a notable pattern where only the six reverse-scored items exhibited moderate loadings; the remaining four item exhibited negligible factor loadings.
-In other words, the ``neglect'' factor appears instead to reflect a reverse-scoring methods factor.
+The specific loadings displayed a troubling pattern where only the six reverse-scored items exhibited moderate loadings on the specific factor; the remaining four item exhibited negligible factor loadings.
+That is, the ``neglect'' factor appears instead to reflect a reverse-scoring methods factor.
 
-As observed with the bifactor model, the overall severity score under the threat- deprivation bifactor S-1 model exhibited excellent reliability (original: $\omega$ = 0.958; replication: $\omega$ = 0.959).
-Unsurprisingly, the $\omega_h$ values indicated the majority of reliable variance in severity score is attributable to the general factor (original: $\omega_h$ = 0.928; replication: $\omega_h$ = 0.922).
-The reliability of the ``neglect'' subscale score was similarly good (original: $\omega$ = 0.877; replication: $\omega$ = 0.921).
-The $\omega_s$ values, however, indicated the subscale score contained less than half reliable variance independent of the general factor (original: $\omega_h$ = 0.384; replication: $\omega_h$ = 0.375).
+Moving on to the reliability measures, we again found that the total score exhibited excellent reliability (original: $\omega$ = 0.958; replication: $\omega$ = 0.959).
+Moreover, the $\omega_h$ values indicated the majority of reliable variance in severity score was attributable to the general factor (original: $\omega_h$ = 0.928; replication: $\omega_h$ = 0.922).
+The reliability of the ``neglect'' subscale score was also good (original: $\omega$ = 0.877; replication: $\omega$ = 0.921).
+The $\omega_s$ values, however, indicated the neglect subscale score contained less than half reliable variance after controlling for the general factor (original: $\omega_h$ = 0.384; replication: $\omega_h$ = 0.375).
 
-In summary then, the results do not support the use of threat and deprivation subscale scores as derived from the abuse and neglect items on the MACE.
-The minority of reliable variance unique from general maltreatment prevents the deprivation score from being interpreted as such.
-Furthermore, the items comprising deprivation subscale seem to be confounded by a reverse-scoring methods factor.
+In summary, the results do not support the use of the MACE to make separate threat and deprivation subscale scores.
+The neglect subscale scores posses only a minority of reliable variance unique from general maltreatment thereby precluding meaningful interpretation of these scores.
+Furthermore, the majority of items measuring neglect on the MACE seem to be contaminated by a reverse-scoring methods factor.
 
@@
 Finally, we inspect the bifactor S-1 model with two specific factors measuring peer victimization and reverse-scoring (model 4).
-In the original sample, the loadings on the general maltreatment factor ranged between 0.158 and 0.864, with an average value of 0.563 (95\% HDI = 0.545 -- 0.580).
-In the replication sample, the general factor loadings ranged between 0.298 and 0.852, with an average value of 0.572 (95\% HDI = 0.550 -- 0.593).
+In the original sample, the loadings on the general maltreatment factor ranged between 0.158 and 0.864 with an average value of 0.563 (95\% HDI = 0.545 -- 0.580).
+In the replication sample, the general factor loadings ranged between 0.298 and 0.852 with an average value of 0.572 (95\% HDI = 0.550 -- 0.593).
 Interestingly, the loadings on the specific factors were of approximately equal magnitude.
 In the original dataset, the specific factor loadings ranged 0.327 to 0.789 with an average value of 0.550 (95\% HDI = 0.522 -- 0.576).
-In the replication dataset, the range was between 0.183 and 0.684, with an average value of 0.526 (95\% HDI = 0.492 -- 0.556).
+In the replication dataset, the range was between 0.183 and 0.684 with an average value of 0.526 (95\% HDI = 0.492 -- 0.556).
 
-As with the other bifactor models, the overall severity score under this bifactor S-1 model exhibited excellent reliability (original: $\omega$ = 0.961; replication: $\omega$ = 0.961).
-Again the majority of reliable variance in severity score was attributable to the general factor.
+As with the other bifactor models, the total score under this model exhibited excellent reliability (original: $\omega$ = 0.961; replication: $\omega$ = 0.961).
+Again the majority of reliable variance in severity score was attributable to the general factor (original: $\omega_h$ = 0.896; replication: $\omega_h$ = 0.904).
 The peer victimization subscale score exhibited good reliability across samples (original: $\omega$ = 0.889; replication: $\omega$ = 0.868).
-Notably, the corresponding $\omega_s$ values were 0.569 for the original sample and 0.493 for the replication sample.
-That is, slightly more than half of the variance in peer victimization scores (original: 64.0\%; replication: 56.8\%) is unique from the general maltreatment factor.
-These results suggest there exists a secondary factor in the MACE, peer victimization, independent of general maltreatment (albeit marginally so).
+The corresponding $\omega_s$ values were 0.569 for the original sample and 0.493 for the replication sample.
+That is, slightly more than half of the variance in peer victimization scores (original: 64.0\%; replication: 56.8\%) is independent of the general maltreatment factor.
+These results suggest the presence a secondary peer victimization factor in the MACE (albeit weakly so) in addition to a primary general maltreatment factor.
 
-Briefly, we note that the subscale scores formed by the reverse-scored items was also reliable (original: $\omega$ = 0.839; replication: $\omega$ = 0.915).
-Moreover, the corresponding $omega_s$ values were similar in magnitude to those observed for the peer victimization scores (original: $\omega_s$ = 0.584; replication: $\omega_s$ = 0.498).
-As such, these scores are also have majority reliable variance independent of general maltreatment.
-It is difficult to interpret these scores, however, as it is unclear if they primarily reflect neglect experiences, a wording effect, or a scoring-induced methods artifact.
+Briefly, we note that the subscale scores formed by the reverse-scored items were also reliable (original: $\omega$ = 0.839; replication: $\omega$ = 0.915).
+Moreover, the corresponding $\omega_s$ values were similar in magnitude to those observed for the peer victimization scores (original: $\omega_s$ = 0.584; replication: $\omega_s$ = 0.498).
+In other words, these scores are also have majority reliable variance independent of general maltreatment.
+It is difficult to interpret these scores, however, as it is unclear what they primarily reflect: neglect experiences, wording effects, and/or scoring-induced methods artifact.
 
@@
 
-The goodness-of-fit of the exploratory factor models to the response data of both samples are summarized in Table \ref{tab:efa_diagnostics}.
-Briefly, the fit indices indicated that all models provided an adequate fit to the data.
-The standardized factor loadings for the 2- and 3-factor solutions produced by the cf-quartimax rotation are summarized in Figure \ref{fig:efa_cf}.
-(The corresponding factor loadings produced by the geomin rotation were qualitatively similar and are presented in Figure \ref{fig:efa_geomin}.) In the 2-factor solution for the original sample, there was a primary adult maltreatment factor and a secondary peer victimization factor.
-In contrast, the replication sample produced a primary adult/peer maltreatment factor with a secondary reverse-scoring factor.
+The goodness-of-fit measures for the exploratory factor models to the response data by sample are summarized in Table \ref{tab:efa_diagnostics}.
+The fit indices indicated that all exploratory models provided an adequate fit to the data.
+The standardized factor loadings for the 2- and 3-factor solutions produced by the cf-quartimax rotation are presented in Figure \ref{fig:efa_cf}.
+(The corresponding factor loadings produced by the geomin rotation were qualitatively similar and are presented in Figure \ref{fig:efa_geomin}.) In the 2-factor solution for the original sample, we observed a primary maladaptive parenting factor and a secondary peer victimization factor.
+In contrast, the replication sample produced a primary maladaptive parenting/peer maltreatment factor with a secondary reverse-scoring factor.
+Neither 2-factor solution precisely resembled the threat-deprivation factor structure of confirmatory model 3.
 
-The three-factor solution for each sample were similar.
-The analysis produced a primary adult maltreatment factor with secondary peer victimization and reverse-scoring factors.
-(To note items concerning sexual abuse loaded more strongly on the peer victimization scale in the replication sample, warranting some caution in interpreting the replicability of factor structures.) With the oblique rotation, we observed only modest inter-factor correlations (original: mean $r$ = 0.371; replication: mean $r$ = 0.269).
-To conclude, the structure of both the two- and three-factor exploratory factor solutions resembled the structure of the bifactor S-1 model with two specific factors (model 4), providing further support for this model of responses on the MACE.
+The 3-factor solution for each sample were similar.
+EFA produced a primary maladaptive parenting factor with secondary peer victimization and reverse-scoring factors.
+(Interestingly, items concerning sexual abuse loaded more strongly on the peer victimization scale in the replication sample, raising some questions about the replicability of the factor structures.) To conclude, the structure of both the 2- and 3-factor exploratory factor models resembled the bifactor S-1 model with two specific factors (model 4), providing further support for that model of the factor structure of the MACE.
 
@@
 
-The MACE is a promising alternative measure of childhood maltreatment \citep{georgieva2022systematic}.
-However, there has been little study of its structural validity \citep{saini2019systematic}, leaving unanswered questions about the scale's underlying factor structure and the validity of its total/subscale scores.
-Here we investigated the factor structure of the MACE using both confirmatory and exploratory item response modeling.
-The fits of a bifactor model (model 2) to the response data from both samples revealed a dominant general factor with uniformly strong loadings from the majority of items, and substantially weaker loadings on the specific subscale factors.
-The general maltreatment factor explained approximately 70\% of the reliable variance in responses across samples.
-When comparing the loadings on the general factor to those from a unidimensonal model, the relative parameter bias was marginal.
+The MACE is a promising measure of childhood maltreatment \citep{georgieva2022systematic}.
+Despite its increasing popularity, there has been little investigation into the MACE's structural validity \citep{saini2019systematic} leaving unanswered questions about its factor structure and the validity of its total/subscale scores.
+Here we investigated the factor structure of the MACE, using both confirmatory and exploratory item response modeling, in two independent samples.
+Bifactor modeling revealed a dominant general maltreatment factor, characterized by uniformly strong loadings across items, and much weaker specific factors.
+The general factor explained approximately 70\% of the reliable variance in the response data across samples.
+Moreover, the general factor loadings from the bifactor models largely resembled the corresponding loadings from a unidimensonal model.
 Together, these results support the conclusion that the MACE is essentially unidimensional.
 
-We also calculated model-based reliability indices of the MACE total and subscale scores using the bifactor model fits.
-We found that the MACE total score has excellent reliability and is a valid measure of general maltreatment.
-In contrast, we found that the MACE subscale scores were not valid measures of specific forms of maltreatment; in all but one example, the majority of reliable variance in subscale scores was attributable to the general factor, thereby precluding meaningful interpretation of the MACE subscale scores as such.
-The same was true of the deprivation subscale scores, derived from the neglect items on the MACE, which also possessed little reliable variance independent of the general factor and was contaminated by a reverse-scoring methods factor.
-In sum, whereas the MACE total score is reliable and valid, subscale scores derived from a subset of items are seldom interpretable as such and instead reflect general maltreatment.
+We also calculated reliability indices for the MACE total and subscale scores.
+We found that the MACE total score has excellent reliability and is a valid measure of overall maltreatment.
+In contrast, we found that virtually no MACE subscale scores were reliable measures of particular types of maltreatment.
+For all but one of the 10 MACE subscales, the majority of reliable variance in scores was attributable to the general factor.
+The same was true of deprivation subscale scores, which were also contaminated by a reverse-scoring methods factor.
+In short, our results further validate the use of the MACE total score, but caution against the use of MACE subscale scores.
+Indeed, MACE subscale scores seldom reflected their intended construct but instead reflected general maltreatment.
 
-Of the four confirmatory factor models we tested, the best-fitting and most parsimonious was a bifactor S-1 model with dominant general maltreatment factor and two smaller specific factors reflecting peer victimization and reverse-scoring (model 4).
-This model of responses on the MACE was further supported by the exploratory factor analyses, where the 3-factor solution found a similar partitioning of the items.
-Notably, across samples, subscale scores derived from the peer victimization items were both reliable and interpretable insofar that the majority of their variance were attributable to the peer victimization factor even after accounting for the general maltreatment factor.
-It therefore may be valid to derive two score measures from the MACE, maltreatment by adults and by peers, but further research is needed to properly validate these measures.
+Of the four confirmatory factor models we tested, the most parsimonious and best- fitting was a bifactor S-1 model with a prominent general maltreatment factor and two specific factors measuring peer victimization and reverse-scoring.
+This model of the MACE factor structure was additionally supported by the exploratory factor analyses, which produced a qualitatively similar partitioning of the items.
+Thus, it appears the MACE is essentially unidimensional with secondary factors affecting a small subset of items.
+Interestingly, we found the peer victimization subscale scores to be reliable; a majority of their variance was attributable to the peer victimization factor even after controlling for general maltreatment.
+It may therefore be appropriate to produce a peer victimization subscale score from the MACE, but additional research is needed to validate this measure.
 
-Our results are consistent with a number of previous studies investigating the factor structure of other childhood maltreatment measures.
-For example, bifactor models of the Childhood Trauma Questionnaire have also revealed a dominant general maltreatment factor and weaker specific factors \citep{spinhoven2014childhood, hollerbach2018main, stagaki2022mediating}.
+Our results are consistent with a number of previous studies investigating the factor structure of other childhood maltreatment questionnaires.
+For example, bifactor models of the Childhood Trauma Questionnaire have also revealed a dominant general maltreatment factor and weaker specific factors \citep{spinhoven2014childhood, stagaki2022mediating}.
 Similar results were found in recent applications of a bifactor model to the International Child Abuse Screening Tool \citep{meinck2021factor} and the Adverse Childhood Experience questionnaire \citep{dobson2021latent}.
-Together, these and our current results underscore the challenge of developing versatile measures of childhood maltreatment.
-When the experience of distinct forms of abuse are correlated, it is difficult to measure any particular type of abuse independent of overall maltreatment.
+Together, these and our current results underscore a challenge in investigating the consequences of childhood maltreatment.
+When maltreatment experiences tend to co-occur, it is difficult to measure the effects of a particular type of adversity unique from overall maltreatment.
 
-Our results are also worth considering alongside previous studies that used subscale scores from the MACE to investigate the link between specific forms of childhood maltreatment and assorted psycho-biological outcomes.
+Our results are also worth considering alongside previous studies that used MACE subscale scores to investigate the link between particular types of childhood maltreatment and assorted psychobiological outcomes.
 For example, researchers have looked at the relationship between MACE subscale scores and risk of depression \citep{gerke2018childhood}, dissociative symptoms, \citep{schalinski2015type}, and cortisol concentration \citep{schalinski2019early}.
-Still others have derived threat and deprivation subscale scores from the MACE to study their associations with cognitive functioning \citep{schalinski2018defining} and hippocampal volume \citep{teicher2018differential}.
-Our results are unequivocal that these measures principally reflect general maltreatment and not particular forms of adversity.
-This presents a challenge for interpreting the results of those studies.
-It is worth remarking that several of these studies attempted to overcome these issues by simultaneously included all of the MACE subscale scores in their models using more sophisticated multivariate analyses (e.g.
+Still others have used threat and deprivation subscale scores from the MACE to study their associations with cognitive functioning \citep{schalinski2018defining} and hippocampal volume \citep{teicher2018differential}.
+Our results are clear that these scores chiefly reflect general maltreatment and not particular types of adversity, which presents a challenge for interpreting the conclusions of those studies.
+It is worth noting that some of these studies attempted to control for general maltreatment by entering all of the MACE subscale scores simultaneously into sophisticated multivariate analyses (e.g.
 random forest regression).
 This approach is unsatisfactory for two reasons.
-First, random forest models can return spurious results when handling collinear predictors \citep{gregorutti2017correlation}.
-Second, even if one were to properly control for general maltreatment, our results indicate that the remaining reliable variance in subscale scores is so low as to render any inference of questionable utility.
+First, random forest models can produce spurious results when dealing with collinear predictors \citep{gregorutti2017correlation}.
+Second, even if one were to properly control for general maltreatment, our results indicate that the remaining reliable variance in most subscale scores is so low as to cast doubt about the utility of these inferences.
 
-One might question whether it is at all appropriate to predict childhood maltreatment data using common factor models (e.g.
+%One might question whether it is appropriate to model childhood maltreatment data using common factor models (e.g.
 bifactor models).
-The core assumption of common factor models is that the direction of causality is from the factor to the indicators.
-That is, variation in item responses (e.g.
-low mood, anergia) are caused by or \emph{relfect} variation in the shared factor (e.g.
-depression).
-This is in contrast to composite variable models where the direction of causality is reversed.
-That is, the indicators (e.g.
-income, education) cause or \emph{form} the latent variable (e.g.
-socioeconomic status).
-As a consequence, the models are differentiable in two keys ways \citep{bollen1991conventional, bollen2011three}.
+The core assumption of these models is that the direction of causality is from the factor to the indicators.
+That is, variation in item responses are caused by (or reflect) variation in a shared factor.
+This is in contrast to composite variable models where the direction of causality is reversed; there the indicators cause (or form) the latent variable.
+Common factor and composite variable models are differentiable in two important ways \citep{bollen1991conventional, bollen2011three}.
 First, covariation among indicators is expected under a common factor model due to the influence of a shared common factor; conversely, covariation among indicators is not required (though may still occur) under a composite variable model.
@@
 
-For the purposes of explaining patterns of responding on the MACE, we think we are justified in using a common factor model.
+%For the purposes of explaining patterns of responding on the MACE, we think we are justified in using a common factor model.
 The vast majority of items on the MACE measure maltreatment by parents or other adults in the home.
@@
 
-This in turn raises another question: whether it is reasonable for indicators of maladaptive parenting and peer victimization on the MACE to share a common factor.
+%This in turn raises another question: whether it is reasonable for indicators of maladaptive parenting and peer victimization on the MACE to share a common factor.
 At first blush, one might reasonably assume that the causes of the former would be independent of the causes of the latter.
@@
 
-The current study revealed other features of the MACE worth remarking on.
-For example, we identified several sources of DIF on the MACE.
-Across samples we found that women experienced more sexual abuse, whereas men experienced more physical abuse by peers.
+The current study revealed other features of the MACE worth highlighting.
+For example, we identified several sources of DIF in the MACE.
+We found that across samples women experienced more sexual abuse whereas men experienced more physical abuse by peers.
 These findings are in line with previous research \citep{radford2013prevalence}.
-We also identified a number of items that exhibited DIF by sample.
-The cause of this is unclear, and may reflect differences in sample population (greater Boston area vs.
+We also identified a number of items with DIF by sample.
+The cause for this is unclear and may reflect differences in sample population (greater Boston area vs.
 US \& Canada), study location (in clinic vs.
@@
 attention checks).
-The causes and impact of violations of measurement invariance on the MACE deserve further study.
+The causes and impact of violations of measurement invariance on the MACE require further study.
 
 We also identified a reverse-scoring methods factor in the MACE.
-In our view, the most plausible explanation is that, under the \cite{teicher2015maltreatment} scoring procedure, affirmative responses on the regular and reverse-scored items on the MACE represent very different responses.
+In our view, the most plausible explanation is that, under the \cite{teicher2015maltreatment} scoring procedure, affirmative responses on the regular and reverse-scored items on the MACE represent very different endorsements.
 For a regular item on the MACE, an endorsement indicates that a person experienced an adverse event during one year of childhood at least.
 In contrast, an endorsement on a reverse-scored item indicates that a person experienced experienced the absence of a positive event for all 18 years of childhood.
-Alternative causes of the reverse-scoring factor, such as participant confusion, acquiescence, or carelessness \citep{weijters2013reversed}, are also possible.
+Alternative causes of the reverse-scoring factor such as participant confusion, acquiescence, or carelessness \citep{weijters2013reversed} are also possible.
 Additional research is needed to see if alternative scoring procedures eliminates this factor.
@@
 The current study is not without limitations.
-The most apparent limitation is that here we only studied the dichotomous response data but not the chronology data also solicited as part of the MACE.
-Though we found that the factor structure of the dichotomous responses are essentially unidimensional, this may belie more complex patterns present in the maltreatment timing data.
-Indeed, modeling the timeseries of adverse events in childhood --- perhaps using dynamic factor analysis \citep{zhang2007bayesian} --- might be more sensitive to more nuanced patterns of maltreatment co-occurrence that are lost when collapsing the data into a measure of any vs.
-no maltreatment.
-Another possibility is to jointly model the MACE endorsement and chronology data using a multidimensional hurdle model \citep{magnus2021symptom}, which measures both a person's susceptibility to maltreatment (i.e.
-number of unique types of maltreatment experienced) and severity of maltreatment (i.e.
-for how long maltreatment was experienced).
-More research is needed to develop valid and reliable MACE scores that integrate both the endorsement and chronology data.
+The most apparent limitation is that we only studied the dichotomous response data, but not the chronology data that is also collected as part of the MACE.
+Though we found that the factor structure of the endorsement responses is essentially unidimensional, this may belie more complex patterns present in the maltreatment chronology data.
+Indeed, longitudinal models of adverse events across childhood --- perhaps using dynamic factor analysis \citep{zhang2007bayesian} --- might be more sensitive to patterns of maltreatment covariation that are lost when collapsing the data into a dichotomous indicator of none vs.
+any maltreatment.
+Another possibility is to jointly model the MACE endorsement and chronology data using a multidimensional hurdle model \citep{magnus2021symptom}, which measures both a person's susceptibility to maltreatment (number of maltreatment experiences) and severity of maltreatment (duration of maltreatment experiences).
+Further research is needed to develop maltreatment scores that integrate the MACE endorsement and chronology data.
 
-A second limitation concerns our sample.
-Though the inclusion of the replication sample increased the overall diversity of our participants with respect to gender and age, other aspects of diversity are notably lacking.
-For example, our combined sample was not racially or ethnically diverse.
-This is important as previous studies have identified DIF for childhood trauma questionnaires by race \citep{thombs2007evaluation, rodriguez2019identification}, which raises the possibility that similar issues could be found for the MACE with more diverse samples.
-Moreover, as noted above, we are not currently able to explain why DIF was identified between the original and replication samples.
-Further research is needed to identify and unpack violations of measurement invariance in the MACE in order to effectively assess childhood maltreatment in many different populations.
+A second limitation is our sample demographics.
+Though the replication sample increased the overall diversity of our sample with respect to gender and age, other types of diversity are notably lacking.
+Our combined sample was neither racially nor ethnically diverse.
+This is important as previous studies have identified DIF for childhood adversity questionnaires by race \citep{rodriguez2019identification}, which raises the question whether similar issues would be found for the MACE in more diverse samples.
+Moreover, we are not currently able to explain why we found DIF for multiple  sample membership in the current study.
+Additional research is needed to study the functioning of the MACE in more diverse samples in order to validate its use in different populations.
 
-To conclude, we found that the factor structure of the MACE is essentially unidimensional.
-Despite the presence of smaller secondary factors, bifactor modeling revealed that the large majority of reliable variance in the dichotomous MACE responses are attributable to a general maltreatment factor.
-Thus, we found that the MACE total score is a valid and reliable of overall childhood maltreatment.
-In contrast, we found that MACE subscale scores are not valid measures of particular types of childhood maltreatment, but instead primarily reflect general maltreatment.
-Moving forward, childhood abuse and neglect researchers utilizing the MACE (or otherwise) should take pains to ensure that any subscale scores they employ are valid, lest they draw spurious inferences.
+To conclude, the results of the current study show the factor structure of the MACE is essentially unidimensional.
+Bifactor models revealed that, despite the presence of smaller secondary factors, the majority of reliable variance in the MACE endorsement responses are attributable to a general maltreatment factor.
+Accordingly, we found that the MACE total score is a reliable and valid of overall maltreatment.
+In contrast, we found that MACE subscale scores are not reliable measures of particular types of childhood maltreatment; instead they mainly reflect general maltreatment.
+Moving forward, childhood maltreatment researchers using the MACE (or any other scale) should ensure that any subscale scores used as part of their analyses are reliable, lest they draw spurious inferences.
 

### 2022-08-08 manuscript/main.tex
manuscipt edits (2022-08-07c)

@@
 \usepackage{setspace}             % adjust caption linespacing 
+\usepackage{multibib}             % references in supplement
 
@@
 
+% setup supplementary references
+\newcites{SM}{Supplementary references}
+
 % decrease size of references (lol)
@@
 \abstract{
+
+\textbf{Background}
+
+\textbf{Methods}
+
+\textbf{Results}
+
+\textbf{Conclusions}
 
@@
 
-%One might question whether it is appropriate to model childhood maltreatment data using common factor models (e.g.
-bifactor models).
-The core assumption of these models is that the direction of causality is from the factor to the indicators.
-That is, variation in item responses are caused by (or reflect) variation in a shared factor.
-This is in contrast to composite variable models where the direction of causality is reversed; there the indicators cause (or form) the latent variable.
-Common factor and composite variable models are differentiable in two important ways \citep{bollen1991conventional, bollen2011three}.
-First, covariation among indicators is expected under a common factor model due to the influence of a shared common factor; conversely, covariation among indicators is not required (though may still occur) under a composite variable model.
-Second, indicators under the common factor model, but not the composite variable model, are expected to share causal antecedents because they reflect a common source.
-Multiple authors have argued that adverse childhood experiences should (often) be treated as formative, not reflective \citep{netland2001assessment, layne2010unpacking}.
-Indeed it is the occurrence of abuse and neglect that defines childhood maltreatment, not the other way around.
-Moreover, theoretical models of the consequences of childhood maltreatment neither require that adverse experiences covary nor assume they share causal antecedents.
-
-%For the purposes of explaining patterns of responding on the MACE, we think we are justified in using a common factor model.
-The vast majority of items on the MACE measure maltreatment by parents or other adults in the home.
-Thus, the homogeneity of these indicators makes it plausible that they reflect a common source (e.g.
-maladaptive parenting, household dysfunction), as does their pattern of covariation as evidenced by strong loadings onto a common factor.
-Moreover, there is theoretical support for a model in which a common source like maladaptive parenting, itself caused by a diversity of parent level (e.g.
-emotional lability, impulse control, history of childhood abuse) and child-parent level factors, resulting in a multiple of forms of maltreatment \citep{belsky1993etiology}.
-This is empirically supported insofar that multiple forms of maltreatment share overlapping risk factors \citep{stith2009risk, mulder2018risk, assink2019risk}.
-Furthermore, the MACE is a retrospective self-report measure.
-Prospective and retrospective measures do not have perfect agreement \citep{patten2015retrospective, reuben2016lest}, and thus must reflect other factors related to the participants including individual differences in current state, memory, appraisal, personality, and biases in discloure \citep{susser2012still, reuben2016lest, colman2016consistency}.
-
-%This in turn raises another question: whether it is reasonable for indicators of maladaptive parenting and peer victimization on the MACE to share a common factor.
-At first blush, one might reasonably assume that the causes of the former would be independent of the causes of the latter.
-However, there is ample evidence that children who have been maltreated are at greater risk for subsequent peer victimization \citep{bolger2001developmental, benedini2016cycle}.
-The reasons for this association are manifold.
-For example, maltreatment may influence a child's sociocognitive development (e.g.
-attachment style, emotion regulation) in ways that may predispose them to peer victimization \citep{goemans2021child, McCrory2022-cj}.
-Parental neglect, or the absence of adult supervision, is also a risk factor for peer victimization \citep{lereya2013parenting}, potentially by increasing the likelihood of children participating in higher risk social situations where victimization may take place.
-Insofar that maladaptive parenting is an indirect cause of peer victimization (mediated by a number of complex psychosocial pathways), then we believe we are justified specifying indicators of peer victimization to load in part onto the common factor.
-This is also highlights the need for future study of the peer victimzation factor on the MACE, to better understand how it is related to and distinct from childhood maltreatment by parents and what its downstream consequences are.
-
 The current study revealed other features of the MACE worth highlighting.
@@
 In addition to the MACE, participants in both samples completed a number of additional self-report measures.
-Participants in the original sample completed multiple measures concerning psychiatric symptoms; these have already been reported \citep{teicher2015maltreatment}.
+Participants in the original sample completed multiple measures concerning psychiatric symptoms; these have already been reported \citepSM{teicher2015maltreatment}.
 The participants in the replication sample completed three additional self-report measures all assessing negative symptoms and motivation for rewards.
-Specifically, participants completed the Motivation and Pleasure scale \citep{llerena2013motivation}, the Self-evaluation of Negative Symptoms scale \citep{dollfus2016self}, and the revised Behavioral Inhibition / Behavioral Activation scale \citep{pagliaccio2016revising}.
+Specifically, participants completed the Motivation and Pleasure scale \citepSM{llerena2013motivation}, the Self-evaluation of Negative Symptoms scale \citepSM{dollfus2016self}, and the revised Behavioral Inhibition / Behavioral Activation scale \citepSM{pagliaccio2016revising}.
 
-Following best practices for online research \citep{zorowitz2021inattentive}, several attention checks were embedded in the secondary self-report measures filled out by the replication sample.
+Following best practices for online research \citepSM{zorowitz2021inattentive}, several attention checks were embedded in the secondary self-report measures filled out by the replication sample.
 Specifically, we used infrequency items which are questions for which all  or virtually all attentive participants should provide the same response.
@@
 Participants were excluded from analysis if they responded incorrectly to one or more of these items.
+
+%One might question whether it is appropriate to model childhood maltreatment data using common factor models (e.g.
+bifactor models).
+The core assumption of these models is that the direction of causality is from the factor to the indicators.
+That is, variation in item responses are caused by (or reflect) variation in a shared factor.
+This is in contrast to composite variable models where the direction of causality is reversed; there the indicators cause (or form) the latent variable.
+Common factor and composite variable models are differentiable in two important ways \citep{bollen1991conventional, bollen2011three}.
+First, covariation among indicators is expected under a common factor model due to the influence of a shared common factor; conversely, covariation among indicators is not required (though may still occur) under a composite variable model.
+Second, indicators under the common factor model, but not the composite variable model, are expected to share causal antecedents because they reflect a common source.
+Multiple authors have argued that adverse childhood experiences should (often) be treated as formative, not reflective \citep{netland2001assessment, layne2010unpacking}.
+Indeed it is the occurrence of abuse and neglect that defines childhood maltreatment, not the other way around.
+Moreover, theoretical models of the consequences of childhood maltreatment neither require that adverse experiences covary nor assume they share causal antecedents.
+
+%For the purposes of explaining patterns of responding on the MACE, we think we are justified in using a common factor model.
+The vast majority of items on the MACE measure maltreatment by parents or other adults in the home.
+Thus, the homogeneity of these indicators makes it plausible that they reflect a common source (e.g.
+maladaptive parenting, household dysfunction), as does their pattern of covariation as evidenced by strong loadings onto a common factor.
+Moreover, there is theoretical support for a model in which a common source like maladaptive parenting, itself caused by a diversity of parent level (e.g.
+emotional lability, impulse control, history of childhood abuse) and child-parent level factors, resulting in a multiple of forms of maltreatment \citep{belsky1993etiology}.
+This is empirically supported insofar that multiple forms of maltreatment share overlapping risk factors \citep{stith2009risk, mulder2018risk, assink2019risk}.
+Furthermore, the MACE is a retrospective self-report measure.
+Prospective and retrospective measures do not have perfect agreement \citep{patten2015retrospective, reuben2016lest}, and thus must reflect other factors related to the participants including individual differences in current state, memory, appraisal, personality, and biases in discloure \citep{susser2012still, reuben2016lest, colman2016consistency}.
+
+%This in turn raises another question: whether it is reasonable for indicators of maladaptive parenting and peer victimization on the MACE to share a common factor.
+At first blush, one might reasonably assume that the causes of the former would be independent of the causes of the latter.
+However, there is ample evidence that children who have been maltreated are at greater risk for subsequent peer victimization \citep{bolger2001developmental, benedini2016cycle}.
+The reasons for this association are manifold.
+For example, maltreatment may influence a child's sociocognitive development (e.g.
+attachment style, emotion regulation) in ways that may predispose them to peer victimization \citep{goemans2021child, McCrory2022-cj}.
+Parental neglect, or the absence of adult supervision, is also a risk factor for peer victimization \citep{lereya2013parenting}, potentially by increasing the likelihood of children participating in higher risk social situations where victimization may take place.
+Insofar that maladaptive parenting is an indirect cause of peer victimization (mediated by a number of complex psychosocial pathways), then we believe we are justified specifying indicators of peer victimization to load in part onto the common factor.
+This is also highlights the need for future study of the peer victimzation factor on the MACE, to better understand how it is related to and distinct from childhood maltreatment by parents and what its downstream consequences are.
 
@@
 
+\bibliographystyleSM{apalike}
+\bibliographySM{main}
+
 \end{document}

### 2022-08-08 manuscript/main.tex
add last paragraph of intro

@@
 Crucially, bifactor models can be used to calculate model-based reliability statistics, including the omega hierarchical ($\omega_h$) and omega subscale ($\omega_s$) indices \citep{reise2013scoring}, which respectively measure the proportion of variance in total and subscale scores attributable to the general factor and specific factors.
-Applied to the MACE, these indices can help determine whether subscale scores measuring particular types of maltreatment are interpretable as such, or if they instead chiefly reflect overall maltreatment.
+Applied to the MACE, these indices can help determine whether subscale scores measuring particular types of maltreatment are interpretable as such, or if they instead primarily reflect overall maltreatment.
 
-summary of paper paragraph tk. data availability statement to conclude
+In this study, we investigated the factor structure of the MACE in two independent samples of participants.
+First, we tested the fits of a series of confirmatory item response models to the maltreatment data from each sample.
+These models varied in the number of latent factors, which allowed us to examine the dimensionality of the MACE.
+We found that the MACE is essentially unidimensional; that is, the majority of reliable variance in participants' responses is attributable to a general maltreatment factor with strong loadings from most items.
+We then calculated the reliability of the MACE total score and a variety of subscale scores.
+We found the MACE total score to be a reliable and valid measure of general maltreatment, whereas most subscale scores were found to be invalid with little reliable variance unique from general maltreatment.
+Finally, we performed exploratory factor analysis to explore the factor structure of the MACE without \textit{a priori} restrictions.
+These analyses corroborated the results of the confirmatory factor analyses.
+All data, code, and analysis materials are publicly available at \url{https://github.com/szorowi1/mace-irt}.
 
@@
 
+Next we inspected the $Q_3$ indices for evidence of local dependence.
+For all models and samples, the maximum observed $Q_3$ index was greater than the critical $Q_3$ value.
+To gauge the extent of local dependence, we visualized the $Q_3$ values for item pairs exceeding the critical value for each model and sample (Figure \ref{fig:local_dependence}).
+The unidimensional model fits exhibited many locally dependent item pairs --- especially among the peer victimization and reverse-scored items --- suggesting possible underfactorization.
+By comparison, the three other models exhibited many fewer locally dependent item pairs.
+Of these pairs, the reasons for dependence were apparent.
+For example, the dependence between items 17 (``Parents made inappropriate sexual comments or suggestions to your sibling'') and 18 (``Parents touched or fondled your sibling in a sexual way'') observed in all models is almost certainly the result of measuring highly similar content.
+Regardless, the estimated item discrimination parameters for each model fit seldom exceeded the normal range ($\alpha \leq 4$).
+That is, the level of local dependence present in the models did not cause substantial bias in parameter estimation \citep{edwards2018diagnostic}.
+Thus, we conclude that all model fits provide at least an acceptable fit to the response data.
+
 \begin{table}[t!]
@@
 \end{table}
-
-Next we inspected the $Q_3$ indices for evidence of local dependence.
-For all models and samples, the maximum observed $Q_3$ index was greater than the critical $Q_3$ value.
-To gauge the extent of local dependence, we visualized the $Q_3$ values for item pairs exceeding the critical value for each model and sample (Figure \ref{fig:local_dependence}).
-The unidimensional model fits exhibited many locally dependent item pairs --- especially among the peer victimization and reverse-scored items --- suggesting possible underfactorization.
-By comparison, the three other models exhibited many fewer locally dependent item pairs.
-Of these pairs, the reasons for dependence were apparent.
-For example, the dependence between items 17 (``Parents made inappropriate sexual comments or suggestions to your sibling'') and 18 (``Parents touched or fondled your sibling in a sexual way'') observed in all models is almost certainly the result of measuring highly similar content.
-Regardless, the estimated item discrimination parameters for each model fit seldom exceeded the normal range ($\alpha \leq 4$).
-That is, the level of local dependence present in the models did not cause substantial bias in parameter estimation \citep{edwards2018diagnostic}.
-Thus, we conclude that all model fits provide at least an acceptable fit to the response data.
 
@@
 Only items from the peer verbal and physical abuse subscales had specific factor loadings of equal or greater magnitude to those on the general factor.
-
-\begin{table}[h!]
+\begin{table}[t!]
 \centering
@@
 
+The goodness-of-fit measures for the exploratory factor models to the response data by sample are summarized in Table \ref{tab:efa_diagnostics}.
+The fit indices indicated that all exploratory models provided an adequate fit to the data.
+The standardized factor loadings for the 2- and 3-factor solutions produced by the cf-quartimax rotation are presented in Figure \ref{fig:efa_cf}.
+(The corresponding factor loadings produced by the geomin rotation were qualitatively similar and are presented in Figure \ref{fig:efa_geomin}.) In the 2-factor solution for the original sample, we observed a primary maladaptive parenting factor and a secondary peer victimization factor.
+In contrast, the replication sample produced a primary maladaptive parenting/peer maltreatment factor with a secondary reverse-scoring factor.
+Neither 2-factor solution precisely resembled the threat-deprivation factor structure of confirmatory model 3.
+
 \begin{figure}[t!]
@@
 \end{figure}
-
-The goodness-of-fit measures for the exploratory factor models to the response data by sample are summarized in Table \ref{tab:efa_diagnostics}.
-The fit indices indicated that all exploratory models provided an adequate fit to the data.
-The standardized factor loadings for the 2- and 3-factor solutions produced by the cf-quartimax rotation are presented in Figure \ref{fig:efa_cf}.
-(The corresponding factor loadings produced by the geomin rotation were qualitatively similar and are presented in Figure \ref{fig:efa_geomin}.) In the 2-factor solution for the original sample, we observed a primary maladaptive parenting factor and a secondary peer victimization factor.
-In contrast, the replication sample produced a primary maladaptive parenting/peer maltreatment factor with a secondary reverse-scoring factor.
-Neither 2-factor solution precisely resembled the threat-deprivation factor structure of confirmatory model 3.
 

### 2022-08-11 analysis/scripts/fit_grmb.py
add back tetrachoric correlation script; add READMEs

@@ -10,7 +10,7 @@ ROOT_DIR = dirname(dirname(os.path.realpath(__file__)))
 #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
 
 ## I/O parameters.
-stan_model = 'baseline'
+stan_model = 'grmb'
 study = sys.argv[1]
 
 ## Sampling parameters.

### 2022-08-11 analysis/scripts/fit_tetrachoric.py

@@ -1,0 +1,72 @@
+import os, sys
+import numpy as np
+from os.path import dirname
+from pandas import DataFrame, read_csv
+from cmdstanpy import CmdStanModel
+from tqdm import tqdm
+ROOT_DIR = dirname(dirname(os.path.realpath(__file__)))
+
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+### Define parameters.
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+
+## I/O parameters.
+stan_model = 'tetrachoric'
+
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+### Load and prepare data.
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+
+## Load data.
+mace = read_csv(os.path.join(ROOT_DIR, 'data', 'mace.csv'))
+
+## Apply rejections.
+covariates = read_csv(os.path.join(ROOT_DIR, 'data', 'covariates.csv')).query('infreq <= 0.5')
+mace = mace.loc[mace.subject.isin(covariates.subject)]
+
+## Pivot data.
+mace = mace.pivot_table('response', ['study','subject'], 'item')
+
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+### Compute tetrachoric correlations.
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
+
+## Load StanModel
+StanModel = CmdStanModel(stan_file=os.path.join('stan_models', f'{stan_model}.stan'))
+
+## Define indices.
+indices = np.tril_indices(mace.columns.size, k=-1)
+
+## Main loop.
+tetrachoric = []
+for i, j in tqdm(np.column_stack(indices)):
+    
+    ## Preallocate space.
+    row = dict(k1=i, k2=j)
+    
+    ## Compute tetrachoric correlation (teicher2015).
+    df = mace.loc['teicher2015', [i+1,j+1]].dropna()
+    dd = dict(J=df.shape[0], Y=df.values.astype(int))
+    StanFit = StanModel.optimize(data=dd, seed=0)
+    row['teicher2015'] = float(StanFit.optimized_params_pd['rho'])
+    
+    ## Compute tetrachoric correlation (tuominen2022).
[…]

### 2022-08-11 analysis/scripts/tetrachoric.R

@@ -1,51 +1,0 @@
-library(tidyverse)
-library(psych)
-
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
-### Define parameters.
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
-
-## Load data.
-data = read.csv('../data/mace.csv')
-
-## Define studies.
-studies = list('teicher2015', 'tuominen2022')
-
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
-### Main loop.
-#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
-
-## Preallocate space.
-corr = c()
-
-for (study in studies) {
-  
-  ## Restrict to study.
-  df = data[data$study == study,]
-  
-  ## Construct pivot table.
-  df <- df %>% 
-    pivot_wider(id_cols = subject, names_from = item, names_prefix= "x", values_from = response) 
-  
-  ## Drop subject column.
-  df <- subset(df, select=-c(subject))
-  
-  ## Compute tetrachoric correlation.
-  df = tetrachoric(df)$rho
-  df = round(df, 6)
-  
-  ## Convert to long-list.
-  df = as.data.frame(as.table(df))
-  df = rename(df, k1 = Var1, k2 = Var2, !!study := Freq)
-  
-  ## Append to list.
-  corr = append(corr, list(df), 0)
-  
-}
-
-## Merge DataFrames.
-corr = corr %>%
-  Reduce(function(dtf1,dtf2) left_join(dtf1,dtf2,by=c("k1","k2")), .)
-
-## Save data.
-write.csv(corr, paste('..', 'stan_results', 'tetrachoric.csv', sep='/'), row.names=FALSE)

### 2022-08-11 manuscript/main.tex
update factor loading thresholds; updates to manuscript (edits + reformatting floats)

--- 2022-08-08 manuscript/main.tex
@@
-\documentclass[letterpaper,man,natbib,floatsintext,longtable]{apa6}
+\documentclass[letterpaper,man,natbib,noextraspace,floatsintext,longtable]{apa6}
 
@@
 \usepackage{multibib}             % references in supplement
+\usepackage{hyperref}             % manage hyperlink colors
 
@@
 
+% change hyperlink colors
+\hypersetup{
+    colorlinks=true,
+    linkcolor=blue,
+    filecolor=magenta,      
+    urlcolor=cyan,
+    citecolor=blue
+}
+
 \title{The MACE is essentially unidimensional: Bifactor models of the `Maltreatment and Abuse Chronology of Exposure' scale}
@@
 
-\textbf{Background}
+\textbf{Background}: Typically measures of childhood maltreatment collect information about several types of abuse and maltreatment.
+The subscales of the retrospective Maltreatment and Abuse Chronology of Exposure (MACE) scale measure 10 such types.
+Since different types of maltreatment often co-occur, it is not clear whether subscale measures can be used to reflect constructs unique from from overall maltreatment.
 
-\textbf{Methods}
+\hfill
 
-\textbf{Results}
+\textbf{Methods}: Here, we investigated the factor structure of the MACE scale in two independent samples (N=1051 \& N=582). We used both confirmatory and exploratory item response modeling.
 
-\textbf{Conclusions}
+\hfill
 
+\textbf{Results}: We found the MACE total score to be a reliable and valid measure of general maltreatment.
+Confirmatory item response models provided strong evidence for unidimensionality of the MACE, with the exception of a potential peer victimization factor.
+That is, most items in the MACE load on general factor and apart from peer victimization very little variance was explained by other subscales.
+Exploratory factor modeling further supported these findings.
+
+\hfill
+
+\textbf{Conclusions}: The MACE is essentially a unidimensional scale.
+Apart from a separate peer victimization factor, researchers should refrain from using subscales data separately since they mostly reflect overall maltreatment.
+\hfill \break
 }
+
+\keywords{child maltreatment, reliability, item response theory, bifactor model}
+
+\authornote{This work was funded in part by an NSF Graduate Research Fellowship. The authors have no conflicts of interest to declare.}
 
@@
 Childhood maltreatment is an unfortunately common experience for children across the globe \citep{stoltenborgh2015prevalence}.
-It is a risk factor for multiple poor outcomes later in life, including the onset of mental illness \citep{kessler2010childhood}, challenges in cognitive functioning \citep{su2019does}, and poor physical health \citep{wegman2009meta}.
+It is a risk factor for multiple poor outcomes later in life including the development of mental illness \citep{kessler2010childhood}, challenges in cognitive functioning \citep{su2019does}, and poor physical health \citep{wegman2009meta}.
 Childhood maltreatment also predicts worse treatment outcomes for several psychiatric disorders \citep{nanni2012childhood, thomas2019childhood}.
@@
 The MACE is a 52-item retrospective self-report scale composed of 10 subscales, each of which measures a unique type of childhood maltreatment.
-Although six of these subscales overlap with other standard measures of abuse (e.g.
-physical abuse, neglect), the remaining four are less commonly represented in other scales (e.g.
+Six of these subscales overlap with other standard measures of abuse (e.g.
+physical abuse, neglect), whereas the remaining four are less commonly represented in other scales (e.g.
 peer victimization, witnessing domestic violence).
-The items in the MACE were carefully selected to measure maltreatment of varying severity in order to finely differentiate people with differing exposures.
+The items in the MACE were carefully selected to measure maltreatment of varying severity in order to finely differentiate people with differing exposure levels.
 The MACE also asks for the onset and duration of maltreatment experiences, which is necessary for investigating how the timing of abuse shapes development.
@@
 These scores have excellent temporal stability, convergent validity with other childhood maltreatment scales, and good predictive validity for a multitude of psychiatric symptoms \citep{teicher2015maltreatment}.
-For these reasons the MACE has been included as among the best measures of childhood maltreatment currently available \citep{saini2019systematic, georgieva2022systematic}.
+For these reasons, the MACE has been included as among the best measures of childhood maltreatment currently available \citep{saini2019systematic, georgieva2022systematic}.
 
 In recent years, there have been calls for childhood maltreatment research to move away from the cumulative risk framework \citep{evans2013cumulative} --- which focuses only on the number of adverse experiences --- towards a dimensional approach, which emphasizes how distinct types of maltreatment uniquely shape developmental trajectories \citep{mclaughlin2016beyond, belsky2012beyond}.
-In response, some researchers have begun to use the MACE to form new indices of maltreatment in lieu of, or in addition to, the total and multiplicity scores.
+In response, some researchers have begun to use the MACE to make new indices of maltreatment in lieu of, or in addition to, the total and multiplicity scores.
 For example, some have used all 10 MACE subscale scores to study the unique contributions of each type of maltreatment to the development of psychopathology \citep{schalinski2015type, gerke2018childhood, schalinski2019early}.
 Inspired by a prominent dimensional model of maltreatment \citep{mclaughlin2014childhood}, others have made new subscale scores indexing threat and deprivation experiences using the items in the MACE that respectively measure abuse and neglect \citep{schalinski2018defining, schalinski2019environmental, teicher2018differential}.
-The psychometric properties of these new subscale scores have not been thoroughly investigated.
+Yet the psychometric properties of these new subscale scores have not been thoroughly investigated.
 
-A critical challenge in studying the consequences of childhood maltreatment is the high co-occurrence of adverse experiences.
+A critical challenge in studying the consequences of childhood maltreatment is the high co-occurrence of these adverse experiences.
 That is, children who have experienced any single form of maltreatment are likely to have experienced multiple other types \citep{dong2004interrelatedness, herrenkohl2009assessing, kessler2010childhood}.
 As a consequence, isolating the unique contribution of one type of maltreatment to some outcome is difficult insofar as any one type of adversity is likely confounded with overall maltreatment.
-Concretely, this means that researchers intending to study the effects of particular types of maltreatment using subscale scores need to verify that those measures actually reflect a construct unique from from overall maltreatment, lest they draw spurious conclusions.
+Concretely, this means that researchers intending to study the effects of particular types of maltreatment using subscale scores need to verify that those measures actually reflect a construct unique from from overall maltreatment, lest they risk drawing spurious conclusions.
 
@@
 Crucially, bifactor models can be used to calculate model-based reliability statistics, including the omega hierarchical ($\omega_h$) and omega subscale ($\omega_s$) indices \citep{reise2013scoring}, which respectively measure the proportion of variance in total and subscale scores attributable to the general factor and specific factors.
-Applied to the MACE, these indices can help determine whether subscale scores measuring particular types of maltreatment are interpretable as such, or if they instead chiefly reflect overall maltreatment.
+Applied to the MACE, these indices can help determine if subscale scores measuring particular types of maltreatment are interpretable as such, or if they instead primarily reflect overall maltreatment.
 
-summary of paper paragraph tk. data availability statement to conclude
+In this study, we investigated the factor structure of the MACE in two independent samples of participants.
+First, we tested the fits of a series of confirmatory item response models to the maltreatment data from each sample.
+These included several bifactor models, which allowed us to quantify the proportion of variance in the data attributable to general maltreatment versus specific types of abuse.
+We found that the MACE is essentially unidimensional; that is, the majority of reliable variance in participants' responses was attributable to a general maltreatment factor.
+We then calculated the reliability of the MACE total score and a host of subscale scores.
+We found the total score was a reliable and valid measure of general maltreatment, whereas most subscale scores contained little-to-no variance unique from general maltreatment.
+The only possible exception to this trend was for a novel peer victimization subscale.
+Finally, we performed exploratory factor analysis to examine the factor structure of the MACE without \textit{a priori} restrictions.
+These analyses supported the results of the confirmatory factor analyses.
 
@@
 
-\subsection{Participants}
+All data, code, and analysis materials are publicly available at \url{https://github.com/szorowi1/mace-irt}.
+
+{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Participants}
 
@@
 
-\subsection{Measures}
+{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Measures}
 
@@
 
-\subsection{Exclusion criteria}
+{\begin{spacing}{0.2} \hfill \\ \end{spacing}}  \subsection{Exclusion criteria}
 
@@
 
-\subsection{Differential item functioning}
+{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Differential item functioning}
 
@@
 
-\subsection{Response dependence}
+{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Response dependence}
 
@@
 
-\subsection{Confirmatory item factor models}
+{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Confirmatory item factor models}
 
@@
 
-\subsection{Goodness of fit \& model comparison}
+{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Goodness of fit \& model comparison}
 
@@
 
-\subsection{Model-based reliability indices}
+{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Model-based reliability indices}
 
 To evaluate the reliability of the MACE total and subscale scores, we calculated several model-based reliability indices using the factor loadings from the bifactor models.
-First, we calculated coefficient omega ($\omega$; \citealt{mcdonald1999test}), which quantifies the proportion of reliable score variance attributable to all factors (i.e.
+First, we calculated coefficient omega ($\omega$; \citealt{mcdonald2013test}), which quantifies the proportion of reliable score variance attributable to all factors (i.e.
 both general and specific factors).
@@
 
-\subsection{Exploratory factor analysis}
+{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Exploratory factor analysis}
 
@@
 We decided to extract two- and three- factor solutions based on the total number of factors in the bifactor S-1 models.
-We were particularly interested to see if a two-factor solution would reproduce the factors from model 3 (threat, deprivation); and if a three- factor solution would reproduce the three factors from model 4 (maladaptive parenting, peer victimization, reverse-scoring).
+We were particularly interested to see if a two-factor solution would reproduce the factors from model 3 (threat, deprivation); and if a three- factor solution would reproduce the three factors from model 4 (parental maltreatment, peer victimization, reverse-scoring).
 
@@
 All RMSEA value were <0.05 and most CFI and TLI values were >0.95.
-The only exception were for the fits of models 1 and 3 to the original sample data where CFI/TLI > 0.90.
+The only exception were for the fits of models 1 and 3 to the original sample data where CFI \& TLI > 0.90.
 In general, model fits were better for the replication sample data than for the original sample.
@@
 \end{adjustbox}
-\captionsetup{width=1.1\textwidth}
+\captionsetup{width=1\textwidth}
 \caption{\normalfont Fit statistics for the confirmatory factor models by sample.
@@
 
-\subsection{Confirmatory item response models}
+{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Confirmatory item response models}
 
@@
 
-\subsubsection{Model 1: One-factor (unidimensional) model of the MACE}
+{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsubsection{Model 1: One-factor (unidimensional) model}
+
+\begin{figure}[tp]
+    \centering
+    \includegraphics[width=1.1\textwidth,center]{figures/fig02.png}
+    \captionsetup{width=1.1\textwidth}
+    \caption{Standardized factor loadings from the four models fit to the original sample.
+Notes: Model 1 = one-factor (unidimensional) model; Model 2 = bifactor model with 10 specific factors; Model 3 = bifactor S-1 model with one specific factor (neglect); Model 4 = bifactor S-1 with two specific factors (peer victimization, reverse-scoring); VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; WSV = witnessing sibling vioence; WIPV = witnessing inter-parental violence; EN = emotional neglect; PN = physical neglect; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.
+Factor loadings $<$ 0.3 are displayed in gray; factor loadings $\geq$ 0.6 are bolded.
+** Reverse-scored items.}
+    \label{fig:loadings_original}
+\end{figure}
+
+\begin{figure}[tp]
+    \centering
+    \includegraphics[width=1.1\textwidth,center]{figures/fig03.png}
+    \captionsetup{width=1.1\textwidth}
+    \caption{Standardized factor loadings from the four models fit to the replication sample.
+Notes: Model 1 = one-factor (unidimensional) model; Model 2 = bifactor model with 10 specific factors; Model 3 = bifactor S-1 model with one specific factor (neglect); Model 4 = bifactor S-1 with two specific factors (peer victimization, reverse-scoring); VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; WSV = witnessing sibling vioence; WIPV = witnessing inter-parental violence; EN = emotional neglect; PN = physical neglect; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.
+Factor loadings $<$ 0.3 are displayed in gray; factor loadings $\geq$ 0.6 are bolded.
+** Reverse-scored items.}
+    \label{fig:loadings_online}
+\end{figure}
 
@@
 
-\begin{figure}[t!]
-    \centering
-    \includegraphics[width=1.1\textwidth,center]{figures/fig02.png}
-    \captionsetup{width=1.1\textwidth}
-    \caption{Standardized factor loadings from the four models estimated from the original sample.
-Notes: VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; WSV = witnessing sibling vioence; WIPV = witnessing inter-parental violence; EN = emotional neglect; PN = physical neglect; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.
-** Reverse-scored items.}
-    \label{fig:loadings_original}
-\end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=1.1\textwidth,center]{figures/fig03.png}
-    \captionsetup{width=1.1\textwidth}
-    \caption{Standardized factor loadings from the four models estimated from the online sample.
-Notes: VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; WSV = witnessing sibling vioence; WIPV = witnessing inter-parental violence; EN = emotional neglect; PN = physical neglect; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.
-** Reverse-scored items.}
-    \label{fig:loadings_online}
-\end{figure}
-
-\subsubsection{Model 2: Bifactor model with 10 specific factors}
+{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsubsection{Model 2: Bifactor model with 10 specific factors}
 
@@
 
-\subsubsection{Model 3: Bifactor S-1 model with one specific factor (neglect)}
+{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsubsection{Model 3: Bifactor S-1 model with one specific factor (neglect)}
 
@@
 
-\subsubsection{Model 4: Bifactor S-1 with two specific factors (peer victimization, reverse-scoring)}
+{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsubsection{Model 4: Bifactor S-1 with two specific factors (peer victimization, reverse-scoring)}
 
@@
 
-\subsection{Exploratory factor analysis}
-
-\begin{figure}[t!]
-    \centering
-    \includegraphics[width=1\textwidth,center]{figures/fig04.png}
-    \caption{\normalfont Standardized factor loadings for the 2- and 3-factor solutions produced by the cf-quartimax rotation for original sample (top row) and replication sample (bottom row).
-Notes: VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; EN = emotional neglect; PN = physical neglect; WSV = witnessing sibling violence; WIPV = witnessing inter-parental violence; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.
-Only factor loadings $\geq$ 0.30 are shown.
-** Reverse-scored items.}
-    \label{fig:efa_cf}
-\end{figure}
+{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Exploratory factor analysis}
 
@@
 Neither 2-factor solution precisely resembled the threat-deprivation factor structure of confirmatory model 3.
+
+\begin{figure}[t!]
+    \centering
+    \includegraphics[width=1\textwidth,center]{figures/fig04.png}
+    \caption{\normalfont Standardized factor loadings for the 2- and 3-factor solutions produced by the cf-quartimax rotation for original sample (top row) and replication sample (bottom row).
+\break \small Notes: VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; EN = emotional neglect; PN = physical neglect; WSV = witnessing sibling violence; WIPV = witnessing inter-parental violence; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.
+Only factor loadings $\geq$ 0.30 are shown.
+** Reverse-scored items.}
+    \label{fig:efa_cf}
+\end{figure}
 
@@
 This is important as previous studies have identified DIF for childhood adversity questionnaires by race \citep{rodriguez2019identification}, which raises the question whether similar issues would be found for the MACE in more diverse samples.
-Moreover, we are not currently able to explain why we found DIF for multiple  sample membership in the current study.
+Moreover, we are not currently able to explain why we found DIF in multiple items by sample membership in the current study.
 Additional research is needed to study the functioning of the MACE in more diverse samples in order to validate its use in different populations.
@@
 In contrast, we found that MACE subscale scores are not reliable measures of particular types of childhood maltreatment; instead they mainly reflect general maltreatment.
-Moving forward, childhood maltreatment researchers using the MACE (or any other scale) should ensure that any subscale scores used as part of their analyses are reliable, lest they draw spurious inferences.
+Moving forward, childhood maltreatment researchers using the MACE (or any other scale) should ensure that any subscale scores used as part of their analyses are reliable, lest they risk drawing spurious inferences.
 
@@
 Thus, the homogeneity of these indicators makes it plausible that they reflect a common source (e.g.
-maladaptive parenting, household dysfunction), as does their pattern of covariation as evidenced by strong loadings onto a common factor.
-Moreover, there is theoretical support for a model in which a common source like maladaptive parenting, itself caused by a diversity of parent level (e.g.
+parental maltreatment, household dysfunction), as does their pattern of covariation as evidenced by strong loadings onto a common factor.
+Moreover, there is theoretical support for a model in which a common source like parental maltreatment, itself caused by a diversity of parent level (e.g.
 emotional lability, impulse control, history of childhood abuse) and child-parent level factors, resulting in a multiple of forms of maltreatment \citep{belsky1993etiology}.
@@
 
-%This in turn raises another question: whether it is reasonable for indicators of maladaptive parenting and peer victimization on the MACE to share a common factor.
+%This in turn raises another question: whether it is reasonable for indicators of parental maltreatment and peer victimization on the MACE to share a common factor.
 At first blush, one might reasonably assume that the causes of the former would be independent of the causes of the latter.
@@
 Parental neglect, or the absence of adult supervision, is also a risk factor for peer victimization \citep{lereya2013parenting}, potentially by increasing the likelihood of children participating in higher risk social situations where victimization may take place.
-Insofar that maladaptive parenting is an indirect cause of peer victimization (mediated by a number of complex psychosocial pathways), then we believe we are justified specifying indicators of peer victimization to load in part onto the common factor.
-This is also highlights the need for future study of the peer victimzation factor on the MACE, to better understand how it is related to and distinct from childhood maltreatment by parents and what its downstream consequences are.
+Insofar that parental maltreatment is an indirect cause of peer victimization (mediated by a number of complex psychosocial pathways), then we believe we are justified specifying indicators of peer victimization to load in part onto the common factor.
+This is also highlights the need for future study of the peer victimization factor on the MACE, to better understand how it is related to and distinct from childhood maltreatment by parents and what its downstream consequences are.
 

### 2022-08-11 analysis/scripts/make_fig03.py

@@ -77,7 +77,7 @@ loadings = loadings.sort_values(['subscale','item']).set_index(['subscale','item
 
 ## Extract loadings.
 X1 = loadings.values
-X2 = loadings.applymap(np.digitize, bins=[0.35, 0.6]).values
+X2 = loadings.applymap(np.digitize, bins=[0.3, 0.6]).values
 
 #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
 ### Define plot parameters.

### 2022-08-11 manuscript/main.tex
manuscript edits (add formative vs. reflective indicator discussion to supplement)

@@
 
-To investigate the factor structure of the MACE, we fit a series of confirmatory item factor models separately to the response data from each sample.
+To investigate the factor structure of the MACE, we fit a series of confirmatory item factor models\footnote{The appropriateness of common factor models with reflective indicators for maltreatment data has been contested.
+We justify our usage of this model of the MACE for the interested reader in the supplement.} separately to the response data from each sample.
 The basis of each model is the graded response model for ordered polytomous data \citep{samejima1997graded}.
@@
     \item I worry about the 2001 Cricket World Cup. (none endorse)
-    \item Please select "Yes" and then select age "17". (instructed item)
+    \item Please select ``Yes'' and then select age ``17''. (instructed item)
 \end{enumerate}
@@
 
-%One might question whether it is appropriate to model childhood maltreatment data using common factor models (e.g.
+\break
+\subsection*{Reflective vs. formative indicator models of maltreatment data}
+
+One might question whether it is appropriate to model childhood maltreatment data using common factor models with reflective indicators (e.g.
 bifactor models).
-The core assumption of these models is that the direction of causality is from the factor to the indicators.
-That is, variation in item responses are caused by (or reflect) variation in a shared factor.
-This is in contrast to composite variable models where the direction of causality is reversed; there the indicators cause (or form) the latent variable.
-Common factor and composite variable models are differentiable in two important ways \citep{bollen1991conventional, bollen2011three}.
-First, covariation among indicators is expected under a common factor model due to the influence of a shared common factor; conversely, covariation among indicators is not required (though may still occur) under a composite variable model.
-Second, indicators under the common factor model, but not the composite variable model, are expected to share causal antecedents because they reflect a common source.
-Multiple authors have argued that adverse childhood experiences should (often) be treated as formative, not reflective \citep{netland2001assessment, layne2010unpacking}.
-Indeed it is the occurrence of abuse and neglect that defines childhood maltreatment, not the other way around.
-Moreover, theoretical models of the consequences of childhood maltreatment neither require that adverse experiences covary nor assume they share causal antecedents.
+The core assumption of these models is that variation in item responses are caused by or reflect variation in a shared latent factor (e.g.
+variation in low mood and anergia reflect latent variation in depression).
+This is in contrast to composite variable models with formative indicators where the direction of causality is reversed \citepSM{bollen1991conventional}.
+In composite variable models, indicators cause or form the latent construct (e.g.
+education and income together make up socioeconomic status).
+Some scholars have argued that measures of adverse childhood experiences often should be treated as formative, not reflective, indicators \citepSM{netland2001assessment, layne2010unpacking}.
+That is, exposure to abuse and neglect defines childhood maltreatment, not the other way around.
 
-%For the purposes of explaining patterns of responding on the MACE, we think we are justified in using a common factor model.
-The vast majority of items on the MACE measure maltreatment by parents or other adults in the home.
-Thus, the homogeneity of these indicators makes it plausible that they reflect a common source (e.g.
-parental maltreatment, household dysfunction), as does their pattern of covariation as evidenced by strong loadings onto a common factor.
-Moreover, there is theoretical support for a model in which a common source like parental maltreatment, itself caused by a diversity of parent level (e.g.
-emotional lability, impulse control, history of childhood abuse) and child-parent level factors, resulting in a multiple of forms of maltreatment \citep{belsky1993etiology}.
-This is empirically supported insofar that multiple forms of maltreatment share overlapping risk factors \citep{stith2009risk, mulder2018risk, assink2019risk}.
-Furthermore, the MACE is a retrospective self-report measure.
-Prospective and retrospective measures do not have perfect agreement \citep{patten2015retrospective, reuben2016lest}, and thus must reflect other factors related to the participants including individual differences in current state, memory, appraisal, personality, and biases in discloure \citep{susser2012still, reuben2016lest, colman2016consistency}.
+For the purposes of explaining covariation in maltreatment endorsements on the MACE, we believe we are justified in using a common factor model.
+Most of the items on the MACE measure maltreatment by parents or other guardians.
+The homogeneity of these indicators suggests they might reflect a common source; this hypothesis is further evidenced by the uniformly strong general factor loadings for this subset of items.
+Furthermore, there is ample theoretical and empirical support for regarding parental maltreatment as a common factor with its own causal antecedents \citepSM{belsky1993etiology}.
+Indeed, many of the same parent-level variables (e.g.
+emotional lability, impulse control, own history of abuse) are risk factors for different types of childhood maltreatment \citepSM{stith2009risk, mulder2018risk, assink2019risk}.
+Separately, that the MACE is a retrospective self-report measure is further cause for a common factor model.
+Retrospective measures of abuse have been shown to reflect additional factors beyond maltreatment exposure including individual differences in current emotional state, personality, and recall ability \citepSM{susser2012still, reuben2016lest, colman2016consistency}.
+Common factor models may better adjust for some of the measurement error in indicators caused by these nuisance factors.
 
-%This in turn raises another question: whether it is reasonable for indicators of parental maltreatment and peer victimization on the MACE to share a common factor.
-At first blush, one might reasonably assume that the causes of the former would be independent of the causes of the latter.
-However, there is ample evidence that children who have been maltreated are at greater risk for subsequent peer victimization \citep{bolger2001developmental, benedini2016cycle}.
+This discussion in turn raises a second question: whether it is reasonable for indicators of parental maltreatment and peer victimization on the MACE to share a common factor.
+One might assume that these two types of maltreatment reflect independent causes, and thus should not be permitted to share a common factor.
+There is considerable evidence, however, that children who have been maltreated are at greater risk for subsequent peer victimization \citepSM{bolger2001developmental, benedini2016cycle}.
 The reasons for this association are manifold.
-For example, maltreatment may influence a child's sociocognitive development (e.g.
-attachment style, emotion regulation) in ways that may predispose them to peer victimization \citep{goemans2021child, McCrory2022-cj}.
-Parental neglect, or the absence of adult supervision, is also a risk factor for peer victimization \citep{lereya2013parenting}, potentially by increasing the likelihood of children participating in higher risk social situations where victimization may take place.
-Insofar that parental maltreatment is an indirect cause of peer victimization (mediated by a number of complex psychosocial pathways), then we believe we are justified specifying indicators of peer victimization to load in part onto the common factor.
-This is also highlights the need for future study of the peer victimization factor on the MACE, to better understand how it is related to and distinct from childhood maltreatment by parents and what its downstream consequences are.
+For example, parental maltreatment may impact a child's sociocognitive functioning (e.g.
+attachment styles, emotion regulation) in ways that may predispose them to later peer victimization \citepSM{goemans2021child, McCrory2022-cj}.
+Separately, parental neglect is a risk factor for peer victimization \citepSM{lereya2013parenting}, possibly by increasing the chance that children enter into risky social situations where victimization may occur.
+Insofar that parental maltreatment is an indirect cause of peer victimization (mediated by multiple complex psychosocial pathways), then we believe we are justified in allowing indicators of peer victimization to load onto the common general maltreatment factor.
+This is also highlights the need for future study of a MACE peer victimization subscale to better understand how it is related to (and distinct from) parental maltreatment, and what its downstream consequences are.
 
@@
     \includegraphics[width=0.82\textwidth,center]{figures/figS01.png}
-    \caption{\normalfont \begin{spacing}{1.25} Differential item functioning by sample.
+    \caption{\normalfont Differential item functioning by sample.
 (A) Proportions of participants endorsing having experienced a maltreatment event by study sample.
@@
 Dotted lines indicate the threshold for large DIF recommended by \cite{hidalgo2014binary}.
-** Reverse-scored items.
-\end{spacing}}
+** Reverse-scored items.}
     \label{fig:dif_study}
@@
     \includegraphics[width=0.82\textwidth,center]{figures/figS02.png}
-    \caption{\normalfont \begin{spacing}{1.25} Differential item functioning by gender.
+    \caption{\normalfont Differential item functioning by gender.
 (A) Proportions of participants endorsing having experienced a maltreatment event by study gender.
@@
 Dotted lines indicate the threshold for large DIF recommended by \cite{hidalgo2014binary}.
-** Reverse-scored items.
-\end{spacing}}
+** Reverse-scored items.}
     \label{fig:dif_gender}

### 2022-08-15 manuscript/main.tex
minor updates: (1) fix efa bug; (2) update efa figures; (3) minor manuscript edits

@@
 
-\textbf{Background}: Typically measures of childhood maltreatment collect information about several types of abuse and maltreatment.
+\noindent \textbf{Background}: Typically measures of childhood maltreatment collect information about several types of abuse and maltreatment.
 The subscales of the retrospective Maltreatment and Abuse Chronology of Exposure (MACE) scale measure 10 such types.
@@
 
-\textbf{Methods}: Here, we investigated the factor structure of the MACE scale in two independent samples (N=1051 \& N=582). We used both confirmatory and exploratory item response modeling.
+\noindent \textbf{Methods}: Here, we investigated the factor structure of the MACE scale in two independent samples (N=1051 \& N=582). We used both confirmatory and exploratory item response modeling.
 
@@
 
-\textbf{Results}: We found the MACE total score to be a reliable and valid measure of general maltreatment.
+\noindent \textbf{Results}: We found the MACE total score to be a reliable and valid measure of general maltreatment.
 Confirmatory item response models provided strong evidence for unidimensionality of the MACE, with the exception of a potential peer victimization factor.
@@
 
-\textbf{Conclusions}: The MACE is essentially a unidimensional scale.
+\noindent \textbf{Conclusions}: The MACE is essentially a unidimensional scale.
 Apart from a separate peer victimization factor, researchers should refrain from using subscales data separately since they mostly reflect overall maltreatment.
@@
 Items on the MACE can be organized according to the source of maltreatment: parents (or other adults in the home) and peers.
-Thus, we hypothesized there would be two factors underlying responses on the MACE: a primary maladaptive parenting factor and a secondary peer victimization factor.
+Thus, we hypothesized there would be two factors underlying responses on the MACE: a primary parental maltreatment factor and a secondary peer victimization factor.
 Based on preliminary model fitting, we also suspected the presence of a tertiary methods factor affecting the six reverse-scored items.
-We initially fit a bifactor model in which all items loaded onto a general maltreatment factor and one of three specific factors (maladaptive parenting, peer victimization, reverse-scoring); however, we observed vanishing loadings on the first specific factor.
+We initially fit a bifactor model in which all items loaded onto a general maltreatment factor and one of three specific factors (parental maltreatment, peer victimization, reverse-scoring); however, we observed vanishing loadings on the first specific factor.
 As such, we instead fit a bifactor S-1 model where all items load onto a general maltreatment factor, and the peer victimization \& reverse-scoring items additionally load onto their own specific factors (Figure \ref{fig:models}d).
@@
 The standardized factor loadings for the 2- and 3-factor solutions produced by the cf-quartimax rotation are presented in Figure \ref{fig:efa_cf}.
-(The corresponding factor loadings produced by the geomin rotation were qualitatively similar and are presented in Figure \ref{fig:efa_geomin}.) In the 2-factor solution for the original sample, we observed a primary maladaptive parenting factor and a secondary peer victimization factor.
-In contrast, the replication sample produced a primary maladaptive parenting/peer maltreatment factor with a secondary reverse-scoring factor.
+(The corresponding factor loadings produced by the geomin rotation were qualitatively similar and are presented in Figure \ref{fig:efa_geomin}.) In the 2-factor solution for the original sample, we observed a primary parental maltreatment factor and a secondary peer victimization factor.
+In contrast, the replication sample produced a primary parental maltreatment/peer maltreatment factor with a secondary reverse-scoring factor.
 Neither 2-factor solution precisely resembled the threat-deprivation factor structure of confirmatory model 3.
@@
 The 3-factor solution for each sample were similar.
-EFA produced a primary maladaptive parenting factor with secondary peer victimization and reverse-scoring factors.
-(Interestingly, items concerning sexual abuse loaded more strongly on the peer victimization scale in the replication sample, raising some questions about the replicability of the factor structures.) To conclude, the structure of both the 2- and 3-factor exploratory factor models resembled the bifactor S-1 model with two specific factors (model 4), providing further support for that model of the factor structure of the MACE.
+EFA produced a primary parental maltreatment factor with secondary peer victimization and reverse-scoring factors.
+Interestingly, items concerning sexual abuse loaded more strongly on the peer victimization scale in the replication sample, raising some questions about the replicability of the factor structures.
+To conclude, the structure of both the 2- and 3-factor exploratory factor models resembled the bifactor S-1 model with two specific factors (model 4), providing further support for that model of the factor structure of the MACE.
 
@@
     \midrule
-    Replication & 2-factor & 1737 (664) & 0.052 & 0.961 & 0.957 & 0.108 \\
-                & 3-factor & 1227 (627) & 0.040 & 0.978 & 0.974 & 0.097 \\
+    Replication & 2-factor & 1612 (664) & 0.051 & 0.961 & 0.957 & 0.121 \\
+                & 3-factor & 1100 (627) & 0.037 & 0.981 & 0.977 & 0.111 \\
     \bottomrule

### 2022-08-15 analysis/scripts/fit_efa.R

@@ -11,6 +11,11 @@ set.seed(47404)
 ## Load data.
 data = read.csv('../data/mace.csv')
 
+## Apply rejections. 
+reject = read.csv('../data/covariates.csv')
+data = merge(data, reject[,c('subject','infreq')], by='subject')
+data = data[data$infreq < 1.0,]
+
 ## Define studies.
 studies = list('teicher2015', 'tuominen2022')
 

### 2022-09-14 manuscript/main.tex
next draft version

@@
 
-\title{The MACE is essentially unidimensional: Bifactor models of the `Maltreatment and Abuse Chronology of Exposure' scale}
-\shorttitle{Bifactor models of MACE}
+\title{Confirmatory factor analysis of the Maltreatment and Abuse Chronology of Exposure (MACE) scale: Evidence for essential unidimensionality}
+\shorttitle{Confirmatory factor analysis of the MACE}
 \author{Samuel Zorowitz$^1$, Lauri Tuominen$^{2}$}
@@
 
-\noindent \textbf{Background}: Typically measures of childhood maltreatment collect information about several types of abuse and maltreatment.
-The subscales of the retrospective Maltreatment and Abuse Chronology of Exposure (MACE) scale measure 10 such types.
-Since different types of maltreatment often co-occur, it is not clear whether subscale measures can be used to reflect constructs unique from from overall maltreatment.
+\noindent \textbf{Background}: The Maltreatment and Abuse Chronology of Exposure (MACE) scale is a retrospective self-report scale that measures 10 distinct types of childhood maltreatment.
+Despite its increasing use, the factor structure of the MACE has not been thoroughly investigated.
+As such, the viability of the MACE total and subscale scores are uncertain.
+The aim of the present study was to investigate the factor structure of the MACE and to quantify the reliability of its total and subscale scores.
 
-\hfill
+{\begin{spacing}{1.25} \hfill \\ \end{spacing}}
 
-\noindent \textbf{Methods}: Here, we investigated the factor structure of the MACE scale in two independent samples (N=1051 \& N=582). We used both confirmatory and exploratory item response modeling.
+\noindent \textbf{Methods}: We evaluated a series of confirmatory item factor models fit to the responses of two independent samples of participants (N=1051 \& N=582) who completed the MACE.
+The models included hierarchical (bifactor) models.
+We used model-based indices to estimate the reliability of the MACE total and subscale scores, and to quantify the essential unidimensionality of the scale.
 
-\hfill
+{\begin{spacing}{1.25} \hfill \\ \end{spacing}}
 
-\noindent \textbf{Results}: We found the MACE total score to be a reliable and valid measure of general maltreatment.
-Confirmatory item response models provided strong evidence for unidimensionality of the MACE, with the exception of a potential peer victimization factor.
-That is, most items in the MACE load on general factor and apart from peer victimization very little variance was explained by other subscales.
-Exploratory factor modeling further supported these findings.
+\noindent \textbf{Results}: We found the MACE total score was a reliable and valid measure of overall childhood maltreatment.
+In contrast, although MACE subscale scores exhibited adequate reliability, we found that the vast majority of reliable variance in these scores reflected general maltreatment and not any particular type of maltreatment as intended.
+We also found that the MACE is essentially unidimensional, and exhibits evidence of differential item functioning by gender.
 
-\hfill
+{\begin{spacing}{1.25} \hfill \\ \end{spacing}}
 
-\noindent \textbf{Conclusions}: The MACE is essentially a unidimensional scale.
-Apart from a separate peer victimization factor, researchers should refrain from using subscales data separately since they mostly reflect overall maltreatment.
+\noindent \textbf{Conclusions}: Our results provide support for a one-factor structure of the MACE, as well as continued use of the MACE total score.
+Our results caution against the use of MACE subscale scores, which have little practical use given that they provide little unique or reliable information above and beyond the total score.
+
 \hfill \break
+
 }
@@
 
-\authornote{This work was funded in part by an NSF Graduate Research Fellowship. The authors have no conflicts of interest to declare.}
+\authornote{This research was funded by the Royal's Emerging Research Innovators in Mental Health program at the University of Ottawa Institute of Mental Health Research and a National Science Foundation Graduate Research Fellowship.
+The authors wish to thank Dr.
+Robyn McQuaid for thoughtful feedback on this manuscript.
+The authors have no conflicts of interest to declare.}
 
@@
 
-Childhood maltreatment is an unfortunately common experience for children across the globe \citep{stoltenborgh2015prevalence}.
-It is a risk factor for multiple poor outcomes later in life including the development of mental illness \citep{kessler2010childhood}, challenges in cognitive functioning \citep{su2019does}, and poor physical health \citep{wegman2009meta}.
+Childhood maltreatment is an unfortunately common experience for children around the world \citep{stoltenborgh2015prevalence}.
+It is a risk factor for multiple poor outcomes later in life including the development of mental illness \citep{kessler2010childhood}, cognitive impairment \citep{su2019does}, and poor physical health \citep{wegman2009meta}.
 Childhood maltreatment also predicts worse treatment outcomes for several psychiatric disorders \citep{nanni2012childhood, thomas2019childhood}.
-Thus, reliable and valid assessment of childhood maltreatment is an important research goal in order to refine our understanding of the link between maltreatment and health outcomes; augment treatment strategies for victims of childhood maltreatment; and ultimately design better interventions to prevent maltreatment in the first place.
+Thus, the reliable and valid assessment of childhood maltreatment is an important research goal in order to refine our understanding of the link between maltreatment and health outcomes; augment treatment strategies for victims of childhood maltreatment; and ultimately design better interventions to prevent maltreatment in the first place.
 
@@
 One recently developed measure, the Maltreatment and Abuse Chronology of Exposure scale (MACE; \citealt{teicher2015maltreatment}), has a number of notable advantages.
-The MACE is a 52-item retrospective self-report scale composed of 10 subscales, each of which measures a unique type of childhood maltreatment.
-Six of these subscales overlap with other standard measures of abuse (e.g.
+First, the MACE is a 52-item retrospective self-report scale composed of 10 subscales, each of which measures a unique type of childhood maltreatment.
+The content of six of these subscales overlap with other childhood maltreatment scales (e.g.
 physical abuse, neglect), whereas the remaining four are less commonly represented in other scales (e.g.
 peer victimization, witnessing domestic violence).
-The items in the MACE were carefully selected to measure maltreatment of varying severity in order to finely differentiate people with differing exposure levels.
-The MACE also asks for the onset and duration of maltreatment experiences, which is necessary for investigating how the timing of abuse shapes development.
-The MACE was originally developed to provide two measures: a total score (number of overall maltreatment experiences) and a multiplicity score (number of types of maltreatment experienced).
+Second, the items in the MACE were carefully selected to measure maltreatment of increasing severity in order to finely differentiate people with differing exposure levels.
+Third, the MACE also asks for the onset and duration of maltreatment experiences, which is necessary for investigating how the timing of abuse shapes development.
+Finally, the MACE was originally developed to provide two measures: a total score (number of overall maltreatment experiences) and a multiplicity score (number of types of maltreatment experienced).
 These scores have excellent temporal stability, convergent validity with other childhood maltreatment scales, and good predictive validity for a multitude of psychiatric symptoms \citep{teicher2015maltreatment}.
@@
 In response, some researchers have begun to use the MACE to make new indices of maltreatment in lieu of, or in addition to, the total and multiplicity scores.
-For example, some have used all 10 MACE subscale scores to study the unique contributions of each type of maltreatment to the development of psychopathology \citep{schalinski2015type, gerke2018childhood, schalinski2019early}.
-Inspired by a prominent dimensional model of maltreatment \citep{mclaughlin2014childhood}, others have made new subscale scores indexing threat and deprivation experiences using the items in the MACE that respectively measure abuse and neglect \citep{schalinski2018defining, schalinski2019environmental, teicher2018differential}.
-Yet the psychometric properties of these new subscale scores have not been thoroughly investigated.
+For example, some researchers have used all 10 MACE subscale scores to study the unique contributions of each type of maltreatment to the development of psychopathology \citep{schalinski2015type, gerke2018childhood, schalinski2019early}.
+Citing a prominent dimensional model of maltreatment \citep{mclaughlin2014childhood}, others have made novel threat and deprivation subscale scores made up of items in the MACE that measure abuse and neglect experiences, respectively \citep{schalinski2018defining, schalinski2019environmental, teicher2018differential}.
+The psychometric properties of these subscale scores have not been thoroughly investigated.
 
 A critical challenge in studying the consequences of childhood maltreatment is the high co-occurrence of these adverse experiences.
-That is, children who have experienced any single form of maltreatment are likely to have experienced multiple other types \citep{dong2004interrelatedness, herrenkohl2009assessing, kessler2010childhood}.
-As a consequence, isolating the unique contribution of one type of maltreatment to some outcome is difficult insofar as any one type of adversity is likely confounded with overall maltreatment.
-Concretely, this means that researchers intending to study the effects of particular types of maltreatment using subscale scores need to verify that those measures actually reflect a construct unique from from overall maltreatment, lest they risk drawing spurious conclusions.
+That is, children who have experienced any single type of maltreatment are likely to have experienced multiple other types \citep{dong2004interrelatedness, herrenkohl2009assessing, kessler2010childhood}.
+As such, isolating the unique contribution of one type of maltreatment to an outcome is difficult insofar as any one type of adversity is likely confounded with overall maltreatment.
+Concretely, this means that researchers intending to study the sequela of particular types of maltreatment using subscale scores need to verify that those scores actually reflect a construct unique from overall maltreatment, lest they risk drawing spurious conclusions.
 
@@
 Crucially, bifactor models can be used to calculate model-based reliability statistics, including the omega hierarchical ($\omega_h$) and omega subscale ($\omega_s$) indices \citep{reise2013scoring}, which respectively measure the proportion of variance in total and subscale scores attributable to the general factor and specific factors.
-Applied to the MACE, these indices can help determine if subscale scores measuring particular types of maltreatment are interpretable as such, or if they instead primarily reflect overall maltreatment.
+Applied to the MACE, these indices can help determine if subscale scores measuring particular types of maltreatment are interpretable as such, or if they instead principally reflect overall maltreatment.
+In addition, bifactor models permit the calculation of other useful indices like the explained common variance (ECV; \citealt{sijtsma2009use}), which measures what fraction of the total reliable variance in a scale can be attributed to a general factor.
+Applied to the MACE, the ECV can quantify the degree to which the MACE is essentially unidimensional (i.e.
+summarizable by a one-factor model).
 
-In this study, we investigated the factor structure of the MACE in two independent samples of participants.
-First, we tested the fits of a series of confirmatory item response models to the maltreatment data from each sample.
-These included several bifactor models, which allowed us to quantify the proportion of variance in the data attributable to general maltreatment versus specific types of abuse.
-We found that the MACE is essentially unidimensional; that is, the majority of reliable variance in participants' responses was attributable to a general maltreatment factor.
-We then calculated the reliability of the MACE total score and a host of subscale scores.
-We found the total score was a reliable and valid measure of general maltreatment, whereas most subscale scores contained little-to-no variance unique from general maltreatment.
-The only possible exception to this trend was for a novel peer victimization subscale.
-Finally, we performed exploratory factor analysis to examine the factor structure of the MACE without \textit{a priori} restrictions.
-These analyses supported the results of the confirmatory factor analyses.
+The overarching aim of the present study was to investigate the factor structure of the MACE, which has received little previous study to date \citep{saini2019systematic}.
+To do so, we fit and evaluated a series of confirmatory item factor models to the responses from two independent samples of participants who completed the MACE.
+The primary goal of our analyses was to quantify the reliability of the MACE total and subscale scores, and to determine whether the latter would provide unique information above and beyond the total score.
+A secondary goal was to examine the dimensionality of the MACE; that is, to determine whether responses on the MACE are best-explained by a hierarchical structure with 10 factors (one per subscale) or by a more parsimonious factor structure.
+All data and analysis materials are publicly available at: \url{https://github.com/szorowi1/mace-irt}.
 
@@
 
-All data, code, and analysis materials are publicly available at \url{https://github.com/szorowi1/mace-irt}.
-
-{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Participants}
+\subsection{Participants}
 
@@
     \hspace{1em} Men                       & 381 (36.3\%) & 258 (44.3\%) & \\
-    \hspace{1em} Trans or NB               &   0  (0.0\%) &   9  (1.5\%) & \\
+    \hspace{1em} Transgender or NB         &   0  (0.0\%) &   9  (1.5\%) & \\
     \hspace{1em} Rather not say            &   0  (0.0\%) &   5  (0.9\%) & \\
@@
 
-{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Measures}
+\subsection{Measures}
 
@@
 For these items we follow the scoring convention set by \cite{teicher2015maltreatment}: a score of 1 is assigned to a response only in the absence of a positive event for all 18 years of childhood.
-This is in contrast to the remaining items in the scale, where a score of 1 is assigned to a response if maltreatment or abuse occurred in at least one year of childhood.
+This is in contrast to all other items, where a score of 1 is assigned if maltreatment was experienced in at least one year of childhood.
 
@@
 
-{\begin{spacing}{0.2} \hfill \\ \end{spacing}}  \subsection{Exclusion criteria}
+\subsection{Exclusion criteria}
 
@@
 
-{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Differential item functioning}
+\subsection{Differential item functioning}
 
 Prior to analysis, we investigated whether it was appropriate to combine response data from the original and replication samples.
-We first compared maltreatment endorsement rates for each item between samples.
+We first compared across samples the maltreatment endorsement rates for each item.
 The Spearman-rank correlation of endorsement rates across samples was large ($\rho$ = 0.951, p < 0.001), indicating excellent correspondence in the relative prevalence of maltreatment types across samples.
@@
 
-An overall shift in overall endorsement rates by sample is not necessarily evidence for a violation of measurement invariance.
+A mean shift in overall endorsement rates by sample is not necessarily evidence for a violation of measurement invariance.
 One possibility is that the replication sample had simply experienced more adversity than the original sample.
@@
 gender, age) explain the differences in endorsement rates.
-To investigate these possibilities, we tested for uniform differential item functioning (DIF) using logistic regression, in which a participant's endorsement response on an item is regressed against their rest score (i.e.
+To investigate these possibilities, we tested for uniform differential item functioning (DIF) using logistic regression, in which a participant's response on an item is regressed against their rest score (i.e.
 their total score for all other items), sample membership, gender, and age.
@@
 We also found an additional 11 items with large DIF by gender (Figure \ref{fig:dif_gender}).
-Five of these items measured sexual abuse (endorsed more frequently by women), and another five measured peer physical abuse (endorsed more frequently by men).
+Five of these items concerned sexual abuse (endorsed more frequently by women), and another five concerned peer physical abuse (endorsed more frequently by men).
 We identified no items that exhibited large DIF by age.
@@
 
-{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Response dependence}
+\subsection{Item response dependence}
 
-The MACE contains a total of 12 item pairs and triplets characterized by response dependence (Table \ref{tab:dependence}).
+Also prior to analysis, we found that the MACE contains a total of 12 item pairs and triplets characterized by response dependence (Table \ref{tab:dependence}).
 The items in some of these sets are structurally dependent, where a response to one item requires or strongly implies a certain response on the second.
@@
 
-We were concerned about response dependence for two reasons.
-First, it can inflate the estimates of item discrimination parameters, making items appear more informative than they actually are \citep{marais2008formalizing}.
-Second, response dependence can inflate the estimates of item loadings on specific factors in bifactor models, giving false support for models with a greater number of factors than needed \citep{reise2013applying}.
-Following \cite{marais2008formalizing}, we therefore converted each set of dependent items into a single polytomous item thereby reducing the MACE from 52 to 39 items in total.
+Structurally-dependent or redundant items can cause serious problems for item factor models.
+First, they can inflate the estimates of item discrimination parameters, making items appear more reliable than they actually are \citep{marais2008formalizing}.
+Second, response dependence can cause overestimation of item loadings on specific factors in bifactor models, making specific factors (and subscores) appear more reliable than they actually are \citep{reise2013applying}.
+Following \cite{marais2008formalizing}, we therefore combined each set of dependent items into a single polytomous item thereby reducing the MACE from 52 to 39 items in total.
 
-{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Confirmatory item factor models}
+\subsection{Confirmatory item factor models}
 
-To investigate the factor structure of the MACE, we fit a series of confirmatory item factor models\footnote{The appropriateness of common factor models with reflective indicators for maltreatment data has been contested.
-We justify our usage of this model of the MACE for the interested reader in the supplement.} separately to the response data from each sample.
-The basis of each model is the graded response model for ordered polytomous data \citep{samejima1997graded}.
-The factor structure for each of the four models is depicted in Figure \ref{fig:models}.
-The first model is a one-factor (undimensional) model where all items load onto a general maltreatment factor (Figure \ref{fig:models}a).
+To investigate the factor structure of the MACE, we fit a series of four confirmatory item factor models separately to the response data from each sample.\footnote{The suitability of common factor models (CFMs) with reflective indicators for explaining childhood maltreatment data has been contested.
+For interested readers, we justify our use of CFMs to explain responses to the MACE in the supplement.} The basis of each model is the graded response model for ordered polytomous data \citep{samejima1997graded}.
+The factor structure for each model is depicted in Figure \ref{fig:models}.
+The first is a one-factor (undimensional) model where all items load onto a general maltreatment factor (Figure \ref{fig:models}a).
 The second is a bifactor model where all items load onto a general maltreatment factor and also one of 10 specific factors based on their subscale membership (Figure \ref{fig:models}b).
@@
 
-The third model was motivated by a theoretical account of childhood maltreatment that distinguishes between threat and deprivation experiences as distinct types of adversity \citep{mclaughlin2014childhood}.
-To explore this model of the MACE, we initially fit a bifactor model where all items loaded onto a general maltreatment factor, and also a threat or neglect specific factor.
+The third model was motivated by a dimensional theory of childhood maltreatment that distinguishes between threat and deprivation experiences as two distinct types of adversity \citep{mclaughlin2014childhood}.
+To examine this model of the MACE, we initially fit a bifactor model where all items loaded onto a general maltreatment factor, and also a threat or neglect specific factor.
 However, we observed vanishing item loadings on the threat specific factor.
-As such, we instead fit a bifactor S-1 model \citep{eid2017anomalous} where all items load onto a general maltreatment factor, and only items measuring neglect load onto a specific factor (Figure \ref{fig:models}c).
-This model will be used to evaluate if the MACE can produce reliable neglect subscale scores distinguishable from general maltreatment.
+Therefore, we instead fit a bifactor S-1 model \citep{eid2017anomalous} where all items load onto a general maltreatment factor, and only items measuring neglect (i.e.
+items from the physical and emotional neglect subscales) load onto a specific neglect factor (Figure \ref{fig:models}c).
+We fit this model to see if the MACE can be used to form neglect subscale scores that are reliable and distinguishable from general maltreatment.
 
@@
 (C) Bifactor S-1 model with 1 specific factor (neglect).
-(D) Bifactor S-1 model with 2 specific factors (peer victimization, reverse-scoring).
+(D) Bifactor S-1 model with 2 specific factors (peer victimization, reverse-scored).
 Notes: G = general maltreatment; VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; WSV = witnessing sibling vioence; WIPV = witnessing inter-parental violence; EN = emotional neglect; PN = physical neglect; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.}
@@
 
-The fourth and final model proposes a more parsimonious structure of the MACE based on the content of its items.
-Items on the MACE can be organized according to the source of maltreatment: parents (or other adults in the home) and peers.
+The fourth and final model we fit proposes a more parsimonious structure of the MACE based on the content of its items.
+Items on the MACE can be categorized according to the source of maltreatment: parents (or other adults in the home) and peers.
 Thus, we hypothesized there would be two factors underlying responses on the MACE: a primary parental maltreatment factor and a secondary peer victimization factor.
-Based on preliminary model fitting, we also suspected the presence of a tertiary methods factor affecting the six reverse-scored items.
-We initially fit a bifactor model in which all items loaded onto a general maltreatment factor and one of three specific factors (parental maltreatment, peer victimization, reverse-scoring); however, we observed vanishing loadings on the first specific factor.
-As such, we instead fit a bifactor S-1 model where all items load onto a general maltreatment factor, and the peer victimization \& reverse-scoring items additionally load onto their own specific factors (Figure \ref{fig:models}d).
-This model will be used to evaluate if the MACE can produce peer victimization reliable subscale scores distinguishable from general maltreatment.
+From preliminary model fits, we also suspected the presence of a tertiary methods factor affecting the six reverse-scored items.
+We initially fit a bifactor model in which all items loaded onto a general maltreatment factor and one of three specific factors (parental maltreatment, peer victimization, reverse-scored); however, we observed vanishing loadings on the first specific factor.
+We instead fit a bifactor S-1 model where all items load onto a general maltreatment factor, and the peer victimization \& reverse-scored items additionally load onto their own specific factors (Figure \ref{fig:models}d).
+We fit this model in part to test a more parsimonious factor structure of the MACE, and to see if the MACE can be used to form peer victimization subscale scores that are reliable and distinguishable from general maltreatment.
 
@@
 
-{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Goodness of fit \& model comparison}
+\subsection{Goodness of fit \& model comparison}
 
@@
 To improve the stability of the $M_2$ statistic, the responses to polytomous items were binarized (i.e.
-none vs.
-any maltreatment).
+any vs.
+no maltreatment).
 Using the $M_2$ statistic, we calculated the root mean square error of approximation (RMSEA), comparative fit index (CFI), and Tucker-Lewis index (TLI) for each model fit.
@@
 
-Given the limitations of SEM-based fit indices for item response models (e.g.
-\citealt{reise2014evaluating}), we calculated two additional fit indices.
+Given the limitations of SEM-based fit indices for item response models \citep{clark2018model}, we calculated two additional fit indices.
 First, we used posterior predictive model checking to calculate a $\chi^2_{NC}$ discrepancy measure based on the total score distribution \citep{sinharay2006posterior}.
@@
 
-{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Model-based reliability indices}
+\subsection{Model-based reliability indices}
 
@@
 
-{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Exploratory factor analysis}
+\subsection{Exploratory factor analysis}
 
-In addition to confirmatory factor analysis, we also performed exploratory factor analysis (EFA) on the response data from each sample separately.
-The goal of the EFA was to examine what factor structure would emerge with no \emph{a priori} restrictions placed on the data.
-We decided to extract two- and three- factor solutions based on the total number of factors in the bifactor S-1 models.
-We were particularly interested to see if a two-factor solution would reproduce the factors from model 3 (threat, deprivation); and if a three- factor solution would reproduce the three factors from model 4 (parental maltreatment, peer victimization, reverse-scoring).
-
-EFA was performed using the \textit{lavaan} (v0.6.11; \citealt{lavaan}) package available in R.
-Factors were extracted using the oblique geomin \citep{yates1987multivariate} and cf-quartimax rotation criteria \citep{crawford1970general}, which were selected in order to extract simple factor structures with few cross-loadings.
-We used two rotation criteria in order to examine the generalizability of the factor solutions across rotations.
-We again relied on traditional fit indices (i.e.
-RMSEA, CFI, TLI) to evaluate the goodness-of-fit of the exploratory factor models to the response data of both samples.
+In addition to confirmatory factor analysis, we also performed exploratory factor analysis (EFA) on the response data for each sample separately.
+The purpose for performing EFA after confirmatory factor analysis was to examine whether any of the hypothesized factor structures emerged from the data with fewer \textit{a priori} restrictions \citep{schmitt2018selecting}.
+For the sake of brevity, the details and results of the EFA are included in the supplement.
 
@@
 The unidimensional model fits exhibited many locally dependent item pairs --- especially among the peer victimization and reverse-scored items --- suggesting possible underfactorization.
-By comparison, the three other models exhibited many fewer locally dependent item pairs.
+By comparison, the remaining three models exhibited many fewer locally-dependent item pairs.
 Of these pairs, the reasons for dependence were apparent.
-For example, the dependence between items 17 (``Parents made inappropriate sexual comments or suggestions to your sibling'') and 18 (``Parents touched or fondled your sibling in a sexual way'') observed in all models is almost certainly the result of measuring highly similar content.
+For example, the residual dependence between items 17 (``Parents made inappropriate sexual comments or suggestions to your sibling'') and 18 (``Parents touched or fondled your sibling in a sexual way'') certainly stems from the fact that these items measure highly similar content.
 Regardless, the estimated item discrimination parameters for each model fit seldom exceeded the normal range ($\alpha \leq 4$).
@@
 These differences in LOO-CV values, however, are both within four standard errors of the mean indicating only weak predictive improvements \citep{vehtari2022cv}.
-Thus the results suggest that models 2 and 4 provide approximately equal fits to the data, which is notable given that model 4 assumes a more parsimonious factor structure of the MACE.
+Thus the results suggest that models 2 and 4 provide approximately equal fits to the data, which is notable given that model 4 assumes a substantially simpler factor structure of the MACE.
 
-{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Confirmatory item response models}
+\subsection{Confirmatory item response models}
 
@@
 
-{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsubsection{Model 1: One-factor (unidimensional) model}
+\subsubsection{Model 1: One-factor (unidimensional) model}
 
@@
     \caption{Standardized factor loadings from the four models fit to the original sample.
-Notes: Model 1 = one-factor (unidimensional) model; Model 2 = bifactor model with 10 specific factors; Model 3 = bifactor S-1 model with one specific factor (neglect); Model 4 = bifactor S-1 with two specific factors (peer victimization, reverse-scoring); VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; WSV = witnessing sibling vioence; WIPV = witnessing inter-parental violence; EN = emotional neglect; PN = physical neglect; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.
+Notes: Model 1 = one-factor (unidimensional) model; Model 2 = bifactor model with 10 specific factors; Model 3 = bifactor S-1 model with one specific factor (neglect); Model 4 = bifactor S-1 with two specific factors (peer victimization, reverse-scored); VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; WSV = witnessing sibling vioence; WIPV = witnessing inter-parental violence; EN = emotional neglect; PN = physical neglect; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.
 Factor loadings $<$ 0.3 are displayed in gray; factor loadings $\geq$ 0.6 are bolded.
@@
     \caption{Standardized factor loadings from the four models fit to the replication sample.
-Notes: Model 1 = one-factor (unidimensional) model; Model 2 = bifactor model with 10 specific factors; Model 3 = bifactor S-1 model with one specific factor (neglect); Model 4 = bifactor S-1 with two specific factors (peer victimization, reverse-scoring); VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; WSV = witnessing sibling vioence; WIPV = witnessing inter-parental violence; EN = emotional neglect; PN = physical neglect; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.
+Notes: Model 1 = one-factor (unidimensional) model; Model 2 = bifactor model with 10 specific factors; Model 3 = bifactor S-1 model with one specific factor (neglect); Model 4 = bifactor S-1 with two specific factors (peer victimization, reverse-scored); VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; WSV = witnessing sibling vioence; WIPV = witnessing inter-parental violence; EN = emotional neglect; PN = physical neglect; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.
 Factor loadings $<$ 0.3 are displayed in gray; factor loadings $\geq$ 0.6 are bolded.
@@
 In the original sample, factor loadings ranged between 0.279 and 0.864 with an average value of 0.597 (95\% HDI = 0.579 -- 0.614); in the replication sample, factor loadings ranged between 0.359 and 0.814 with an average value of 0.600 (95\% HDI = 0.578 -- 0.621).
-Most items had loadings greater than 0.50 (original: 80.0\%, replication: 80.0\%), and thereby appear to be good indicators of general maltreatment.
+Most items had loadings greater than 0.50 (original: 80.0\%, replication: 80.0\%), and therefore appear to be good indicators of general maltreatment.
 The rank correlation of general factor loadings between samples was $\rho$ = 0.565 (95\% HDI = 0.414 -- 0.711), indicating only moderate agreement between samples.
 
-To begin characterizing the dimensionality of the MACE, we inspected the first six eigenvalues of the polychoric item correlation matrix for each sample.
+As an initial characterization the dimensionality of the MACE, we inspected the first six eigenvalues of the polychoric item correlation matrix for each sample.
 In the original sample, these values were 14.62, 3.77, 2.64, 2.11, 1.79, and 1.35; in the replication sample, they were 13.27, 4.43, 2.48, 2.04, 1.81, and 1.56.
@@
 Virtually all items exhibited moderate-to-large loadings onto the common factor.
-Moreover, the eigenvalues of the polychoric correlation matrix suggested one primary factor.
+Moreover, the eigenvalues of the polychoric correlation matrix suggested the presence of one primary factor.
 However, the degree of local dependence present among in the unidimensional model fits (especially among the peer victimization items) suggest there may be additional dimensions to the MACE.
@@
 
-{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsubsection{Model 2: Bifactor model with 10 specific factors}
+\subsubsection{Model 2: Bifactor model with 10 specific factors}
 
@@
 In the replication dataset, the range was between 0.043 and 0.698, with an average value of 0.343 (95\% HDI = 0.315 -- 0.369).
-Only items from the peer verbal and physical abuse subscales had specific factor loadings of equal or greater magnitude to those on the general factor.
+Only items from the two peer abuse subscales had specific factor loadings of equal or greater magnitude to those on the general factor.
+
 \begin{table}[t!]
@@
 
-We found that the MACE total score had excellent reliability (original: $\omega$ = 0.963; replication: $\omega$ = 0.965).
+We now consider the reliability of the MACE total score.
+We found the MACE total score had excellent reliability (original: $\omega$ = 0.963; replication: $\omega$ = 0.965).
 The $\omega_h$ value was 0.924 for both samples indicating that 96.0\% and 95.8\% of the reliable variance in the total score is attributable to general maltreatment.
-(This result is not very surprising given the large number of subscales composed of few items.) Importantly, the ECV values were 0.716 and 0.697 for the original and replication samples, respectively.
-In conjunction with a large PUC value of 0.912, these results provide strong evidence for a dominant general maltreatment factor in the MACE.
+Importantly, the ECV values were 0.716 and 0.697 for the original and replication samples, respectively.
+In conjunction with a PUC value of 0.912, the ECV values suggest the MACE is essentially unidimensional despite the multidimensionality of its contents.
+Together, the results support the interpretation of the MACE total score is a reliable and valid measure of overall maltreatment.
 
-The reliability of the MACE subscale scores were more variable in comparison.
-The $\omega$ values ranged between 0.590 and 0.909 in the original sample, and between 0.595 and 0.893 in the replication sample, with smaller $\omega$ values for subscales with fewer items.
-Crucially, the $\omega_s$ values were small for both the original sample (mean $\omega_s$ = 0.181, range = 0.027 -- 0.621) and replication sample (mean $\omega_s$ = 0.205, range = 0.037 -- 0.565).
-That is, the MACE subscale scores contain almost no reliable variance after controlling for the general maltreatment factor.
-They should not therefore be interpreted as reflecting specific forms of maltreatment.
+Next we consider the reliability of the MACE subscale scores, where the results are in direct contrast to the MACE total score.
+As would be expected given the lower number of items, the reliability of the subscales was both lower and more variable.
+The $\omega$ values ranged between 0.590 and 0.909 in the original sample, and between 0.595 and 0.893 in the replication sample.
+Most subscale scores exhibited at least adequate reliability (>0.7).
+Crucially, subscale scores for both samples contained little reliable variance after controlling for the general maltreatment factor (original: mean $\omega_s$ = 0.181, range = 0.027 -- 0.621; replication: mean $\omega_s$ = 0.205, range = 0.037 -- 0.565).
+That is, the reliability of MACE subscale scores primarily reflect general maltreatment and not any specific type of maltreatment.
 The only exception was for the peer verbal abuse subscale, where more than half the reliable variance in subscale scores was independent of general maltreatment (original = 70.7\%; replication = 66.2\%).
 
-The results of the full bifactor model were clear: responses on the MACE are explained by a dominant general maltreatment factor characterized by large loadings from the majority of items.
-In contrast, specific factors corresponding to each MACE subscale are weak and characterized mostly by small factor loadings.
-Whereas the MACE total score is a reliable and valid measure of overall maltreatment, the MACE subscale scores are less reliable and cannot be meaningfully interpreted as uniquely reflecting exposure to particular types of abuse.
-The question remains, however, if there are other subscale scores to be derived from the MACE.
+To summarize, the results of the full bifactor model (model 2) indicate responses on the MACE are explained by a dominant general maltreatment factor characterized by large loadings across items.
+In contrast, specific factors representing each MACE subscale are characterized by mostly small factor loadings that contribute little reliable variance to total and subscale scores.
+Where the MACE total score is a reliable and valid measure of overall maltreatment, the majority of MACE subscale scores contain little reliable variance unique from general maltreatment.
+Therefore, they should not be interpreted as reflecting specific types of maltreatment.
+One question that remains, however, is if other reliable subscale scores can be derived from the MACE.
 
-{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsubsection{Model 3: Bifactor S-1 model with one specific factor (neglect)}
+\subsubsection{Model 3: Bifactor S-1 model with one specific factor (neglect)}
 
@@
 The specific loadings displayed a troubling pattern where only the six reverse-scored items exhibited moderate loadings on the specific factor; the remaining four item exhibited negligible factor loadings.
-That is, the ``neglect'' factor appears instead to reflect a reverse-scoring methods factor.
+That is, the ``neglect'' factor appears instead to reflect a reverse-scored methods factor.
 
@@
 The reliability of the ``neglect'' subscale score was also good (original: $\omega$ = 0.877; replication: $\omega$ = 0.921).
-The $\omega_s$ values, however, indicated the neglect subscale score contained less than half reliable variance after controlling for the general factor (original: $\omega_h$ = 0.384; replication: $\omega_h$ = 0.375).
+As indicated by the $\omega_s$ values, however, reliable variance in the ``neglect'' subscale scores primarily reflected the general maltreatment factor (original: $\omega_h$ = 0.384; replication: $\omega_h$ = 0.375).
 
-In summary, the results do not support the use of the MACE to make separate threat and deprivation subscale scores.
-The neglect subscale scores posses only a minority of reliable variance unique from general maltreatment thereby precluding meaningful interpretation of these scores.
-Furthermore, the majority of items measuring neglect on the MACE seem to be contaminated by a reverse-scoring methods factor.
+In sum, the results do not support the use of the MACE to compute separate threat and neglect subscale scores.
+The neglect subscale scores posses an unsatisfactory amount of reliable variance unique from general maltreatment; therefore, they cannot be interpreted as meaningfully reflecting neglect.
+Furthermore, the majority of items measuring neglect on the MACE seem to be contaminated by a reverse-scored methods factor.
 
-{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsubsection{Model 4: Bifactor S-1 with two specific factors (peer victimization, reverse-scoring)}
+\subsubsection{Model 4: Bifactor S-1 with two specific factors (peer victimization, reverse- scoring)}
 
-Finally, we inspect the bifactor S-1 model with two specific factors measuring peer victimization and reverse-scoring (model 4).
+Finally, we turn to the bifactor S-1 model with two specific factors (peer victimization, reverse-scoring; model 4).
 In the original sample, the loadings on the general maltreatment factor ranged between 0.158 and 0.864 with an average value of 0.563 (95\% HDI = 0.545 -- 0.580).
@@
 
-As with the other bifactor models, the total score under this model exhibited excellent reliability (original: $\omega$ = 0.961; replication: $\omega$ = 0.961).
-Again the majority of reliable variance in severity score was attributable to the general factor (original: $\omega_h$ = 0.896; replication: $\omega_h$ = 0.904).
-The peer victimization subscale score exhibited good reliability across samples (original: $\omega$ = 0.889; replication: $\omega$ = 0.868).
-The corresponding $\omega_s$ values were 0.569 for the original sample and 0.493 for the replication sample.
-That is, slightly more than half of the variance in peer victimization scores (original: 64.0\%; replication: 56.8\%) is independent of the general maltreatment factor.
-These results suggest the presence a secondary peer victimization factor in the MACE (albeit weakly so) in addition to a primary general maltreatment factor.
+As before, the MACE total score again exhibited excellent reliability (original: $\omega$ = 0.961; replication: $\omega$ = 0.961).
+Similarly, the majority of reliable variance in severity score was attributable to the general factor (original: $\omega_h$ = 0.896; replication: $\omega_h$ = 0.904).
+The peer victimization subscale scores exhibited good reliability across samples (original: $\omega$ = 0.889; replication: $\omega$ = 0.868).
+The corresponding $\omega_s$ values were 0.569 for the original sample and 0.493 for the replication sample indicating that slightly more than half of the variance in peer victimization scores were independent of the general maltreatment factor (original: 64.0\%; replication: 56.8\%).
+Thus, the results suggest (albeit weakly so) the presence a secondary peer victimization factor in the MACE in addition to the general maltreatment factor.
 
@@
 Moreover, the corresponding $\omega_s$ values were similar in magnitude to those observed for the peer victimization scores (original: $\omega_s$ = 0.584; replication: $\omega_s$ = 0.498).
-In other words, these scores are also have majority reliable variance independent of general maltreatment.
-It is difficult to interpret these scores, however, as it is unclear what they primarily reflect: neglect experiences, wording effects, and/or scoring-induced methods artifact.
-
-{\begin{spacing}{0.2} \hfill \\ \end{spacing}} \subsection{Exploratory factor analysis}
-
-The goodness-of-fit measures for the exploratory factor models to the response data by sample are summarized in Table \ref{tab:efa_diagnostics}.
-The fit indices indicated that all exploratory models provided an adequate fit to the data.
-The standardized factor loadings for the 2- and 3-factor solutions produced by the cf-quartimax rotation are presented in Figure \ref{fig:efa_cf}.
-(The corresponding factor loadings produced by the geomin rotation were qualitatively similar and are presented in Figure \ref{fig:efa_geomin}.) In the 2-factor solution for the original sample, we observed a primary parental maltreatment factor and a secondary peer victimization factor.
-In contrast, the replication sample produced a primary parental maltreatment/peer maltreatment factor with a secondary reverse-scoring factor.
-Neither 2-factor solution precisely resembled the threat-deprivation factor structure of confirmatory model 3.
-
-\begin{figure}[t!]
-    \centering
-    \includegraphics[width=1\textwidth,center]{figures/fig04.png}
-    \caption{\normalfont Standardized factor loadings for the 2- and 3-factor solutions produced by the cf-quartimax rotation for original sample (top row) and replication sample (bottom row).
-\break \small Notes: VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; EN = emotional neglect; PN = physical neglect; WSV = witnessing sibling violence; WIPV = witnessing inter-parental violence; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.
-Only factor loadings $\geq$ 0.30 are shown.
-** Reverse-scored items.}
-    \label{fig:efa_cf}
-\end{figure}
-
-The 3-factor solution for each sample were similar.
-EFA produced a primary parental maltreatment factor with secondary peer victimization and reverse-scoring factors.
-Interestingly, items concerning sexual abuse loaded more strongly on the peer victimization scale in the replication sample, raising some questions about the replicability of the factor structures.
-To conclude, the structure of both the 2- and 3-factor exploratory factor models resembled the bifactor S-1 model with two specific factors (model 4), providing further support for that model of the factor structure of the MACE.
+In other words, these scores also exhibit majority reliable variance unique from general maltreatment.
+It is difficult to interpret these scores, however, as it is unclear what they primarily reflect (e.g.
+neglect, wording effects, and/or scoring-induced methods artifact).
 
@@
 The MACE is a promising measure of childhood maltreatment \citep{georgieva2022systematic}.
-Despite its increasing popularity, there has been little investigation into the MACE's structural validity \citep{saini2019systematic} leaving unanswered questions about its factor structure and the validity of its total/subscale scores.
-Here we investigated the factor structure of the MACE, using both confirmatory and exploratory item response modeling, in two independent samples.
-Bifactor modeling revealed a dominant general maltreatment factor, characterized by uniformly strong loadings across items, and much weaker specific factors.
-The general factor explained approximately 70\% of the reliable variance in the response data across samples.
-Moreover, the general factor loadings from the bifactor models largely resembled the corresponding loadings from a unidimensonal model.
-Together, these results support the conclusion that the MACE is essentially unidimensional.
+Despite its increasing popularity, there has been little investigation of its structural validity \citep{saini2019systematic}, leaving unanswered questions about its factor structure and the interpretation of its total \& subscale scores.
+Here we investigated the factor structure of the MACE, using a series confirmatory item response models, in two independent samples.
+The primary goal of our analyses was to evaluate the reliability of the MACE total and subscale scores, and to determine whether the latter would provide unique information above and beyond the total score.
+A secondary goal was to examine the dimensionality of the MACE; that is, whether its inter-item covariance was best-explained by a hierarchical structure with 10 factors (one per subscale) or by a more parsimonious structure.
 
-We also calculated reliability indices for the MACE total and subscale scores.
-We found that the MACE total score has excellent reliability and is a valid measure of overall maltreatment.
-In contrast, we found that virtually no MACE subscale scores were reliable measures of particular types of maltreatment.
-For all but one of the 10 MACE subscales, the majority of reliable variance in scores was attributable to the general factor.
-The same was true of deprivation subscale scores, which were also contaminated by a reverse-scoring methods factor.
-In short, our results further validate the use of the MACE total score, but caution against the use of MACE subscale scores.
-Indeed, MACE subscale scores seldom reflected their intended construct but instead reflected general maltreatment.
+Regarding our first goal, we found the MACE total score has excellent reliability and overwhelmingly reflects general maltreatment.
+That is, despite being a composite of items measuring 10 distinct types of maltreatment, the MACE total score is a univocal index of overall childhood maltreatment.
+In stark contrast, although we found the MACE subscale scores exhibited adequate reliability on average, we also found that these scores mostly reflected general maltreatment and not any particular type of maltreatment as intended.
+The same was true of neglect subscale scores (composed of the physical and emotional neglect subscales), which were additionally contaminated by a reverse-scored methods factor.
+In short, our results support the continued use of the MACE total score but caution against the use of MACE subscale scores.
+Indeed the latter have little practical use given the low reliability of the portion of subscale scores independent of the total score.
 
-Of the four confirmatory factor models we tested, the most parsimonious and best- fitting was a bifactor S-1 model with a prominent general maltreatment factor and two specific factors measuring peer victimization and reverse-scoring.
-This model of the MACE factor structure was additionally supported by the exploratory factor analyses, which produced a qualitatively similar partitioning of the items.
-Thus, it appears the MACE is essentially unidimensional with secondary factors affecting a small subset of items.
-Interestingly, we found the peer victimization subscale scores to be reliable; a majority of their variance was attributable to the peer victimization factor even after controlling for general maltreatment.
-It may therefore be appropriate to produce a peer victimization subscale score from the MACE, but additional research is needed to validate this measure.
+Regarding our second goal, confirmatory bifactor models of the MACE revealed a dominant general maltreatment factor, characterized by uniformly strong loadings across items, and much weaker specific factors.
+Importantly, general factor loadings from the bifactor models strongly resembled their corresponding loadings from a one-factor model.
+Moreover, the general factor explained approximately 70\% of the reliable variance in participants' responses.
+Together, these results support the conclusion that, while the item content of the MACE is multidimensional, its factor structure is essentially unidimensional.
+In other words, researchers can safely interpret the MACE total score as an index of overall maltreatment; they can also expect little bias in factor loading estimates if they were to specify a one-factor model to explain responses on the MACE.
+
+Of the four confirmatory factor models we tested, the most parsimonious and best- fitting was a bifactor S-1 model with a general maltreatment factor and two specific factors measuring peer victimization and reverse-scoring.
+This model provided a fit to the response data as good as the more complex full bifactor model with 10 specific factors.
+(This bifactor S-1 model was also partially supported by exploratory factor analyses; see the supplement for details.) Thus, our results suggest the factor structure of the MACE is essentially unidimensional with two secondary factors affecting a subset of items.
+Notably, we found that the scores from the peer victimization subscale (composed of items from the peer physical and emotional abuse subscales) were viable.
+That is, they were reliable and, after controlling for general maltreatment, a majority of their reliable variance was attributable to peer victimization.
+Researchers may therefore be justified in using peer victimization subscale scores from the MACE, but additional research is needed to validate this index.
 
@@
 Similar results were found in recent applications of a bifactor model to the International Child Abuse Screening Tool \citep{meinck2021factor} and the Adverse Childhood Experience questionnaire \citep{dobson2021latent}.
-Together, these and our current results underscore a challenge in investigating the consequences of childhood maltreatment.
-When maltreatment experiences tend to co-occur, it is difficult to measure the effects of a particular type of adversity unique from overall maltreatment.
+Together, these and our current results underscore a challenge in investigating the consequences of childhood maltreatment: when maltreatment experiences tend to co-occur, it is difficult to measure the effects of a particular type of adversity unique from overall maltreatment.
 
@@
 This approach is unsatisfactory for two reasons.
-First, random forest models can produce spurious results when dealing with collinear predictors \citep{gregorutti2017correlation}.
-Second, even if one were to properly control for general maltreatment, our results indicate that the remaining reliable variance in most subscale scores is so low as to cast doubt about the utility of these inferences.
+First, random forest models can produce spurious results when given collinear predictors \citep{gregorutti2017correlation}.
+Second, even if one were to properly control for general maltreatment, our results indicate that the remaining reliable variance in most subscale scores would be so low as to cause doubt about the meaning of any subsequent inferences.
 
@@
 
-We also identified a reverse-scoring methods factor in the MACE.
+We also identified a reverse-scored methods factor in the MACE.
 In our view, the most plausible explanation is that, under the \cite{teicher2015maltreatment} scoring procedure, affirmative responses on the regular and reverse-scored items on the MACE represent very different endorsements.
@@
 In contrast, an endorsement on a reverse-scored item indicates that a person experienced experienced the absence of a positive event for all 18 years of childhood.
-Alternative causes of the reverse-scoring factor such as participant confusion, acquiescence, or carelessness \citep{weijters2013reversed} are also possible.
-Additional research is needed to see if alternative scoring procedures eliminates this factor.
+However, alternative causes of the reverse-scored factor (e.g.
+participant confusion, acquiescence, or carelessness; \citealt{weijters2013reversed}) are possible.
+Additional research is needed to see if alternative scoring procedures would eliminate this factor.
 
@@
 Bifactor models revealed that, despite the presence of smaller secondary factors, the majority of reliable variance in the MACE endorsement responses are attributable to a general maltreatment factor.
-Accordingly, we found that the MACE total score is a reliable and valid of overall maltreatment.
-In contrast, we found that MACE subscale scores are not reliable measures of particular types of childhood maltreatment; instead they mainly reflect general maltreatment.
-Moving forward, childhood maltreatment researchers using the MACE (or any other scale) should ensure that any subscale scores used as part of their analyses are reliable, lest they risk drawing spurious inferences.
+Accordingly, we found that the MACE total score is a reliable and valid measure of overall childhood maltreatment.
+In contrast, we found that MACE subscale scores are unreliable measures of their intended constructs; that is, they primarily reflect general maltreatment and not any specific type of maltreatment.
+Moving forward, childhood maltreatment researchers using the MACE (or any other scale) should be careful to ensure that any total or subscale scores they use as part of their analyses are reliable, lest they risk drawing spurious inferences.
 
@@
 
+\subsection*{Reflective vs. formative indicator models of maltreatment data}
+
+One might question whether it is appropriate to explain childhood maltreatment data using common factor models with reflective indicators (e.g.
+bifactor models).
+The core assumption of these models is that variation in item responses are caused by or reflect variation in a shared latent factor (e.g.
+variation in low mood and anergia reflect latent individual differences in depression).
+This is in contrast to composite variable models with formative indicators where the direction of causality is reversed \citepSM{bollen1991conventional}.
+In composite variable models, indicators cause or form the latent construct (e.g.
+education and income together make up socioeconomic status).
+Some scholars have argued that measures of adverse childhood experiences should often be treated as formative, not reflective, indicators \citepSM{netland2001assessment, layne2010unpacking}.
+That is, exposure to abuse and neglect defines childhood maltreatment, not the other way around.
+
+For the purposes of explaining covariation in participants' responses on the MACE, we believe we are justified in using a common factor model.
+Most of the items on the MACE measure maltreatment by parents or other guardians.
+The homogeneity of these indicators suggests they might reflect a common source --- a hypothesis that is further supported by the uniformly strong (general) factor loadings we observed for this subset of items.
+Furthermore, there is ample theoretical and empirical support for regarding parental maltreatment as a common factor with its own causal antecedents \citepSM{belsky1993etiology}.
+Indeed, many of the same parent-level variables (e.g.
+emotional lability, impulse control, own history of abuse) are risk factors for different types of childhood maltreatment \citepSM{stith2009risk, mulder2018risk, assink2019risk}.
+
+Separately, that the MACE is a retrospective self-report measure is further justification for a common factor model.
+Retrospective measures of abuse have been shown to reflect additional factors beyond exposure to maltreatment including individual differences in current emotional state, personality, and recall ability \citepSM{susser2012still, reuben2016lest, colman2016consistency}.
+Common factor models are better suited to adjust the measurement error in indicators caused in part by these nuisance factors.
+
+This discussion in turn raises a second question: is it reasonable for the items measuring parental maltreatment and peer victimization on the MACE to share a common factor?
+One might assume that these two types of maltreatment reflect independent causes and should therefore not be permitted to share a common factor.
+There is considerable evidence, however, that children who have been maltreated are at greater risk for subsequent peer victimization \citepSM{bolger2001developmental, benedini2016cycle}.
+The reasons for this association are manifold.
+For example, parental maltreatment may impact a child's sociocognitive functioning (e.g.
+attachment styles, emotion regulation) in ways that may predispose them to later peer victimization \citepSM{goemans2021child, McCrory2022-cj}.
+Separately, parental neglect is a risk factor for peer victimization \citepSM{lereya2013parenting}, possibly by increasing the chance that children enter into risky social situations where victimization may occur.
+Insofar that parental maltreatment is an indirect cause of peer victimization (mediated by multiple complex psychosocial pathways), then we believe we are justified in allowing indicators of peer victimization to load onto the common general maltreatment factor.
+This is also highlights the need for future study of MACE peer victimization subscale scores in order to better understand how it is similar to (and distinct from) parental maltreatment, and what their downstream sequalae are.
+
 \subsection*{Additional self-report measures}
@@
 Participants in the original sample completed multiple measures concerning psychiatric symptoms; these have already been reported \citepSM{teicher2015maltreatment}.
-The participants in the replication sample completed three additional self-report measures all assessing negative symptoms and motivation for rewards.
+The participants in the replication sample completed three additional self-report measures assessing negative symptoms and motivation for rewards.
 Specifically, participants completed the Motivation and Pleasure scale \citepSM{llerena2013motivation}, the Self-evaluation of Negative Symptoms scale \citepSM{dollfus2016self}, and the revised Behavioral Inhibition / Behavioral Activation scale \citepSM{pagliaccio2016revising}.
@@
 
-\break
-\subsection*{Reflective vs. formative indicator models of maltreatment data}
+\subsection{Exploratory factor analysis}
 
-One might question whether it is appropriate to model childhood maltreatment data using common factor models with reflective indicators (e.g.
-bifactor models).
-The core assumption of these models is that variation in item responses are caused by or reflect variation in a shared latent factor (e.g.
-variation in low mood and anergia reflect latent variation in depression).
-This is in contrast to composite variable models with formative indicators where the direction of causality is reversed \citepSM{bollen1991conventional}.
-In composite variable models, indicators cause or form the latent construct (e.g.
-education and income together make up socioeconomic status).
-Some scholars have argued that measures of adverse childhood experiences often should be treated as formative, not reflective, indicators \citepSM{netland2001assessment, layne2010unpacking}.
-That is, exposure to abuse and neglect defines childhood maltreatment, not the other way around.
+In addition to confirmatory factor analysis, we also performed exploratory factor analysis (EFA) on the response data from each sample separately.
+The goal of the EFA was to examine what factor structure would emerge with no \emph{a priori} restrictions placed on the data.
+We decided to extract two- and three- factor solutions based on the total number of factors in the bifactor S-1 models.
+We were particularly interested to see if a two-factor solution would reproduce the factors from model 3 (threat, deprivation); and if a three- factor solution would reproduce the three factors from model 4 (parental maltreatment, peer victimization, reverse-scored).
 
-For the purposes of explaining covariation in maltreatment endorsements on the MACE, we believe we are justified in using a common factor model.
-Most of the items on the MACE measure maltreatment by parents or other guardians.
-The homogeneity of these indicators suggests they might reflect a common source; this hypothesis is further evidenced by the uniformly strong general factor loadings for this subset of items.
-Furthermore, there is ample theoretical and empirical support for regarding parental maltreatment as a common factor with its own causal antecedents \citepSM{belsky1993etiology}.
-Indeed, many of the same parent-level variables (e.g.
-emotional lability, impulse control, own history of abuse) are risk factors for different types of childhood maltreatment \citepSM{stith2009risk, mulder2018risk, assink2019risk}.
-Separately, that the MACE is a retrospective self-report measure is further cause for a common factor model.
-Retrospective measures of abuse have been shown to reflect additional factors beyond maltreatment exposure including individual differences in current emotional state, personality, and recall ability \citepSM{susser2012still, reuben2016lest, colman2016consistency}.
-Common factor models may better adjust for some of the measurement error in indicators caused by these nuisance factors.
+EFA was performed using the \textit{lavaan} (v0.6.11; \citealt{lavaan}) package available in R.
+Factors were extracted using the oblique geomin \citep{yates1987multivariate} and cf-quartimax rotation criteria \citep{crawford1970general}, which were selected in order to extract simple factor structures with few cross-loadings.
+We used two rotation criteria in order to examine the generalizability of the factor solutions across rotations.
+We again relied on traditional fit indices (i.e.
+RMSEA, CFI, TLI) to evaluate the goodness-of-fit of the exploratory factor models to the response data of both samples.
 
-This discussion in turn raises a second question: whether it is reasonable for indicators of parental maltreatment and peer victimization on the MACE to share a common factor.
-One might assume that these two types of maltreatment reflect independent causes, and thus should not be permitted to share a common factor.
-There is considerable evidence, however, that children who have been maltreated are at greater risk for subsequent peer victimization \citepSM{bolger2001developmental, benedini2016cycle}.
-The reasons for this association are manifold.
-For example, parental maltreatment may impact a child's sociocognitive functioning (e.g.
-attachment styles, emotion regulation) in ways that may predispose them to later peer victimization \citepSM{goemans2021child, McCrory2022-cj}.
-Separately, parental neglect is a risk factor for peer victimization \citepSM{lereya2013parenting}, possibly by increasing the chance that children enter into risky social situations where victimization may occur.
-Insofar that parental maltreatment is an indirect cause of peer victimization (mediated by multiple complex psychosocial pathways), then we believe we are justified in allowing indicators of peer victimization to load onto the common general maltreatment factor.
-This is also highlights the need for future study of a MACE peer victimization subscale to better understand how it is related to (and distinct from) parental maltreatment, and what its downstream consequences are.
+The goodness-of-fit measures for the exploratory factor models to the response data by sample are summarized in Table \ref{tab:efa_diagnostics}.
+The fit indices indicated that all exploratory models provided adequate fits to the data.
+The standardized factor loadings for the 2- and 3-factor solutions produced by the cf-quartimax rotation are presented in Figure \ref{fig:efa_cf}.
+The corresponding factor loadings produced by the geomin rotation were qualitatively similar and are presented in Figure \ref{fig:efa_geomin}.
+In the 2-factor solution for the original sample, we observed a primary parental maltreatment factor and a secondary peer victimization factor.
+In contrast, the replication sample produced an unclear factor solution characterized by strong loadings from the everse-scored items (F1) and peer victimization (F2).
+The parental maltreatment items then weakly loaded on both factors.
+Neither 2-factor solution precisely resembled the threat-deprivation factor structure of confirmatory model 3.
+
+The 3-factor solutions for each sample were similar.
+EFA produced a primary parental maltreatment factor with secondary peer victimization and reverse-scored factors.
+Interestingly, items concerning sexual abuse loaded less strongly on the parental maltreatment factor in the replication sample, raising some questions about the replicability of the factor structures.
+Nevertheless, the structure of both the 2- and 3-factor exploratory factor models resembled most closely the bifactor S-1 model with two specific factors (model 4), providing further support for that model of the factor structure of the MACE.
 
@@
     \includegraphics[width=1\textwidth,center]{figures/figS04.png}
+    \caption{\normalfont Standardized factor loadings for the 2- and 3-factor solutions produced by the cf-quartimax rotation for original sample (top row) and replication sample (bottom row).
+\break \small Notes: VA = verbal abuse; PA = physical abuse; NVEA = nonverbal emotional abuse; SA = sexual abuse; EN = emotional neglect; PN = physical neglect; WSV = witnessing sibling violence; WIPV = witnessing inter-parental violence; PeerVA = peer verbal abuse; PeerPA = peer physical abuse.
+Only factor loadings $\geq$ 0.30 are shown.
+** Reverse-scored items.}
+    \label{fig:efa_cf}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=1\textwidth,center]{figures/figS05.png}
     \caption{\normalfont Standardized factor loadings for the 2- and 3-factor solutions produced by the geomin rotation for original sample (top row) and replication sample (bottom row).

### 2022-09-14 analysis/scripts/make_figS04.py

@@ -23,7 +23,7 @@ order = ['VA','PA','NVEA','SA','WSV','WIPV','EN','PN','PeerVA','PeerPA']
 thresh = 0.3
 
 ## Define rotation.
-rotation = 'geomin'
+rotation = 'cf-quartimax'
 
 #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
 ### Load and prepare design matrix.

### 2022-09-14 analysis/scripts/make_figS05.py

@@ -23,7 +23,7 @@ order = ['VA','PA','NVEA','SA','WSV','WIPV','EN','PN','PeerVA','PeerPA']
 thresh = 0.3
 
 ## Define rotation.
-rotation = 'cf-quartimax'
+rotation = 'geomin'
 
 #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#
 ### Load and prepare design matrix.
@@ -110,4 +110,4 @@ for i, study in enumerate(['teicher2015','tuominen2022']):
             cbar.set_label(r'Factor Loadings', fontsize=16, labelpad=10)
             cbar.ax.tick_params(labelsize=11) 
             
-plt.savefig(os.path.join(ROOT_DIR, '..', 'figures', 'fig04.png'), dpi=100)
+plt.savefig(os.path.join(ROOT_DIR, '..', 'figures', 'figS05.png'), dpi=100)

### 2022-09-20 manuscript/main.tex
ready for submission

@@
-\documentclass[letterpaper,man,natbib,noextraspace,floatsintext,longtable]{apa6}
+\documentclass[letterpaper,man,natbib,noextraspace,floatsintext,longtable,12pt]{apa6}
 
@@
 \usepackage{hyperref}             % manage hyperlink colors
+\usepackage{hanging}
 
@@
 
+% define title page
 \title{Confirmatory factor analysis of the Maltreatment and Abuse Chronology of Exposure (MACE) scale: Evidence for essential unidimensionality}
 \shorttitle{Confirmatory factor analysis of the MACE}
-\author{Samuel Zorowitz$^1$, Lauri Tuominen$^{2}$}
+\author{Samuel Zorowitz$^{1*}$, Lauri Tuominen$^{2}$}
 \affiliation{$^1$Princeton Neuroscience Institute, Princeton University, USA\\$^2$The Royal’s Institute of Mental Health Research, University of Ottawa, Canada}
 
+% add author note
+\authornote{
+
+\noindent *Corresponding author: zorowitz@princeton.edu
+
+{\begin{hangparas}{.25in}{1} Acknowledgments: The authors wish to thank Dr. Robyn McQuaid for thoughtful feedback on this manuscript. \end{hangparas}}
+
+{\begin{hangparas}{.25in}{1} \noindent Ethics approval: The Institutional Review Board of the University of Ottawa (\#2021003) approved this study. \end{hangparas}}
+
+{\begin{hangparas}{.25in}{1} \noindent Funding sources: This research was funded by the Royal's Emerging Research Innovators in Mental Health program at the University of Ottawa Institute of Mental Health Research and a National Science Foundation Graduate Research Fellowship.
+\end{hangparas}}
+
+\noindent Conflict of interest: The authors have no conflicts of interest to declare.
+
+}
+
+% add abstract
 \abstract{
@@
 As such, the viability of the MACE total and subscale scores are uncertain.
-The aim of the present study was to investigate the factor structure of the MACE and to quantify the reliability of its total and subscale scores.
 
@@
 
-\noindent \textbf{Methods}: We evaluated a series of confirmatory item factor models fit to the responses of two independent samples of participants (N=1051 \& N=582) who completed the MACE.
-The models included hierarchical (bifactor) models.
+\noindent \textbf{Objectives}: The current study investigated the factor structure of the MACE in order to quantify the reliability of its total and subscale scores.
+
+{\begin{spacing}{1.25} \hfill \\ \end{spacing}}
+
+\noindent \textbf{Participants and setting}: Two independent samples of participants (N=1051 \& N=582) who completed the MACE were included in this study.
+
+{\begin{spacing}{1.25} \hfill \\ \end{spacing}}
+
+\noindent \textbf{Methods}: Using confirmatory item response models, we tested one-factor and several bifactor models of participants' responses.
 We used model-based indices to estimate the reliability of the MACE total and subscale scores, and to quantify the essential unidimensionality of the scale.
@@
 \noindent \textbf{Results}: We found the MACE total score was a reliable and valid measure of overall childhood maltreatment.
-In contrast, although MACE subscale scores exhibited adequate reliability, we found that the vast majority of reliable variance in these scores reflected general maltreatment and not any particular type of maltreatment as intended.
-We also found that the MACE is essentially unidimensional, and exhibits evidence of differential item functioning by gender.
+In contrast, although we found MACE subscale scores exhibited adequate reliability, the vast majority of their reliable variance reflected general maltreatment and not any particular type of maltreatment as intended.
+We also found that the MACE is essentially unidimensional and some of its items exhibit of differential item functioning by gender.
 
@@
 \noindent \textbf{Conclusions}: Our results provide support for a one-factor structure of the MACE, as well as continued use of the MACE total score.
-Our results caution against the use of MACE subscale scores, which have little practical use given that they provide little unique or reliable information above and beyond the total score.
+Our results caution against the use of MACE subscale scores, which have little practical use insofar that they provide little unique, reliable information above and beyond the total score.
 
-\hfill \break
+{\begin{spacing}{1.15} \hfill \\ \end{spacing}}
 
@@
 \keywords{child maltreatment, reliability, item response theory, bifactor model}
-
-\authornote{This research was funded by the Royal's Emerging Research Innovators in Mental Health program at the University of Ottawa Institute of Mental Health Research and a National Science Foundation Graduate Research Fellowship.
-The authors wish to thank Dr.
-Robyn McQuaid for thoughtful feedback on this manuscript.
-The authors have no conflicts of interest to declare.}
 
@@
 peer victimization, witnessing domestic violence).
-Second, the items in the MACE were carefully selected to measure maltreatment of increasing severity in order to finely differentiate people with differing exposure levels.
+Second, using item response theory, the items in the MACE were carefully selected to measure maltreatment experiences of increasing severity in order to finely differentiate people with differing exposure levels.
 Third, the MACE also asks for the onset and duration of maltreatment experiences, which is necessary for investigating how the timing of abuse shapes development.
@@
 Citing a prominent dimensional model of maltreatment \citep{mclaughlin2014childhood}, others have made novel threat and deprivation subscale scores made up of items in the MACE that measure abuse and neglect experiences, respectively \citep{schalinski2018defining, schalinski2019environmental, teicher2018differential}.
-The psychometric properties of these subscale scores have not been thoroughly investigated.
+However, the psychometric properties of these subscale scores have not been thoroughly investigated.
 
@@
 
-To investigate the factor structure of the MACE, we fit a series of four confirmatory item factor models separately to the response data from each sample.\footnote{The suitability of common factor models (CFMs) with reflective indicators for explaining childhood maltreatment data has been contested.
-For interested readers, we justify our use of CFMs to explain responses to the MACE in the supplement.} The basis of each model is the graded response model for ordered polytomous data \citep{samejima1997graded}.
+To investigate the factor structure of the MACE, we fit a series of four confirmatory item factor models separately to the response data from each sample.
+(The suitability of common factor models (CFMs) with reflective indicators for explaining childhood maltreatment data has been contested.
+For interested readers, we justify our use of CFMs to explain responses to the MACE in the supplement.) The basis of each model is the graded response model for ordered polytomous data \citep{samejima1997graded}.
 The factor structure for each model is depicted in Figure \ref{fig:models}.
@@
 
+We identified 12 item pairs and triplets in the MACE with response dependence; that is, sets of items where a response to one item requires or strongly implies a certain response to another.
+These items violate the local independence assumption of item response theory models and, in the present analyses, were therefore collapsed into polytomous items.
+However, future studies might consider abbreviating the MACE by removing some of these dependent or redundant items altogether (possibly without a significant loss of test information).
 We also identified a reverse-scored methods factor in the MACE.
@@
 For a regular item on the MACE, an endorsement indicates that a person experienced an adverse event during one year of childhood at least.
-In contrast, an endorsement on a reverse-scored item indicates that a person experienced experienced the absence of a positive event for all 18 years of childhood.
+In contrast, an endorsement on a reverse-scored item indicates that a person experienced the absence of a positive event for all 18 years of childhood.
 However, alternative causes of the reverse-scored factor (e.g.

### 2022-10-24 manuscript/main.tex
fix typo
resubmission

@@
 % define title page
-\title{Confirmatory factor analysis of the Maltreatment and Abuse Chronology of Exposure (MACE) scale: Evidence for essential unidimensionality}
-\shorttitle{Confirmatory factor analysis of the MACE}
+\title{Examination and replication of the factorial validity and subscale reliability of the Maltreatment and Abuse Chronology of Exposure (MACE) scale}
+\shorttitle{Factorial validity of the MACE}
 \author{Samuel Zorowitz$^{1*}$, Lauri Tuominen$^{2}$}
@@
 
-\noindent \textbf{Background}: The Maltreatment and Abuse Chronology of Exposure (MACE) scale is a retrospective self-report scale that measures 10 distinct types of childhood maltreatment.
-Despite its increasing use, the factor structure of the MACE has not been thoroughly investigated.
-As such, the viability of the MACE total and subscale scores are uncertain.
+\noindent \textbf{Objective}: The Maltreatment and Abuse Chronology of Exposure (MACE) scale is a retrospective self-report measure of childhood maltreatment.
+Despite increasing use, the factor structure of the MACE has not been previously investigated.
+As such, the psycho- metric properties of the MACE total and subscale scores are indeterminate.
+The aims of the current study was to examine the factorial validity of the MACE and the reliability of its total and subscale scores.
 
@@
 
-\noindent \textbf{Objectives}: The current study investigated the factor structure of the MACE in order to quantify the reliability of its total and subscale scores.
+\noindent \textbf{Methods}: We analyzed the response data from two independent samples of participants (N=1051 \& N=582) using confirmatory factor analysis.
+We fit a series of models with increasing dimensionality to explore factorial validity.
+Furthermore, we used model-based indices to estimate the reliability of the MACE total and subscale scores, and to test for essential unidimensionality.
 
@@
 
-\noindent \textbf{Participants and setting}: Two independent samples of participants (N=1051 \& N=582) who completed the MACE were included in this study.
+\noindent \textbf{Results}: Confirmatory factor analysis indicated that a bifactor S-1 model with two specific factors (peer victimization, reverse-scoring) provided the best and most parsimonious fit to the data.
+Despite this, model-based indices suggested the MACE scale is essentially unidimensional.
+We found the MACE total score was a reliable and valid measure of overall childhood maltreatment.
+In contrast, the MACE subscale scores exhibited little reliable variance after controlling for general maltreatment.
 
@@
 
-\noindent \textbf{Methods}: Using confirmatory item response models, we tested one-factor and several bifactor models of participants' responses.
-We used model-based indices to estimate the reliability of the MACE total and subscale scores, and to quantify the essential unidimensionality of the scale.
-
-{\begin{spacing}{1.25} \hfill \\ \end{spacing}}
-
-\noindent \textbf{Results}: We found the MACE total score was a reliable and valid measure of overall childhood maltreatment.
-In contrast, although we found MACE subscale scores exhibited adequate reliability, the vast majority of their reliable variance reflected general maltreatment and not any particular type of maltreatment as intended.
-We also found that the MACE is essentially unidimensional and some of its items exhibit of differential item functioning by gender.
-
-{\begin{spacing}{1.25} \hfill \\ \end{spacing}}
-
-\noindent \textbf{Conclusions}: Our results provide support for a one-factor structure of the MACE, as well as continued use of the MACE total score.
-Our results caution against the use of MACE subscale scores, which have little practical use insofar that they provide little unique, reliable information above and beyond the total score.
+\noindent \textbf{Conclusions}: Our results suggest the MACE is essentially unidimensional and support the continued use of the MACE total score.
+However, our results caution against the use of MACE subscores, which provide little unique or reliable information above and beyond the total score.
 

### 2023-02-18 manuscript/main.tex
abbreviated & submitted

@@
-\documentclass[letterpaper,man,natbib,noextraspace,floatsintext,longtable,12pt]{apa6}
+% set document class
+\documentclass[letterpaper,man,natbib,longtable,floatsintext,12pt]{apa6}
 
+% load packages
 \usepackage[english]{babel}
 \usepackage[utf8x]{inputenc}
+\usepackage{float}
 \usepackage{amsmath}
@@
 % define title page
-\title{Examination and replication of the factorial validity and subscale reliability of the Maltreatment and Abuse Chronology of Exposure (MACE) scale}
-\shorttitle{Factorial validity of the MACE}
+\title{Evaluation of the factor structure and structural validity of the Maltreatment and Abuse Chronology of Exposure (MACE) scale}
+\shorttitle{Structural validity of the MACE}
 \author{Samuel Zorowitz$^{1*}$, Lauri Tuominen$^{2}$}
-\affiliation{$^1$Princeton Neuroscience Institute, Princeton University, USA\\$^2$The Royal’s Institute of Mental Health Research, University of Ottawa, Canada}
+\affiliation{$^1$Princeton Neuroscience Institute, Princeton University, USA\\$^2$The Royal’s Institute of Mental Health Research, University of Ottawa, Canada\\$^*$Corresponding author: zorowitz@princeton.edu}
 
@@
 
-\noindent *Corresponding author: zorowitz@princeton.edu
+{\begin{hangparas}{.25in}{1} \noindent Data availability: All data and analysis materials are publicly available at: \url{https://github.com/szorowi1/mace-irt}. \end{hangparas}}
 
@@
 
-\noindent \textbf{Objective}: The Maltreatment and Abuse Chronology of Exposure (MACE) scale is a retrospective self-report measure of childhood maltreatment.
+\noindent \textbf{Purpose}: The Maltreatment and Abuse Chronology of Exposure (MACE) scale is a retrospective self-report measure of childhood maltreatment.
 Despite increasing use, the factor structure of the MACE has not been previously investigated.
-As such, the psycho- metric properties of the MACE total and subscale scores are indeterminate.
-The aims of the current study was to examine the factorial validity of the MACE and the reliability of its total and subscale scores.
+As such, the validity of the MACE total and subscale scores are unknown.
+The objective of the current study was to evaluate the factor structure of the MACE, including the structural validity of its total and subscale scores.
 
@@
 
-\noindent \textbf{Methods}: We analyzed the response data from two independent samples of participants (N=1051 \& N=582) using confirmatory factor analysis.
-We fit a series of models with increasing dimensionality to explore factorial validity.
-Furthermore, we used model-based indices to estimate the reliability of the MACE total and subscale scores, and to test for essential unidimensionality.
+\noindent \textbf{Methods}: We analyzed the response data from two independent samples of participants who completed the MACE (N=1051 \& N=582) using confirmatory factor analysis.
+We fit a series of models of increasing dimensionality to explore its factor structure.
+We calculated model-based indices to quantify the reliability of the MACE total and subscale scores, and to test for essential unidimensionality.
 
@@
 \noindent \textbf{Results}: Confirmatory factor analysis indicated that a bifactor S-1 model with two specific factors (peer victimization, reverse-scoring) provided the best and most parsimonious fit to the data.
-Despite this, model-based indices suggested the MACE scale is essentially unidimensional.
-We found the MACE total score was a reliable and valid measure of overall childhood maltreatment.
-In contrast, the MACE subscale scores exhibited little reliable variance after controlling for general maltreatment.
+Model-based indices suggested the MACE scale is essentially unidimensional.
+The MACE total score was determined to be a reliable and valid measure of overall childhood maltreatment.
+In contrast, MACE subscale scores exhibited little reliable variance after controlling for overall maltreatment.
 
@@
 
-Childhood maltreatment is an unfortunately common experience for children around the world \citep{stoltenborgh2015prevalence}.
-It is a risk factor for multiple poor outcomes later in life including the development of mental illness \citep{kessler2010childhood}, cognitive impairment \citep{su2019does}, and poor physical health \citep{wegman2009meta}.
-Childhood maltreatment also predicts worse treatment outcomes for several psychiatric disorders \citep{nanni2012childhood, thomas2019childhood}.
-Thus, the reliable and valid assessment of childhood maltreatment is an important research goal in order to refine our understanding of the link between maltreatment and health outcomes; augment treatment strategies for victims of childhood maltreatment; and ultimately design better interventions to prevent maltreatment in the first place.
+Childhood maltreatment is a common experience \citep{stoltenborgh2015prevalence} that increases the risk for mental and physical illness  later in life \citep{kessler2010childhood, wegman2009meta} and predicts worse treatment response for multiple psychiatric disorders \citep{nanni2012childhood, thomas2019childhood}.
+Reliable and valid assessment of childhood maltreatment is an important research goal in order to better understand the link between maltreatment and health outcomes; improve treatment strategies for victims of childhood maltreatment; and design better interventions to prevent maltreatment before it occurs.
 
-There already exist many measures of childhood abuse and maltreatment for researchers to choose from \citep{saini2019systematic}.
+There already exist many measures of childhood abuse and maltreatment for researchers to use \citep{saini2019systematic}.
 One recently developed measure, the Maltreatment and Abuse Chronology of Exposure scale (MACE; \citealt{teicher2015maltreatment}), has a number of notable advantages.
-First, the MACE is a 52-item retrospective self-report scale composed of 10 subscales, each of which measures a unique type of childhood maltreatment.
-The content of six of these subscales overlap with other childhood maltreatment scales (e.g.
-physical abuse, neglect), whereas the remaining four are less commonly represented in other scales (e.g.
-peer victimization, witnessing domestic violence).
-Second, using item response theory, the items in the MACE were carefully selected to measure maltreatment experiences of increasing severity in order to finely differentiate people with differing exposure levels.
-Third, the MACE also asks for the onset and duration of maltreatment experiences, which is necessary for investigating how the timing of abuse shapes development.
-Finally, the MACE was originally developed to provide two measures: a total score (number of overall maltreatment experiences) and a multiplicity score (number of types of maltreatment experienced).
-These scores have excellent temporal stability, convergent validity with other childhood maltreatment scales, and good predictive validity for a multitude of psychiatric symptoms \citep{teicher2015maltreatment}.
-For these reasons, the MACE has been included as among the best measures of childhood maltreatment currently available \citep{saini2019systematic, georgieva2022systematic}.
+First, the MACE measures 10 distinct types of childhood maltreatment, including four types that are seldom included in other scales (e.g., peer victimization, witnessing domestic violence).
+Second, the items in the MACE were selected using item response theory in order to ensure that each subscale assesses maltreatment experiences of increasing severity.
+Third, the MACE also measures the timing of maltreatment experiences, which is necessary for investigating how the chronology of abuse shapes development.
+Finally, the MACE total score (total number of unique maltreatment experiences) exhibits good psychometric properties, including excellent temporal stability, convergent validity with other childhood maltreatment scales, and predictive validity for many psychiatric symptoms \citep{teicher2015maltreatment}.
+For these reasons, the MACE has been appraised as among the best measures of childhood maltreatment currently available \citep{saini2019systematic, georgieva2022systematic}.
 
-In recent years, there have been calls for childhood maltreatment research to move away from the cumulative risk framework \citep{evans2013cumulative} --- which focuses only on the number of adverse experiences --- towards a dimensional approach, which emphasizes how distinct types of maltreatment uniquely shape developmental trajectories \citep{mclaughlin2016beyond, belsky2012beyond}.
-In response, some researchers have begun to use the MACE to make new indices of maltreatment in lieu of, or in addition to, the total and multiplicity scores.
-For example, some researchers have used all 10 MACE subscale scores to study the unique contributions of each type of maltreatment to the development of psychopathology \citep{schalinski2015type, gerke2018childhood, schalinski2019early}.
-Citing a prominent dimensional model of maltreatment \citep{mclaughlin2014childhood}, others have made novel threat and deprivation subscale scores made up of items in the MACE that measure abuse and neglect experiences, respectively \citep{schalinski2018defining, schalinski2019environmental, teicher2018differential}.
-However, the psychometric properties of these subscale scores have not been thoroughly investigated.
+One critical issue for using the MACE in childhood maltreatment research is that its structural validity --- defined as the degree to which an instrument's scores adequately reflect the dimensionality of the construct being measured --- is unknown \citep{saini2019systematic}.
+This issue is pressing as some researchers have begun to use scores from each of the 10 MACE subscales to study the unique contributions of distinct maltreatment types to developmental outcomes (e.g., \citealt{schalinski2015type, gerke2018childhood, schalinski2019early}).
+Others have calculated bespoke subscale scores of threat and deprivation exposure using composites of MACE items that respectively measure abuse and neglect experiences (e.g., \citealt{schalinski2018defining, schalinski2019environmental, teicher2018differential}).
+Because the factor structure of the MACE has not been investigated, the validity of these subscale scores (i.e., the extent to which they reflect particular types of maltreatment as intended) is indeterminate.
 
-A critical challenge in studying the consequences of childhood maltreatment is the high co-occurrence of these adverse experiences.
+Though the MACE measures 10 distinct types of maltreatment, the factor structure that best explains the patterns of responses on the MACE may have fewer than 10 factors given the high co-occurrence of maltreatment experiences.
 That is, children who have experienced any single type of maltreatment are likely to have experienced multiple other types \citep{dong2004interrelatedness, herrenkohl2009assessing, kessler2010childhood}.
-As such, isolating the unique contribution of one type of maltreatment to an outcome is difficult insofar as any one type of adversity is likely confounded with overall maltreatment.
-Concretely, this means that researchers intending to study the sequela of particular types of maltreatment using subscale scores need to verify that those scores actually reflect a construct unique from overall maltreatment, lest they risk drawing spurious conclusions.
+As such, it is possible that MACE subscale scores may primarily reflect overall maltreatment rather than any specific type of maltreatment.
+Bifactor models, a type of hierarchical factor models, are well-suited for resolving this type of concern \citep{bornovalova2020appropriate}.
+Under a bifactor model, covariation in responses to a set of items are accounted for by a \emph{general} factor, reflecting shared variance across all items, and one or more \emph{specific} factors, reflecting additional shared variance among subsets of items \citep{Reise2012-ql}.
+Applied to the MACE, bifactor models can aid in determining whether its subscale scores are interpretable as measures of distinct types of maltreatment, or if they instead chiefly reflect general maltreatment.
 
-Bifactor models, a type of hierarchical factor model, are well-suited for addressing this issue \citep{bornovalova2020appropriate}.
-They assume that the covariation in responses to a set of items can be accounted for by a \emph{general} factor, reflecting shared variance across all items, and one or more \emph{specific} factors, reflecting additional shared variance among subsets of items \citep{Reise2012-ql}.
-Crucially, bifactor models can be used to calculate model-based reliability statistics, including the omega hierarchical ($\omega_h$) and omega subscale ($\omega_s$) indices \citep{reise2013scoring}, which respectively measure the proportion of variance in total and subscale scores attributable to the general factor and specific factors.
-Applied to the MACE, these indices can help determine if subscale scores measuring particular types of maltreatment are interpretable as such, or if they instead principally reflect overall maltreatment.
-In addition, bifactor models permit the calculation of other useful indices like the explained common variance (ECV; \citealt{sijtsma2009use}), which measures what fraction of the total reliable variance in a scale can be attributed to a general factor.
-Applied to the MACE, the ECV can quantify the degree to which the MACE is essentially unidimensional (i.e.
-summarizable by a one-factor model).
-
-The overarching aim of the present study was to investigate the factor structure of the MACE, which has received little previous study to date \citep{saini2019systematic}.
-To do so, we fit and evaluated a series of confirmatory item factor models to the responses from two independent samples of participants who completed the MACE.
-The primary goal of our analyses was to quantify the reliability of the MACE total and subscale scores, and to determine whether the latter would provide unique information above and beyond the total score.
-A secondary goal was to examine the dimensionality of the MACE; that is, to determine whether responses on the MACE are best-explained by a hierarchical structure with 10 factors (one per subscale) or by a more parsimonious factor structure.
-All data and analysis materials are publicly available at: \url{https://github.com/szorowi1/mace-irt}.
+The main objective of the present study was to investigate the structural validity of the MACE.
+To do so, we fit a series of confirmatory factor models to the response data from two independent samples of participants who completed the MACE.
+The first objective of our analyses was to examine the dimensionality of the MACE; that is, to determine whether responses on the MACE are best-explained by a factor structure with 10 factors (one per subscale) or by a more parsimonious factor structure.
+The second objective was to quantify the reliability of, and sources of variability in, the MACE total and subscale scores in order to determine whether each score adequately measured its construct as intended.
 
@@
 The participants in the current study (N=1633 total) belonged to one of two samples.
-The first sample of participants (N=1051) are from the original development of the MACE \citep{teicher2015maltreatment}.
-The full recruitment details for this sample have been reported elsewhere \citep{teicher2015maltreatment}.
-Briefly, these participants were recruited from the greater Boston area between 2010 and 2013.
-Inclusion criteria included being medically healthy, unmedicated, and between 18–25 years of age.
+The first sample of participants (N=1051) are from the original development of the MACE and the details of their recruitment have been reported elsewhere \citep{teicher2015maltreatment}.
+Briefly, inclusion criteria included being medically healthy, unmedicated, and between 18–25 years of age.
 This sample will hereafter be referred to as the \textit{original} sample.
 
-The second sample of participants (N=687) were recruited from the Prolific Academic platform (\url{https://www.prolific.co}) to participate in an online experiment in August -- October, 2021.
+The second sample of participants (N=687) were recruited from Prolific Academic (\url{https://www.prolific.co}) to participate in an experiment in August -- October, 2021.
 Of these, N=582 were retained for analysis (see Exclusion Criteria below).
-Participants were eligible for participation if they were 18 years or older and currently resided in the United States or Canada.
+Participants were eligible for participation if they were 18 years or older and resided in the United States or Canada.
 Participants received monetary compensation for their time (rate: \$12 USD/hr).
@@
 This sample will hereafter be referred to as the \textit{replication} sample.
+
+The demographics of the samples are summarized in Table \ref{tab:demographics}.
+The majority of participants in both samples identified as women, but the original sample was composed of proportionally more woman ($z$ = 4.142, $p$ < 0.001).
+Participants in the original sample were also younger on average ($t$ = -19.313, $p$ < 0.001).
+The majority of participants identified as White and not as Hispanic or Latino.
+The proportion of participants identifying as White was not significantly different across samples ($z = 0.417$, $p = 0.677$), nor was the proportion of participants identifying as Hispanic or Latino ($z$ = -1.241, $p$ = 0.214).
 
@@
     \hspace{1em} Men                       & 381 (36.3\%) & 258 (44.3\%) & \\
-    \hspace{1em} Transgender or NB         &   0  (0.0\%) &   9  (1.5\%) & \\
+    \hspace{1em} Transgender or nonbinary         &   0  (0.0\%) &   9  (1.5\%) & \\
     \hspace{1em} Rather not say            &   0  (0.0\%) &   5  (0.9\%) & \\
@@
 
-The demographics of the two samples are summarized in Table \ref{tab:demographics}.
-The majority of participants in both samples identified as women, though the original sample was composed of proportionally more woman than the replication sample ($z$ = 4.142, $p$ < 0.001).
-Participants in the original sample were also younger on average ($t$ = -19.313, $p$ < 0.001).
-The majority of participants in both samples identified as White and not as Hispanic or Latino.
-The proportion of participants identifying as White was not significantly different across samples ($z = 0.417$, $p = 0.677$), nor was the proportion of participants identifying as Hispanic or Latino ($z$ = -1.241, $p$ = 0.214).
-
 \subsection{Measures}
@@
 All participants completed the 52-item version of the MACE \citep{teicher2015maltreatment}.
-The MACE is a retrospective self-report scale that measures 10 unique types of childhood maltreatment (i.e.
-parental verbal abuse, physical abuse, nonverbal emotional abuse, sexual abuse, emotional neglect, physical neglect, witnessing violence to siblings, witnessing interparental violence, peer verbal abuse, peer physical abuse).
-For each item, participants endorse whether they ever experienced a particular event during childhood (e.g.
-``Parents or guardians intentionally pushed, grabbed, shoved, slapped, pinched, punched or kicked you'') and, if so, at what ages (up to age 18).
-Here we only analyze the dichotomous endorsement responses (Yes = 1, No = 0) and not the chronology data.
-Six of the 52-items --- evenly divided between the emotional and physical neglect subscales --- are positively valenced and reverse-scored (e.g.
-``One or more individuals in your family helped you feel important or special'').
-For these items we follow the scoring convention set by \cite{teicher2015maltreatment}: a score of 1 is assigned to a response only in the absence of a positive event for all 18 years of childhood.
-This is in contrast to all other items, where a score of 1 is assigned if maltreatment was experienced in at least one year of childhood.
+The MACE is a retrospective self-report scale that measures 10 unique types of childhood maltreatment (i.e., parental verbal abuse, physical abuse, nonverbal emotional abuse, sexual abuse, emotional neglect, physical neglect, witnessing violence to siblings, witnessing interparental violence, peer verbal abuse, peer physical abuse).
+For each item, participants endorsed if they ever experienced a particular event during childhood (e.g., ``Parents or guardians intentionally pushed, grabbed, shoved, slapped, pinched, punched or kicked you'') and, if so, at what ages (up to age 18).
+Here we only analyze the dichotomous response data (Yes = 1, No = 0).
+Six of the 52-items --- evenly divided between the emotional and physical neglect subscales --- are reverse-scored (e.g., ``One or more individuals in your family helped you feel important or special'').
+For these items we followed the scoring procedure set by \cite{teicher2015maltreatment}: a score of 1 is assigned to a response only in the absence of a positive event for all 18 years of childhood.
+This is in contrast to all remaining items, where a score of 1 is assigned if maltreatment was experienced in at least one year of childhood.
 
 In both the original and replication samples, participants completed a number of additional self-report symptom measures.
-We do not include these in our analyses and therefore do not discuss these measures further in the main text.
-For completeness, we report these secondary self-report measures in the supplementary materials.
+We did not analyze these data and therefore do not discuss these measures further.
+For completeness, the additional self-report measures that were collected are reported in the Supplementary Materials.
 
@@
 
-To ensure data quality in the replication sample, we relied on attention checks embedded in the self-report measures to identify and remove participants engaging in careless or insufficient effort responding (\citealt{zorowitz2021inattentive}; see the supplementary materials for the attention checks).
-The data from N=64 (9.9\%) participants in the replication sample were excluded for failing one or more of these attention checks, leaving the data from N=582 participants for analysis.
-
-\subsection{Differential item functioning}
-
-Prior to analysis, we investigated whether it was appropriate to combine response data from the original and replication samples.
-We first compared across samples the maltreatment endorsement rates for each item.
-The Spearman-rank correlation of endorsement rates across samples was large ($\rho$ = 0.951, p < 0.001), indicating excellent correspondence in the relative prevalence of maltreatment types across samples.
-When we compared the distribution of person-level total scores across samples, however, we found that scores were greater on average in the replication sample than in the original sample ($p_1$ = 0.185, $p_2$ = 0.219, $t$ = -4.611, p < 0.001).
-
-A mean shift in overall endorsement rates by sample is not necessarily evidence for a violation of measurement invariance.
-One possibility is that the replication sample had simply experienced more adversity than the original sample.
-A second is that the differences in the demographic composition of the two samples (e.g.
-gender, age) explain the differences in endorsement rates.
-To investigate these possibilities, we tested for uniform differential item functioning (DIF) using logistic regression, in which a participant's response on an item is regressed against their rest score (i.e.
-their total score for all other items), sample membership, gender, and age.
-
-Using the cutoffs recommended by \cite{hidalgo2014binary}, we identified 11 of 52 items on the MACE with large DIF by study (Figure \ref{fig:dif_study}).
-These included all six reverse-scored items.
-We also found an additional 11 items with large DIF by gender (Figure \ref{fig:dif_gender}).
-Five of these items concerned sexual abuse (endorsed more frequently by women), and another five concerned peer physical abuse (endorsed more frequently by men).
-We identified no items that exhibited large DIF by age.
-Given the number of items with DIF by sample membership, we decided to analyze each sample separately.
-
-\subsection{Item response dependence}
-
-Also prior to analysis, we found that the MACE contains a total of 12 item pairs and triplets characterized by response dependence (Table \ref{tab:dependence}).
-The items in some of these sets are structurally dependent, where a response to one item requires or strongly implies a certain response on the second.
-For example, a person endorsing item 8 (``Parents or guardians hit you so hard that you received medical attention'') is all but certain to endorse item 7 (``Parents or guardians hit you so hard that it left marks for more than a few minutes'').
-Other items are dependent by virtue of redundancy.
-One such example is items 36 (``Peers forced you to engage in sexual activity against your will'') and 37 (``Peers forced you to do things sexually that you did not want to do``); here the same question is asked twice using slightly different language.
-
-Structurally-dependent or redundant items can cause serious problems for item factor models.
-First, they can inflate the estimates of item discrimination parameters, making items appear more reliable than they actually are \citep{marais2008formalizing}.
-Second, response dependence can cause overestimation of item loadings on specific factors in bifactor models, making specific factors (and subscores) appear more reliable than they actually are \citep{reise2013applying}.
-Following \cite{marais2008formalizing}, we therefore combined each set of dependent items into a single polytomous item thereby reducing the MACE from 52 to 39 items in total.
+To ensure data quality in the replication sample, we relied on attention checks embedded in the self-report measures to identify and remove participants engaging in inattentive responding (\citealt{zorowitz2021inattentive}; see the Supplementary Materials for details).
+The data from N=64 (9.9\%) participants were excluded for failing one or more attention checks, leaving the data from N=582 participants for analysis.
 
@@
 
-To investigate the factor structure of the MACE, we fit a series of four confirmatory item factor models separately to the response data from each sample.
-(The suitability of common factor models (CFMs) with reflective indicators for explaining childhood maltreatment data has been contested.
-For interested readers, we justify our use of CFMs to explain responses to the MACE in the supplement.) The basis of each model is the graded response model for ordered polytomous data \citep{samejima1997graded}.
+To investigate the factor structure of the MACE, we tested four confirmatory item factor models.
+The basis of each model is the graded response model for ordered polytomous data \citep{samejima1997graded}.
 The factor structure for each model is depicted in Figure \ref{fig:models}.
@@
 The second is a bifactor model where all items load onto a general maltreatment factor and also one of 10 specific factors based on their subscale membership (Figure \ref{fig:models}b).
-This model will be used to evaluate if the MACE can produce reliable subscale scores for each of the 10 types of maltreatment assessed by the MACE.
+We used this model to evaluate the reliability of the MACE subscale scores.
 
-The third model was motivated by a dimensional theory of childhood maltreatment that distinguishes between threat and deprivation experiences as two distinct types of adversity \citep{mclaughlin2014childhood}.
-To examine this model of the MACE, we initially fit a bifactor model where all items loaded onto a general maltreatment factor, and also a threat or neglect specific factor.
-However, we observed vanishing item loadings on the threat specific factor.
-Therefore, we instead fit a bifactor S-1 model \citep{eid2017anomalous} where all items load onto a general maltreatment factor, and only items measuring neglect (i.e.
-items from the physical and emotional neglect subscales) load onto a specific neglect factor (Figure \ref{fig:models}c).
-We fit this model to see if the MACE can be used to form neglect subscale scores that are reliable and distinguishable from general maltreatment.
+The third model was motivated by previous research that used the MACE to calculate bespoke subscale scores measuring threat and deprivation experiences (e.g., \citealt{teicher2018differential}).
+To evaluate those scores, we initially fit a bifactor model where all items loaded onto a general maltreatment factor, and also a threat or neglect specific factor.
+After observing vanishing loadings on the threat specific factor, however, we instead fit a bifactor S-1 model \citep{eid2017anomalous} where all items loaded onto a general maltreatment factor, and only items measuring neglect load onto a specific neglect factor (Figure \ref{fig:models}c).
+We used this model to evaluate the reliability of the neglect subscale scores.
 
@@
 
-The fourth and final model we fit proposes a more parsimonious structure of the MACE based on the content of its items.
-Items on the MACE can be categorized according to the source of maltreatment: parents (or other adults in the home) and peers.
-Thus, we hypothesized there would be two factors underlying responses on the MACE: a primary parental maltreatment factor and a secondary peer victimization factor.
-From preliminary model fits, we also suspected the presence of a tertiary methods factor affecting the six reverse-scored items.
-We initially fit a bifactor model in which all items loaded onto a general maltreatment factor and one of three specific factors (parental maltreatment, peer victimization, reverse-scored); however, we observed vanishing loadings on the first specific factor.
-We instead fit a bifactor S-1 model where all items load onto a general maltreatment factor, and the peer victimization \& reverse-scored items additionally load onto their own specific factors (Figure \ref{fig:models}d).
-We fit this model in part to test a more parsimonious factor structure of the MACE, and to see if the MACE can be used to form peer victimization subscale scores that are reliable and distinguishable from general maltreatment.
+The fourth model offers a more parsimonious structure of the MACE where items are organized according to the source of maltreatment: parents (or other adults in the home) and peers.
+That is, this model represents the hypothesis that two factors are sufficient to summarize responses on the MACE: a primary parental maltreatment factor and a secondary peer victimization factor.
+Based on preliminary model fits, we suspected the presence of a tertiary methods factor affecting the six reverse-scored items.
+We initially fit a bifactor model in which all items loaded onto a general maltreatment factor and one of three specific factors (i.e., parental maltreatment, peer victimization, reverse-scoring).
+After observing vanishing loadings on the first specific factor, however, we instead fit a bifactor S-1 model where all items load onto a general maltreatment factor, and the peer victimization and reverse-scored items additionally load onto their own specific factors (Figure \ref{fig:models}d).
+We fit this model to test a more parsimonious factor structure of the MACE, and to evaluate the reliability of peer victimization subscale scores.
+
+We fit each model separately to the data from the original and replication samples.
+We did so because, prior to analysis, we detected uniform differential item functioning (DIF) by sample for 11 items (see the Supplementary Materials for details).
+Also prior to analysis, we identified a total of 12 pairs and triplets of MACE items with response dependence.
+For example, a person endorsing item 8 (``Parents or guardians hit you so hard that you received medical attention'') is all but certain to endorse item 7 (``Parents or guardians hit you so hard that it left marks for more than a few minutes'').
+To prevent bias in the estimation of factor loadings due to response dependence \citep{reise2013applying}, we followed best-practice recommendations and combined each set of dependent items into a single polytomous item \citep{marais2008formalizing}.
+Thus, we reduced the MACE from 52 to 39 independent items (Table \ref{tab:dependence}).
 
@@
 We relied on multiple indices to judge the fit of the models to the data.
-First we calculated three traditional goodness-of-fit statistics using the ordinal $M_2$ statistic \citep{cai2013limited}.
-$M_2$ is asymptotically $\chi^2$-distributed with $\kappa − \nu$ degrees of freedom ($df$), where $\kappa$ is the reduced number of first- and second-order marginal residuals (i.e.
-$\kappa = n(n + 1)/2$ for $n$ items) and $\nu$ denotes the number of free parameters.
-To improve the stability of the $M_2$ statistic, the responses to polytomous items were binarized (i.e.
-any vs.
-no maltreatment).
-Using the $M_2$ statistic, we calculated the root mean square error of approximation (RMSEA), comparative fit index (CFI), and Tucker-Lewis index (TLI) for each model fit.
-Following convention \citep{hu1999cutoff}, the benchmarks for adequate model fit are RMSEA < 0.08, CFI > 0.90, and TLI > 0.90; in turn, the benchmarks for good model fit are RMSEA < 0.05, CFI > 0.95, and TLI > 0.95.
+First we calculated three traditional goodness-of-fit statistics based on the ordinal $M_2$ statistic \citep{cai2013limited}.
+Specifically, we calculated the root mean square error of approximation (RMSEA), comparative fit index (CFI), and Tucker-Lewis index (TLI) for each model fit.
+Following convention \citep{hu1999cutoff}, we defined the benchmarks for adequate model fit as RMSEA < 0.08, CFI > 0.90, and TLI > 0.90, and the benchmarks for good model fit as RMSEA < 0.05, CFI > 0.95, and TLI > 0.95.
 
-Given the limitations of SEM-based fit indices for item response models \citep{clark2018model}, we calculated two additional fit indices.
+Given the limitations of SEM-based fit indices for item response models (e.g., \citealt{clark2018model}), we calculated two additional fit indices.
 First, we used posterior predictive model checking to calculate a $\chi^2_{NC}$ discrepancy measure based on the total score distribution \citep{sinharay2006posterior}.
 This measure compares the observed and model-predicted proportion of participants at each total score level.
-Second, to test for local dependence in each pair of items, we calculated Yen's $Q_3$ statistic \citep{yen1984effects} and also the critical $Q_3$-value (i.e.
-the value above which local dependence is indicated; \citealt{christensen2017critical}).
-For each model and sample, we simulated 1000 locally independent datasets using the posterior distribution of the model parameters.
+Second, to test for local dependence in each pair of items, we calculated Yen's $Q_3$ statistic \citep{yen1984effects} and also the critical $Q_3$-value (i.e., the value above which local dependence is indicated; \citealt{christensen2017critical}).
+Briefly, for each model and sample, we simulated 1000 locally-independent datasets using the posterior distribution of model parameters.
 We then recorded the maximum $Q_3$ value per simulation.
@@
 
-The goodness-of-fit of the models was compared using Bayesian leave-one out cross- validation (LOO-CV; \citealt{vehtari2017practical}).
-The LOO-CV measure quantifies the discrepancy between the model and data while taking into account model complexity.
-LOO-CV values are presented in deviance scale, i.e.
-smaller values indicate better fit.
+The goodness-of-fit of the models was compared using Bayesian leave-one-out cross-validation (LOO-CV; \citealt{vehtari2017practical}).
+LOO-CV quantifies the discrepancy between the model and data while taking into account model complexity.
+LOO-CV values are reported here in deviance scale (i.e., smaller values indicate better fit).
 
@@
 
-To evaluate the reliability of the MACE total and subscale scores, we calculated several model-based reliability indices using the factor loadings from the bifactor models.
-First, we calculated coefficient omega ($\omega$; \citealt{mcdonald2013test}), which quantifies the proportion of reliable score variance attributable to all factors (i.e.
-both general and specific factors).
+To evaluate the MACE total and subscale scores, we calculated several model-based reliability indices.
+First, we calculated coefficient omega ($\omega$; \citealt{mcdonald2013test}), which quantifies the proportion of reliable score variance attributable to all factors (i.e., both general and specific).
 Next we calculated coefficient omega hierarchical ($\omega_h$) and coefficient omega subscale ($\omega_s$;  \citealt{reise2013scoring}).
-Whereas $\omega_h$ measures the fraction of total score variance attributable to the general factor, $\omega_s$ measures the fraction of subscale score variance attributable to a specific factor after controlling for the general factor.
-The values of $\omega_h$/$\omega_s$ facilitate interpretation of total and subscale scores.
-Values of $\omega_h$ approaching 1 indicate the general factor is the primary source of reliable variance in a total score.
-When $\omega_h$ > 0.80, the total score may be considered an essentially unidimensional reflection of the general factor \citep{rodriguez2016applying}.
-Values of $\omega_s$ approaching 1 indicate the specific factor, not the general factor, is the primary source of reliable variance in a subscale score.
+Whereas $\omega_h$ measures the fraction of total score variance attributable to the general factor, $\omega_s$ measures the fraction of subscale score variance attributable to a specific factor.
+Larger values of $\omega_h$ indicate the general factor is the primary source of reliable variance in a total score.
+When $\omega_h > 0.80$, the total score is an essentially unidimensional reflection of the general factor \citep{rodriguez2016applying}.
+In turn, larger values of $\omega_s$ indicate the specific factor is the primary source of reliable variance in a subscale score.
 \cite{canivez2016bifactor} suggested that an acceptable $\omega_s$ value is 0.50 and that >0.70 is desirable.
-In contrast, values of $\omega_s$ < 0.50 preclude the interpretation of subscale scores as primarily reflecting a specific factor \citep{gignac2013bifactor}.
+When $\omega_s < 0.50$, a subscale score instead primarily reflects the general factor \citep{gignac2013bifactor}.
 
-To assess the unidimensionality of the MACE, we calculated the explained common variance (ECV; \citealt{sijtsma2009use}) and the proportion of uncontaminated correlations (PUC; \citealt{reise2013multidimensionality}).
-The ECV is the ratio of the variance explained by the general factor divided by the total common variance of a scale.
-It thus measures the contribution of the general factor relative to the specific factors.
-As ECV values approach 1, the general factor loadings of a bifactor model are expected to resemble the item loadings that would be obtained by a one-dimensional model.
-ECV values greater than 0.70 often indicate a scale is essentially unidimensional \citep{rodriguez2016applying}.
+To evaluate the dimensionality of the MACE, we calculated the explained common variance (ECV; \citealt{sijtsma2009use}) and the proportion of uncontaminated correlations (PUC; \citealt{reise2013multidimensionality}).
+The ECV is the ratio of the variance explained by the general factor divided by the total common variance.
+It measures the contribution of the general factor relative to the specific factors.
+Larger ECV values indicate the general factor loadings of a bifactor model should resemble the loadings that would be obtained under a unidimensional model.
+When ECV > 0.70, a scale is essentially unidimensional \citep{rodriguez2016applying}.
 In turn, PUC quantifies how many inter-item correlations are accounted for only by the general factor.
-PUC is an important moderator of the ECV; as PUC increases, ECV is less important for evaluating the risk of bias when fitting a unidimensional model to data with a latent bifactor structure.
-When PUC > 0.80, there is a low risk of bias when a multi- dimensional scale is treated as unidimensional \citep{reise2013multidimensionality}.
+When PUC > 0.80, there is low risk of bias when a multidimensional scale is treated as unidimensional \citep{reise2013multidimensionality}.
 We also calculated the relative parameter bias (RPB), which is the difference between an item's loading in a unidimensional model and its corresponding general factor loading in a bifactor model, divided by the general factor loading from the bifactor model.
@@
 
-Finally, we calculated the H index to assess the construct replicability of each factor \citep{hancock2001rethinking}.
-The H index is an estimate of how well a set of items represents a latent factor.
-H values closer to 1 indicate a well-defined latent factor with item loadings more likely to be stable across studies.
-In contrast, small H values suggest a poorly defined latent factor where item loadings are liable to change across studies.
-Because we are presently focused on evaluating the reliability of MACE total and subscale scores, we report the H values below but do not interpret them further.
+We also calculated the H index \citep{hancock2001rethinking}, which measures how well a set of items represents a latent factor.
+Larger H values indicate a well-defined latent factor with item loadings more likely to be stable across studies.
+Because we are presently focused on evaluating the MACE total and subscale scores, we report H values but do not interpret them further.
 
@@
 
-In addition to confirmatory factor analysis, we also performed exploratory factor analysis (EFA) on the response data for each sample separately.
-The purpose for performing EFA after confirmatory factor analysis was to examine whether any of the hypothesized factor structures emerged from the data with fewer \textit{a priori} restrictions \citep{schmitt2018selecting}.
-For the sake of brevity, the details and results of the EFA are included in the supplement.
+In addition to the confirmatory factor models, we also performed exploratory factor analysis (EFA) on each sample's data separately.
+The motivation for performing EFA was to examine whether any of the confirmatory factor structures emerged from the data with fewer \textit{a priori} restrictions \citep{schmitt2018selecting}.
+The details and results of the EFA are included in the Supplementary Materials.
 
@@
 
-The goodness-of-fit indices for each model and sample are summarized in Table \ref{table:cfa_diagnostics}.
+The goodness-of-fit indices for each sample and model are summarized in Table \ref{table:cfa_diagnostics}.
 According to the traditional indices, each model provided at least an acceptable fit to the data.
 All RMSEA value were <0.05 and most CFI and TLI values were >0.95.
-The only exception were for the fits of models 1 and 3 to the original sample data where CFI \& TLI > 0.90.
-In general, model fits were better for the replication sample data than for the original sample.
-Furthermore, none of the posterior predictive p-values corresponding to the $\chi^2_{NC}$ discrepancy measure exceeded the critical values indicating that all models were able to reproduce the observed distribution of total scores for each sample.
+The only exception were for the fits of models 1 and 3 to the original sample data where CFI and TLI > 0.90.
+Furthermore, none of the posterior predictive $p$-values from the $\chi^2_{NC}$ discrepancy measure exceeded the critical values, indicating that all models were able to reproduce the observed total score distribution for each sample.
+
+Next we inspected the $Q_3$ indices for evidence of local dependence.
+For all models and samples, the maximum observed $Q_3$ index was greater than the critical $Q_3$ value.
+To measure the degree of local dependence, we visualized the $Q_3$ values for item pairs exceeding the critical value for each model and sample (Figure \ref{fig:local_dependence}).
+The unidimensional model fits exhibited many locally dependent item pairs, especially among the peer victimization and reverse-scored items, suggesting underfactorization.
+By comparison, the three bifactor models exhibited fewer locally-dependent item pairs.
+The cause for dependence among these item pairs were often self-evident.
+For example, the residual dependence between items 17 (``Parents made inappropriate sexual comments or suggestions to your sibling'') and 18 (``Parents touched or fondled your sibling in a sexual way'') is likely because these items measure similar content.
+Regardless, the estimated item discrimination parameters for each model fit seldom exceeded the normal range ($\alpha \leq 4$; \citealt{edwards2018diagnostic}); that is, the degree of local-dependence observed here did not result in substantial bias in parameter estimation.
+Thus, we conclude that all model fits provide at least an acceptable fit to the response data.
+
+The results of the model comparison are also summarized in Table \ref{table:cfa_diagnostics}.
+Across samples, Models 2 and 4 provided better fits to the data than models 1 and 3.
+Model 2 provided a numerically better fit to the data than Model 4 in the original sample ($\Delta \text{LOO}$ = 7.9, se = 56.3), but the opposite was observed in the replication sample ($\Delta \text{LOO}$ = -24.0, se = 52.0).
+These differences in LOO-CV values, however, are both within four standard errors of the mean indicating only weak predictive improvements \citep{vehtari2022cv}.
+The results of the model comparison therefore suggest that Models 2 and 4 provide approximately equivalent fits to the data.
+This is noteworthy insofar that Model 4 assumes a substantially simpler factor structure of the MACE.
 
@@
 \caption{\normalfont Fit statistics for the confirmatory factor models by sample.
-LOO-CV values are presented in deviance scale (i.e.
-smaller values indicate better fit).
+LOO-CV values are presented in deviance scale (i.e., smaller values indicate better fit).
 Notes: RMSEA = root mean square error of approximation; CFI = comparative fit index; TLI = Tucker-Lewis index; PPP = posterior predictive p-value; $Q3_{\max}$ = maximum $Q_3$ value in the sample for the model; $Q3_{\text{crit}}$ = critical $Q_3$ value for the model; LOO-CV = leave-one-out cross-validation.}
@@
 
-Next we inspected the $Q_3$ indices for evidence of local dependence.
-For all models and samples, the maximum observed $Q_3$ index was greater than the critical $Q_3$ value.
-To gauge the extent of local dependence, we visualized the $Q_3$ values for item pairs exceeding the critical value for each model and sample (Figure \ref{fig:local_dependence}).
-The unidimensional model fits exhibited many locally dependent item pairs --- especially among the peer victimization and reverse-scored items --- suggesting possible underfactorization.
-By comparison, the remaining three models exhibited many fewer locally-dependent item pairs.
-Of these pairs, the reasons for dependence were apparent.
-For example, the residual dependence between items 17 (``Parents made inappropriate sexual comments or suggestions to your sibling'') and 18 (``Parents touched or fondled your sibling in a sexual way'') certainly stems from the fact that these items measure highly similar content.
-Regardless, the estimated item discrimination parameters for each model fit seldom exceeded the normal range ($\alpha \leq 4$).
-That is, the level of local dependence present in the models did not cause substantial bias in parameter estimation \citep{edwards2018diagnostic}.
-Thus, we conclude that all model fits provide at least an acceptable fit to the response data.
-
-The results of the model comparison are also summarized in Table \ref{table:cfa_diagnostics}.
-Across samples, models 2 and 4 provided better fits to the data than models 1 and 3.
-Model 2 provided a numerically better fit to the data than model 4 in the original sample ($\Delta \text{LOO}$ = 7.9, se = 56.3), but the opposite was observed in the replication sample ($\Delta \text{LOO}$ = -24.0, se = 52.0).
-These differences in LOO-CV values, however, are both within four standard errors of the mean indicating only weak predictive improvements \citep{vehtari2022cv}.
-Thus the results suggest that models 2 and 4 provide approximately equal fits to the data, which is notable given that model 4 assumes a substantially simpler factor structure of the MACE.
-
 \subsection{Confirmatory item response models}
 
-In the following sections, we interpret the fit of each confirmatory item response model.
+In the following, we interpret the fit of each confirmatory item response model.
 To facilitate interpretation, we have reproduced the standardized factor loadings of each model for the original sample in Figure \ref{fig:loadings_original} and for the replication sample in Figure \ref{fig:loadings_online}.
@@
 \subsubsection{Model 1: One-factor (unidimensional) model}
+
+The standardized factor loadings of the unidimensional model were, on average, large in magnitude.
+In the original sample, factor loadings ranged between 0.279 and 0.864 with an average value of 0.597 (95\% HDI = 0.579 -- 0.614); in the replication sample, factor loadings ranged between 0.359 and 0.814 with an average value of 0.600 (95\% HDI = 0.578 -- 0.621).
+Most items had loadings greater than 0.50 (original: 80.0\%, replication: 80.0\%), and therefore appear to be good indicators of general maltreatment.
+The rank correlation of general factor loadings between samples was $\rho$ = 0.565 (95\% HDI = 0.414 -- 0.711), indicating moderate agreement between samples.
+
+In sum, the fits of the one-factor models provide partial evidence in support of a unidimensional model of the MACE.
+For example, most items exhibited moderate-to-large loadings on the unitary factor.
+However, the degree of local dependence present among in the unidimensional model fits (especially among the peer victimization items) suggest there may be secondary dimensions to the MACE.
+We turn next to the multidimensional confirmatory models to explore this possibility.
 
@@
 
-The item discrimination parameters of the unidimensional model were, on average, large for both the original sample (mean $\alpha$ = 1.357, 95\% HDI = 1.295 -- 1.414) and replication sample (mean $\alpha$ = 1.349, 95\% HDI = 1.272 -- 1.422).
-The standardized factor loadings were correspondingly large in magnitude.
-In the original sample, factor loadings ranged between 0.279 and 0.864 with an average value of 0.597 (95\% HDI = 0.579 -- 0.614); in the replication sample, factor loadings ranged between 0.359 and 0.814 with an average value of 0.600 (95\% HDI = 0.578 -- 0.621).
-Most items had loadings greater than 0.50 (original: 80.0\%, replication: 80.0\%), and therefore appear to be good indicators of general maltreatment.
-The rank correlation of general factor loadings between samples was $\rho$ = 0.565 (95\% HDI = 0.414 -- 0.711), indicating only moderate agreement between samples.
-
-As an initial characterization the dimensionality of the MACE, we inspected the first six eigenvalues of the polychoric item correlation matrix for each sample.
-In the original sample, these values were 14.62, 3.77, 2.64, 2.11, 1.79, and 1.35; in the replication sample, they were 13.27, 4.43, 2.48, 2.04, 1.81, and 1.56.
-The first-to-second eigenvalue ratios of 3.84 and 3.02 are just above the 3-to-1 criterion often cited in determining whether item response data are essentially unidimensional \citep{embretson2013item}.
-
-In sum, the fits of the one-factor models provide partial evidence in support of a unidimensional model of the MACE.
-Virtually all items exhibited moderate-to-large loadings onto the common factor.
-Moreover, the eigenvalues of the polychoric correlation matrix suggested the presence of one primary factor.
-However, the degree of local dependence present among in the unidimensional model fits (especially among the peer victimization items) suggest there may be additional dimensions to the MACE.
-We turn next to the multidimensional confirmatory models to explore these possibilities.
-
 \subsubsection{Model 2: Bifactor model with 10 specific factors}
 
-Next we inspect the bifactor model with 10 specific factors, i.e.
-one per subscale on the MACE (model 2).
+Next we inspect the bifactor model with 10 specific factors (i.e., one per MACE subscale).
 In the original sample, the loadings on the general maltreatment factor ranged between 0.235 and 0.864 with an average value of 0.575 (95\% HDI = 0.556 -- 0.593).
 In the replication sample, the general factor loadings ranged between 0.299 and 0.827 with an average value of 0.580 (95\% HDI = 0.560 -- 0.601).
-The relative parameter bias was marginal for both samples (original = 2.1\%; replication = 2.7\%), suggesting there was little bias in the factor loadings of the one-factor models due to multidimensionality.
+The relative parameter bias was marginal for both samples (original = 2.1\%; replication = 2.7\%), indicating an acceptable level of bias in the loadings of the one-factor models due to multidimensionality.
 
-In contrast, the loadings on the specific factors were smaller and seldom exceeded loadings on the general factor.
+In contrast, the loadings on the specific factors were small on average and rarely exceeded their corresponding loadings on the general factor.
 In the original dataset, the specific factor loadings ranged between 0.014 to 0.797 with an average value of 0.313 (95\% HDI = 0.285 -- 0.341).
 In the replication dataset, the range was between 0.043 and 0.698, with an average value of 0.343 (95\% HDI = 0.315 -- 0.369).
-Only items from the two peer abuse subscales had specific factor loadings of equal or greater magnitude to those on the general factor.
+Only items from the peer abuse subscales had specific factor loadings of greater magnitude to those on the general factor.
 
@@
 
-We now consider the reliability of the MACE total score.
-We found the MACE total score had excellent reliability (original: $\omega$ = 0.963; replication: $\omega$ = 0.965).
+The MACE total score had excellent reliability (original: $\omega$ = 0.963; replication: $\omega$ = 0.965).
 The $\omega_h$ value was 0.924 for both samples indicating that 96.0\% and 95.8\% of the reliable variance in the total score is attributable to general maltreatment.
 Importantly, the ECV values were 0.716 and 0.697 for the original and replication samples, respectively.
-In conjunction with a PUC value of 0.912, the ECV values suggest the MACE is essentially unidimensional despite the multidimensionality of its contents.
-Together, the results support the interpretation of the MACE total score is a reliable and valid measure of overall maltreatment.
+Together with a PUC = 0.912, the ECV values suggest the MACE is essentially unidimensional.
+In sum, the results indicate the MACE total score is a reliable and structurally valid measure of overall maltreatment.
 
-Next we consider the reliability of the MACE subscale scores, where the results are in direct contrast to the MACE total score.
-As would be expected given the lower number of items, the reliability of the subscales was both lower and more variable.
+In turn, the reliability of the MACE subscale scores were smaller and more variable.
 The $\omega$ values ranged between 0.590 and 0.909 in the original sample, and between 0.595 and 0.893 in the replication sample.
-Most subscale scores exhibited at least adequate reliability (>0.7).
-Crucially, subscale scores for both samples contained little reliable variance after controlling for the general maltreatment factor (original: mean $\omega_s$ = 0.181, range = 0.027 -- 0.621; replication: mean $\omega_s$ = 0.205, range = 0.037 -- 0.565).
-That is, the reliability of MACE subscale scores primarily reflect general maltreatment and not any specific type of maltreatment.
-The only exception was for the peer verbal abuse subscale, where more than half the reliable variance in subscale scores was independent of general maltreatment (original = 70.7\%; replication = 66.2\%).
+Most subscale scores exhibited at least adequate reliability ($\omega > 0.7$).
+Crucially, subscale scores contained little reliable variance after controlling for the general maltreatment factor (original: mean $\omega_s$ = 0.181, range = 0.027 -- 0.621; replication: mean $\omega_s$ = 0.205, range = 0.037 -- 0.565).
+That is, MACE subscale scores primarily reflect general maltreatment and not specific types of maltreatment.
+The only exception was for the peer verbal abuse subscale, where the majority of reliable variance in subscale scores was unique from general maltreatment (original = 70.7\%; replication = 66.2\%).
 
-To summarize, the results of the full bifactor model (model 2) indicate responses on the MACE are explained by a dominant general maltreatment factor characterized by large loadings across items.
-In contrast, specific factors representing each MACE subscale are characterized by mostly small factor loadings that contribute little reliable variance to total and subscale scores.
-Where the MACE total score is a reliable and valid measure of overall maltreatment, the majority of MACE subscale scores contain little reliable variance unique from general maltreatment.
-Therefore, they should not be interpreted as reflecting specific types of maltreatment.
+To summarize, the fits of the full bifactor model (Model 2) indicate responses on the MACE are largely explained by a general maltreatment factor with large loadings across items.
+In contrast, the specific factors corresponding to each MACE subscale were mostly characterized by small factor loadings.
+Whereas the MACE total score is a reliable and structurally valid measure of overall maltreatment, the majority of MACE subscale scores contain little reliable variance unique from general maltreatment.
+Thus, they cannot be said to reflect specific types of maltreatment.
 One question that remains, however, is if other reliable subscale scores can be derived from the MACE.
@@
 
-We turn our attention next to the bifactor S-1 model with one specific factor (neglect; model 3).
-The loadings on the general factor largely resembled those of the bifactor model.
+Next we inspect the bifactor S-1 model with one specific neglect factor (Model 3).
+The loadings on the general factor largely resembled those from Model 2.
 In the original sample, the loadings on the general maltreatment factor ranged between 0.150 and 0.864 with an average value of 0.577 (95\% HDI = 0.558 -- 0.594).
@@
 In the original sample, the average specific loading was 0.410 (95\% HDI = 0.373 -- 0.454); in the replication sample, it was 0.454 (95\% HDI = 0.409 -- 0.499).
-The specific loadings displayed a troubling pattern where only the six reverse-scored items exhibited moderate loadings on the specific factor; the remaining four item exhibited negligible factor loadings.
-That is, the ``neglect'' factor appears instead to reflect a reverse-scored methods factor.
+Troublingly, only the six reverse-scored items exhibited moderate loadings on the specific factor; the remaining four items exhibited negligible factor loadings.
+In other words, the ``neglect'' factor appears instead to reflect a reverse-scored methods factor.
 
-Moving on to the reliability measures, we again found that the total score exhibited excellent reliability (original: $\omega$ = 0.958; replication: $\omega$ = 0.959).
+The MACE total score exhibited excellent reliability (original: $\omega$ = 0.958; replication: $\omega$ = 0.959).
 Moreover, the $\omega_h$ values indicated the majority of reliable variance in severity score was attributable to the general factor (original: $\omega_h$ = 0.928; replication: $\omega_h$ = 0.922).
 The reliability of the ``neglect'' subscale score was also good (original: $\omega$ = 0.877; replication: $\omega$ = 0.921).
-As indicated by the $\omega_s$ values, however, reliable variance in the ``neglect'' subscale scores primarily reflected the general maltreatment factor (original: $\omega_h$ = 0.384; replication: $\omega_h$ = 0.375).
+As indicated by the $\omega_s$ values, however, reliable variance in the ``neglect'' subscale scores principally reflected the general maltreatment factor and not its intended construct (original: $\omega_h$ = 0.384; replication: $\omega_h$ = 0.375).
 
-In sum, the results do not support the use of the MACE to compute separate threat and neglect subscale scores.
-The neglect subscale scores posses an unsatisfactory amount of reliable variance unique from general maltreatment; therefore, they cannot be interpreted as meaningfully reflecting neglect.
-Furthermore, the majority of items measuring neglect on the MACE seem to be contaminated by a reverse-scored methods factor.
+In sum, the fits of Model 3 to the data do not support using the MACE to calculate independent threat and neglect subscale scores.
+As previously defined, the neglect subscale scores contain little reliable variance unique from general maltreatment and cannot, therefore, be said to measure neglect.
+Furthermore, most MACE items purportedly measuring neglect seem instead to reflect a reverse-scored methods artifact.
 
@@
 
-Finally, we turn to the bifactor S-1 model with two specific factors (peer victimization, reverse-scoring; model 4).
+Finally, we inspect the bifactor S-1 model with two specific factors (peer victimization, reverse-scoring; Model 4).
 In the original sample, the loadings on the general maltreatment factor ranged between 0.158 and 0.864 with an average value of 0.563 (95\% HDI = 0.545 -- 0.580).
 In the replication sample, the general factor loadings ranged between 0.298 and 0.852 with an average value of 0.572 (95\% HDI = 0.550 -- 0.593).
-Interestingly, the loadings on the specific factors were of approximately equal magnitude.
+The loadings on the specific factors were similar in magnitude to those on the general factor.
 In the original dataset, the specific factor loadings ranged 0.327 to 0.789 with an average value of 0.550 (95\% HDI = 0.522 -- 0.576).
@@
 
-As before, the MACE total score again exhibited excellent reliability (original: $\omega$ = 0.961; replication: $\omega$ = 0.961).
+The MACE total score again exhibited excellent reliability (original: $\omega$ = 0.961; replication: $\omega$ = 0.961).
 Similarly, the majority of reliable variance in severity score was attributable to the general factor (original: $\omega_h$ = 0.896; replication: $\omega_h$ = 0.904).
 The peer victimization subscale scores exhibited good reliability across samples (original: $\omega$ = 0.889; replication: $\omega$ = 0.868).
-The corresponding $\omega_s$ values were 0.569 for the original sample and 0.493 for the replication sample indicating that slightly more than half of the variance in peer victimization scores were independent of the general maltreatment factor (original: 64.0\%; replication: 56.8\%).
-Thus, the results suggest (albeit weakly so) the presence a secondary peer victimization factor in the MACE in addition to the general maltreatment factor.
+The corresponding $\omega_s$ values were 0.569 for the original sample and 0.493 for the replication sample indicating that the majority of variance in peer victimization scores were unique from the general maltreatment factor (original: 64.0\%; replication: 56.8\%).
+Thus, the results suggest (albeit weakly so) a secondary peer victimization factor in the MACE in addition to the general maltreatment factor.
 
@@
 Moreover, the corresponding $\omega_s$ values were similar in magnitude to those observed for the peer victimization scores (original: $\omega_s$ = 0.584; replication: $\omega_s$ = 0.498).
-In other words, these scores also exhibit majority reliable variance unique from general maltreatment.
-It is difficult to interpret these scores, however, as it is unclear what they primarily reflect (e.g.
-neglect, wording effects, and/or scoring-induced methods artifact).
+These scores therefore also exhibit majority reliable variance unique from general maltreatment.
+It is difficult to interpret these scores, however, as it is unclear what they primarily reflect (e.g., neglect, wording effects, reverse-scored methods artifact).
 
@@
 
-The MACE is a promising measure of childhood maltreatment \citep{georgieva2022systematic}.
-Despite its increasing popularity, there has been little investigation of its structural validity \citep{saini2019systematic}, leaving unanswered questions about its factor structure and the interpretation of its total \& subscale scores.
-Here we investigated the factor structure of the MACE, using a series confirmatory item response models, in two independent samples.
-The primary goal of our analyses was to evaluate the reliability of the MACE total and subscale scores, and to determine whether the latter would provide unique information above and beyond the total score.
-A secondary goal was to examine the dimensionality of the MACE; that is, whether its inter-item covariance was best-explained by a hierarchical structure with 10 factors (one per subscale) or by a more parsimonious structure.
+The MACE is a promising and increasingly popular measure of childhood maltreatment \citep{georgieva2022systematic}.
+Despite this, there has been no investigation into its factor structure \citep{saini2019systematic}, leaving unanswered questions about its structural validity and the interpretation of its total and subscale scores.
+Here we investigated the factor structure of the MACE through a series of confirmatory item response models.
+Our objectives were to evaluate the dimensionality of the MACE, and to quantify the degree and sources of reliable variance in its total and subscale scores.
 
-Regarding our first goal, we found the MACE total score has excellent reliability and overwhelmingly reflects general maltreatment.
-That is, despite being a composite of items measuring 10 distinct types of maltreatment, the MACE total score is a univocal index of overall childhood maltreatment.
-In stark contrast, although we found the MACE subscale scores exhibited adequate reliability on average, we also found that these scores mostly reflected general maltreatment and not any particular type of maltreatment as intended.
-The same was true of neglect subscale scores (composed of the physical and emotional neglect subscales), which were additionally contaminated by a reverse-scored methods factor.
-In short, our results support the continued use of the MACE total score but caution against the use of MACE subscale scores.
-Indeed the latter have little practical use given the low reliability of the portion of subscale scores independent of the total score.
+Of the confirmatory factor models we tested, the most parsimonious and best-fitting was the bifactor S-1 model with a general maltreatment factor and two smaller specific factors reflecting peer victimization and reverse item scoring (Model 4).
+This model provided a fit to the data equivalent to that for the more complex bifactor model with a specific factor for each MACE subscale.
+(This model was partially supported by EFA; see the Supplementary Materials for details.) Moreover, across all the models tested, a general maltreatment factor explained at least 70\% of the reliable variance in participants' responses.
+Together, the results support the conclusion that, while the content of the MACE is multidimensional, its factor structure is essentially unidimensional.
 
-Regarding our second goal, confirmatory bifactor models of the MACE revealed a dominant general maltreatment factor, characterized by uniformly strong loadings across items, and much weaker specific factors.
-Importantly, general factor loadings from the bifactor models strongly resembled their corresponding loadings from a one-factor model.
-Moreover, the general factor explained approximately 70\% of the reliable variance in participants' responses.
-Together, these results support the conclusion that, while the item content of the MACE is multidimensional, its factor structure is essentially unidimensional.
-In other words, researchers can safely interpret the MACE total score as an index of overall maltreatment; they can also expect little bias in factor loading estimates if they were to specify a one-factor model to explain responses on the MACE.
+We also found the MACE total score has excellent reliability.
+Given the essential unidimensionality of the MACE, the total score can be regarded as a valid and univocal index of overall childhood maltreatment.
+In stark contrast, we found that the majority of reliable variance for most MACE subscale scores was attributable to the general maltreatment and not specific types of maltreatment as intended.
+The same was true of the neglect subscale scores, which were additionally contaminated by a reverse-scored methods artifact.
+In sum, our results support the continued use of the MACE total score but caution against the use of MACE subscale scores.
+Indeed, the latter are unreliable measures of specific maltreatment types and instead predominantly reflect overall maltreatment.
+The only exception was the composite peer victimization subscale scores where the majority of reliable variance was attributable to the peer victimization factor and not general maltreatment.
+Regardless, additional research is needed to validate these subscores before they should be used in childhood maltreatment research.
 
-Of the four confirmatory factor models we tested, the most parsimonious and best- fitting was a bifactor S-1 model with a general maltreatment factor and two specific factors measuring peer victimization and reverse-scoring.
-This model provided a fit to the response data as good as the more complex full bifactor model with 10 specific factors.
-(This bifactor S-1 model was also partially supported by exploratory factor analyses; see the supplement for details.) Thus, our results suggest the factor structure of the MACE is essentially unidimensional with two secondary factors affecting a subset of items.
-Notably, we found that the scores from the peer victimization subscale (composed of items from the peer physical and emotional abuse subscales) were viable.
-That is, they were reliable and, after controlling for general maltreatment, a majority of their reliable variance was attributable to peer victimization.
-Researchers may therefore be justified in using peer victimization subscale scores from the MACE, but additional research is needed to validate this index.
+Our results are consistent with previous studies that investigated the factor structure of other childhood maltreatment questionnaires.
+For example, bifactor models have revealed dominant general maltreatment factors and weaker specific factors for the Childhood Trauma Questionnaire \citep{spinhoven2014childhood, stagaki2022mediating}, International Child Abuse Screening Tool \citep{meinck2021factor}, and Adverse Childhood Experience questionnaire \citep{dobson2021latent}.
+Together, these and our results underscore a central challenge in childhood maltreatment research: when maltreatment experiences tend to co-occur, it is difficult to isolate the unique contribution of a particular type of adversity from overall maltreatment.
 
-Our results are consistent with a number of previous studies investigating the factor structure of other childhood maltreatment questionnaires.
-For example, bifactor models of the Childhood Trauma Questionnaire have also revealed a dominant general maltreatment factor and weaker specific factors \citep{spinhoven2014childhood, stagaki2022mediating}.
-Similar results were found in recent applications of a bifactor model to the International Child Abuse Screening Tool \citep{meinck2021factor} and the Adverse Childhood Experience questionnaire \citep{dobson2021latent}.
-Together, these and our current results underscore a challenge in investigating the consequences of childhood maltreatment: when maltreatment experiences tend to co-occur, it is difficult to measure the effects of a particular type of adversity unique from overall maltreatment.
-
-Our results are also worth considering alongside previous studies that used MACE subscale scores to investigate the link between particular types of childhood maltreatment and assorted psychobiological outcomes.
+Our results are also worth considering alongside previous studies that used the MACE subscale scores to investigate the link between particular types of childhood maltreatment and assorted psychobiological outcomes.
 For example, researchers have looked at the relationship between MACE subscale scores and risk of depression \citep{gerke2018childhood}, dissociative symptoms, \citep{schalinski2015type}, and cortisol concentration \citep{schalinski2019early}.
-Still others have used threat and deprivation subscale scores from the MACE to study their associations with cognitive functioning \citep{schalinski2018defining} and hippocampal volume \citep{teicher2018differential}.
-Our results are clear that these scores chiefly reflect general maltreatment and not particular types of adversity, which presents a challenge for interpreting the conclusions of those studies.
-It is worth noting that some of these studies attempted to control for general maltreatment by entering all of the MACE subscale scores simultaneously into sophisticated multivariate analyses (e.g.
-random forest regression).
-This approach is unsatisfactory for two reasons.
-First, random forest models can produce spurious results when given collinear predictors \citep{gregorutti2017correlation}.
-Second, even if one were to properly control for general maltreatment, our results indicate that the remaining reliable variance in most subscale scores would be so low as to cause doubt about the meaning of any subsequent inferences.
+Still others have used bespoke threat and deprivation subscale scores from the MACE to study their associations with cognitive functioning \citep{schalinski2018defining} and hippocampal volume \citep{teicher2018differential}.
+Our results are unequivocal that these subscale scores mostly reflect general maltreatment and not particular types of adversity, which poses a challenge for interpreting the results of those studies.
 
 The current study revealed other features of the MACE worth highlighting.
-For example, we identified several sources of DIF in the MACE.
-We found that across samples women experienced more sexual abuse whereas men experienced more physical abuse by peers.
-These findings are in line with previous research \citep{radford2013prevalence}.
-We also identified a number of items with DIF by sample.
-The cause for this is unclear and may reflect differences in sample population (greater Boston area vs.
-US \& Canada), study location (in clinic vs.
-online), recruitment years (early 2010s vs.
-early 2020s), inclusion criteria (healthy \& unmedicated vs.
-none), and/or exclusion criteria (none vs.
-attention checks).
-The causes and impact of violations of measurement invariance on the MACE require further study.
-
-We identified 12 item pairs and triplets in the MACE with response dependence; that is, sets of items where a response to one item requires or strongly implies a certain response to another.
-These items violate the local independence assumption of item response theory models and, in the present analyses, were therefore collapsed into polytomous items.
-However, future studies might consider abbreviating the MACE by removing some of these dependent or redundant items altogether (possibly without a significant loss of test information).
-We also identified a reverse-scored methods factor in the MACE.
-In our view, the most plausible explanation is that, under the \cite{teicher2015maltreatment} scoring procedure, affirmative responses on the regular and reverse-scored items on the MACE represent very different endorsements.
-For a regular item on the MACE, an endorsement indicates that a person experienced an adverse event during one year of childhood at least.
+For example, we identified a number of items with DIF by sample.
+The cause for this is unclear and may reflect differences in sample populations, study locations, recruitment years, or inclusion/exclusion criteria.
+The causes and consequences of measurement noninvariance on the MACE deserve further study.
+We also identified a reverse-scoring methods factor in the MACE.
+The most plausible explanation for this is that, under the \cite{teicher2015maltreatment} scoring procedure, affirmative responses to the regular- and reverse-scored items on the MACE constitute very different answers.
+For a regular item on the MACE, an endorsement indicates that a person experienced an adverse event during at least one year of childhood.
 In contrast, an endorsement on a reverse-scored item indicates that a person experienced the absence of a positive event for all 18 years of childhood.
-However, alternative causes of the reverse-scored factor (e.g.
-participant confusion, acquiescence, or carelessness; \citealt{weijters2013reversed}) are possible.
-Additional research is needed to see if alternative scoring procedures would eliminate this factor.
+However, alternative causes of this methods factor are plausible (e.g., participant confusion, acquiescence, or carelessness; \citealt{weijters2013reversed}).
+Further research is needed to see if alternative scoring procedures would mitigate this artifact.
 
 The current study is not without limitations.
-The most apparent limitation is that we only studied the dichotomous response data, but not the chronology data that is also collected as part of the MACE.
-Though we found that the factor structure of the endorsement responses is essentially unidimensional, this may belie more complex patterns present in the maltreatment chronology data.
-Indeed, longitudinal models of adverse events across childhood --- perhaps using dynamic factor analysis \citep{zhang2007bayesian} --- might be more sensitive to patterns of maltreatment covariation that are lost when collapsing the data into a dichotomous indicator of none vs.
-any maltreatment.
+The most apparent limitation is that we only studied the dichotomous response data but not the chronology data that is also collected as part of the MACE.
+Though we found that the factor structure of the dichotomous responses is essentially unidimensional, this may belie more complex patterns present in the chronology data.
+Indeed, longitudinal models of adverse events across childhood (e.g., using dynamic factor analysis; \citealt{zhang2007bayesian}) might be more sensitive to patterns of maltreatment covariation that are lost when collapsing the data into simple yes/no responses.
 Another possibility is to jointly model the MACE endorsement and chronology data using a multidimensional hurdle model \citep{magnus2021symptom}, which measures both a person's susceptibility to maltreatment (number of maltreatment experiences) and severity of maltreatment (duration of maltreatment experiences).
-Further research is needed to develop maltreatment scores that integrate the MACE endorsement and chronology data.
+Further research is needed to develop maltreatment scores that jointly model the MACE endorsement and chronology data.
 
 A second limitation is our sample demographics.
-Though the replication sample increased the overall diversity of our sample with respect to gender and age, other types of diversity are notably lacking.
+Though the replication sample increased the overall diversity of our sample with respect to gender and age, other types of diversity were notably lacking.
 Our combined sample was neither racially nor ethnically diverse.
-This is important as previous studies have identified DIF for childhood adversity questionnaires by race \citep{rodriguez2019identification}, which raises the question whether similar issues would be found for the MACE in more diverse samples.
-Moreover, we are not currently able to explain why we found DIF in multiple items by sample membership in the current study.
-Additional research is needed to study the functioning of the MACE in more diverse samples in order to validate its use in different populations.
+This is important as previous studies have identified DIF by race for other childhood adversity questionnaires \citep{rodriguez2019identification}.
+Additional research is needed to study the functioning of the MACE in more diverse samples in order to validate its use in those populations.
 
 To conclude, the results of the current study show the factor structure of the MACE is essentially unidimensional.
-Bifactor models revealed that, despite the presence of smaller secondary factors, the majority of reliable variance in the MACE endorsement responses are attributable to a general maltreatment factor.
-Accordingly, we found that the MACE total score is a reliable and valid measure of overall childhood maltreatment.
-In contrast, we found that MACE subscale scores are unreliable measures of their intended constructs; that is, they primarily reflect general maltreatment and not any specific type of maltreatment.
-Moving forward, childhood maltreatment researchers using the MACE (or any other scale) should be careful to ensure that any total or subscale scores they use as part of their analyses are reliable, lest they risk drawing spurious inferences.
+Confirmatory bifactor models revealed that, despite the presence of smaller secondary factors, the majority of reliable variance in responses on the MACE is attributable to a general maltreatment factor.
+Accordingly, the MACE total score is a reliable and valid measure of overall childhood maltreatment.
+In contrast, the MACE subscale scores are invalid measures of their intended constructs; that is, they primarily reflect general maltreatment and not any specific type of maltreatment.
+Moving forward, we caution against the use of MACE subscale scores lest future childhood maltreatment researchers risk drawing invalid inferences.
 
@@
 
-One might question whether it is appropriate to explain childhood maltreatment data using common factor models with reflective indicators (e.g.
-bifactor models).
-The core assumption of these models is that variation in item responses are caused by or reflect variation in a shared latent factor (e.g.
-variation in low mood and anergia reflect latent individual differences in depression).
+One might question whether it is appropriate to explain childhood maltreatment data using common factor models with reflective indicators (e.g., bifactor models).
+The core assumption of these models is that variation in item responses are caused by or reflect variation in a shared latent factor (e.g., variation in low mood and anergia reflect latent individual differences in depression).
 This is in contrast to composite variable models with formative indicators where the direction of causality is reversed \citepSM{bollen1991conventional}.
-In composite variable models, indicators cause or form the latent construct (e.g.
-education and income together make up socioeconomic status).
+In composite variable models, indicators cause or form the latent construct (e.g., education and income together make up socioeconomic status).
 Some scholars have argued that measures of adverse childhood experiences should often be treated as formative, not reflective, indicators \citepSM{netland2001assessment, layne2010unpacking}.
@@
 Furthermore, there is ample theoretical and empirical support for regarding parental maltreatment as a common factor with its own causal antecedents \citepSM{belsky1993etiology}.
-Indeed, many of the same parent-level variables (e.g.
-emotional lability, impulse control, own history of abuse) are risk factors for different types of childhood maltreatment \citepSM{stith2009risk, mulder2018risk, assink2019risk}.
+Indeed, many of the same parent-level variables (e.g., emotional lability, impulse control, own history of abuse) are risk factors for different types of childhood maltreatment \citepSM{stith2009risk, mulder2018risk, assink2019risk}.
 
@@
 The reasons for this association are manifold.
-For example, parental maltreatment may impact a child's sociocognitive functioning (e.g.
-attachment styles, emotion regulation) in ways that may predispose them to later peer victimization \citepSM{goemans2021child, McCrory2022-cj}.
+For example, parental maltreatment may impact a child's sociocognitive functioning (e.g., attachment styles, emotion regulation) in ways that may predispose them to later peer victimization \citepSM{goemans2021child, McCrory2022-cj}.
 Separately, parental neglect is a risk factor for peer victimization \citepSM{lereya2013parenting}, possibly by increasing the chance that children enter into risky social situations where victimization may occur.
@@
 
-In addition to the MACE, participants in both samples completed a number of additional self-report measures.
-Participants in the original sample completed multiple measures concerning psychiatric symptoms; these have already been reported \citepSM{teicher2015maltreatment}.
+In addition to the MACE, participants in both the original and replication samples completed a number of additional self-report measures.
+Participants in the original sample completed multiple measures of psychiatric symptoms, which have been reported elsewhere \citepSM{teicher2015maltreatment}.
 The participants in the replication sample completed three additional self-report measures assessing negative symptoms and motivation for rewards.
@@
 
+\subsection*{Differential item functioning}
+
+Prior to analysis, we investigated whether it was appropriate to combine response data from the original and replication samples.
+We first compared across samples the maltreatment endorsement rates for each item.
+The Spearman-rank correlation of endorsement rates across samples was large ($\rho$ = 0.951, p < 0.001), indicating excellent correspondence in the relative prevalence of maltreatment types across samples.
+When we compared the distribution of person-level total scores across samples, however, we found that scores were greater on average in the replication sample than in the original sample ($p_1$ = 0.185, $p_2$ = 0.219, $t$ = -4.611, p < 0.001).
+
+A mean shift in overall endorsement rates by sample is not necessarily evidence for a violation of measurement invariance.
+One possibility is that the replication sample had simply experienced more adversity than the original sample.
+A second is that the differences in the demographic composition of the two samples (e.g., gender, age) explain the differences in endorsement rates.
+To investigate these possibilities, we tested for uniform differential item functioning (DIF) using logistic regression, in which a participant's response on an item is regressed against their rest score (i.e., their total score for all other items), sample membership, gender, and age.
+
+Using the cutoffs recommended by \citeSM{hidalgo2014binary}, we identified 11 of 52 items on the MACE with large DIF by study (Figure \ref{fig:dif_study}).
+These included all six reverse-scored items.
+We also found an additional 11 items with large DIF by gender (Figure \ref{fig:dif_gender}).
+Five of these items concerned sexual abuse (endorsed more frequently by women), and another five concerned peer physical abuse (endorsed more frequently by men).
+These findings are in line with previous research \citepSM{radford2013prevalence}.
+We identified no items that exhibited large DIF by age.
+Given the number of items with DIF by sample membership, we decided to analyze each sample separately.
+
+\subsection*{Item response dependence}
+
+Also prior to analysis, we found that the MACE contains a total of 12 item pairs and triplets characterized by response dependence (Table \ref{tab:dependence}).
+The items in some of these sets are structurally dependent, where a response to one item requires or strongly implies a certain response on the second.
+For example, a person endorsing item 8 (``Parents or guardians hit you so hard that you received medical attention'') is all but certain to endorse item 7 (``Parents or guardians hit you so hard that it left marks for more than a few minutes'').
+Other items are dependent by virtue of redundancy.
+One such example is items 36 (``Peers forced you to engage in sexual activity against your will'') and 37 (``Peers forced you to do things sexually that you did not want to do``); here the same question is asked twice using slightly different language.
+
+Structurally-dependent or redundant items can cause serious problems for item factor models.
+First, they can inflate the estimates of item discrimination parameters, making items appear more reliable than they actually are \citepSM{marais2008formalizing}.
+Second, response dependence can cause overestimation of item loadings on specific factors in bifactor models, making specific factors (and subscores) appear more reliable than they actually are \citepSM{reise2013applying}.
+Following \citepSM{marais2008formalizing}, we therefore combined each set of dependent items into a single polytomous item thereby reducing the MACE from 52 to 39 items in total.
+
 \subsection{Exploratory factor analysis}
@@
 We used two rotation criteria in order to examine the generalizability of the factor solutions across rotations.
-We again relied on traditional fit indices (i.e.
-RMSEA, CFI, TLI) to evaluate the goodness-of-fit of the exploratory factor models to the response data of both samples.
+We again relied on traditional fit indices (i.e., RMSEA, CFI, TLI) to evaluate the goodness-of-fit of the exploratory factor models to the response data of both samples.
 
@@
 \pagebreak
-\subsection*{Supplementary tables \& figures}
 
@@
 
+\pagebreak
+
 \begin{table}[H]
@@
 
+\pagebreak
+
 \begin{figure}[H]
@@
     \caption{\normalfont Differential item functioning by sample.
-(A) Proportions of participants endorsing having experienced a maltreatment event by study sample.
-(B) The log odds that the proportion of endorsements are larger in one or the other sample.
+(A) Proportions of participants that endorsed experiencing a maltreatment event by study sample.
+(B) The log-odds that the proportion of endorsements are larger in one group.
 Dotted lines indicate the threshold for large DIF recommended by \cite{hidalgo2014binary}.
@@
     \caption{\normalfont Differential item functioning by gender.
-(A) Proportions of participants endorsing having experienced a maltreatment event by study gender.
-(B) The log odds that the proportion of endorsements are larger in one or the other gender.
+(A) Proportions of participants that endorsed experiencing a maltreatment event by gender.
+(B) The log-odds that the proportion of endorsements are larger in one group.
 Dotted lines indicate the threshold for large DIF recommended by \cite{hidalgo2014binary}.
@@
 \end{figure}
+
+\pagebreak
 
@@
 
+\pagebreak
+
 \bibliographystyleSM{apalike}
